### 2024-08-04 documentation/arxiv/arxiv.tex
add overleaf files

\documentclass{article}

\usepackage{amsfonts, amsmath, graphicx}
\usepackage[margin=1in]{geometry}
\usepackage{natbib}
\bibliographystyle{apa-good}

\title{The Two-Foot Rule: A game theoretic analysis of the pickoff limit in Major League Baseball}

\begin{document}

  \maketitle

  \section{Introduction}

    In baseball, the stolen base is an exciting play in which a runner will take off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out. Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \citep{tango_book_2007}. On the micro level, it improves a team's win probability to reduce stolen base attempts that have negative expected outcomes, but on the macro level, the reduction in stolen base attempts affects the entertainment value of the game.

    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting \citep{castrovince_pitch_2023}. With these changes, a pitch clock limits the time between pitches, and there are restrictions on infielder positioning, curtailing defensive shifts. Two changes are particularly relevant to stolen base attempts: Bases are now wider (reducing the distance between first and second base by 4.5 inches), and there are now count limitations on pickoff attempts, a previously unlimited deterrent of stolen bases.

    Whether attempting a stolen base or not, runners typically take a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, in the direction of the next base. The longer the lead, the shorter the distance need to run on a stolen base attempt. On a pickoff attempt, instead of throwing a pitch, the pitcher will disengage from the pitching rubber and throw the ball behind the runner to a fielder covering the base, attempting to tag the runner out. Under the new rules, if the pitcher disengages from the pitching rubber three times without successfully picking off the runner, then the runner automatically advances to the next base.

    This disengagement limit creates an interesting cat-and-mouse game between the pitcher and the runner. Each unsuccessful pickoff attempt disincentivizes the pitcher from attempting another pickoff, which incentivizes the runner to increase their lead distance. But how much should the runner extend their lead? And how does the answer change depending on runner speed and other context?

    In the present work, we conduct the first public analysis of this game. We model the baseball inning as a Markov decision process and solve for (a) the optimal policy for runner lead distance given current pitcher pickoff behavior and (b) the game theoretic equilibrium behavior in the two-player game between runner and pitcher. We present the practically actionable rule-of-thumb that an average runner should increase their lead distance by two feet after each disengagement.

    \subsection{Related work}

    \cite{downey_pick_2015} previously analyzed pickoff throws and stolen base attempts as a mixed strategy game without using lead distance data. They modeled the two-player game as a simultaneous decision in which the pitcher chooses whether to attempt a pickoff and the runner chooses whether to attempt a stolen base, showing that this theoretical game has a mixed-strategy Nash equilibrium. Under this equilibrium, left-handed pitchers (possessing better pickoff moves than their right-handed counterparts) would attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data. In a later analysis, \cite{downey_pressure_2019} argued that pitchers can effectively implement a mixed strategy in low-pressure situations but struggle to randomize their decisions as pressure increases, showing a higher auto-correlation for sequential pickoff decisions in high-pressure situations. Both of these analyses covered a game in which pickoff attempts were unlimited.

    Several studies in sports economics literature have examined the extent to which professional athletes play minimax strategy. \cite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save. \cite{palacios-huerta_experientia_2008} followed this with a laboratory experiment featuring professional soccer players and college students simulating the penalty kick decision using cards, showing that the professionals played minimax but the students did not. On the other hand, \cite{kovash_professionals_2009} found evidence inconsistent with minimax equilibrium behavior in the National Football League (teams called more run plays and fewer pass plays than theoretically optimal) and in MLB (pitchers threw more fastballs than theoretically optimal).

    The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model \citep{bukiet_markov_1997}. \cite{hirotsu_using_2002} used a four-state Markov model for soccer to inform substitution decision-making based on how substitutions impact transition probabilities between states. \cite{hirotsu_markov_2003} used similar methodology to inform substitution (pinch hitting) decision-making in baseball. Recently, \cite{chan_points_2021} used a Markov model to evaluate team performance in football. More specifically, the Markov decision process has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \citep{nadimpalli_when_2013}, baseball \citep{hirotsu_using_2019}, basketball \citep{sandholtz_markov_2020}, darts \citep{baird_optimising_2020} and football \citep{biro_reinforcement_2022}.

    An intermediate result of the present work is a probability model for stolen base attempts and successes (as well as an analysis how those probabilities changed with the 2023 MLB rule changes), a topic which has not received much attention. \cite{loughin_assessing_2008} estimated linear mixed-effects models for stolen base attempts and successes, using random effects for pitcher and catcher (not the runner, notably) and fixed effects for balls, strikes, outs, score, venue and pitcher hand. They showed that the spread of pitcher effects is slightly greater than the spread of catcher effects. That analysis did not have data on runner sprint speed and catcher arm strength. More recently, \cite{stanley_modeling_2023} modeled stolen base success probability using both logistic regression and a random forest, including runner sprint speed and catcher arm strength as features but excluding player identities. That analysis was limited to data prior to the 2023 MLB rule changes that incentivized base stealing.

  \section{Data}

    \subsection{MLB Stats API}

    \subsection{Baseball Savant}

    \subsection{Proprietary Statcast Data}

  \section{Methods}

    \subsection{Notation}

    Throughout this section, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
    \begin{align*}
      b_i & \in \{0, 1\}^3 \mbox{ denotes which bases (1B, 2B, 3B, respectively) are occupied;}\\
      (c_i^B, c_i^S) \equiv c_i & \in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ denotes the ball-strike count;}\\
      d_i & \in \{0, 1, 2\} \mbox{ denotes the number of disengagements already used; and}\\
      o_i & \in \{0, 1, 2\} \mbox{ denotes the number of outs.}
    \end{align*}
    We use $\mathcal{H}^R$, $\mathcal{H}^P$, and $\mathcal{H}^C$ to denote the sets of runners, pitchers and catchers, respectively. On play $i$:
    \begin{align*}
      h_i^R & \in \mathcal{H}^R \mbox{ is the runner;}\\
      h_i^P & \in \mathcal{H}^P \mbox{ is the pitcher;}\\
      h_i^C & \in \mathcal{H}^C \mbox{ is the catcher;}\\
      z_i^R & \in \mathbb{R^+} \mbox{ is the sprint speed of runner $h_i^R$; and}\\
      z_i^C & \in \mathbb{R^+} \mbox{ is the arm strength of catcher $h_i^C$.}
    \end{align*}
    When the play begins, we use the following notation to describe the actions taken by the players:
    \begin{align}
    \label{eqn:runner-outcome}
      \begin{split}
        \ell_i  \in &~ \mathbb{R}^+ \mbox{ denotes the lead distance (in feet) taken by the first-base runner;}\\
        p_i \in &~ \{0, 1\} \mbox{ denotes whether the pitcher attempts (1) or does not attempt (0) a pickoff; and}\\
        r_i \in &~ \{\mbox{P}^+,\, \mbox{P}^-,\, \mbox{S}^+,\, \mbox{S}^-,\, \mbox{N}\} \equiv \mathcal{R} \mbox{ denotes the runner outcome, where}\\
          & \mbox{P}^+  \mbox{ represents a successful pickoff attempt (runner is out);}\\
          & \mbox{P}^-  \mbox{ represents an unsuccessful pickoff attempt (runner is safe);}\\
          & \mbox{S}^+  \mbox{ represents a successful stolen base attempt (runner is safe);}\\
          & \mbox{S}^-  \mbox{ represents an usuccessful stolen base attempt (runner is out); and}\\
          & \mbox{N}~   \mbox{ represents no runner action (a pitch with no stolen base attempt).}
%        p_i & \in \{0, 1\} \mbox{ indicates whether the pitcher attempts a pickoff on the first-base runner; and}\\
%        s_i & \in \{0, 1\} \mbox{ indicates whether the first-base runner attempts a stolen base.}
      \end{split}
    \end{align}
    If there is no runner on first base, then $\ell_i$ and $r_i$ are undefined.% We use a superscript to denote the success/failure of pickoff and stolen base attempts: 
%    \begin{align*}
%      p_i^+ & \in \{0, 1\} \mbox{ denotes whether the pickoff attempt is successful (i.e. the runner is tagged out); and}\\
%      s_i^+ & \in \{0, 1\} \mbox{ denotes whether the stolen base attempt is successful.}
%    \end{align*}

    \subsection{Probability model for runner outcomes}
    \label{sec:prob-runner-outcome}

      As itemized in (\ref{eqn:runner-outcome}), there are five possible outcomes for the runner with respect to the run game. To estimate each of these probabilities conditional on player identities, count and lead distance, we model four probabilities. Using $R_i$ to denote the random variable governing the probability distribution over $r_i$,
      \begin{align*}
        \pi_i    &= \mathbb{P}(R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff attempt probability;}\\
        \pi_i^+  &= \mathbb{P}(R_i \in \{\mbox{P}^+\} \mid R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff success probability;}\\
        \psi_i   &= \mathbb{P}(R_i \in \{\mbox{S}^+,\, \mbox{S}^-\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-, N\}) \mbox{ is the stolen base attempt probability; and}\\
        \psi_i^+ &= \mathbb{P}(R_i \in \{\mbox{S}^+\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-\}) \mbox{ is the stolen base success probability.}
      \end{align*}
      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \citep{bates_fitting_2015}.

      \subsubsection{Pickoff attempt probability}
      \label{sec:prob-po-attempt}

        We model $\pi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), and lead distance; and with a random effect for the pitcher.
        \begin{align}
          \label{eqn:prob-po-attempt}
          \begin{split}
            \log\left(\frac{\pi_i}{1 - \pi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + \beta^L \ell_i + \gamma^P_{h_i^P}\\
            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
          \end{split}
        \end{align}
        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$ and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$. The number of random effects is $|\mathcal{H}^R| + |\mathcal{H}^P|$.

      \subsubsection{Pickoff success probability}
      \label{sec:prob-po-success}

        We model $\pi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher. These were the only variables for which we found substantive effects.
        \begin{align}
          \label{eqn:prob-po-success}
          \begin{split}
            \log\left(\frac{\pi_i^+}{1 - \pi_i^+}\right) &= \alpha + \beta^L \ell_i + \gamma^P_{h_i^P},\\
            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
          \end{split}
        \end{align}
        This model has three fixed, unknown parameters: the intercept $\alpha$, the slope $\beta^L$, and the variance parameters $\sigma^2_P$. The number of random effects is $|\mathcal{H}^P|$.

      \subsubsection{Stolen base attempt probability}
      \label{sec:prob-so-attempt}

        Modeling the runner's decision to attempt a stolen base is complicated. One may think of this as a deterministic decision made by the runner based on the expected run value with and without attempting a steal. Such a model aligns very poorly with real decisions made by runners. We posit that the runner's ability to attempt a stolen base is influenced by external factors such as the runner's perception of the pitcher's body language timing between pitches. For this reason, we model stolen base attempts as a stochastic process rather than a deterministic decision.

        Intentionally, we exclude lead distance from the stolen base attempt model. While lead distance does help explain the probability of a stolen base attempt, this can cause problems with the framing of the runner's decision. When including lead distance, we find situations in which the expected value of a stolen base attempt is negative, and the optimal lead distance is zero feet to minimize the probability of a stolen base attempt. This is a poor reflection of reality because the runner always has the power to choose not to attempt a stolen base. We obtain a more fit-for-purpose model by excluding lead distance.

        We model $\psi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
        \begin{align}
          \label{eqn:prob-sb-attempt}
          \begin{split}
            \log\left(\frac{\psi_i}{1 - \psi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
          \end{split}
        \end{align}

        We interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the effect of runner $h_i^R$ on the attempt probability, combining the fixed effect of the runner's sprint speed with the random effect of their identity. By including a fixed effect for sprint speed, we are effectively regularizing the runner effects toward priors based on their sprint speeds. The interpretation of $(\beta^C z_i^C + \gamma^C_{h_i^C})$ for catchers is similar: Each catcher has an estimated effect regularized toward a prior based on their arm strength.

      \subsubsection{Stolen base success probability}
      \label{sec:prob-so-success}

        We model $\psi_i^+$ using a mixed-effects logistic regression with fixed effects for lead distance, runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
        \begin{align}
          \label{eqn:prob-sb-success}
          \begin{split}
            \log\left(\frac{\psi_i^+}{1 - \psi_i^+}\right) &= \alpha + \beta^L \ell_i + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
          \end{split}
        \end{align}
        As in (\ref{eqn:prob-sb-attempt}), we interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the runner effect regularized toward a prior based on sprint speed and $(\beta^C z_i^C + \gamma^C_{h_i^C})$ as the catcher effect regularized toward a prior based on arm strength.

    \subsection{The baseball inning as a Markov decision process}

      \subsubsection{Markov chain model}

        Modeling the progression of a baseball inning is a well-trodden application of Markov chains in sports.(citation needed) Traditionally, the state space is defined by the bases occupied and the number of outs (24 non-terminal states, e.g., citation needed) or by those and the ball-strike count (288 non-terminal states, e.g., citation needed). To understand the game theory under the new pickoff limit rule, we take this a step further and include the number of disengagements in the state space.

        {\it State space.} The game state is given by $s = (b, c, d, o)$, where:
        \begin{align*}
          b &\in \{0, 1\}^3 \mbox{ represents which bases are occupied by runners};\\
          c &\in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ represents the ball-strike count};\\
          d &\in \{0, 1, 2\} \mbox{ represents the number of disengagements already used}; and\\
          o &\in \{0, 1, 2\} \mbox{ represents the number of outs}.
        \end{align*}
        The total number of non-terminal states is $2^3 \times (4 \times 3) \times 3 \times 3 = 864$. There is one terminal state (end of inning), bringing the total number of states to 865. We use $\mathcal{S}$ to denote the set of all states, including the terminal state; we use $s_i = (b_i, c_i, d_i, o_i)$ to denote the starting state for play $i$; and we use $S_i$ to denote the random variable governing the probability distribution over $s_i$.

        {\it Transition probabilities.} The probability of transition to state $s'$ from starting state $s$ is given by the transition probability function $\mathbb{P}(S_{i+1} = s' \mid S_i = s)$. Because the present work focuses on the consequences of decisions surrounding the runner on first base, it is helpful to break down these transition probabilities by conditioning on the runner outcome:
        \begin{align}
          \label{eqn:transition-prob}
          \mathbb{P}(S_{i+1} = s' \mid S_i = s) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s, R_i = r).
        \end{align}

        To estimate $\mathbb{P}(R_i = r \mid S_i = s)$, we may use the empirical frequency of each runner outcome $r$ in each state $s$. In Section \ref{sec:single-agent-mdp} below, we will replace these empirical frequencies with the probabilities estimated in Section \ref{sec:prob-runner-outcome}.

        To estimate $\mathbb{P}(S_{i+1} = s' \mid S_i = s, R_i = r)$, we use conditional empirical frequencies of the ending state $s'$ given the starting state $s$ and the runner outcome $r$. However, we pool across the starting disengagements $d$ when calculating these empirical frequencies. Our assumption is that, given the runner event, the starting disengagements $d$ does not impact the transition probabilities between the reduced states $\tilde s \equiv (b, c, \cdot, o)$ and $\tilde s' \equiv (b', c', \cdot, o')$. By pooling data across different values of $d$, we mitigate small sample size issues when estimating transition probabilities from rare states, such as having a runner on third base with an 0-0 count, two disengagements and zero outs.

        The pooled empirical transition probability conditions on $e = d' - d$ instead of conditioning on $d$ and $d'$. Using $E_{i}$ to denote the random variable governing the probability distribution over $d_{i + 1} - d_i$,
        \begin{align*}
          Q(\tilde s, r, \tilde s', e)
            &= \hat{\mathbb{P}}\left(\tilde S_{i+1} = \tilde s',\, E_i = e \mid \tilde S_i = \tilde s,\, R_i = r\right)\\
            &= \frac
              {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\} \cdot \mathbb{I}\left\{\tilde s_{i+1} = \tilde s',\, d_{i+1} - d_i = e\right\}}
              {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\}}.
        \end{align*}
        Finally, we estimate the full-state transition probabilities as
        \begin{align*}
          \hat{\mathbb{P}}(S_{i+1} = (b', c', d', o') \mid S_i = (b, c, d, o), R_i = r) = Q\left\big((b, c, \cdot, o),\, r,\, (b', c', \cdot, o'),\, d' - d\right\big).
        \end{align*}

% NOTE (SP): I don't think we need the paragraph below, but I'm holding onto it for now in case we decide we do need it.
%        To convert the reduced-state estimated transition probabilities to full-state transition probabilities, we leverage the fact that the number of disengagements $d'$ in the ending state is a deterministic function of the starting disengagements $d$, the runner outcome $r$ and  the ending count $c'$:
%        \begin{align}
%          \label{eqn:disengagement-transition}
%          d' = D(d, c', r)
%            \equiv \begin{cases}
%            \hfil d & \mbox{if } r \in \{\mbox{N}\} \mbox{ and } c' \ne (0, 0)\\
%            \hfil 0 & \mbox{if } r \in \{\mbox{N}\} \mbox{ and } c' = (0, 0)\\
%            d + 1   & \mbox{if } r \in \{\mbox{P}^-\} \mbox{ and } d < 2\\
%            \hfil 0 & \mbox{if } r \in \{\mbox{P}^-\} \mbox{ and } d = 2\\
%            \hfil 0 & \mbox{if } r \in \{\mbox{P}^+,\, \mbox{S}^+,\, \mbox{S}^-\}\\
%          \end{cases}.
%        \end{align}

      \subsubsection{Single-agent Markov decision process}
      \label{sec:single-agent-mdp}

        The Markov decision process is an extension of the Markov chain model that allows for deterministic actions which change transition probabilities between states. In this section we allow for the runner to choose a lead distance $\ell$, and we model the pitcher's decision whether to attempt a pickoff as a stochastic process (see Section \ref{sec:prob-po-attempt}). We call this the {\it single-agent} model because only the runner has agency in this model of the game. We limit this action to states in which first base is occupied and the other bases are unoccupied, i.e. $b = (1, 0, 0)$. This limits our scope to situations in which the runner might steal second base (encompassing XX\%\footnote{TODO: fill this in} of stolen base attempts in 2023) and excludes the first-and-third scenario $b = (1, 0, 1)$, in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.

        {\it Action space.} In states for which $b = (1, 0, 0)$, the first-base runner must choose a lead distance $\ell \in [0, 20] \equiv \mathcal{L}$, representing distance in feet from first base. For computational purposes, we discretize this action space by rounding $\ell$ to the nearest tenth of a foot, meaning that there are 201 possible actions.

        {\it Transition probabilities.} The probability of transition to ending state $s'$ from starting state $s$ given lead distance $\ell$ is $P(s' \mid s,\, \ell) = \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = \ell)$. To estimate state transition probabilities that depend on lead distance, we modify (\ref{eqn:transition-prob}) by conditioning on the lead distance:
        \begin{align}
          \label{eqn:transition-prob-lead}
          P(s' \mid s,\, \ell) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r).
        \end{align}
        In this equation, the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance. The procedure for estimating $\mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r)$ is described in Section \ref{sec:single-agent-mdp}. To estimate $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell)$, we use the models from Section \ref{sec:prob-runner-outcome}. Specifically, $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$ is given by
        \begin{align*}
          \begin{cases}
              \hfil \pi_i(\ell) \cdot \pi_i^+(\ell)                             & \mbox{if } r = \mbox{P}^+\\
              \hfil \pi_i(\ell) \cdot (1 - \pi_i^+(\ell))                       & \mbox{if } r = \mbox{P}^-\\
              \hfil (1 - \pi_i(\ell)) \cdot \psi_i \cdot \psi_i^+(\ell)         & \mbox{if } r = \mbox{S}^+\\
              \hfil (1 - \pi_i(\ell)) \cdot \psi_i \cdot (1 - \psi_i^+(\ell))   & \mbox{if } r = \mbox{S}^-\\
              \hfil (1 - \pi_i(\ell)) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
          \end{cases}.
        \end{align*}
        Recall that the stolen base attempt probability $\psi_i$ does not depend on $\ell$, as described in Section \ref{eqn:prob-sb-attempt}.

        {\it Reward function.} The reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition. Note that this reward function does not depend on the action taken by the runner, so we can write $R(s, \ell, s') = R(s, s')$. If $s = (b, c, d, o)$ and $s' = (b', c', d', o')$, then:
        \begin{align}
          \label{eqn:reward}
          R(s, s') = (g(b') + o') - (g(b) + o) - \mathbb{I}\{c' = (0, 0),\, d' = 0\},
        \end{align}
        where $g: \{0, 1\}^3 \rightarrow \{0, 1, 2, 3\}$ counts the number of runners on base. This expression for the reward function counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance. For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state. To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state. These intermediary states transition to the terminal end-of-inning state with probability 1.

        {\it Policy iteration.} A policy function $\pi : \mathcal{S} \rightarrow \mathcal{L}$ maps the state space to the action space, defining the lead distance to be taken by the runner in each state. The policy function $\pi$ specifies transition probabilities $P(s' \mid s,\, \ell = \pi(s))$ and induces a probability distribution over the sequence of states $\{S_t\}_{t=1}^\infty$. The value function $V_\pi : \mathcal{S} \rightarrow \mathbb{R}^+$ is the expected sum of rewards with respect to this probability distribution, which we interpret as the expected number of runs scored in the inning, starting from state $s$:
        \begin{align}
          \label{eqn:run-expectancy}
          V_\pi(s) = \mathbb{E}_\pi\left[\left.\sum_{t = 0}^\infty R(S_{t}, S_{t+1})~\right|~S_0 = s\right].
        \end{align}

        The relevant question for the runner is, ``What lead distance should I choose in each situation to maximize run scoring overall?'' We find the policy function $\pi^*$ which maximizes (\ref{eqn:run-expectancy}) via policy iteration (citation needed) by initializing $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, and then iteratively applying the updates
        \begin{align}
          \label{eqn:update-value-single-agent}
          \hat V_{i+1}(s) &= \sum_{s' \in \mathcal{S}} P(s' | s,\, \hat\pi_i(s)) [R(s, s') + \hat V_i(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
          \label{eqn:update-policy-single-agent}
          \hat\pi_{i+1}(s) &= \arg\max_{\ell \in \mathcal{L}} \sum_{s' \in \mathcal{S}} P(s' | s,\, \ell) [R(s, s') + \hat V_{i+1}(s')] && \forall s \in \mathcal{S},
        \end{align}
        until $\max_{s\in\mathcal{S}}|\pi_{i+1}(s) - \pi_i(s)|$ converges to zero.\footnote{Are we guaranteed convergence? Are we guaranteed convergence to the optimal policy?}

      \subsubsection{Two-agent Markov decision process}
      \label{sec:two-agent-mdp}

        In this section we extend the single-agent model by allowing the pitcher to choose whether to attempt a pickoff. We call this the {\it two-agent} model because both the runner and the pitcher have agency in this model of the game. Specifically, we model this as a two-stage game, where the runner first chooses a lead distance, and then the pitcher decides whether to attempt a pickoff after observing the runner's lead distance.

        {\it Action space.} As in the single-agent model, we limit actions to states in which only first base is occupied. The runner chooses a lead distance $\ell \in \mathcal{L}$, and then the pitcher chooses a pickoff action $p \in \{0, 1\} \equiv \mathcal{P}$, where $p = 1$ corresponds to a pickoff attempt.

        {\it Transition probabilities.} The probability of transition to ending state $s'$ from starting state $s$ given lead distance $\ell$ and pickoff decision $p$ is $P(s' \mid s,\, \ell,\, p) = \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$. We modify (\ref{eqn:transition-prob-lead}) by conditioning additionally on the pickoff decision:
        \begin{align}
          \label{eqn:transition-prob-lead-pickoff}
          P(s' \mid s, \ell, p) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r),
        \end{align}
        where $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$ is given by
        \begin{align*}
            \begin{cases}
                \hfil p \cdot \pi_i^+(\ell)                             & \mbox{if } r = \mbox{P}^+\\
                \hfil p \cdot (1 - \pi_i^+(\ell))                       & \mbox{if } r = \mbox{P}^-\\
                \hfil (1 - p) \cdot \psi_i \cdot \psi_i^+(\ell)         & \mbox{if } r = \mbox{S}^+\\
                \hfil (1 - p) \cdot \psi_i \cdot (1 - \psi_i^+(\ell))   & \mbox{if } r = \mbox{S}^-\\
                \hfil (1 - p) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
            \end{cases}.
        \end{align*}

        {\it Reward function.} The reward function (\ref{eqn:reward}) is the same for the two-agent model as for the single-agent model. In this game, however, the runner wants to maximize the reward, and the pitcher wants to minimize the reward.

        {\it Policy iteration.} In the the policy update step (\ref{eqn:update-policy-single-agent}) of policy iteration for the single-agent model, we seek the policy which maximizes the expected one-step sum of reward and resulting value, given an estimated value function. In the two-agent model, there are game theoretic considerations. Given the estimated value function $\hat V$ and the runner's chosen lead distance $\ell$, the pitcher chooses $p$ to minimize the expected one-step sum of reward and resulting value:
        \begin{align*}
          \arg\min_{p\in\mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' \mid s,\, \ell,\, p) [R(s, s') + \hat V(s')].
        \end{align*}
        Knowing that this is the optimal behavior for the pitcher, the runner chooses their lead distance to maximize the resulting run value:
        \begin{align*}
          \arg\max_{\ell \in \mathcal{L}}\left\{\min_{p\in\mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' \mid s,\, \ell,\, p) [R(s, s') + \hat V(s')]\right\}.
        \end{align*}

        The optimal policy for the pitcher is clear: between the choices of attempting a pickoff or not, make the choice that results in the lower run expectation. To find the optimal policy $\pi^* : \mathcal{S} \rightarrow \mathcal{L}$ for the runner, we again apply policy iteration. Starting with $\hat V_0(s) = 0$ and $\hat pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{Are we guaranteed convergence to the optimal policy?}

        \begin{align}
          \label{eqn:update-value-two-agent}
          \hat V_{i+1}(s) &= \min_{p \in \mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' | s,\, \hat\pi_i(s)) [R(s, s') + \hat V_i(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
          \label{eqn:update-policy-two-agent}
          \hat\pi_{i+1}(s) &= \arg\max_{\ell \in \mathcal{L}} \left\{\min_{p \in \mathcal{P}}\sum_{s' \in \mathcal{S}} P(s' | s,\, \ell,\, p) [R(s, s') + \hat V_{i+1}(s')]\right\} && \forall s \in \mathcal{S}.
        \end{align}

  \section{Results}

    \subsection{Runner outcome probabilities given lead distance}

    First, let's look at how runners' behavior changed in MLB with the new pickoff rules introduced before the 2023 season. Figure \ref{fig:leads-overall} shows how runners' lead distances at 1st base changes from 2022 to 2023 once disengagements were limited.

    \begin{figure}[b]
      \centering
      \includegraphics[width = 0.5\textwidth]{figures/leads_overall.png}
      \caption{The curves show the distribution of the lead distances at first base in feet. The red curve shows all leads from 2022, while the blue curves show the leads from 2023 under different levels of disengagements. The vertical lines represent the mean in each situation.}
      \label{fig:leads-overall}
    \end{figure}

    In 2022, we found that the number of disengagements did not have much of an  effect on lead distances at first base. In 2023, we saw that the distribution of leads by runners under 0 disengagement situations was virtually identical to all situations from 2022. With each successive disengagement, runners' leads tended to increase by about 0.6-0.8 feet. Unsurprisingly, the rule changes did make runners slightly more aggressive when taking their leads at first base. However, runners did not increase their aggression to the level that we would have recommended.

     Next, we will examine the results of our runner outcome probability models. These logistic regression models help us update our transition probabilities based on the runner's lead distance, the game situation, and the players involved in a play.

        Starting with the pickoff attempt probability model, Figure \ref{fig:prob-pickoff-attempt} shows how our estimated probability changes based on lead distance and the the number of prior disengagements:

        \begin{figure}
          \label{fig:prob-pickoff-attempt}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_attempt.png}
          \caption{The y-axis shows the probability of a pickoff attempt based on the lead distance on the x-axis. The blue lines represent the number of prior disengagements in our 2023 data, and the red line compares that to data from 2022.}
        \end{figure}

        The number of disengagements has a major effect on the likelihood of a pickoff attempt. With each successive disengagement, the probability is cut by about 50\%. Even at large lead distances, the probability of a pickoff remains very low after a pitcher has performed a disengagement or two. Notably, the 2023 rule changes also caused a significant decline from 2022's frequency of pickoffs. Even when they had 0 prior disengagements, pitchers in 2023 attempted pickoffs less often than in 2022. Other factors affecting pickoff attempt probability include the count and the pitcher's tendency to attempt pickoffs. However, prior disengagements has the most significant effect.

        Next, we'll take a look at the pickoff success probability model. Figure \ref{fig:prob-pickoff-success}  shows how our estimated probability changes based on lead distance and the pitcher's skill at picking runners off:

        \begin{figure}
          \label{fig:prob-pickoff-success}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_success.png}
          \caption{The y-axis shows the likelihood of success on a pickoff attempt based on the lead distance on the x-axis. The blue lines represent the skill level of the pitcher at picking off runners, which was calculated in our random effects model}
        \end{figure}

        Unsurprisingly, the probability of a successful pickoff is fairly low for most lead distances. However, once leads approach 13-15 feet, the success rate of pickoffs increases quickly. The gap between the median pitcher and 90th percentile pitcher indicates that certain pitchers are uniquely talented at picking runners off at first base, but even the most talented pitchers' success rates remain low. We found that runner's skill level and the game situation do not have any significant effect on the success rate of a pickoff.

        Now taking a look at the stolen base success probability model, Figure \ref{fig:prob-sb-success}  shows how our estimated probability of success changes based on lead distance and the skill level of the runner. The runner skill level is derived from the combined effects of the runner's sprint speed and his ability to steal bases:

        \begin{figure}
          \label{fig:prob-sb-success}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/prob_sb_success.png}
          \caption{The y-axis shows the likelihood of success on a stolen base attempt based on the lead distance on the x-axis. The blue lines represent the skill of the runner in our 2023 model, and the red line compares that to a median runner from our 2022 model.}
        \end{figure}

        Clearly, longer leads make stolen bases much more likely to be successful. The 2023 rule changes have made stealing bases significantly easier. A 10th percentile runner in 2023 is now even more likely to be successful while stealing as a median runner was in 2022. Other factors affecting stolen base success include the pitcher and catcher skill levels and the arm strength of the catcher.

        We elected not to include lead distance in our stolen base attempt probability model. We found that the results obtained by solely modeling stolen base attempt probability based on game situation and player's involved made much more sense than if we used lead distance. For each possible combination of situation and players, we model the probability that the runner will steal and then allow the runner to choose a lead to maximize their likelihood of stealing successfully while avoiding the risk of being picked off.

    Finally, let's also look at how our random effect models assign skill levels to runners and catchers. Figure \ref{fig:random-effect-catcher} shows how correlated our catcher skill effect is with their arm strength.

        \begin{figure}
          \label{fig:random-effect-catcher}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/catcher_effect.png}
          \caption{The x-axis shows the catcher's arm strength in miles per hour, while the y-axis shows their skill at preventing stolen bases according to our random effects model. A lower y-value means a catcher is better at preventing stolen bases.}
        \end{figure}

        There's clearly a strong correlation between arm strength and catcher skill. However, there is also a decent amount of variance in catcher skill that arm strength cannot explain. This likely comes from a catcher's arm accuracy and pop time.

        Figure \ref{fig:random-effect-runner} shows how correlated our runner skill effect is with their sprint speed.

         \begin{figure}
          \label{fig:random-effect-runner}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/runner_effect.png}
          \caption{The x-axis shows the runner's sprint speed in feet per second, while the y-axis shows their skill at stealing bases according to our random effects model. A higher y-value means a runner is better at stealing bases.}
        \end{figure}

        Sprint speed definitely is a major factor in how good a runner is at stealing bases. However, it cannot explain all of the variance in skill here. Getting good jumps and skilfully avoiding tags also likely help differentiate runners in this area.

    \subsection{Optimal lead distance under current conditions}

      We use the models described in the previous section to update our transition probabilities based on the runner's lead distance. Then, for each game state, we select the lead distance that maximizes the run expectancy, update the transition probabilities again, and repeat until convergence.

        Figure \ref{fig:finding-optimal-lead} illustrates how we find the optimal lead distance for a runner on first in a given situation. This shows how the run expectancy of the inning changes based on the runner's lead distance and the number of prior disengagements. This chart assumes a 0-0 count, 0 outs, an average runner, pitcher, and catcher, and nobody else on base:

        \begin{figure}
          \label{fig:finding-optimal-lead}
          \centering
          \includegraphics[width = 0.5\textwidth]{figures/finding_optimal_lead.png}
          \caption{The y-axis shows the run expectancy of each lead distance on the x-axis. The blue curves represent the number of prior disengagements in the plate appearance, and the vertical lines mark the lead that maximizes the run expectancy in each situation.}
        \end{figure}

        For each situation, we select the lead that maximizes the run expectancy for the inning. In this situation, along with most others we will examine, each successive disengagement increases the optimal lead distance by about two feet.

        Table \ref{tab:count} shows the optimal lead distance for a runner on first base based on the count and prior disengagements. This table assumes an average runner and battery, 0 outs, and no other runners on base:

        \begin{table}
          \centering
          \begin{tabular}{c|ccc}
            Count & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
            \hline
            \input{figures/lead_by_count}
          \end{tabular}
          \caption{Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements}
          \label{tab:count}
        \end{table}

        Each successive disengagement leads to an increase in recommended lead distance of approximately two feet. As pitchers approach the two disengagement threshold, the likelihood of a pickoff plummets, so runners can be more aggressive. Also, we notice that recommended leads tend to be a bit higher as the count becomes more advantageous for the batter. This is mainly because pitchers are more likely to attempt a pickoff in these counts.

        Table \ref{tab:runners-outs} shows the optimal lead distance for a runner on first base based on the number of outs and prior disengagements. This table assumes an average runner and battery and a 0-0 count:

        \begin{table}
          \centering
          \begin{tabular}{c|ccc}
             Outs & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
            \hline
            \input{figures/lead_by_runners_outs}
          \end{tabular}
          \caption{Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements}
          \label{tab:runners-outs}
        \end{table}

        Again, we see that each successive disengagement leads to an increase in recommended lead distance of approximately two feet. We also see that the recommended leads for 0 and 1 out are virtually identical, and are slightly longer with 2 outs. With 2 outs, stolen bases are more likely and pickoff attempts are less likely, so runners can be more aggressive.

        Table \ref{tab:players} shows the optimal lead distance for a runner on first base based on the players involved in the play. The runner skill level is based on sprint speed and their ability to steal bases. The battery skill level includes the pitcher and catcher's abilities to pick runners off and throw out potential basestealers, as well as the catcher's arm strength. This table assumes a 0-0 count, 0 outs, and nobody else on base:

        \begin{table}
          \centering
          \begin{tabular}{cc|ccc}
            Battery Skill  & Runner Skill & 0 Disengagements & 1 Disengagements & 2 Disengagements\\
            \hline
            \input{figures/lead_by_players}
          \end{tabular}
          \caption{Based on the skill of the battery (pitcher and catcher) and the skill of the runner in each row, each entry shows the optimal lead distance for a runner on first base depending on the number of disengagements}
          \label{tab:players}
        \end{table}

        The effect of the battery seems to be more significant than the effect of the runner. Facing a stronger pitcher and catcher cause the runner's optimal lead distance to shrink by several feet. We again see consistent results that runners should increase their lead by 1.5-2.5 feet upon each successive disengagement. However, better runners don't need to increase their leads by as much, likely because their stolen base success rates will be high to begin with.

        Currently, MLB runners are not doing a great job of putting pressure on the defense by maximizing their lead distances. Table \ref{tab:actual-vs-rec} compares the average actual and recommended leads for runners at 1st base with 2nd base open. With 0 disengagements, runners' actual leads come close to our recommendation, but they could often be adding an extra foot. As disengagements increase, we recommend about a 2 foot increase, but runners only tend to increase their leads by 0.6-0.7 feet. Runners overestimate the danger of a pickoff attempt under these situations, and could be increasing their team's run expectancy by adding a few feet to their lead distances here. Runners rarely meet or exceed the lead distance we recommend, especially with 1 or 2 disengagements.

        \begin{table}
          \centering
          \begin{tabular}{c|ccc}
            Disengagements & Average Actual Lead & Average Recommended Lead & Actual Exceeds Recommendation\\
            \hline
            \input{figures/actual_vs_rec_lead}
          \end{tabular}
          \caption{Each entry shows the average lead at 1st base based on the number of disengagements prior to the pitch. The recommended leads in the final column are based on our single-agent model. The final column shows the percent of pitches where the actual lead exceeds our recommendation.}
          \label{tab:actual-vs-rec}
        \end{table}

        In order to quantify the effect of optimizing baserunners' lead distances, we compared the run expectancy at the start of an inning before and after iterating through the Bellman Equation. In the 2023 season, the run expectancy of an inning was 0.5163 runs. After our policy iteration, the run expectancy jumped to 0.5277 runs. This increase of 0.0114 runs per inning may seem very small, but this difference adds up quickly. Over a 162 game season, adding 0.0114 runs per inning increases a team's run total by 16.68 runs. Using the typical valuation of \$1 Million per run, optimizing baserunners' lead distances could be worth nearly \$17 Million to a team under the current conditions. This is an incredibly valuable result that does not require much effort for a team to implement.

      \subsection{Game theoretic equilibrium}

        All of the results shown so far correspond to our single-actor model, where a runner choose a lead distance and all other outcomes are probabilistic based on their choice. Now, we expand our model to a two-actor model, where the runner first chooses a lead distance, then the pitcher elects whether to attempt a pickoff or throw a pitch. The pitcher makes whichever decision minimizes the run expectancy for the offensive team. Therefore, in order for the runner to maximize their run expectancy, they should choose a lead distance such that the pitcher is indifferent as to whether to attempt a pickoff or throw a pitch. Pitchers should always attempt a pickoff if the runner's lead extends beyond this distance, and should always throw a pitch if the lead is shorter. This is the game theoretic equilibrium.

        Table \ref{tab:count-twoagent} shows the optimal lead by count under the two-agent model:

        \begin{table}
          \centering
          \begin{tabular}{c|ccc}
            Count & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
            \hline
            \input{figures/count_two_agent}
          \end{tabular}
          \caption{Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements for our two-agent model}
          \label{tab:count-twoagent}
        \end{table}

        Under this game-theoretic equilibrium, we see that runners should be taking even longer leads in most situations. This is likely because pickoffs are often wasteful at lower lead distances. If pitchers were to optimize their pickoff behavior, then runners would be able to get longer leads before a pickoff becomes worthwhile for the pitcher. Our two-foot rule of thumb from the single-agent model doesn't hold here. Instead, we see a recommended increase of about 0.5-1 feet after 1 disengagement, and a recommended increase of around 3.5-5 feet after the 2nd disengagement. Once a pitcher reaches 2 disengagements, it becomes extremely risky to attempt a pickoff, so runners can take advantage of that by extending their leads beyond 15 feet.

        In order to evaluate how well pitchers follow our recommendations, we looked back at all situations from the 2023 season with a runner on first base only. Our two-agent model suggests that pitchers should have attempted a pickoff in 8.3\% of these situations. In reality, they only attempted pickoffs 5.6\% of the time.

        In situations where our model suggested a pickoff, pitchers actually attempted a pickoff 10.8\% of the time in 2023.

        In situations where our model suggests throwing a pitch, pitchers actually attempted a pickoff 5.0\% of the time.

        Overall, it seems that pitchers should be attempting more pickoffs. If runners start increasing their lead distances in the way we recommend, then pitchers should respond by picking off even more often.

        Under this equilibrium of optimal pitcher and baserunner behavior, runners still hold a slight advantage compared to the status quo, but not by much. We calculated that runners could increase their team's run expectancy by about 0.0036 runs per inning by following these lead recommendations if pitchers also optimized their behavior. Over a 162 game season, this adds up to 5.2 runs. Ultimately, the current behavior of pitchers and runners is not far from the Nash Equilibrium. However, both sides could improve their team's performance by several runs by optimizing their behavior surrounding the running game.

  \section{Discussion}

  \section{Assumptions}

    \begin{itemize}
      \item NOTE (SP): Let's try including pitcher handedness in our model
      \item Successful pickoff eliminates runner with no other runner movement
      \item Number of disengagements does not affect batter outcomes
      \item Lead distance does not affect outcomes on batted balls (note: domain knowledge tells us that secondary lead is more important than the lead distance we observe, which is first pitcher movement)
      \item We are ignoring the possibility that the runner takes off for 2B on a pickoff attempt
      \item Other than pickoff attempts, no other disengagements occur
    \end{itemize}

  \bibliography{arxiv}

\end{document}

### 2024-08-04 documentation/old/main.tex
add overleaf files

\documentclass{article}

\usepackage{amsfonts, amsmath, graphicx, tikz}

\begin{document}

\section{Potential results for publication}

\subsection{From an OR perspective}

Prove analytically that:
\begin{enumerate}
    \item Pitcher follows a threshold strategy in lead length. 
    \item Lead length is increasing in pickoff count. 
    \item Runs monontonic in components of state. 
\end{enumerate}

\subsection{From a baseball perspective}

Target journal: Journal of Sports Economics

\subsubsection*{Desired Results}

\begin{enumerate}
    \item What is the optimal lead distance based on current pitcher behavior? What is the corresponding run expectation?
    \item How does runner lead distance behavior differ from what would be optimal based on current pitcher behavior? How many runs are they leaving on the table?
    \item How does pitcher pickoff behavior differ from what would be optimal based on current runner behavior? How many runs are they leaving on the table?
    \item What is the optimal lead distance in the Nash equilibrium (assuming pitchers and runners behave optimally)? What is the corresponding run expectation?
    \item What would be a rule of thumb for how much you should extend your lead as the number of remaining pickoff attempts decreases? How does this interplay with count? What if you are a fast runner ($>$ 90th percentile)? Average runner? Slow runner?
\end{enumerate}

\subsubsection*{Necessary Steps}

{\bf Run game probability model.} Using 2023 data, we need to estimate regression models for the probabilities for (a) a pickoff attempt, (b) a stolen base attempt, and (c) a successful stolen base, given count, lead distance, pickoffs remaining and catcher arm strength (possibly a random effect for pitcher).\\
~\\
{\bf State space model.} We need a Markov model for the progression of an inning. {\it Note: Kritin and Mallesh shared a draft of this in August.}\\
~\\
{\bf Game solver.} Once we have the state space model and the estimated transition probabilities, we use either dynamic programming or value iteration to find the optimal policies.\\

  \section{Problem Statement}

  There is a runner on first base (second and third are empty) with no outs. The runner chooses their lead distance $d$ from first base. Knowing $d$, the pitcher decides whether to attempt a pickoff or to throw a pitch. The runner's objective is to maximize $E[R]$, the expected number of runs scored for the rest of the inning. The pitcher's objective is to minimize $E[R]$. The number of balls in the count is $b$; the number of strikes in the count is $s$; and the number of pickoffs already attempted is $p$. Once $p$ reaches 3, if the runner has not been tagged out on a pickoff attempt, they go to second base for free. The question is: How should the runner choose their lead distance $d$ (as a function of $b$, $s$, and $p$), to maximize $E[R]$?

  We can make two different assumptions about the pitcher's decision whether to attempt a pickoff. Both assumptions lead to worthwhile interpretations.
  \begin{enumerate}
      \item The decision is deterministic. Each choice yields a different run expectation, and the pitcher makes the choice to minimize run expectation.
      \item The decision is probabilistic (and we model it with data).
  \end{enumerate}

  \section{Transition Model}

  \includegraphics[width = \textwidth]{figures/flowchart.png}
  Note (SP): I think we want to expand the start state to include $p$.

\section{``Formal Model'' (Kritin)}
\section{New Model}
N = $<1,2>$, where 1 = baseman, 2 = pitcher \\\\
H = $<l, b, s, p, o, r>$ \\ \\
$l$ = $\{0, 1\}^3$ represents which bases are occupied by runners\\
$b$ = ball count \\
$s$ = strike count \\
$p$ = pickoff count \\
$o$ = out count \\
$r$ = runs scored in the inning \\\\
Z, the set of terminal states, $\subseteq$ H: \\
$\forall$ $t$ $\in$ Z, $o$ = 3. \\\\
P $\forall$ H $\setminus$ Z: P($\emptyset$) = 1 (?),\\P(h) = 2 (pitcher choosing to pickoff or pitch), $\forall$ h $\neq$ $\emptyset$\\\\
Preference Relation: 
$\forall$ $t \in Z$, $t \succ_2$ \\
Let $r_1$ and $r_2$ signify the runs scored in terminal states $t_1$ and $t_2$, respectively. Let $r_1$ $\geq$ $r_2$. Given $t_1$ and $t_2$, $t_2$ $\succ_2$

\section{Subgame}
$\Gamma$ = $<$ N, H$|_h$, P$|_h$, ($\succeq_i$) $>$ \\\\
N = $<1,2>$, 1 = baseman, 2 = pitcher \\\\
H$|_h$ = $< \{1,0,0,0\}, b, s, p, o, r >$, $\forall$ $b,s,p,o$ \\\\
Z$|_h$, the set of terminal states, $\subseteq$ H$|_h$: \\
$\forall$ $t$ $\in$ Z$|_h$, $o$ = 3 $\lor$ p = 2 \\\\
P$|_h$ $\forall$ H$|_h$ $\setminus$ Z$|_h$: \\
P$|_h$($\emptyset$) = 1 (baseman choosing lead distance), \\P(h) = 2 (?) \\\\
Preference Relation: \\
$\forall$ $t \in Z|_h$, where $o \in t = 3$, $t \succeq_2$\\
Let $r_1$ and $r_2$ signify the runs scored in terminal states $t_1$ and $t_2$, respectively. Let $r_1$ $\geq$ $r_2$. Given $t_1$ and $t_2$, where $o \in t = 3$, $t_2$ $\succeq_2$ \\\\
$\forall$ $t \in Z|_h$, where $p \in t = 2$, $t \succeq_1$\\
Let $r_3$ and $r_4$ signify the runs scored in terminal states $t_3$ and $t_4$, respectively. Let $r_3$ $\geq$ $r_4$. Given $t_3$ and $t_4$, where $p \in t = 2$, $t_3$ $\succeq_1$ \\

\section{Proof}
Prove cutoff strategy for pitcher: \\\\
More outs are preferable to less\\
probability of out increases as lead distance increases \\
given a lead distance, there is a threshold at which the pitcher prefers an attempt to pickoff rather than pitching to the batter. \\
\subsection{Formal Assumptions (everyone) }
\begin{enumerate}
    \item Expected runs not expected wins
\end{enumerate}

  \section{Possible Solutions}

    The problem we face is that the optimal lead distance $d^*(b, s, p)$ depends on the expected value of the start state $E[R | \{1, 0, 0, 0, b, s, p\}]$ while the expected value of the start state depends on the chosen lead distance. We have discussed three possible solutions:

    \subsection{Backward Induction}
    Starting from \{1, 0, 0, 0, 3, 2, 2\}, there are four possible subsequent states:
    \begin{enumerate}
        \item {\bf \{0, 0, 0, 1\}} {\it (terminal state, value is assumed known)}
        \item {\bf \{0, 1, 0, 0\}} {\it (terminal state, value is assumed known)}
        \item {\bf \{1, 0, 0, 0, end\}} {\it (terminal state, value is assumed known)}
        \item {\bf \{1, 0, 0, 0, 3, 2, 2\}} {\it (same as starting point, happens on foul ball)}
    \end{enumerate}
    If the pitcher attempts a pickoff, $E[R]$ becomes:
    $$
      f_1(d) \equiv P(\mbox{safe}\, |\, \mbox{pickoff},\, d,\, 2) \cdot E[R\, |\, \{0, 1, 0, 0\}]~+~(1 - P(\mbox{safe}\, |\, \mbox{pickoff},\, d,\, 2)) \cdot E[R\, |\, {0, 0, 0, 1}]
    $$
    If the pitcher does not attempt a pickoff, $E[R]$ becomes:
    \begin{align*}
      f_0(d) \equiv~& P(\mbox{steal}\, |\, d,\, 3,\, 2,\, 2) \cdot P(\mbox{safe}\, |\, \mbox{steal},\, d,\, 2) \cdot E[R\, |\, \{0, 1, 0, 0\}]\\
      & + P(\mbox{steal}\, |\, d,\, 3,\, 2,\, 2) \cdot (1 - P(\mbox{safe}\, |\, \mbox{steal},\, d,\, 2)) \cdot E[R\, |\, \{0, 0, 0, 1\}]\\
      & + (1 - P(\mbox{steal}\, |\, d,\, 3,\, 2,\, 2)) \cdot P(\mbox{PA Over}\, |\, 3,\, 2) \cdot E[R\, |\, \{1, 0, 0, 0\}, \mbox{end}]\\
      & + (1 - P(\mbox{steal}\, |\, d,\, 3,\, 2,\, 2)) \cdot (1 - P(\mbox{PA Over}\, |\, 3,\, 2)) \cdot E[R\, |\, \{1, 0, 0, 0, 3, 2, 2\}]
    \end{align*}
    Between $f_1$ and $f_0$, there is only one unknown: $x = E[R\, |\, \{1, 0, 0, 0, 3, 2, 2\}]$, which is itself a function of $f_1$ and $f_0$. SP posits that we can therefore solve for $x$---but there are some details to work out here that will require more carefully chosen notation. This will involve a one-dimensional grid search over $d$.

    Once we have $E[R\, |\, \{1, 0, 0, 0, 3, 2, 2\}]$, the induction moves backward from there. For example, the state $\{1, 0, 0, 0, 3, 2, 1\}$ can transition to a terminal state, to itself, or to $\{1, 0, 0, 0, 3, 2, 2\}$, whose value (and optimal lead distance $d^*(3, 2, 2)$) was derived in the previous paragraph. Therefore, we can use the same logic to derive $E[R\, |\, \{1, 0, 0, 0, 3, 2, 1\}]$ and $d^*(3, 2, 1)$. The same is true for $\{1, 0, 0, 0, 3, 1, 2\}$ and $\{1, 0, 0, 0, 2, 2, 2\}$---each of those states can only transition to a terminal state, to itself, or to $\{1, 0, 0, 0, 3, 2, 2\}$. Once we determine the optimal lead distance and run expectancy for those three states which are one step away from $\{1, 0, 0, 0, 3, 2, 2\}$, we continue the induction backward from there.

    \subsection{Iterative Optimization}

    If $E[R\, |\, \{1, 0, 0, 0, b, s, p\}]$ is known for each $(b, s, p)$, then finding the optimal lead distance $d^*(b, s, p)$ is easy. If the lead distance $d(b, s, p)$ is known, then calculating $E[R\, |\, \{1, 0, 0, 0, b, s, p\}]$ is easy. Begin by initializing $E[R\, |\, \{1, 0, 0, 0, b, s, p\}]$ according to the observed transitions in the dataset---corresponding to runners taking their observed leads, not the optimal leads. Using these run expectancies, solve for $d^*(b, s, p)$ (this will involve a one-dimensional grid search over $d$). Iterate these two steps until convergence:
    \begin{enumerate}
        \item Solve for $d^*(b, s, p)$ given $E[R\, |\, \{1, 0, 0, 0, b, s, p\}]$.\\
        (This will be a one-dimensional grid search for each $d^*(b, s, p)$.)
        \item Calculate $E[R\, |\, \{1, 0, 0, 0, b, s, p\}]$ given $d^*(b, s, p)$.
    \end{enumerate}
    Question: Does this converge to the globally optimal $d^*(b, s, p)$?

    \subsection{Full Grid Search}
    Given $d(b, s, p)$, we can calculate the transition probabilities between states and therefore the expected value of each state. Under full grid search, we try a whole bunch of possible values for $d(b, s, p)$, and we choose the value which yields the highest run expectancy for $\{1, 0, 0, 0, 0, 0, 0\}$. This full grid search is a 36-dimensional grid search, which may not be feasible.

  \newpage

  \section{Old Stuff}

    \subsection{Notation}

    $b \in \{0, 1\}^3$ represents which bases are occupied by runners\\
    $o \in \{0, 1, 2\}$ is the number of outs\\
    $c \equiv (c_b, c_s) \in \{0, 1, 2, 3\} \times \{0, 1, 2\}$ is the ball-strike count\\
    $p \in \{0, 1, 2\}$ is the number of pickoffs already attempted in this plate appearance\\
    ~\\
    $f_+: \{0, 1\}^3 \rightarrow \{0, 1\}^3$ gives occupied bases if 1B runner moves to 2B\\
    \indent e.g. $f_1((1, 0, 0)) = (0, 1, 0)$\\
    ~\\
    $f_-: \{0, 1\}^3 \rightarrow \{0, 1\}^3$ gives occupied bases if 1B runner is out\\
    \indent e.g. $f_0((1, 0, 0)) = (0, 0, 0)$\\
    ~\\
    {\it
      Note: We'll restrict ourselves to situations in which 1B is occupied and 2B is empty, meaning that $b \in \{(1, 0, 0), (1, 0 1)\}$; $f_+(b) \in \{(0, 1, 0), (0, 1, 1)\}$; and $f_-(b) \in \{(0, 0, 0), (0, 0, 1)\}$.
    }\\
    ~\\
    $A$ is the event that the pitcher attempts a pickoff\\
    $B$ is the event that the runner is safe on a pickoff attempt\\
    $C$ is the event that the runner attempts a steal\\
    $D$ is the event that the runner is safe on a steal attempt

    \subsection{Markov chain model}

    Let's define the state of the game to be $(b, o, c, p)$, which allows us to model expected runs scored to end of inning. We can use empirically observed transition probabilities between states to derive the run value of each state. We'll denote this by $R(b, o, c, p)$. However, these transition probabilities will change as we optimize lead distance.

    \subsection{The game within the pitch}

    Consider first the case where $p = 2$, i.e. the pitcher has already attempted two pickoffs since the last reset. In this case, if the pitcher attempts another pickoff and fails to catch the runner, then the runner advances for free. The decision tree below shows how the choice of lead distance affects the transition probabilities out of this state. We can construct the same decision tree for $p = 0$ and $p = 1$.

    \begin{tikzpicture}[
      box/.style={draw, rectangle},
    ]
      \node[box] (1) at (4.5, 8) {runner takes lead};
      \node[box] (2) at (1.5, 6) {pickoff attempt};
      \node[box] (3) at (7.5, 6) {pitch};
      \node[box] (4) at (0, 4) {safe};
      \node[box] (5) at (3, 4) {out};
      \node[box] (6) at (6, 4) {no steal};
      \node[box] (7) at (9, 4) {steal};
      \node[box] (8) at (9, 2) {safe};
      \node[box] (9) at (12, 2) {out};
      \node (10) at (0, 2) {$R(f_+(b), o, c, 0)$};
      \node (11) at (3, 2) {$R(f_-(b), o + 1, c, 0)$};
      \node (12) at (6, 2) {$R(b, o, c, 2)^*$};
      \node (13) at (9, 0) {$R(f_+(b), o, c, 0)^*$};
      \node (14) at (12, 0) {$R(f_-(b), o + 1, c, 0)^*$};
      \node at (4.5, 7) {\small deterministic};
      \node at (1.5, 4.5) {\small $\mathbb{P}(B| \ell, A)$};
      \node at (7.5, 4.5) {\small $\mathbb{P}(C| \ell, -A)$};
      \node at (10, 2.5) {\small $\mathbb{P}(D| \ell, C)$};
      \draw[->] (1) -- (2);
      \draw[->] (1) -- (3);
      \draw[->] (2) -- (4);
      \draw[->] (2) -- (5);
      \draw[->] (3) -- (6);
      \draw[->] (3) -- (7);
      \draw[->] (7) -- (8);
      \draw[->] (7) -- (9);
      \draw[->] (4) -- (10);
      \draw[->] (5) -- (11);
      \draw[->] (6) -- (12);
      \draw[->] (8) -- (13);
      \draw[->] (9) -- (14);
    \end{tikzpicture}
    {\small
      * This isn't exactly right because the outcome state is actually a distribution over states depending on the outcome of the pitch. Hard to write down but easy to compute.
    }\\
    ~\\
    If the pitcher attempts a pickoff, the expected resulting run value is
    $$
      R_1(\ell) = \mathbb{P}(B| \ell, A) \cdot R(f_+(b), o, c, 0) + (1 - \mathbb{P}(B| \ell, A)) \cdot R(f_-(b), o + 1, c, 0).
    $$
    If the pitcher does not attempt a pickoff, the expected resulting run value is
    \begin{align*}
      R_0(\ell) = (1 & - \mathbb{P}(C| \ell, -A)) \cdot R(b, o, c, 2)^*\\
      & + \mathbb{P}(C| \ell, -A) \cdot \left(\mathbb{P}(D| \ell, C) \cdot R(f_+(b), o, c, 0)^* + (1 - \mathbb{P}(D| \ell, C)) \cdot R(f_-(b), o + 1, c, 0)^*\right)
    \end{align*}
    Logically, the pitcher should attempt the pickoff if $R_1(\ell) < R_0(\ell)$, so the runner wants to choose $\ell$ to maximize $\min\{R_0(\ell), R_1(\ell)\}$. Assuming that $\mathbb{P}(B| \ell, A)$, $\mathbb{P}(C| \ell, -A)$ and $\mathbb{P}(D| \ell, C)$ are all monotone, continuous functions of $\ell$, the optimal lead distance $\tilde \ell$ for the runner will satisfy $R_0(\tilde\ell) = R_1(\tilde\ell)$, meaning that the pitcher is indifferent to attempting a pickoff or not.

    However! These choice have been made to maximize $R(\cdot)$, which assumes lead distances matching empirical observation, not optimal lead distances. If the runner knows they can choose the optimal lead for the next pitch, then they want to maximize the expected runs under this optimality assumption, which we'll call $\tilde R(\cdot)$. Unfortunately, we've reached a circle in our logic, but $\tilde R(\cdot)$ would be used to calculate $\tilde \ell$, and $\tilde \ell$ would be used to calculate $\tilde R(\cdot)$.

    \subsection{Solving the game}

    I admit that I'm at a little bit of a loss here. The hack I would try is to initialize $\tilde R(\cdot)$ at $R(\cdot)$, and then alternate between using $\tilde R(\cdot)$ to calculate $\tilde \ell$ and vice versa. I imagine this will probably work well because I don't think $\tilde R(\cdot)$ will be very different from $R(\cdot)$, but I don't know whether we can guarantee convergence. Perhaps there is a way to evaluate $\tilde \ell$ and $\tilde R(\cdot)$ simultaneously, and this is what Andrew refers to as ``solving the stochastic game''?

    \subsection{Extensions}

    Right now, lead distance is the only covariate that influences baserunner advancement/elimination probabilities. But we'll also want to include information about the players involved (pitcher, catcher, runner) and additional context (count, etc.).

\end{document}

### 2024-08-04 documentation/old/outline.tex
add overleaf files

\documentclass{article}

\usepackage[margin=1in]{geometry}

\begin{document}
    \section{Introduction}
    \section{Data}
    \section{Methods}
        \subsection{Probability model for runner outcomes}
            \subsubsection{Linear mixed-effects regression}
        \subsection{The baseball inning as a Markov decision process}
            \subsubsection{The single-agent model}
            \subsubsection{The two-agent model}
            \subsubsection{Policy iteration}
    \section{Results}
        \subsection{Runner outcome probabilities given lead distance}
            \begin{itemize}
                \item What is the relationship between lead distance and:
                \begin{itemize}
                    \item the probability of a pickoff attempt,
                    \item the success probability of a pickoff attempt,
                    \item the success probability of a stolen base attempt?
                \end{itemize}
            \end{itemize}
        \subsection{Optimal lead distance under current conditions}
            \begin{itemize}
                \item What is the optimal lead distance for the average runner against an average battery, as a function of disengagements?
                \item How does optimal lead distance change for 10$^{th}$-/90$^{th}$-percentile runner?
                How does optimal lead distance change for 10$^{th}$-/90$^{th}$-percentile battery?
                \item How many more runs could an offense expect to score (in the current environment) if they successfully implemented the optimal policy?
            \end{itemize}
        \subsection{Game theoretic equilibrium}
            \begin{itemize}
                \item How much more frequently does our model think pitchers should be attempting pickoffs? Under the game theoretic equilibrium, how frequently would pitchers be attempting pickoffs?
                \item Under the game theoretic equilibrium, what is the optimal lead distance for an average runner against an average battery, given disengagements?
                \item Relative to current conditions, would the game theoretic equilibrium favor offense or defense, and by how much?
            \end{itemize}
    \section{Discussion}
\end{document}

### 2024-12-11 documentation/arxiv/arxiv.tex
update table filepaths
draft Data section

@@
     \subsection{MLB Stats API}
+    \label{sec:statsapi}
+
+      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \citep{powers_sabrmetrics_2024}.
+The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
+For each play, we observe the identities of the pitcher, catcher and runners involved.
+Before and after each play, we observe the state of the game, described by the number of outs in the inning; which bases are occupied by runners; the ball-strike count on the batter; and the number of disengagements already used by the pitcher.
+On each pickoff attempt, we observe whether it is successful (runner is out) or unsuccessful (runner is safe).
+On each pitch, we observe whether the runner attempts a stolen base; and if so, we observe whether they are successful (runner is safe) or unsuccessful (runner is out).
+On each pitch terminating a plate appearance (a batting attempt by a batter), we observe the plate appearance outcome, such as strikeout, groundout, single or home run.
 
     \subsection{Baseball Savant}
+    \label{sec:baseballsavant}
+
+      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant \citep{mlb_advanced_media_notitle_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
+This is defined as the average of the top 10 percent of throw speeds when the catcher is attempting to catch a base stealer.
+For each runner-season, we observe {\it sprint speed} (feet per second), which measures the runner's top running speed in their fastest one-second window.
+This is defined as the average of the top 67 percent of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
 
     \subsection{Proprietary Statcast Data}
+
+      In addition to the publicly available data described in Sections \ref{sec:statsapi} and \ref{sec:baseballsavant}, this article relies on one key piece of proprietary information provided by MLB Advanced Media, estimated using computer vision on video data from several camera angles.
+For each play, observe each runner's {\it lead distance} (feet), which is how far they are ranging from their base at the time of the pitcher's first movement.
+This is defined as the distance between the runner's center of mass (projected onto the line between the runner's current base and their next base) and the nearest edge of the runner's current base.
+When the pitcher throws a pitch, runners will typically extend their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a {\it secondary lead}), which is relevant for runner advancement on batted balls but not on stolen base attempts.
+Our data include only the lead distance at the time of the pitcher's first movement, not the secondary lead.
 
@@
 
-        {\it Policy iteration.} In the the policy update step (\ref{eqn:update-policy-single-agent}) of policy iteration for the single-agent model, we seek the policy which maximizes the expected one-step sum of reward and resulting value, given an estimated value function.
+        {\it Policy iteration.} In the policy update step (\ref{eqn:update-policy-single-agent}) of policy iteration for the single-agent model, we seek the policy which maximizes the expected one-step sum of reward and resulting value, given an estimated value function.
 In the two-agent model, there are game theoretic considerations.
@@
 To find the optimal policy $\pi^* : \mathcal{S} \rightarrow \mathcal{L}$ for the runner, we again apply policy iteration.
-Starting with $\hat V_0(s) = 0$ and $\hat pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{Are we guaranteed convergence to the optimal policy?}
+Starting with $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{TODO: Are we guaranteed convergence to the optimal policy?}
 
@@
         \end{align}
+
+        Given this runner policy, the pitcher is indifferent between throwing a pitch and attempting a pickoff.
 
@@
 
-        Starting with the pickoff attempt probability model, Figure \ref{fig:prob-pickoff-attempt} shows how our estimated probability changes based on lead distance and the the number of prior disengagements:
+        Starting with the pickoff attempt probability model, Figure \ref{fig:prob-pickoff-attempt} shows how our estimated probability changes based on lead distance and the number of prior disengagements:
 
@@
             \hline
-            \input{figures/lead_by_count}
+            \input{tables/lead_by_count.tex}
           \end{tabular}
@@
             \hline
-            \input{figures/lead_by_runners_outs}
+            \input{tables/lead_by_runners_outs.tex}
           \end{tabular}
@@
             \hline
-            \input{figures/lead_by_players}
+            \input{tables/lead_by_players.tex}
           \end{tabular}
@@
             \hline
-            \input{figures/actual_vs_rec_lead}
+            \input{tables/actual_vs_rec_lead.tex}
           \end{tabular}
@@
             \hline
-            \input{figures/count_two_agent}
+            \input{tables/count_two_agent.tex}
           \end{tabular}
@@
       \item NOTE (SP): Let's try including pitcher handedness in our model
+      \item NOTE (SP): Should we use catcher pop time instead of arm strength?
       \item Successful pickoff eliminates runner with no other runner movement
@@
 
+  \section{Acknowledgments}
+
+    Major League Baseball trademarks and copyrights are used with permission of MLB Advanced Media, L.P. All rights reserved.
+
   \bibliography{arxiv}

### 2024-12-11 documentation/arxiv/arxiv.tex
standardize indentation in Results section

@@
 
-    \begin{figure}[b]
+      \begin{figure}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/leads_overall.png}
-      \caption{The curves show the distribution of the lead distances at first base in feet.
+        \caption{
+          \it The curves show the distribution of the lead distances at first base in feet.
 The red curve shows all leads from 2022, while the blue curves show the leads from 2023 under different levels of disengagements.
-The vertical lines represent the mean in each situation.}
+The vertical lines represent the mean in each situation.
+        }
         \label{fig:leads-overall}
@@
       \begin{figure}
-          \label{fig:prob-pickoff-attempt}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_attempt.png}
-          \caption{The y-axis shows the probability of a pickoff attempt based on the lead distance on the x-axis.
-The blue lines represent the number of prior disengagements in our 2023 data, and the red line compares that to data from 2022.}
+        \caption{
+          \it The y-axis shows the probability of a pickoff attempt based on the lead distance on the x-axis.
+The blue lines represent the number of prior disengagements in our 2023 data, and the red line compares that to data from 2022.
+        }
+        \label{fig:prob-pickoff-attempt}
       \end{figure}
@@
       \begin{figure}
-          \label{fig:prob-pickoff-success}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_success.png}
-          \caption{The y-axis shows the likelihood of success on a pickoff attempt based on the lead distance on the x-axis.
-The blue lines represent the skill level of the pitcher at picking off runners, which was calculated in our random effects model}
+        \caption{
+          \it The y-axis shows the likelihood of success on a pickoff attempt based on the lead distance on the x-axis.
+The blue lines represent the skill level of the pitcher at picking off runners, which was calculated in our random effects model.
+        }
+        \label{fig:prob-pickoff-success}
       \end{figure}
@@
       \begin{figure}
-          \label{fig:prob-sb-success}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/prob_sb_success.png}
-          \caption{The y-axis shows the likelihood of success on a stolen base attempt based on the lead distance on the x-axis.
-The blue lines represent the skill of the runner in our 2023 model, and the red line compares that to a median runner from our 2022 model.}
+        \caption{
+          \it The y-axis shows the likelihood of success on a stolen base attempt based on the lead distance on the x-axis.
+The blue lines represent the skill of the runner in our 2023 model, and the red line compares that to a median runner from our 2022 model.
+        }
+        \label{fig:prob-sb-success}
       \end{figure}
@@
       \begin{figure}
-          \label{fig:random-effect-catcher}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/catcher_effect.png}
-          \caption{The x-axis shows the catcher's arm strength in miles per hour, while the y-axis shows their skill at preventing stolen bases according to our random effects model.
-A lower y-value means a catcher is better at preventing stolen bases.}
+        \caption{
+          \it The x-axis shows the catcher's arm strength in miles per hour, while the y-axis shows their skill at preventing stolen bases according to our random effects model.
+A lower y-value means a catcher is better at preventing stolen bases.
+        }
+        \label{fig:random-effect-catcher}
       \end{figure}
@@
       \begin{figure}
-          \label{fig:random-effect-runner}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/runner_effect.png}
-          \caption{The x-axis shows the runner's sprint speed in feet per second, while the y-axis shows their skill at stealing bases according to our random effects model.
-A higher y-value means a runner is better at stealing bases.}
+        \caption{
+          \it The x-axis shows the runner's sprint speed in feet per second, while the y-axis shows their skill at stealing bases according to our random effects model.
+A higher y-value means a runner is better at stealing bases.
+        }
+        \label{fig:random-effect-runner}
       \end{figure}
@@
       \begin{figure}
-          \label{fig:finding-optimal-lead}
         \centering
         \includegraphics[width = 0.5\textwidth]{figures/finding_optimal_lead.png}
-          \caption{The y-axis shows the run expectancy of each lead distance on the x-axis.
-The blue curves represent the number of prior disengagements in the plate appearance, and the vertical lines mark the lead that maximizes the run expectancy in each situation.}
+        \caption{
+          \it The y-axis shows the run expectancy of each lead distance on the x-axis.
+The blue curves represent the number of prior disengagements in the plate appearance, and the vertical lines mark the lead that maximizes the run expectancy in each situation.
+        }
+        \label{fig:finding-optimal-lead}
       \end{figure}
@@
         \end{tabular}
-          \caption{Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements}
+        \caption{
+          \it Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements.
+        }
         \label{tab:count}
@@
         \end{tabular}
-          \caption{Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements}
+        \caption{
+          \it Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements.
+        }
         \label{tab:runners-outs}
@@
         \end{tabular}
-          \caption{Based on the skill of the battery (pitcher and catcher) and the skill of the runner in each row, each entry shows the optimal lead distance for a runner on first base depending on the number of disengagements}
+        \caption{
+          \it Based on the skill of the battery (pitcher and catcher) and the skill of the runner in each row, each entry shows the optimal lead distance for a runner on first base depending on the number of disengagements.
+        }
         \label{tab:players}
@@
         \end{tabular}
-          \caption{Each entry shows the average lead at 1st base based on the number of disengagements prior to the pitch.
+        \caption{
+          \it Each entry shows the average lead at 1st base based on the number of disengagements prior to the pitch.
 The recommended leads in the final column are based on our single-agent model.
-The final column shows the percent of pitches where the actual lead exceeds our recommendation.}
+The final column shows the percent of pitches where the actual lead exceeds our recommendation.
+        }
         \label{tab:actual-vs-rec}
@@
         \end{tabular}
-          \caption{Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements for our two-agent model}
+        \caption{
+          \it Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements for our two-agent model.
+        }
         \label{tab:count-twoagent}

### 2025-08-01 scripts/estimate_models.R
rename sprint speed player name column

@@ -322,7 +322,7 @@ dev.off()
 pitch22 <- read_csv("data/pitch/2022.csv") %>% select(play_id, description)
 event22 <- read_csv("data/event/2022.csv")
 arm_strength_2022 <- read_csv("data/catcher_throwing22.csv") %>% select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed_2022 <- read_csv("data/sprint_speed22.csv") %>% select(player_id, `last_name, first_name`, sprint_speed, competitive_runs)
+sprint_speed_2022 <- read_csv("data/sprint_speed22.csv") %>% select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
 
 
 

### 2025-08-01 scripts/estimate_models.R
put back in the random effects I removed

@@ -458,8 +458,70 @@ m4 <- readRDS("models/m4.rds")
 
 
 
-# PLOTTING MODEL RESULTS ----
+# EXTRACT PLAYER EFFECTS FROM MODELS ----
 
+# M1 pitcher effects
+pitcher_effects <- ranef(fit_po_success)$pitcher_id %>%
+  arrange(`(Intercept)`)
+median_pitcher_m1 <- rownames(pitcher_effects)[0.5 * nrow(pitcher_effects)]
+pct90_pitcher_m1 <- rownames(pitcher_effects)[0.9 * nrow(pitcher_effects)]
+pct10_pitcher_m1 <- rownames(pitcher_effects)[0.1 * nrow(pitcher_effects)]
+
+# M2 pitcher effects
+pitcher_effects <- ranef(fit_po_attempt)$pitcher_id %>%
+  arrange(`(Intercept)`)
+pct10_pitcher_m2 <- rownames(pitcher_effects)[0.1 * nrow(pitcher_effects)]
+median_pitcher_m2 <- rownames(pitcher_effects)[0.5 * nrow(pitcher_effects)]
+pct90_pitcher_m2 <- rownames(pitcher_effects)[0.9 * nrow(pitcher_effects)]
+
+# M3 pitcher effects
+pitcher_effects <- ranef(fit_sb_success)$pitcher_id %>%
+  mutate(player_id = as.numeric(rownames(pitcher_effects))) %>%
+  arrange(`(Intercept)`) 
+
+# M3 runner effects
+runner_effects <- ranef(fit_sb_success)$run1b
+runner_combined <- runner_effects %>%
+  mutate(player_id = as.numeric(rownames(runner_effects))) %>%
+  left_join(sprint_speed, by = "player_id") %>%
+  mutate(combined_effect = sprint_speed_coef * sprint_speed + `(Intercept)`) %>%
+  arrange(combined_effect) %>%
+  filter(!is.na(sprint_speed))
+pct10_runner_m3 <- runner_combined[0.1 * nrow(runner_combined),"player_id"]
+median_runner_m3 <- runner_combined[0.5 * nrow(runner_combined),"player_id"]
+pct90_runner_m3 <- runner_combined[0.9 * nrow(runner_combined),"player_id"]
+run1b_m3 <- c(pct10_runner_m3, median_runner_m3, pct90_runner_m3)
+pct10_ss_m3 <- runner_combined[0.1 * nrow(runner_combined),"sprint_speed"]
[…]

### 2025-08-02 scripts/estimate_models.R
squish remaining bugs

@@ -476,11 +476,13 @@ pct90_pitcher_m2 <- rownames(pitcher_effects)[0.9 * nrow(pitcher_effects)]
 
 # M3 pitcher effects
 pitcher_effects <- ranef(fit_sb_success)$pitcher_id %>%
-  mutate(player_id = as.numeric(rownames(pitcher_effects))) %>%
+  tibble::rownames_to_column(var = "player_id") %>%
   arrange(`(Intercept)`) 
 
 # M3 runner effects
-runner_effects <- ranef(fit_sb_success)$run1b
+runner_effects <- ranef(fit_sb_success)$run1b %>%
+  arrange(`(Intercept)`)
+sprint_speed_coef <- fixef(fit_sb_success)[3]
 runner_combined <- runner_effects %>%
   mutate(player_id = as.numeric(rownames(runner_effects))) %>%
   left_join(sprint_speed, by = "player_id") %>%
@@ -496,7 +498,9 @@ median_ss_m3 <- runner_combined[0.5 * nrow(runner_combined),"sprint_speed"]
 pct90_ss_m3 <- runner_combined[0.9 * nrow(runner_combined),"sprint_speed"]
 
 # M3 catcher effects
-catcher_effects <- ranef(fit_sb_success)$fielder_2_id
+catcher_effects <- ranef(fit_sb_success)$fielder_2_id %>%
+  arrange(`(Intercept)`)
+arm_strength_coef <- fixef(fit_sb_success)[4]
 catcher_combined <- catcher_effects %>%
   mutate(player_id = as.numeric(rownames(catcher_effects))) %>%
   left_join(arm_strength, by = "player_id") %>%
@@ -521,7 +525,7 @@ battery_combined <- catcher_effects %>%
   filter(!is.na(arm_strength))
 pct10_battery_m3_pitcher <- battery_combined[0.9 * nrow(battery_combined),"player_id.y"]
 median_battery_m3_pitcher <- battery_combined[0.5 * nrow(battery_combined),"player_id.y"]
-pct90_battery_m3_pitcher <-  battery_combined[0.1 * nrow(battery_combined),"player_id.y"]
+pct90_battery_m3_pitcher <- battery_combined[0.1 * nrow(battery_combined),"player_id.y"]
 
 ## M4 PITCHER EFFECTS
 pitcher_effects <- ranef(m4)$pitcher_id

### 2025-08-02 scripts/estimate_models.R
Fix bugs introduced by previous re-factor (#2)
* rename sprint speed player name column

* put back in the random effects I removed

* squish remaining bugs

@@ -322,7 +322,7 @@ dev.off()
 pitch22 <- read_csv("data/pitch/2022.csv") %>% select(play_id, description)
 event22 <- read_csv("data/event/2022.csv")
 arm_strength_2022 <- read_csv("data/catcher_throwing22.csv") %>% select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed_2022 <- read_csv("data/sprint_speed22.csv") %>% select(player_id, `last_name, first_name`, sprint_speed, competitive_runs)
+sprint_speed_2022 <- read_csv("data/sprint_speed22.csv") %>% select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
 
 
 
@@ -458,8 +458,74 @@ m4 <- readRDS("models/m4.rds")
 
 
 
-# PLOTTING MODEL RESULTS ----
+# EXTRACT PLAYER EFFECTS FROM MODELS ----
 
+# M1 pitcher effects
+pitcher_effects <- ranef(fit_po_success)$pitcher_id %>%
+  arrange(`(Intercept)`)
+median_pitcher_m1 <- rownames(pitcher_effects)[0.5 * nrow(pitcher_effects)]
+pct90_pitcher_m1 <- rownames(pitcher_effects)[0.9 * nrow(pitcher_effects)]
+pct10_pitcher_m1 <- rownames(pitcher_effects)[0.1 * nrow(pitcher_effects)]
+
+# M2 pitcher effects
+pitcher_effects <- ranef(fit_po_attempt)$pitcher_id %>%
+  arrange(`(Intercept)`)
+pct10_pitcher_m2 <- rownames(pitcher_effects)[0.1 * nrow(pitcher_effects)]
+median_pitcher_m2 <- rownames(pitcher_effects)[0.5 * nrow(pitcher_effects)]
+pct90_pitcher_m2 <- rownames(pitcher_effects)[0.9 * nrow(pitcher_effects)]
+
+# M3 pitcher effects
+pitcher_effects <- ranef(fit_sb_success)$pitcher_id %>%
+  tibble::rownames_to_column(var = "player_id") %>%
+  arrange(`(Intercept)`) 
+
+# M3 runner effects
+runner_effects <- ranef(fit_sb_success)$run1b %>%
+  arrange(`(Intercept)`)
+sprint_speed_coef <- fixef(fit_sb_success)[3]
+runner_combined <- runner_effects %>%
[…]

### 2025-08-02 scripts/estimate_models.R
update file paths for data CSVs

@@ -3,24 +3,40 @@
 library(tidyverse)
 library(lme4)
 
-event23 <- read_csv("data/event/2023.csv")
-lead23 <- read_csv("data/lead_distance/2023.csv")
-game23 <- read_csv("data/game/2023.csv")
-pitch23 <- read_csv("data/pitch/2023.csv") %>% select(play_id, description)
-play23 <- read_csv("data/play/2023.csv")
-event_map <- read_csv("data/batter_event.csv")
-pitch_map <- read_csv("data/batter_pitch.csv")
-arm_strength <- data.table::fread("data/catcher_throwing.csv") |>
-  dplyr::select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed <- data.table::fread("data/sprint_speed.csv") |>
-  dplyr::select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+event_2022 <- read_csv("input/data/event/2022.csv")
+event_2023 <- read_csv("input/data/event/2023.csv")
+
+lead_2022 <- read_csv("input/data/lead_distance/2022.csv")
+lead_2023 <- read_csv("input/data/lead_distance/2023.csv")
+
+pitch_2022 <- read_csv("input/data/pitch/2022.csv") |>
+  select(play_id, description)
+pitch_2023 <- read_csv("input/data/pitch/2023.csv") |>
+  select(play_id, description)
+
+play_2022 <- read_csv("input/data/play/2022.csv")
+play_2023 <- read_csv("input/data/play/2023.csv")
+
+event_map <- read_csv("input/data/batter_event.csv")
+pitch_map <- read_csv("input/data/batter_pitch.csv")
+
+arm_strength_2023 <- read_csv("input/data/arm_strength/2023.csv") |>
+  select(player_id, player_name, arm_strength, sb_attempts)
+arm_strength_2022 <- read_csv("input/data/arm_strength/2022.csv") |>
+  select(player_id, player_name, arm_strength, sb_attempts)
+
+sprint_speed_2023 <- read_csv("input/data/sprint_speed/2023.csv") |>
+  select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+sprint_speed_2022 <- read_csv("input/data/sprint_speed/2022.csv") |>
+  select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+
 
 # Remove Duplicate Lead Distances
[…]

### 2025-08-02 scripts/generate_results.R
add input/output dirs to filepaths in generate_results.R

@@ -5,9 +5,9 @@ fig_mode <- "dark"
 
 # Read data ----
 
-arm_strength <- data.table::fread("data/catcher_throwing.csv") |>
+arm_strength <- data.table::fread("input/data/arm_strength/2023.csv") |>
   dplyr::select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed <- data.table::fread("data/sprint_speed.csv") |>
+sprint_speed <- data.table::fread("input/data/sprint_speed/2023.csv") |>
   dplyr::select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
 
 
@@ -17,7 +17,7 @@ breaks <- seq(from = 4, to = 16, by = 2)
 
 if (fig_make) {
 
-  sputil::open_device(glue::glue("figures/leads_overall_{fig_mode}.pdf"), height = 4, width = 7)
+  sputil::open_device(glue::glue("output/figures/leads_overall_{fig_mode}.pdf"), height = 4, width = 7)
   plot <- lead_distance_1b |>
     ggplot2::ggplot(ggplot2::aes(`lead_distance_1st Base`, col = year_dis, linetype = year_dis)) +
     ggplot2::stat_density(geom = "line", position = "identity") +
@@ -55,7 +55,7 @@ lead_distance_grid <- tibble::tibble(lead1b = seq(from = 0, to = 20, by = 0.1))
 
 if (fig_make) {
 
-  sputil::open_device(glue::glue("figures/prob_po_attempt_{fig_mode}.pdf"), height = 4, width = 4)
+  sputil::open_device(glue::glue("output/figures/prob_po_attempt_{fig_mode}.pdf"), height = 4, width = 4)
   covariate_grid <- covariate_baseline |>
     dplyr::cross_join(lead_distance_grid) |>
     dplyr::cross_join(tibble::tibble(subset = c("2022_0", "2023_0", "2023_1", "2023_2"))) |>
@@ -115,7 +115,7 @@ pitcher_grid <- ranef(fit_po_success)$pitcher_id |>
   )
 
 if (fig_make) {
-  sputil::open_device(glue::glue("figures/prob_po_success_{fig_mode}.pdf"), height = 4, width = 4)
+  sputil::open_device(glue::glue("output/figures/prob_po_success_{fig_mode}.pdf"), height = 4, width = 4)
   covariate_grid <- covariate_baseline |>
[…]

### 2025-08-02 articles/arxiv/paper.tex
rename latex source files

--- 2024-12-11 documentation/arxiv/arxiv.tex
@@
   \section{Introduction}
+  \label{sec:introduction}
 
@@
     In the present work, we conduct the first public analysis of this game.
-We model the baseball inning as a Markov decision process and solve for (a) the optimal policy for runner lead distance given current pitcher pickoff behavior and (b) the game theoretic equilibrium behavior in the two-player game between runner and pitcher.
+We model the baseball inning as a Markov decision process (MDP) and solve for (a) the optimal policy for runner lead distance given current pitcher pickoff behavior and (b) the game theoretic equilibrium behavior in the two-player game between runner and pitcher.
 We present the practically actionable rule-of-thumb that an average runner should increase their lead distance by two feet after each disengagement.
@@
 Recently, \cite{chan_points_2021} used a Markov model to evaluate team performance in football.
-More specifically, the Markov decision process has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \citep{nadimpalli_when_2013}, baseball \citep{hirotsu_using_2019}, basketball \citep{sandholtz_markov_2020}, darts \citep{baird_optimising_2020} and football \citep{biro_reinforcement_2022}.
+More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \citep{nadimpalli_when_2013}, baseball \citep{hirotsu_using_2019}, basketball \citep{sandholtz_markov_2020}, darts \citep{baird_optimising_2020} and football \citep{biro_reinforcement_2022}.
 
@@
       \subsubsection{Stolen base attempt probability}
-      \label{sec:prob-so-attempt}
+      \label{sec:prob-sb-attempt}
 
@@
 
-        Intentionally, we exclude lead distance from the stolen base attempt model.
+        Deliberately, we exclude lead distance from the stolen base attempt model.
 While lead distance does help explain the probability of a stolen base attempt, this can cause problems with the framing of the runner's decision.
@@
       \subsubsection{Stolen base success probability}
-      \label{sec:prob-so-success}
+      \label{sec:prob-sb-success}
 
@@
 
-        The Markov decision process is an extension of the Markov chain model that allows for deterministic actions which change transition probabilities between states.
+        The MDP is an extension of the Markov chain model that allows for deterministic actions which change transition probabilities between states.
 In this section we allow for the runner to choose a lead distance $\ell$, and we model the pitcher's decision whether to attempt a pickoff as a stochastic process (see Section \ref{sec:prob-po-attempt}).
@@
 
-    \subsection{Runner outcome probabilities given lead distance}
+    \subsection{Observed runner behavior}
 
-      First, let's look at how runners' behavior changed in MLB with the new pickoff rules introduced before the 2023 season.
-Figure \ref{fig:leads-overall} shows how runners' lead distances at 1st base changes from 2022 to 2023 once disengagements were limited.
+      As detailed in Section \ref{sec:introduction}, MLB made several rule changes between the 2022 season and the 2023 season, most notably (for the present work) widening the bases and limiting the number of pickoffs that can be attempted by the pitcher.
+Figure \ref{fig:leads-overall} illustrates the difference in runner lead distance behavior (from first base) between 2022 and 2023.
+With zero disengagements in 2023, the distribution of lead distances is very similar to the overall distribution of lead distances in 2022.
+With each successive disengagement, we observe a positive shift in the distribution of lead distance in the 2023 data.
+In 2022, the average lead distance was XX feet.
+In 2023, the average lead distance under 0, 1 and 2 disengagements was XX feet, XX feet and XX feet, respectively.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/leads_overall.png}
+        \includegraphics[width = 0.8\textwidth]{figures/leads_overall.png}
         \caption{
-          \it The curves show the distribution of the lead distances at first base in feet.
-The red curve shows all leads from 2022, while the blue curves show the leads from 2023 under different levels of disengagements.
-The vertical lines represent the mean in each situation.
+          \it The league-wide distribution of lead distance from first base when second base is unoccupied, comparing 2022 with 2023.
+The 2023 data are split by prior disengagements because, beginning in 2023, pitchers were limited to three disengagements, influencing pitcher and runner behavior.
+In 2022, there was no limit on the number of disengagements.
+The vertical lines mark the means of the distributions.
         }
@@
 
-      In 2022, we found that the number of disengagements did not have much of an  effect on lead distances at first base.
-In 2023, we saw that the distribution of leads by runners under 0 disengagement situations was virtually identical to all situations from 2022.
-With each successive disengagement, runners' leads tended to increase by about 0.6-0.8 feet.
-Unsurprisingly, the rule changes did make runners slightly more aggressive when taking their leads at first base.
-However, runners did not increase their aggression to the level that we would have recommended.
+    \subsection{Probability model for runner outcomes}
 
-      Next, we will examine the results of our runner outcome probability models.
-These logistic regression models help us update our transition probabilities based on the runner's lead distance, the game situation, and the players involved in a play.
+      Figures \ref{fig:prob-pickoff-attempt}-\ref{fig:prob-sb-success} illustrate the relationship between lead distance and runner outcome probabilities, as estimated using the mixed-effects logistic regression models detailed in Section \ref{sec:prob-runner-outcome}.
+For all three models, the direction of the relationship between lead distance and outcome probability is as expected.
 
-      Starting with the pickoff attempt probability model, Figure \ref{fig:prob-pickoff-attempt} shows how our estimated probability changes based on lead distance and the number of prior disengagements:
+      Figure \ref{fig:prob-pickoff-attempt} shows pickoff attempt probability modeled as a function of lead distance, split by season and number of prior disengagements.
+We observe that pickoff attempt rates were lower in 2023 than in 2022, especially as the number of prior disengagement increased.
+The modeled probability remains low even for extremely long leads, such as 20 feet.
+We have very limited data on lead distances beyond 14 feet, especially with fewer than two prior disengagements, so extrapolation much beyond 14 feet is likely not reliable.
+The ball-strike count and the identity of the pitcher also influence the pickoff attempt probability, but the primary drivers are lead distance and the number prior disengagements.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_attempt.png}
+        \includegraphics[width = 0.8\textwidth]{figures/prob_pickoff_attempt.png}
         \caption{
-          \it The y-axis shows the probability of a pickoff attempt based on the lead distance on the x-axis.
-The blue lines represent the number of prior disengagements in our 2023 data, and the red line compares that to data from 2022.
+          \it Pickoff attempt probability modeled as function of lead distance, split by season and number of prior disengagements.
+This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-po-attempt}, assuming 0 balls, 0 strikes, 0 outs and a median pitcher random effect.
         }
@@
 
-      The number of disengagements has a major effect on the likelihood of a pickoff attempt.
-With each successive disengagement, the probability is cut by about 50\%.
-Even at large lead distances, the probability of a pickoff remains very low after a pitcher has performed a disengagement or two.
-Notably, the 2023 rule changes also caused a significant decline from 2022's frequency of pickoffs.
-Even when they had 0 prior disengagements, pitchers in 2023 attempted pickoffs less often than in 2022.
-Other factors affecting pickoff attempt probability include the count and the pitcher's tendency to attempt pickoffs.
-However, prior disengagements has the most significant effect.
-
-      Next, we'll take a look at the pickoff success probability model.
-Figure \ref{fig:prob-pickoff-success}  shows how our estimated probability changes based on lead distance and the pitcher's skill at picking runners off:
+      Figure \ref{fig:prob-pickoff-success} shows pickoff success probability modeled as a function of lead distance, split by pitcher pickoff skill (based on their random effect from the mixed-effect logistic regression).
+We observe that, except for the most skilled pitchers, pickoff attempt probability is very low within the range covering the vast majority of lead distances.
+Toward the high end of this range, the success rate increases quickly as lead distance increases.
+Interestingly, there is a large gap between the 90$^{th}$-percentile pitcher and the median pitcher but a small gap between the median pitcher and the 10$^{th}$-percentile pitcher, demonstrating a right skew in the distribution of pitcher pickoff skill.
+Earlier exploration indicated that game state and runner identity did not provide predictive value for pickoff success given lead distance.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/prob_pickoff_success.png}
+        \includegraphics[width = 0.8\textwidth]{figures/prob_pickoff_success.png}
         \caption{
-          \it The y-axis shows the likelihood of success on a pickoff attempt based on the lead distance on the x-axis.
-The blue lines represent the skill level of the pitcher at picking off runners, which was calculated in our random effects model.
+          \it Pickoff success probability modeled as function of lead distance, split by pitcher random effect.
+This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-po-success}, which includes only an intercept, a fixed effect for lead distance and a random effect for pitcher identity.
         }
@@
 
-      Unsurprisingly, the probability of a successful pickoff is fairly low for most lead distances.
-However, once leads approach 13-15 feet, the success rate of pickoffs increases quickly.
-The gap between the median pitcher and 90th percentile pitcher indicates that certain pitchers are uniquely talented at picking runners off at first base, but even the most talented pitchers' success rates remain low.
-We found that runner's skill level and the game situation do not have any significant effect on the success rate of a pickoff.
+      The stolen base attempt probability model does not include lead distance, as explained in Section \ref{sec:prob-sb-attempt}.
+As a result, this model has only a secondary effect on the lead distance policies learned in later sections.
 
-      Now taking a look at the stolen base success probability model, Figure \ref{fig:prob-sb-success}  shows how our estimated probability of success changes based on lead distance and the skill level of the runner.
-The runner skill level is derived from the combined effects of the runner's sprint speed and his ability to steal bases:
+      Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as function of lead distance, split by season and by runner skill (based on sprint speed and runner random effect).
+We observe that, conditioned on lead distance, even 10$^{th}$-percentile runners in 2023 outperformed the median runner in 2022, which is likely attributable to rule changes making it easier to successfully steal bases in 2023.
+Within the range covering most lead distances, the difference between the 90$^{th}$-percentile runner and the median runner is just under 5 percentage points of success probability, similar to the difference between the median runner and the 10$^{th}$-percentile runner.
+Base-out state and {\it battery} (i.e.
+pitcher and catcher) skill also influence stolen base success probability although they are not shown in this visualization.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/prob_sb_success.png}
+        \includegraphics[width = 0.8\textwidth]{figures/prob_sb_success.png}
         \caption{
-          \it The y-axis shows the likelihood of success on a stolen base attempt based on the lead distance on the x-axis.
-The blue lines represent the skill of the runner in our 2023 model, and the red line compares that to a median runner from our 2022 model.
+          \it Stolen base success probability modeled as function of lead distance, split by season and runner effect (combining the fixed effect of runner sprint speed and the random effect of runner identity).
+This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-sb-success}, assuming 0 balls, 0 strikes, 0 outs and a median value for the combined effect of pitcher identity (random effect), catcher arm strength (fixed effect) and catcher identity (random effect).
         }
@@
 
-      Clearly, longer leads make stolen bases much more likely to be successful.
-The 2023 rule changes have made stealing bases significantly easier.
-A 10th percentile runner in 2023 is now even more likely to be successful while stealing as a median runner was in 2022.
-Other factors affecting stolen base success include the pitcher and catcher skill levels and the arm strength of the catcher.
+      A novel aspect of these runner outcome probability models is the combination of fixed effects for measurable player characteristics (catcher arm strength and runner sprint speed).
+This feature allows us to answer questions such as: How important are these measurable characteristics for player performance outcomes?
+And to what extent do players over- or under-perform what is expected from their measurable characteristics?
+The combination of the fixed effect and the random effect gives us an estimate (which we call the {\it combined effect}) of each player's skill, regularized toward what is expected from their measurable characteristics.
 
-      We elected not to include lead distance in our stolen base attempt probability model.
-We found that the results obtained by solely modeling stolen base attempt probability based on game situation and player's involved made much more sense than if we used lead distance.
-For each possible combination of situation and players, we model the probability that the runner will steal and then allow the runner to choose a lead to maximize their likelihood of stealing successfully while avoiding the risk of being picked off.
-
-      Finally, let's also look at how our random effect models assign skill levels to runners and catchers.
-Figure \ref{fig:random-effect-catcher} shows how correlated our catcher skill effect is with their arm strength.
+      For the stolen base success probability model, Figure \ref{fig:random-effect} shows the relationship between the combined player effect and the corresponding measurable characteristic, for both catchers and runners.
+We observe that arm strength explains 65\% of the variance in the catcher combined effect and that sprint speed explains 86\% of the variance in the runner combined effect.
+In other words, sprint speed tells us more about runner skill than arm strength tells us about catcher skill, with regard to stolen base outcomes.
+Beyond arm strength, the catcher can reduce stolen base success by releasing the ball more quickly or by throwing it more accurately to the target base.
+Beyond sprint speed, the runner can increase stolen base success by starting to run earlier (i.e., getting a better jump) or by accelerating more quickly to their top speed.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/catcher_effect.png}
+        \includegraphics[width = 0.49\textwidth]{figures/catcher_effect.png}
+        \includegraphics[width = 0.49\textwidth]{figures/runner_effect.png}
         \caption{
-          \it The x-axis shows the catcher's arm strength in miles per hour, while the y-axis shows their skill at preventing stolen bases according to our random effects model.
-A lower y-value means a catcher is better at preventing stolen bases.
+          \it
+          (Left) The relationship between arm strength and catcher combined effect on stolen base success probability (relative to average, on the log-odds scale).
+The combined effect is the sum of the fixed effect due to arm strength and the catcher random effect.
+          (Right) The relationship between sprint speed and runner combined effect on stolen base success probability (relative to average, on the log-odds scale).
+The combined effect is the sum of the fixed effect due to sprint speed and the runner random effect.
         }
-        \label{fig:random-effect-catcher}
+        \label{fig:random-effect}
       \end{figure}
 
-      There's clearly a strong correlation between arm strength and catcher skill.
-However, there is also a decent amount of variance in catcher skill that arm strength cannot explain.
-This likely comes from a catcher's arm accuracy and pop time.
+    \subsection{Single-agent Markov decision process}
 
-      Figure \ref{fig:random-effect-runner} shows how correlated our runner skill effect is with their sprint speed.
+      In the single-agent MDP, the only decision-maker is the runner, and they choose a policy for lead distance as a function of the game state, with the objective of maximizing the expected number of runs in the inning.
+In practice, the purpose of this MDP is to optimize runner decision-making given current conditions of pitcher behavior, anticipating that pitcher behavior will take some time to react to changes in runner behavior.
+
+      Figure \ref{fig:finding-optimal-lead} illustrates how the runner would choose an optimal lead distance for different numbers of prior disengagements in an example scenario.
+Short lead distances have very low pickoff probabilities and low stolen base probabilities.
+As lead distance increases, stolen base probability increases, but so does pickoff probability.
+For very long lead distances, the run expectancy drops off sharply as the pickoff probability increases sharply.
+Between those two extremes, the runner strikes a balance to maximize run expectancy.
+The optimal lead distances for 0, 1 and 2 disengagements are 10.0 feet, 11.5 feet and 14.1 feet, respectively.
 
@@
         \centering
-        \includegraphics[width = 0.5\textwidth]{figures/runner_effect.png}
+        \includegraphics[width = 0.8\textwidth]{figures/finding_optimal_lead.png}
         \caption{
-          \it The x-axis shows the runner's sprint speed in feet per second, while the y-axis shows their skill at stealing bases according to our random effects model.
-A higher y-value means a runner is better at stealing bases.
-        }
-        \label{fig:random-effect-runner}
-      \end{figure}
-
-      Sprint speed definitely is a major factor in how good a runner is at stealing bases.
-However, it cannot explain all of the variance in skill here.
-Getting good jumps and skilfully avoiding tags also likely help differentiate runners in this area.
-
-    \subsection{Optimal lead distance under current conditions}
-
-      We use the models described in the previous section to update our transition probabilities based on the runner's lead distance.
-Then, for each game state, we select the lead distance that maximizes the run expectancy, update the transition probabilities again, and repeat until convergence.
-
-      Figure \ref{fig:finding-optimal-lead} illustrates how we find the optimal lead distance for a runner on first in a given situation.
-This shows how the run expectancy of the inning changes based on the runner's lead distance and the number of prior disengagements.
-This chart assumes a 0-0 count, 0 outs, an average runner, pitcher, and catcher, and nobody else on base:
-
-      \begin{figure}
-        \centering
-        \includegraphics[width = 0.5\textwidth]{figures/finding_optimal_lead.png}
-        \caption{
-          \it The y-axis shows the run expectancy of each lead distance on the x-axis.
-The blue curves represent the number of prior disengagements in the plate appearance, and the vertical lines mark the lead that maximizes the run expectancy in each situation.
+          \it Rest-of-inning run expectancy for different choices of lead distance.
+Three separate curves show the run expectancy for differing numbers of prior disengagements, assuming 0 balls; 0 strikes; 0 outs; third base unoccupied; and median player effects for pitcher, catcher and runner.
+Run expectancy is defined as the expectation of the value function $V_{\pi^*}$ (evaluated at the resulting state) corresponding to the optimal policy $\pi^*$ defined by (\ref{eqn:run-expectancy}).
+The vertical line corresponding to each curve marks the value at which the curve is maximized.
         }
@@
 
-      For each situation, we select the lead that maximizes the run expectancy for the inning.
-In this situation, along with most others we will examine, each successive disengagement increases the optimal lead distance by about two feet.
+      Policy iteration, as detailed by equations (\ref{eqn:update-value-single-agent}) and (\ref{eqn:update-policy-single-agent}), yields a policy $\pi^*$ optimizing rest-of-inning run expectancy.
+Tables \ref{tab:lead-by-count}, \ref{tab:lead-by-runners-outs} and \ref{tab:lead-by-players} report the optimal lead distance policy $\pi^*$ for different game states and player skills.
 
-      Table \ref{tab:count} shows the optimal lead distance for a runner on first base based on the count and prior disengagements.
-This table assumes an average runner and battery, 0 outs, and no other runners on base:
+      Table \ref{tab:lead-by-count} shows how optimal lead distance varies by ball-strike count and number of prior disengagements, for an average runner facing an average battery (pitcher and catcher) with no outs.
+We observe that count has a mild influence on optimal lead distance, and lead distance has a strong influence.
+As the pitcher nears the disengagement limit, the runner has more incentive to extend their lead distance because pickoff attempts are less likely and a free base advancement (from a third unsuccessful pickoff attempt) is more likely.
 
@@
         \begin{tabular}{c|ccc}
-          Count & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
+          Count & \multicolumn{3}{c}{Disengagements}\\
+                & 0 & 1 & 2\\
           \hline
@@
         \caption{
-          \it Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements.
+          \it The optimal lead distance policy (in feet) for an average runner facing an average battery (pitcher and catcher) with no outs.
         }
-        \label{tab:count}
+        \label{tab:lead-by-count}
       \end{table}
 
-      Each successive disengagement leads to an increase in recommended lead distance of approximately two feet.
-As pitchers approach the two disengagement threshold, the likelihood of a pickoff plummets, so runners can be more aggressive.
-Also, we notice that recommended leads tend to be a bit higher as the count becomes more advantageous for the batter.
-This is mainly because pitchers are more likely to attempt a pickoff in these counts.
+      Across counts, the recommended increase in lead distance is generally close to 1.5 feet after the first disengagement and 2.5 feet after the second disengagement.
+We observed this to be a good approximation across different runner/battery skills and different numbers of outs.
+Based on these results, we recommend the rule of thumb that runners increase their lead by two feet after each disengagement, a very simple and actionable guideline for athletes to follow.
 
-      Table \ref{tab:runners-outs} shows the optimal lead distance for a runner on first base based on the number of outs and prior disengagements.
-This table assumes an average runner and battery and a 0-0 count:
+      Table \ref{tab:lead-by-runners-outs} shows how optimal lead distance varies by number of outs, for an average runner facing an average batter with a 0-0 count, and Table \ref{tab:lead-by-players} shows how optimal lead distance varies by runner and battery skill, with 0 outs and a 0-0 count.
 
@@
         \begin{tabular}{c|ccc}
-           Outs & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
+          Outs  & \multicolumn{3}{c}{Disengagements}\\
+                & 0 & 1 & 2 \\
           \hline
@@
         \caption{
-          \it Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements.
+          \it Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on first depending on the number of disengagements.
         }
-        \label{tab:runners-outs}
+        \label{tab:lead-by-runners-outs}
       \end{table}
@@
         }
-        \label{tab:players}
+        \label{tab:lead-by-players}
       \end{table}
@@
 
-    \subsection{Game theoretic equilibrium}
+    \subsection{Two-agent Markov decision process}
 

### 2025-08-02 scripts/estimate_models.R
move data wrangling code for first figure to generate_results.R

@@ -284,54 +284,6 @@ with_leads22$lead3b <-with_leads22$`lead_distance_3rd Base`
 
 
 
-second_open_22 <- baserunners_22 %>% left_join(play_2022, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2022")
-
-second_open_23 <- baserunners %>% left_join(play_2023, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2023")
-
-lead_distance_1b <- rbind(second_open_22, second_open_23) |>
-  dplyr::mutate(
-    year_dis = case_when(
-      year == 2022 ~ "2022 - All Situations",
-      year == 2023 & pre_disengagements == 0 ~ "2023 - 0 Disengagements",
-      year == 2023 & pre_disengagements == 1 ~ "2023 - 1 Disengagements",
-      year == 2023 & pre_disengagements == 2 ~ "2023 - 2 Disengagements"
-    )
-  ) |>
-  dplyr::filter(!is.na(year_dis))
-
-mean_leads <- lead_distance_1b %>% group_by(year_dis) %>% summarize(meanlead = mean(`lead_distance_1st Base`))
-mean_lead22 <- mean_leads[1,2]$meanlead
-mean_lead23_0 <- mean_leads[2,2]$meanlead
-mean_lead23_1 <- mean_leads[3,2]$meanlead
-mean_lead23_2 <- mean_leads[4,2]$meanlead
-
-# Overal 2022 vs 2023 lead distance density plot
-leads_overall <- lead_distance_1b |>
-  ggplot2::ggplot() +
-  ggplot2::aes(`lead_distance_1st Base`, col = year_dis, group = year_dis) +
-  ggplot2::geom_density() +
-  ggplot2::xlim(3, 16) +
-  ggplot2::theme_classic() +
-  ggplot2::labs(
-    x = "Lead Distance",
-    y = "Density",
-    title = "Lead Distance at 1st Base - 2022/2023",
-    color = "Year and Prior Disengagements"
-  ) +
[…]

### 2025-08-02 scripts/generate_results.R

@@ -13,6 +13,21 @@ sprint_speed <- data.table::fread("input/data/sprint_speed/2023.csv") |>
 
 # Plot data summary ----
 
+second_open_22 <- baserunners_22 %>% left_join(play_2022, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2022")
+
+second_open_23 <- baserunners %>% left_join(play_2023, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2023")
+
+lead_distance_1b <- rbind(second_open_22, second_open_23) |>
+  dplyr::mutate(
+    year_dis = case_when(
+      year == 2022 ~ "2022 - All Situations",
+      year == 2023 & pre_disengagements == 0 ~ "2023 - 0 Disengagements",
+      year == 2023 & pre_disengagements == 1 ~ "2023 - 1 Disengagements",
+      year == 2023 & pre_disengagements == 2 ~ "2023 - 2 Disengagements"
+    )
+  ) |>
+  dplyr::filter(!is.na(year_dis))
+
 breaks <- seq(from = 4, to = 16, by = 2)
 
 if (fig_make) {

### 2025-08-02 scripts/estimate_models.R
remove code for unused runners_outs_two_agent table

@@ -893,9 +893,6 @@ while(change > threshold || iterations < 2) {
   print(old_re_table_two[1,2])
 }     
 
-sorted_two <- old_run_1b_two %>% mutate(bases = substr(State, 1, 3), dis = substr(State, 5,5), countouts = substr(State, 7, 9)) %>% arrange(bases, countouts, dis)
-
-table1_twoagent <- sorted_two %>% ungroup() %>%  filter(substr(countouts, 1,2) == "00") %>% mutate(Runners = ifelse(bases == "100", "Man on 1st", "Men on 1st and 3rd"), Outs = substr(countouts,3,3)) %>% select(Runners, Outs, dis, lead1b) %>%  pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "Disengagements_")
 
 
 
@@ -927,15 +924,6 @@ mean(no_pick$isPickAttempt) # When pick not recommended, they actually pick 5.0%
 old_re_table_two
 
 
-
-print(
-  xtable::xtable(table1_twoagent, digits = 1),
-  hline.after = NULL,
-  include.colnames = FALSE,
-  include.rownames = FALSE,
-  only.contents = TRUE,
-  file = "figures/runners_outs_two_agent.tex"
-)
 
 
 # MONOTONICITY CHECKS ----

### 2025-08-02 scripts/estimate_models.R
update output filepath for lead_by_runners_outs.tex

@@ -845,7 +845,7 @@ print(
   include.colnames = FALSE,
   include.rownames = FALSE,
   only.contents = TRUE,
-  file = "figures/lead_by_runners_outs.tex"
+  file = "output/figures/lead_by_runners_outs.tex"
 )
 
 ## How well do runners select lead distance?

### 2025-08-02 scripts/generate_results.R
fix style in generate_results.R

@@ -13,9 +13,25 @@ sprint_speed <- data.table::fread("input/data/sprint_speed/2023.csv") |>
 
 # Plot data summary ----
 
-second_open_22 <- baserunners_22 %>% left_join(play_2022, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2022")
+second_open_22 <- baserunners_22 |>
+  dplyr::left_join(play_2022, by = "play_id") |>
+  dplyr::filter(
+    is.na(`lead_distance_2nd Base`),
+    is.na(`lead_distance_3rd Base`),
+    !is.na(`lead_distance_1st Base`)
+  ) |>
+  dplyr::select(`lead_distance_1st Base`, pre_disengagements) |>
+  dplyr::mutate(year = "2022")
 
-second_open_23 <- baserunners %>% left_join(play_2023, by = "play_id") %>% filter(is.na(`lead_distance_2nd Base`), is.na(`lead_distance_3rd Base`), !is.na(`lead_distance_1st Base`)) %>% select(`lead_distance_1st Base`, pre_disengagements) %>% mutate(year = "2023")
+second_open_23 <- baserunners |>
+  dplyr::left_join(play_2023, by = "play_id") |>
+  dplyr::filter(
+    is.na(`lead_distance_2nd Base`),
+    is.na(`lead_distance_3rd Base`),
+    !is.na(`lead_distance_1st Base`)
+  ) |>
+  dplyr::select(`lead_distance_1st Base`, pre_disengagements) |>
+  dplyr::mutate(year = "2023")
 
 lead_distance_1b <- rbind(second_open_22, second_open_23) |>
   dplyr::mutate(

### 2025-08-02 scripts/estimate_models.R
update output filepaths for models

@@ -360,15 +360,15 @@ all_pickoff_var1b$pre_disengagements <- as.factor(all_pickoff_var1b$pre_disengag
 #m4 <- glmer(isSBAttempt ~ pre_balls + pre_strikes + pre_outs + pre_disengagements + (1|pitcher_id) + (1|fielder_2_id) + (1|run1b) + sprint_speed + arm_strength, data = sb_var_1b_threats, family = binomial)
 #summary(m4)
 
-#saveRDS(fit_po_success, "models/fit_po_success.rds")
-#saveRDS(fit_po_attempt, "models/fit_po_attempt.rds")
-#saveRDS(fit_sb_success, "models/fit_sb_success.rds")
+#saveRDS(fit_po_success, "output/models/fit_po_success.rds")
+#saveRDS(fit_po_attempt, "output/models/fit_po_attempt.rds")
+#saveRDS(fit_sb_success, "output/models/fit_sb_success.rds")
 #saveRDS(m4, "m4model")
 
-fit_po_success <- readRDS("models/fit_po_success.rds")
-fit_po_attempt <- readRDS("models/fit_po_attempt.rds")
-fit_sb_success <- readRDS("models/fit_sb_success.rds")
-m4 <- readRDS("models/m4.rds")
+fit_po_success <- readRDS("output/models/fit_po_success.rds")
+fit_po_attempt <- readRDS("output/models/fit_po_attempt.rds")
+fit_sb_success <- readRDS("output/models/fit_sb_success.rds")
+m4 <- readRDS("output/models/m4.rds")
 
 # 
 # # Pitcher Pickoff Ratio vs Effect

### 2025-08-02 articles/arxiv/paper.tex
Re-organize file structure (#3)
* update file paths for data CSVs

* add input/output dirs to filepaths in generate_results.R

* rename latex source files

* ignore input/output and latex auxiliary files

* update file paths in latex code

* fix bug with underscore in url

* gitignore *.blg and *.synctex.gz

* move tracked figures and tables to output folder

* update folder structure in README

* move data wrangling code for first figure to generate_results.R

* remove code for unused runners_outs_two_agent table

* update output filepath for lead_by_runners_outs.tex

* fix style in generate_

@@
         \begin{align*}
-          \hat{\mathbb{P}}(S_{i+1} = (b', c', d', o') \mid S_i = (b, c, d, o), R_i = r) = Q\left\big((b, c, \cdot, o),\, r,\, (b', c', \cdot, o'),\, d' - d\right\big).
+          \hat{\mathbb{P}}(S_{i+1} = (b', c', d', o') \mid S_i = (b, c, d, o), R_i = r) = Q\big((b, c, \cdot, o),\, r,\, (b', c', \cdot, o'),\, d' - d\big).
         \end{align*}
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{figures/leads_overall.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/leads_overall.png}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{figures/prob_pickoff_attempt.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_pickoff_attempt.png}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{figures/prob_pickoff_success.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_pickoff_success.png}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{figures/prob_sb_success.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_sb_success.png}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.49\textwidth]{figures/catcher_effect.png}
-        \includegraphics[width = 0.49\textwidth]{figures/runner_effect.png}
+        \includegraphics[width = 0.49\textwidth]{../../output/figures/catcher_effect.png}
+        \includegraphics[width = 0.49\textwidth]{../../output/figures/runner_effect.png}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{figures/finding_optimal_lead.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/finding_optimal_lead.png}
         \caption{
@@
           \hline
-          \input{tables/lead_by_count.tex}
+          \input{../../output/tables/lead_by_count.tex}
         \end{tabular}
@@
           \hline
-          \input{tables/lead_by_runners_outs.tex}
+          \input{../../output/tables/lead_by_runners_outs.tex}
         \end{tabular}
@@
           \hline
-          \input{tables/lead_by_players.tex}
+          \input{../../output/tables/lead_by_players.tex}
         \end{tabular}
@@
           \hline
-          \input{tables/actual_vs_rec_lead.tex}
+          \input{../../output/tables/actual_vs_rec_lead.tex}
         \end{tabular}
@@
           \hline
-          \input{tables/count_two_agent.tex}
+          \input{../../output/tables/count_two_agent.tex}
         \end{tabular}
@@
 
-  \bibliography{arxiv}
+  \bibliography{../pickoff-game-theory}
 

### 2025-08-02 output/tables/lead_by_players.tex
Re-organize file structure (#3)
* update file paths for data CSVs

* add input/output dirs to filepaths in generate_results.R

* rename latex source files

* ignore input/output and latex auxiliary files

* update file paths in latex code

* fix bug with underscore in url

* gitignore *.blg and *.synctex.gz

* move tracked figures and tables to output folder

* update folder structure in README

* move data wrangling code for first figure to generate_results.R

* remove code for unused runners_outs_two_agent table

* update output filepath for lead_by_runners_outs.tex

* fix style in generate_

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Sun Jul 21 17:31:33 2024
 10th Percentile Battery & 10th Percentile Runner & 10.5 & 11.9 & 14.5 \\ 
  10th Percentile Battery & Median Runner & 11.3 & 12.5 & 14.7 \\ 
  10th Percentile Battery & 90th Percentile Runner & 12.0 & 13.0 & 14.8 \\ 
  Median Battery & 10th Percentile Runner & 9.6 & 11.3 & 14.1 \\ 
  Median Battery & Median Runner & 10.3 & 11.7 & 14.1 \\ 
  Median Battery & 90th Percentile Runner & 11.0 & 12.1 & 14.2 \\ 
  90th Percentile Battery & 10th Percentile Runner & 7.8 & 9.8 & 12.6 \\ 
  90th Percentile Battery & Median Runner & 8.2 & 10.0 & 12.7 \\ 
  90th Percentile Battery & 90th Percentile Runner & 8.8 & 10.4 & 12.8 \\

### 2025-08-02 scripts/estimate_models.R

@@ -3,24 +3,40 @@
 library(tidyverse)
 library(lme4)
 
-event23 <- read_csv("data/event/2023.csv")
-lead23 <- read_csv("data/lead_distance/2023.csv")
-game23 <- read_csv("data/game/2023.csv")
-pitch23 <- read_csv("data/pitch/2023.csv") %>% select(play_id, description)
-play23 <- read_csv("data/play/2023.csv")
-event_map <- read_csv("data/batter_event.csv")
-pitch_map <- read_csv("data/batter_pitch.csv")
-arm_strength <- data.table::fread("data/catcher_throwing.csv") |>
-  dplyr::select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed <- data.table::fread("data/sprint_speed.csv") |>
-  dplyr::select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+event_2022 <- read_csv("input/data/event/2022.csv")
+event_2023 <- read_csv("input/data/event/2023.csv")
+
+lead_2022 <- read_csv("input/data/lead_distance/2022.csv")
+lead_2023 <- read_csv("input/data/lead_distance/2023.csv")
+
+pitch_2022 <- read_csv("input/data/pitch/2022.csv") |>
+  select(play_id, description)
+pitch_2023 <- read_csv("input/data/pitch/2023.csv") |>
+  select(play_id, description)
+
+play_2022 <- read_csv("input/data/play/2022.csv")
+play_2023 <- read_csv("input/data/play/2023.csv")
+
+event_map <- read_csv("input/data/batter_event.csv")
+pitch_map <- read_csv("input/data/batter_pitch.csv")
+
+arm_strength_2023 <- read_csv("input/data/arm_strength/2023.csv") |>
+  select(player_id, player_name, arm_strength, sb_attempts)
+arm_strength_2022 <- read_csv("input/data/arm_strength/2022.csv") |>
+  select(player_id, player_name, arm_strength, sb_attempts)
+
+sprint_speed_2023 <- read_csv("input/data/sprint_speed/2023.csv") |>
+  select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+sprint_speed_2022 <- read_csv("input/data/sprint_speed/2022.csv") |>
+  select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
+
 
[…]

### 2025-08-02 scripts/generate_results.R

@@ -5,19 +5,50 @@ fig_mode <- "dark"
 
 # Read data ----
 
-arm_strength <- data.table::fread("data/catcher_throwing.csv") |>
+arm_strength <- data.table::fread("input/data/arm_strength/2023.csv") |>
   dplyr::select(player_id, player_name, arm_strength, sb_attempts)
-sprint_speed <- data.table::fread("data/sprint_speed.csv") |>
+sprint_speed <- data.table::fread("input/data/sprint_speed/2023.csv") |>
   dplyr::select(player_id, player_name = `last_name, first_name`, sprint_speed, competitive_runs)
 
 
 # Plot data summary ----
 
+second_open_22 <- baserunners_22 |>
+  dplyr::left_join(play_2022, by = "play_id") |>
+  dplyr::filter(
+    is.na(`lead_distance_2nd Base`),
+    is.na(`lead_distance_3rd Base`),
+    !is.na(`lead_distance_1st Base`)
+  ) |>
+  dplyr::select(`lead_distance_1st Base`, pre_disengagements) |>
+  dplyr::mutate(year = "2022")
+
+second_open_23 <- baserunners |>
+  dplyr::left_join(play_2023, by = "play_id") |>
+  dplyr::filter(
+    is.na(`lead_distance_2nd Base`),
+    is.na(`lead_distance_3rd Base`),
+    !is.na(`lead_distance_1st Base`)
+  ) |>
+  dplyr::select(`lead_distance_1st Base`, pre_disengagements) |>
+  dplyr::mutate(year = "2023")
+
+lead_distance_1b <- rbind(second_open_22, second_open_23) |>
+  dplyr::mutate(
+    year_dis = case_when(
+      year == 2022 ~ "2022 - All Situations",
+      year == 2023 & pre_disengagements == 0 ~ "2023 - 0 Disengagements",
+      year == 2023 & pre_disengagements == 1 ~ "2023 - 1 Disengagements",
+      year == 2023 & pre_disengagements == 2 ~ "2023 - 2 Disengagements"
+    )
+  ) |>
+  dplyr::filter(!is.na(year_dis))
+
 breaks <- seq(from = 4, to = 16, by = 2)
 
 if (fig_make) {
 
-  sputil::open_device(glue::glue("figures/leads_overall_{fig_mode}.pdf"), height = 4, width = 7)
[…]

### 2025-08-02 articles/arxiv/paper.tex
switch from natbib to biblatex

@@
 
+\usepackage[alldates=year, block=space, style=apa]{biblatex}
+\setlength\bibitemsep{0.5\baselineskip}
+\addbibresource{../bibliography.bib}
+
+\usepackage[margin=1in]{geometry}
+\linespread{1.2}
+
+\usepackage[colorlinks, allcolors=blue]{hyperref}
+
 \usepackage{amsfonts, amsmath, graphicx}
-\usepackage[margin=1in]{geometry}
-\usepackage{natbib}
-\bibliographystyle{apa-good}
 
@@
     In baseball, the stolen base is an exciting play in which a runner will take off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out.
-Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \citep{tango_book_2007}.
+Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring (\cite{tango_book_2007}).
 On the micro level, it improves a team's win probability to reduce stolen base attempts that have negative expected outcomes, but on the macro level, the reduction in stolen base attempts affects the entertainment value of the game.
 
-    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting \citep{castrovince_pitch_2023}.
+    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting (\cite{castrovince_pitch_2023}).
 With these changes, a pitch clock limits the time between pitches, and there are restrictions on infielder positioning, curtailing defensive shifts.
@@
 
-    \cite{downey_pick_2015} previously analyzed pickoff throws and stolen base attempts as a mixed strategy game without using lead distance data.
+    \textcite{downey_pick_2015} previously analyzed pickoff throws and stolen base attempts as a mixed strategy game without using lead distance data.
 They modeled the two-player game as a simultaneous decision in which the pitcher chooses whether to attempt a pickoff and the runner chooses whether to attempt a stolen base, showing that this theoretical game has a mixed-strategy Nash equilibrium.
 Under this equilibrium, left-handed pitchers (possessing better pickoff moves than their right-handed counterparts) would attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data.
-In a later analysis, \cite{downey_pressure_2019} argued that pitchers can effectively implement a mixed strategy in low-pressure situations but struggle to randomize their decisions as pressure increases, showing a higher auto-correlation for sequential pickoff decisions in high-pressure situations.
+In a later analysis, \textcite{downey_pressure_2019} argued that pitchers can effectively implement a mixed strategy in low-pressure situations but struggle to randomize their decisions as pressure increases, showing a higher auto-correlation for sequential pickoff decisions in high-pressure situations.
 Both of these analyses covered a game in which pickoff attempts were unlimited.
@@
     Several studies in sports economics literature have examined the extent to which professional athletes play minimax strategy.
-\cite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save.
-\cite{palacios-huerta_experientia_2008} followed this with a laboratory experiment featuring professional soccer players and college students simulating the penalty kick decision using cards, showing that the professionals played minimax but the students did not.
-On the other hand, \cite{kovash_professionals_2009} found evidence inconsistent with minimax equilibrium behavior in the National Football League (teams called more run plays and fewer pass plays than theoretically optimal) and in MLB (pitchers threw more fastballs than theoretically optimal).
+\textcite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save.
+\textcite{palacios-huerta_experientia_2008} followed this with a laboratory experiment featuring professional soccer players and college students simulating the penalty kick decision using cards, showing that the professionals played minimax but the students did not.
+On the other hand, \textcite{kovash_professionals_2009} found evidence inconsistent with minimax equilibrium behavior in the National Football League (teams called more run plays and fewer pass plays than theoretically optimal) and in MLB (pitchers threw more fastballs than theoretically optimal).
 
-    The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model \citep{bukiet_markov_1997}.
-\cite{hirotsu_using_2002} used a four-state Markov model for soccer to inform substitution decision-making based on how substitutions impact transition probabilities between states.
-\cite{hirotsu_markov_2003} used similar methodology to inform substitution (pinch hitting) decision-making in baseball.
-Recently, \cite{chan_points_2021} used a Markov model to evaluate team performance in football.
-More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \citep{nadimpalli_when_2013}, baseball \citep{hirotsu_using_2019}, basketball \citep{sandholtz_markov_2020}, darts \citep{baird_optimising_2020} and football \citep{biro_reinforcement_2022}.
+    The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model (\cite{bukiet_markov_1997}).
+\textcite{hirotsu_using_2002} used a four-state Markov model for soccer to inform substitution decision-making based on how substitutions impact transition probabilities between states.
+\textcite{hirotsu_markov_2003} used similar methodology to inform substitution (pinch hitting) decision-making in baseball.
+Recently, \textcite{chan_points_2021} used a Markov model to evaluate team performance in football.
+More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis (\cite{nadimpalli_when_2013}), baseball (\cite{hirotsu_using_2019}), basketball (\cite{sandholtz_markov_2020}), darts (\cite{baird_optimising_2020}) and football (\cite{biro_reinforcement_2022}).
 
     An intermediate result of the present work is a probability model for stolen base attempts and successes (as well as an analysis how those probabilities changed with the 2023 MLB rule changes), a topic which has not received much attention.
-\cite{loughin_assessing_2008} estimated linear mixed-effects models for stolen base attempts and successes, using random effects for pitcher and catcher (not the runner, notably) and fixed effects for balls, strikes, outs, score, venue and pitcher hand.
+\textcite{loughin_assessing_2008} estimated linear mixed-effects models for stolen base attempts and successes, using random effects for pitcher and catcher (not the runner, notably) and fixed effects for balls, strikes, outs, score, venue and pitcher hand.
 They showed that the spread of pitcher effects is slightly greater than the spread of catcher effects.
 That analysis did not have data on runner sprint speed and catcher arm strength.
-More recently, \cite{stanley_modeling_2023} modeled stolen base success probability using both logistic regression and a random forest, including runner sprint speed and catcher arm strength as features but excluding player identities.
+More recently, \textcite{stanley_modeling_2023} modeled stolen base success probability using both logistic regression and a random forest, including runner sprint speed and catcher arm strength as features but excluding player identities.
 That analysis was limited to data prior to the 2023 MLB rule changes that incentivized base stealing.
@@
 
-      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \citep{powers_sabrmetrics_2024}.
+      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics (\cite{powers_sabrmetrics_2024}).
 The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
@@
 
-      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant \citep{mlb_advanced_media_notitle_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant (\cite{mlb_advanced_media_notitle_2024}) describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
 For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
@@
       \end{align*}
-      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \citep{bates_fitting_2015}.
+      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 (\cite{bates_fitting_2015}).
 
@@
 
-  \bibliography{../pickoff-game-theory}
+  \printbibliography
 

### 2025-08-02 scripts/generate_results.R
add code missing from generate_results.R

@@ -367,6 +367,13 @@ skill_grid |>
     hline.after = c(0, 3, 6)
   )
 
+sorted_two <- old_run_1b_two |>
+  dplyr::mutate(
+    bases = substr(State, 1, 3),
+    dis = substr(State, 5,5), countouts = substr(State, 7, 9)
+  ) |>
+  dplyr::arrange(bases, countouts, dis)
+
 sorted_two |>
   dplyr::ungroup() |>
   dplyr::filter(substr(countouts, 3, 3) == "0", bases == "100") |>
@@ -397,6 +404,3 @@ actual_vs_recommended_leads |>
     digits = c(NA, 0, 1, 1, NA)
   )
 
-
-
-#TACOCAT GOAT CHEESE PIZZA

### 2025-08-02 articles/arxiv/paper.tex
switch paper to new light mode figures

@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/leads_overall.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/leads_overall_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_pickoff_attempt.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_po_attempt_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_pickoff_success.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_po_success_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_sb_success.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_sb_success_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.49\textwidth]{../../output/figures/catcher_effect.png}
-        \includegraphics[width = 0.49\textwidth]{../../output/figures/runner_effect.png}
+        \includegraphics[width = 0.49\textwidth]{../../output/figures/effect_catcher_light.pdf}
+        \includegraphics[width = 0.49\textwidth]{../../output/figures/effect_runner_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/finding_optimal_lead.png}
+        \includegraphics[width = 0.8\textwidth]{../../output/figures/finding_optimal_lead_light.pdf}
         \caption{

### 2025-08-02 scripts/generate_results.R

@@ -1,6 +1,6 @@
 
 fig_make <- TRUE
-fig_mode <- "dark"
+fig_mode <- "light"
 
 
 # Read data ----
@@ -63,7 +63,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(x = "Lead Distance", y = "Density") +
     ggplot2::coord_cartesian(xlim = c(3, 16)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.2, 0.8))
   print(plot)
   dev.off()
@@ -126,7 +126,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(title = "Pickoff Attempt", x = "Lead Distance", y = "Probability") +
     ggplot2::coord_cartesian(xlim = c(3, 16), ylim = c(0, 0.7)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.4, 0.7))
   print(plot)
   dev.off()
@@ -177,7 +177,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(title = "Pickoff Success", x = "Lead Distance", y = "Probability") +
     ggplot2::coord_cartesian(xlim = c(3, 16), ylim = c(0, 0.7)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.3, 0.7))
   print(plot)
   dev.off()
@@ -204,7 +204,7 @@ if (fig_make) {
       y = "Effect on SB Success (log-odds)"
     ) +
     ggplot2::coord_cartesian(ylim = c(-0.6, 0.6)) +
-    sputil::theme_sleek()
+    sputil::theme_sleek(mode = fig_mode)
   print(plot)
   dev.off()
 }
@@ -230,7 +230,7 @@ if (fig_make) {
       y = "Effect on SB Success (log-odds)"
     ) +
     ggplot2::coord_cartesian(ylim = c(-0.6, 0.6)) +
-    sputil::theme_sleek()
+    sputil::theme_sleek(mode = fig_mode)
   print(plot)
   dev.off()
 }
@@ -281,7 +281,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(x = "Lead Distance", y = "Probability") +
[…]

### 2025-08-02 articles/arxiv/paper.tex
switch paper to new tables

@@
         \centering
-        \begin{tabular}{c|ccc}
-          Count & \multicolumn{3}{c}{Disengagements}\\
-                & 0 & 1 & 2\\
-          \hline
         \input{../../output/tables/lead_by_count.tex}
-        \end{tabular}
         \caption{
@@
         \centering
-        \begin{tabular}{cc|ccc}
-          Battery Skill  & Runner Skill & 0 Disengagements & 1 Disengagements & 2 Disengagements\\
-          \hline
         \input{../../output/tables/lead_by_players.tex}
-        \end{tabular}
         \caption{
@@
         \centering
-        \begin{tabular}{c|ccc}
-          Disengagements & Average Actual Lead & Average Recommended Lead & Actual Exceeds Recommendation\\
-          \hline
         \input{../../output/tables/actual_vs_rec_lead.tex}
-        \end{tabular}
         \caption{
@@
         \centering
-        \begin{tabular}{c|ccc}
-          Count & 0 Disengagements & 1 Disengagement & 2 Disengagements\\
-          \hline
         \input{../../output/tables/count_two_agent.tex}
-        \end{tabular}
         \caption{

### 2025-08-02 output/tables/lead_by_players.tex
switch paper to new tables

@@
-% latex table generated in R 4.3.1 by xtable 1.8-4 package
-% Sun Jul 21 17:31:33 2024
- 10th Percentile Battery & 10th Percentile Runner & 10.5 & 11.9 & 14.5 \\ 
-  10th Percentile Battery & Median Runner & 11.3 & 12.5 & 14.7 \\ 
-  10th Percentile Battery & 90th Percentile Runner & 12.0 & 13.0 & 14.8 \\ 
-  Median Battery & 10th Percentile Runner & 9.6 & 11.3 & 14.1 \\ 
-  Median Battery & Median Runner & 10.3 & 11.7 & 14.1 \\ 
-  Median Battery & 90th Percentile Runner & 11.0 & 12.1 & 14.2 \\ 
-  90th Percentile Battery & 10th Percentile Runner & 7.8 & 9.8 & 12.6 \\ 
-  90th Percentile Battery & Median Runner & 8.2 & 10.0 & 12.7 \\ 
-  90th Percentile Battery & 90th Percentile Runner & 8.8 & 10.4 & 12.8 \\
+\begin{tabular}{cc|rrr}
+\multicolumn{2}{c|}{Skill Percentile} & \multicolumn{3}{c}{Disengagements}\\
+Battery & Runner & 0 & 1 & 2 \\
+% latex table generated in R 4.5.0 by xtable 1.8-4 package
+% Sat Aug  2 17:11:21 2025
+  \hline
+10th & 10th & 10.5 & 11.8 & 14.1 \\ 
+  10th & 50th & 11.2 & 12.3 & 14.3 \\ 
+  10th & 90th & 11.8 & 12.7 & 14.5 \\ 
+   \hline
+50th & 10th & 9.4 & 10.9 & 13.7 \\ 
+  50th & 50th & 9.9 & 11.4 & 13.8 \\ 
+  50th & 90th & 10.6 & 11.9 & 14.0 \\ 
+   \hline
+90th & 10th & 8.1 & 9.8 & 12.5 \\ 
+  90th & 50th & 8.6 & 10.1 & 12.6 \\ 
+  90th & 90th & 9.2 & 10.5 & 12.7 \\ 
+\end{tabular}

### 2025-08-03 scripts/generate_results.R
rotate lead by count tables wide

@@ -331,14 +331,18 @@ if (fig_make) {
 sorted |>
   dplyr::ungroup() |>
   dplyr::filter(substr(countouts, 3, 3) == "0", bases == "100") |>
+  dplyr::arrange(substr(countouts, 2, 2), substr(countouts, 1, 1)) |>   # sort by strikes then balls
   dplyr::mutate(count = paste0(substr(countouts, 1, 1), "-", substr(countouts, 2, 2))) |>
   dplyr::select(count, dis, lead1b) |>
-  tidyr::pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "disengagements_") |>
+  tidyr::pivot_wider(names_from = count, values_from = lead1b, names_prefix = "count") |>
   sputil::write_latex_table(
     file = "output/tables/lead_by_count.tex",
-    prefix_rows = " & \\multicolumn{3}{c}{Disengagements}",
-    colnames = c("Count", "0", "1", "2"),
-    align = "c|rrr",
+    prefix_rows = "Prior & \\multicolumn{12}{c}{Count}",
+    colnames = c(
+      "Disengagements",
+      "0-0", "1-0", "2-0", "3-0", "0-1", "1-1", "2-1", "3-1", "0-2", "1-2", "2-2", "3-2"
+    ),
+    align = "r|cccc|cccc|cccc",
     digits = 1
   )
 
@@ -377,14 +381,18 @@ sorted_two <- old_run_1b_two |>
 sorted_two |>
   dplyr::ungroup() |>
   dplyr::filter(substr(countouts, 3, 3) == "0", bases == "100") |>
-  dplyr::mutate(Count = paste0(substr(countouts, 1, 1), "-", substring(countouts, 2, 2))) |>
-  dplyr::select(Count, dis, lead1b) |>
-  pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "Disengagements_") |>
+  dplyr::arrange(substr(countouts, 2, 2), substr(countouts, 1, 1)) |>   # sort by strikes then balls
+  dplyr::mutate(count = paste0(substr(countouts, 1, 1), "-", substring(countouts, 2, 2))) |>
+  dplyr::select(count, dis, lead1b) |>
+  tidyr::pivot_wider(names_from = count, values_from = lead1b, names_prefix = "count") |>
   sputil::write_latex_table(
     file = "output/tables/count_two_agent.tex",
[…]

### 2025-08-03 scripts/estimate_models.R
move lead-by-outs table to generate_results.R

@@ -831,22 +831,6 @@ for (i in 1:nrow(skill_grid)) {
 }
 
 
-# SUMMARY TABLES ----
-
-
-sorted <- old_run_1b %>% mutate(bases = substr(State, 1, 3), dis = substr(State, 5,5), countouts = substr(State, 7, 9)) %>% arrange(bases, countouts, dis)
-
-
-table1 <- sorted %>% ungroup() %>%  filter(substr(countouts, 1,2) == "00") %>% mutate(Runners = ifelse(bases == "100", "Man on 1st", "Men on 1st and 3rd"), Outs = substr(countouts,3,3)) %>% select(Outs, dis, lead1b) %>%  pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "Disengagements_")
-
-print(
-  xtable::xtable(table1, digits = 1),
-  hline.after = NULL,
-  include.colnames = FALSE,
-  include.rownames = FALSE,
-  only.contents = TRUE,
-  file = "output/figures/lead_by_runners_outs.tex"
-)
 
 ## How well do runners select lead distance?
 

### 2025-08-03 scripts/generate_results.R

@@ -328,6 +328,28 @@ if (fig_make) {
 
 # Write results tables ----
 
+sorted <- old_run_1b |>
+  dplyr::mutate(
+    bases = substr(State, 1, 3),
+    dis = substr(State, 5,5),
+    countouts = substr(State, 7, 9)
+  ) |>
+  arrange(bases, countouts, dis)
+
+sorted |>
+  dplyr::ungroup() |>
+  dplyr::filter(substr(countouts, 1,2) == "00") |>
+  dplyr::mutate(outs = substr(countouts, 3, 3)) |>
+  dplyr::select(outs, dis, lead1b) |>
+  tidyr::pivot_wider(names_from = outs, values_from = lead1b, names_prefix = "outs_") |>
+  sputil::write_latex_table(
+    file = "output/tables/lead_by_outs.tex",
+    prefix_rows = "Prior & \\multicolumn{3}{c}{Outs}",
+    colnames = c("Disengagements", "0", "1", "2"),
+    align = "r|ccc",
+    digits = 1
+  )
+
 sorted |>
   dplyr::ungroup() |>
   dplyr::filter(substr(countouts, 3, 3) == "0", bases == "100") |>

### 2025-08-03 scripts/generate_results.R
transpose tab:lead-by-outs

@@ -338,20 +338,6 @@ sorted <- old_run_1b |>
 
 sorted |>
   dplyr::ungroup() |>
-  dplyr::filter(substr(countouts, 1,2) == "00") |>
-  dplyr::mutate(outs = substr(countouts, 3, 3)) |>
-  dplyr::select(outs, dis, lead1b) |>
-  tidyr::pivot_wider(names_from = outs, values_from = lead1b, names_prefix = "outs_") |>
-  sputil::write_latex_table(
-    file = "output/tables/lead_by_outs.tex",
-    prefix_rows = "Prior & \\multicolumn{3}{c}{Outs}",
-    colnames = c("Disengagements", "0", "1", "2"),
-    align = "r|ccc",
-    digits = 1
-  )
-
-sorted |>
-  dplyr::ungroup() |>
   dplyr::filter(substr(countouts, 3, 3) == "0", bases == "100") |>
   dplyr::arrange(substr(countouts, 2, 2), substr(countouts, 1, 1)) |>   # sort by strikes then balls
   dplyr::mutate(count = paste0(substr(countouts, 1, 1), "-", substr(countouts, 2, 2))) |>
@@ -365,6 +351,20 @@ sorted |>
       "0-0", "1-0", "2-0", "3-0", "0-1", "1-1", "2-1", "3-1", "0-2", "1-2", "2-2", "3-2"
     ),
     align = "r|cccc|cccc|cccc",
+    digits = 1
+  )
+
+sorted |>
+  dplyr::ungroup() |>
+  dplyr::filter(substr(countouts, 1,2) == "00") |>
+  dplyr::mutate(outs = substr(countouts, 3, 3)) |>
+  dplyr::select(outs, dis, lead1b) |>
+  tidyr::pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "dis_") |>
+  sputil::write_latex_table(
+    file = "output/tables/lead_by_outs.tex",
+    prefix_rows = " & \\multicolumn{3}{c}{Disengagements}",
+    colnames = c("Outs", "0", "1", "2"),
+    align = "r|ccc",
     digits = 1
   )
 

### 2025-08-03 scripts/estimate_models.R
remove deprecated code for unsuccuessful pickoff transitions

@@ -160,21 +160,6 @@ succesful_pick <- function(State) {
   return(paste0("00", substr(State, 3, 3), " 0 ", substr(State, 7, 8), as.numeric(outs)+1))
 }
 
-# Function for what happens on unsuccesful pickoff
-# Assumes no pickoff errors or other runners moving
-unsuccesful_pick <- function(State) {
-  if (substr(State, 1, 2) != "10") {
-    return(State)
-  }
-  dis <- substr(State, 5, 5) 
-  if (dis == 2) {
-    return(paste0("010 0 ", substr(State, 7, 9)))
-  } else {
-    return(paste0(substr(State, 1, 4), as.numeric(dis) + 1, substr(State, 6, 9)))
-  }
-  
-}
-
 # Call function on all states
 all_states <- data.frame(State = all_possible_states)
 T_sp <- all_states %>% mutate(New_State = sapply(State, succesful_pick), Freq = 1)
@@ -185,19 +170,6 @@ all_transitions_tsp <- state_grid %>% left_join(T_sp, by = c("State", "New_State
 # Create final transition prob matrix
 P_SP <-  all_transitions_tsp %>% group_by(State) %>% mutate(Total = sum(Freq)) %>% ungroup() %>% mutate(Prob = Freq / Total)
 P_SP <- P_SP %>% mutate(Prob = ifelse(is.na(Prob), ifelse(New_State == State, 1, 0), Prob)) 
-
-
-# P^UP MATRIX (Failed Pick)
-
-# Call function on all states
-T_up <- all_states %>% mutate(New_State = sapply(State, unsuccesful_pick), Freq = 1)
-
-# join state grid with transition probs
-all_transitions_tup <- state_grid %>% left_join(T_up, by = c("State", "New_State")) %>% mutate(Freq = ifelse(is.na(Freq), 0 , Freq))
-
-# Create final transition prob matrix
-P_UP <-  all_transitions_tup %>% group_by(State) %>% mutate(Total = sum(Freq)) %>% ungroup() %>% mutate(Prob = Freq / Total)
-P_UP <- P_UP %>% mutate(Prob = ifelse(is.na(Prob), ifelse(New_State == State, 1, 0), Prob)) 
 
 
 # P^UP MATRIX (Failed Pick)

### 2025-08-03 scripts/generate_results.R
add code for reporting average lead distance by scenario

@@ -44,6 +44,10 @@ lead_distance_1b <- rbind(second_open_22, second_open_23) |>
   ) |>
   dplyr::filter(!is.na(year_dis))
 
+lead_distance_1b |>
+  dplyr::group_by(year_dis) |>
+  dplyr::summarize(mean(`lead_distance_1st Base`))
+
 breaks <- seq(from = 4, to = 16, by = 2)
 
 if (fig_make) {

### 2025-08-03 articles/arxiv/paper.tex
force locations of figures and tables
finish the first full draft of the paper

@@
 
-\usepackage{amsfonts, amsmath, graphicx}
+\usepackage{amsfonts, amsmath, float, graphicx}
 
@@
 
+  \begin{abstract}
+    In 2023, Major League Baseball (MLB) introduced new several rule changes to incentivize base stealing and more exciting plays.
+Among these changes, bases became wider (slighly reducing running distance for a stolen base), and the pitcher was limited to three pickoff attempts.
+The pickoff limit creates an interesting cat-and-mouse game between pitcher and runner because each failed pickoff attempt adds pressure on the pitcher to avoid using the next pickoff attempt.
+The runner can increase this pressure by extending the distance of their leadoff toward the next base, increasing their chances of successfully stealing the base.
+Using pitch-by-pitch lead distance data obtained from MLB, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
+Using these conditional transition probabilities, we estimate optimal lead distance via two separate Markov decision process models: single-agent (runner only) and two-agent (game theoretic).
+From these results, we present an actionable rule of thumb: the Two-Foot Rule.
+We recommend that a runner on first base increase their lead distance by two feet after each successive pickoff attempt.
+  \end{abstract}
+
   \section{Introduction}
@@
 
-    In baseball, the stolen base is an exciting play in which a runner will take off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out.
-Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring (\cite{tango_book_2007}).
+    On May 17, 2025, the Chicago Cubs hosted the Chicago White Sox in the second game of their Crosstown Series.
+In the bottom of the first inning, White Sox pitcher Sean Burke faced Cubs batter Michael Bush with two outs and Cubs runner Kyle Tucker on first base.
+Before the 1-1 pitch, Burke pivoted and threw to first base behind Tucker, who retreated safely before a tag.
+At the time, Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball among base stealers who had not been caught \parencite{fangraphs_major_2025}.
+On the White Sox broadcast, color commentator Steve Stone explained to the audience, ``The only problem is now there's two throws over there, and Tucker's gonna get a much bigger lead.'' On the next pitch, Tucker stole his 11$^{th}$ base of the season, the first Burke had allowed.
+``I don't think anybody in the room could have thrown him out on that play,'' said Stone.
+
+    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out.
+Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
 On the micro level, it improves a team's win probability to reduce stolen base attempts that have negative expected outcomes, but on the macro level, the reduction in stolen base attempts affects the entertainment value of the game.
 
-    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting (\cite{castrovince_pitch_2023}).
+    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting \parencite{castrovince_pitch_2023}.
 With these changes, a pitch clock limits the time between pitches, and there are restrictions on infielder positioning, curtailing defensive shifts.
@@
 Each unsuccessful pickoff attempt disincentivizes the pitcher from attempting another pickoff, which incentivizes the runner to increase their lead distance.
+As explained by Stone before the Tucker stolen base against the White Sox, players and coaches understand this intuitively.
 But how much should the runner extend their lead?
@@
 
-    The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model (\cite{bukiet_markov_1997}).
+    The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model \parencite{bukiet_markov_1997}.
 \textcite{hirotsu_using_2002} used a four-state Markov model for soccer to inform substitution decision-making based on how substitutions impact transition probabilities between states.
@@
 Recently, \textcite{chan_points_2021} used a Markov model to evaluate team performance in football.
-More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis (\cite{nadimpalli_when_2013}), baseball (\cite{hirotsu_using_2019}), basketball (\cite{sandholtz_markov_2020}), darts (\cite{baird_optimising_2020}) and football (\cite{biro_reinforcement_2022}).
+More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \parencite{nadimpalli_when_2013}, baseball \parencite{hirotsu_using_2019}, basketball \parencite{sandholtz_markov_2020}, darts \parencite{baird_optimising_2020} and football \parencite{biro_reinforcement_2022}.
 
@@
 
-      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics (\cite{powers_sabrmetrics_2024}).
+      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
 The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
@@
 
-      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant (\cite{mlb_advanced_media_notitle_2024}) describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
 For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
@@
       \end{align*}
-      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 (\cite{bates_fitting_2015}).
+      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
 
@@
 
-        Modeling the progression of a baseball inning is a well-trodden application of Markov chains in sports.(citation needed) Traditionally, the state space is defined by the bases occupied and the number of outs (24 non-terminal states, e.g., citation needed) or by those and the ball-strike count (288 non-terminal states, e.g., citation needed).
+        Modeling the progression of a baseball inning is a well-trodden application of Markov chains in sports \parencite{bukiet_markov_1997}.
+Traditionally, the state space is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states).
 To understand the game theory under the new pickoff limit rule, we take this a step further and include the number of disengagements in the state space.
@@
 $b = (1, 0, 0)$.
-This limits our scope to situations in which the runner might steal second base (encompassing XX\%\footnote{TODO: fill this in} of stolen base attempts in 2023) and excludes the first-and-third scenario $b = (1, 0, 1)$, in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
+This limits our scope to situations in which the runner might steal second base (encompassing 71\% of stolen base attempts in 2023) and excludes the first-and-third scenario $b = (1, 0, 1)$, in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
 
@@
 
-        The relevant question for the runner is, ``What lead distance should I choose in each situation to maximize run scoring overall?'' We find the policy function $\pi^*$ which maximizes (\ref{eqn:run-expectancy}) via policy iteration (citation needed) by initializing $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, and then iteratively applying the updates
+        The relevant question for the runner is, ``What lead distance should I choose in each situation to maximize run scoring overall?'' We find the policy function $\pi^*$ which maximizes (\ref{eqn:run-expectancy}) via policy iteration \parencite{bellman_dynamic_1957} by initializing $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, and then iteratively applying the updates
         \begin{align}
@@
         \end{align}
-        until $\max_{s\in\mathcal{S}}|\pi_{i+1}(s) - \pi_i(s)|$ converges to zero.\footnote{Are we guaranteed convergence? Are we guaranteed convergence to the optimal policy?}
+        until $\max_{s\in\mathcal{S}}|\pi_{i+1}(s) - \pi_i(s)|$ converges to zero.\footnote{Do we need to prove that this will converge to the globally optimal policy?}
 
@@
 To find the optimal policy $\pi^* : \mathcal{S} \rightarrow \mathcal{L}$ for the runner, we again apply policy iteration.
-Starting with $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{TODO: Are we guaranteed convergence to the optimal policy?}
+Starting with $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{TODO: Are we guaranteed convergence to the optimal policy in the two-agent case?}
 
@@
 With each successive disengagement, we observe a positive shift in the distribution of lead distance in the 2023 data.
-In 2022, the average lead distance was XX feet.
-In 2023, the average lead distance under 0, 1 and 2 disengagements was XX feet, XX feet and XX feet, respectively.
+In 2022, the average lead distance was 9.6 feet.
+In 2023, the average lead distance under zero, one, and two disengagements was 9.6 feet, 10.3 feet and 11.0 feet, respectively.
 
-      \begin{figure}
+      \begin{figure}[H]
         \centering
@@
 
-      Figures \ref{fig:prob-pickoff-attempt}-\ref{fig:prob-sb-success} illustrate the relationship between lead distance and runner outcome probabilities, as estimated using the mixed-effects logistic regression models detailed in Section \ref{sec:prob-runner-outcome}.
+      Figures \ref{fig:prob-pickoff} and \ref{fig:prob-sb-success} illustrate the relationship between lead distance and runner outcome probabilities, as estimated using the mixed-effects logistic regression models detailed in Section \ref{sec:prob-runner-outcome}.
 For all three models, the direction of the relationship between lead distance and outcome probability is as expected.
 
-      Figure \ref{fig:prob-pickoff-attempt} shows pickoff attempt probability modeled as a function of lead distance, split by season and number of prior disengagements.
+      Figure \ref{fig:prob-pickoff} illustrates how pickoff attempt probability and success probability vary with distance.
 We observe that pickoff attempt rates were lower in 2023 than in 2022, especially as the number of prior disengagement increased.
-The modeled probability remains low even for extremely long leads, such as 20 feet.
+The modeled probability remains low even for extremely long leads.
 We have very limited data on lead distances beyond 14 feet, especially with fewer than two prior disengagements, so extrapolation much beyond 14 feet is likely not reliable.
 The ball-strike count and the identity of the pitcher also influence the pickoff attempt probability, but the primary drivers are lead distance and the number prior disengagements.
-
-      \begin{figure}
-        \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_po_attempt_light.pdf}
-        \caption{
-          \it Pickoff attempt probability modeled as function of lead distance, split by season and number of prior disengagements.
-This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-po-attempt}, assuming 0 balls, 0 strikes, 0 outs and a median pitcher random effect.
-        }
-        \label{fig:prob-pickoff-attempt}
-      \end{figure}
-
-      Figure \ref{fig:prob-pickoff-success} shows pickoff success probability modeled as a function of lead distance, split by pitcher pickoff skill (based on their random effect from the mixed-effect logistic regression).
-We observe that, except for the most skilled pitchers, pickoff attempt probability is very low within the range covering the vast majority of lead distances.
+We observe that, except for the most skilled pitchers, pickoff success probability is very low within the range covering the vast majority of lead distances.
 Toward the high end of this range, the success rate increases quickly as lead distance increases.
@@
 
-      \begin{figure}
+      \begin{figure}[H]
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_po_success_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{../../output/figures/prob_po_attempt_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{../../output/figures/prob_po_success_light.pdf}
         \caption{
-          \it Pickoff success probability modeled as function of lead distance, split by pitcher random effect.
-This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-po-success}, which includes only an intercept, a fixed effect for lead distance and a random effect for pitcher identity.
+          \it Pickoff attempt probability (left) and success probability (right) modeled as functions of lead distance.
+The left figure corresponds to the model detailed from Section \ref{sec:prob-po-attempt}, assuming 0 balls, 0 strikes, 0 outs and a median pitcher random effect.
+The right figure corresponds to the model from Section \ref{sec:prob-po-success}, which includes only an intercept, a fixed effect for lead distance and a random effect for pitcher identity.
         }
-        \label{fig:prob-pickoff-success}
+        \label{fig:prob-pickoff}
       \end{figure}
@@
       Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as function of lead distance, split by season and by runner skill (based on sprint speed and runner random effect).
-We observe that, conditioned on lead distance, even 10$^{th}$-percentile runners in 2023 outperformed the median runner in 2022, which is likely attributable to rule changes making it easier to successfully steal bases in 2023.
+We observe that, conditioned on lead distance, even 10$^{th}$-percentile runners in 2023 had higher success probabilities than the median runner in 2022, which is likely attributable to rule changes making it easier to steal bases in 2023.
 Within the range covering most lead distances, the difference between the 90$^{th}$-percentile runner and the median runner is just under 5 percentage points of success probability, similar to the difference between the median runner and the 10$^{th}$-percentile runner.
@@
 
-      \begin{figure}
+      \begin{figure}[H]
         \centering
@@
 
-      A novel aspect of these runner outcome probability models is the combination of fixed effects for measurable player characteristics (catcher arm strength and runner sprint speed).
+      A novel aspect of these runner outcome probability models is the combination of fixed effects for measurable player characteristics (catcher arm strength and runner sprint speed) and random effects for player identities.
 This feature allows us to answer questions such as: How important are these measurable characteristics for player performance outcomes?
@@
       Table \ref{tab:lead-by-count} shows how optimal lead distance varies by ball-strike count and number of prior disengagements, for an average runner facing an average battery (pitcher and catcher) with no outs.
-We observe that count has a mild influence on optimal lead distance, and lead distance has a strong influence.
+We observe that count has a mild influence on optimal lead distance, and the number of prior disengagements has a strong influence.
 As the pitcher nears the disengagement limit, the runner has more incentive to extend their lead distance because pickoff attempts are less likely and a free base advancement (from a third unsuccessful pickoff attempt) is more likely.
 
-      \begin{table}
+      \begin{table}[H]
         \centering
@@
         \caption{
-          \it The optimal lead distance policy (in feet) for an average runner facing an average battery (pitcher and catcher) with no outs.
+          \it The optimal lead distance (in feet) by count and prior disengagements for an average runner on first base, facing an average battery (pitcher and catcher) with no outs and third base empty.
+This lead distance policy maximizes the expected runs to end of inning as modeled by a the single-agent MDP.
         }
@@
       Across counts, the recommended increase in lead distance is generally close to 1.5 feet after the first disengagement and 2.5 feet after the second disengagement.
-We observed this to be a good approximation across different runner/battery skills and different numbers of outs.
-Based on these results, we recommend the rule of thumb that runners increase their lead by two feet after each disengagement, a very simple and actionable guideline for athletes to follow.
+We observe this to be a good approximation across different runner/battery skills and different numbers of outs.
+Based on these results, we recommend the {\it Two-Foot Rule}: that runners increase their lead by two feet after each disengagement, a very simple and actionable guideline for athletes to follow.
 
-      Table \ref{tab:lead-by-runners-outs} shows how optimal lead distance varies by number of outs, for an average runner facing an average batter with a 0-0 count, and Table \ref{tab:lead-by-players} shows how optimal lead distance varies by runner and battery skill, with 0 outs and a 0-0 count.
+      Table \ref{tab:lead-by-runners-outs} shows how optimal lead distance varies by number of outs, for an average runner facing an average battery with a 0-0 count.
+Again, we observe that each successive disengagement leads to an increase in recommended lead distance of approximately two feet.
+The recommended leads for zero outs and one out are virtually identical, and for two outs they are slightly longer.
+With two outs, stolen bases are more likely and pickoff attempts are less likely, so runners can be more aggressive.
 
-      \begin{table}
+      \begin{table}[H]
         \centering
-        \begin{tabular}{c|ccc}
-          Outs  & \multicolumn{3}{c}{Disengagements}\\
-                & 0 & 1 & 2 \\
-          \hline
-          \input{../../output/tables/lead_by_runners_outs.tex}
-        \end{tabular}
+        \input{../../output/tables/lead_by_outs.tex}
         \caption{
-          \it Based on the number of outs in each row, each entry shows the optimal lead distance (in feet) for a runner on first depending on the number of disengagements.
+          \it The optimal lead distance (in feet) by outs and prior disengagements for an average runner on first base, facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
+This lead distance policy maximizes the expected runs to end of inning as modeled by the single-agent MDP.
         }
@@
 
-      Again, we see that each successive disengagement leads to an increase in recommended lead distance of approximately two feet.
-We also see that the recommended leads for 0 and 1 out are virtually identical, and are slightly longer with 2 outs.
-With 2 outs, stolen bases are more likely and pickoff attempts are less likely, so runners can be more aggressive.
+      Table \ref{tab:lead-by-players} shows how optimal lead distance varies by battery and runner skill, with no outs and a 0-0 count.
+The battery skill accounts for catcher arm strength and for catcher and pitcher random effects.
+The composite runner skill accounts for sprint speed and for runner random effect.
+Against batteries more skilled at controlling the run game, the optimal lead distance is shorted.
+For more skilled runners, the optimal lead distance is longer.
+Regardless of the skill levels, the optimal policy for the runner is to increase their lead by 1.5--2.5 feet with each successive disengagement.
+We observe that the Two-Foot Rule is a good approximation for different combinations of battery and runner skill, not just for average players.
 
-      Table \ref{tab:players} shows the optimal lead distance for a runner on first base based on the players involved in the play.
-The runner skill level is based on sprint speed and their ability to steal bases.
-The battery skill level includes the pitcher and catcher's abilities to pick runners off and throw out potential basestealers, as well as the catcher's arm strength.
-This table assumes a 0-0 count, 0 outs, and nobody else on base:
-
-      \begin{table}
+      \begin{table}[H]
         \centering
@@
         \caption{
-          \it Based on the skill of the battery (pitcher and catcher) and the skill of the runner in each row, each entry shows the optimal lead distance for a runner on first base depending on the number of disengagements.
+          \it Optimal lead distance (in feet) by battery/runner skill and prior disengagements, for an 0-0 count with no outs and third base empty.
+Player skill is held constant across runner outcome models; for example, the hypothetical 90$^{th}$-percentile battery represents the 90$^{th}$ percentile at attempting pickoffs, successfully executing pickoffs, suppressing stolen base attempts, and catching would-be base stealers.
         }
@@
 
-      The effect of the battery seems to be more significant than the effect of the runner.
-Facing a stronger pitcher and catcher cause the runner's optimal lead distance to shrink by several feet.
-We again see consistent results that runners should increase their lead by 1.5-2.5 feet upon each successive disengagement.
-However, better runners don't need to increase their leads by as much, likely because their stolen base success rates will be high to begin with.
+      Table \ref{tab:actual-vs-rec} compares observed runner behavior against the optimal policy learned by the MDP.
+For each pitch with a runner on first base and second base empty, we compare the runner's lead distance with the recommended lead distance for an average runner against an average battery.
+With zero disengagements, the average lead is slightly less than the average recommended lead, and most leads are shorter than recommended.
+As the disengagements increase, this gap widens.
+On average, batters increase their leads by 0.6--0.7 feet afer each disengagement, significantly less than the Two-Foot Rule would recommend.
 
-      Currently, MLB runners are not doing a great job of putting pressure on the defense by maximizing their lead distances.
-Table \ref{tab:actual-vs-rec} compares the average actual and recommended leads for runners at 1st base with 2nd base open.
-With 0 disengagements, runners' actual leads come close to our recommendation, but they could often be adding an extra foot.
-As disengagements increase, we recommend about a 2 foot increase, but runners only tend to increase their leads by 0.6-0.7 feet.
-Runners overestimate the danger of a pickoff attempt under these situations, and could be increasing their team's run expectancy by adding a few feet to their lead distances here.
-Runners rarely meet or exceed the lead distance we recommend, especially with 1 or 2 disengagements.
-
-      \begin{table}
+      \begin{table}[H]
         \centering
@@
         \caption{
-          \it Each entry shows the average lead at 1st base based on the number of disengagements prior to the pitch.
-The recommended leads in the final column are based on our single-agent model.
-The final column shows the percent of pitches where the actual lead exceeds our recommendation.
+          \it Comparison of actual versus recommended lead distance by number of prior disengagements.
+The recommended lead distance is based on the single-agent MDP model and accounts for count, outs and prior disengagements, assuming average battery and runner skill.
         }
@@
 
-      In order to quantify the effect of optimizing baserunners' lead distances, we compared the run expectancy at the start of an inning before and after iterating through the Bellman Equation.
-In the 2023 season, the run expectancy of an inning was 0.5163 runs.
-After our policy iteration, the run expectancy jumped to 0.5277 runs.
-This increase of 0.0114 runs per inning may seem very small, but this difference adds up quickly.
-Over a 162 game season, adding 0.0114 runs per inning increases a team's run total by 16.68 runs.
-Using the typical valuation of \$1 Million per run, optimizing baserunners' lead distances could be worth nearly \$17 Million to a team under the current conditions.
-This is an incredibly valuable result that does not require much effort for a team to implement.
+      Our recommendations differ from runner behavior, and we can approximate the difference in run value between runner behavior and our recommendations.
+In 2023, the scoring expectation was 0.516 runs per inning.
+Under the optimal lead distance policy $\pi^*$, the corresponding value of the start-of-inning state is 0.528 runs, an increase of 0.012 runs per inning.
+While that may seem insigificant at first, it adds up to approximately 17 runs over 162 games of nine innings each.
+To add that many expected runs in free agency would cost roughly \$17 million \parencite{clemens_what_2021}, which speaks to a significant (but not outlandish) impact.
 
@@
 
-      All of the results shown so far correspond to our single-actor model, where a runner choose a lead distance and all other outcomes are probabilistic based on their choice.
-Now, we expand our model to a two-actor model, where the runner first chooses a lead distance, then the pitcher elects whether to attempt a pickoff or throw a pitch.
-The pitcher makes whichever decision minimizes the run expectancy for the offensive team.
-Therefore, in order for the runner to maximize their run expectancy, they should choose a lead distance such that the pitcher is indifferent as to whether to attempt a pickoff or throw a pitch.
-Pitchers should always attempt a pickoff if the runner's lead extends beyond this distance, and should always throw a pitch if the lead is shorter.
-This is the game theoretic equilibrium.
+      All of the results shown so far correspond to our single-agent model, in which the runner chooses a lead distance and all other outcomes are probabilistic based on their choice.
+While the single-agent model is the most immediately useful framework for runner decision-making in the current environment, we also consider how pitchers might respond using the same model to optimize their decision-making.
+In the two-agent model, the runner first chooses a lead distance, and the pitcher then decides whether to attempt a pickoff, as detailed in Section \ref{sec:two-agent-mdp}.
+The optimal strategy for the runner is maximin because they choose the lead distance which maximizes the minimum run expectancy across the pitcher's two options (pickoff or pitch).
 
-      Table \ref{tab:count-twoagent} shows the optimal lead by count under the two-agent model:
+      In the two-agent model, pitchers would be more aggressive with attempting pickoffs when runners take long leads.
+Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 8.3\% of opportunities with only first base occupied.
+When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.8\% of the time---this would be 100\% in the two-agent model.
+When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 5.0\% of the time---this would be 0\% in the two-agend model.
+Overall, pitchers attempted pickoffs on 5.8\% of opportunities, less than the recommmended 8.3\%.
 
-      \begin{table}
+      Table \ref{tab:count-twoagent} reports the optimal lead distance policy by count when both runner and pitcher are behaving optimally under the two-agent model.
+In this setting, the optimal lead distance is longer than in the single-agent model (Table \ref{tab:lead-by-count}), especially early in the count and with two prior disengagements.
+The longer lead distances are somewhat counterintuitive because we expect more pickoff attempts from pitchers when behaving optimally.
+The explanation is that in the two-agent model, the pitcher will never attempt a pickoff when it is suboptimal to do so.
+The optimal behavior for the runner is to force their hand (or rather, to make the pitcher indifferent to their two choices).
+Interestingly, the Two-Foot Rule breaks down in this setting.
+Averaged across counts, we recommend extending the lead by roughly 0.7 feet after the first disengagement and by much more after the second disengagement---roughly 4.2 feet!
+However, this advice will only be actionable if pitchers change their behavior to be more in line with the modeled optimal policy.
+
+      \begin{table}[H]
         \centering
@@
 
-      Under this game-theoretic equilibrium, we see that runners should be taking even longer leads in most situations.
-This is likely because pickoffs are often wasteful at lower lead distances.
-If pitchers were to optimize their pickoff behavior, then runners would be able to get longer leads before a pickoff becomes worthwhile for the pitcher.
-Our two-foot rule of thumb from the single-agent model doesn't hold here.
-Instead, we see a recommended increase of about 0.5-1 feet after 1 disengagement, and a recommended increase of around 3.5-5 feet after the 2nd disengagement.
-Once a pitcher reaches 2 disengagements, it becomes extremely risky to attempt a pickoff, so runners can take advantage of that by extending their leads beyond 15 feet.
-
-      In order to evaluate how well pitchers follow our recommendations, we looked back at all situations from the 2023 season with a runner on first base only.
-Our two-agent model suggests that pitchers should have attempted a pickoff in 8.3\% of these situations.
-In reality, they only attempted pickoffs 5.6\% of the time.
-
-      In situations where our model suggested a pickoff, pitchers actually attempted a pickoff 10.8\% of the time in 2023.
-
-      In situations where our model suggests throwing a pitch, pitchers actually attempted a pickoff 5.0\% of the time.
-
-      Overall, it seems that pitchers should be attempting more pickoffs.
-If runners start increasing their lead distances in the way we recommend, then pitchers should respond by picking off even more often.
-
-      Under this equilibrium of optimal pitcher and baserunner behavior, runners still hold a slight advantage compared to the status quo, but not by much.
-We calculated that runners could increase their team's run expectancy by about 0.0036 runs per inning by following these lead recommendations if pitchers also optimized their behavior.
-Over a 162 game season, this adds up to 5.2 runs.
-Ultimately, the current behavior of pitchers and runners is not far from the Nash Equilibrium.
-However, both sides could improve their team's performance by several runs by optimizing their behavior surrounding the running game.
+      If both runners and pitchers adapted their behavior to match the game theoretic optimal policy under the two-agent model, the corresponding run value of the start-of-inning state is 0.520 runs per inning.
+Relative to the current status quo, this reflects an increase of 0.004 runs per inning, meaning a very slight increase to overall offense.
+Over 162 games of nine innings each, we would expect each team to score approximately 5 more runs over a full season.
 
@@
 
-  \section{Assumptions}
+    Under the new pickoff rules, baseball players and coaches understand intuitively that runners want to increase their lead distance after each successive disengagement by the pitcher---but by how much?
+In this paper, we formalize the cat-and-mouse game between pitcher and runner as a Markov decision process and estimate the optimal lead distance policy for a runner on first base with other bases empty, given the number of outs, the ball-strike count, and the number of prior disengagements by the pitcher.
+We model the probabilities of runner outcomes (pickoff attempt/success, stolen base attempt/success) as functions of lead distance, context, and player skill.
+This allows us to estimate the optimal lead distance policy for any combination of pitcher skill and runner skill.
 
-    \begin{itemize}
-      \item NOTE (SP): Let's try including pitcher handedness in our model
-      \item NOTE (SP): Should we use catcher pop time instead of arm strength?
-      \item Successful pickoff eliminates runner with no other runner movement
-      \item Number of disengagements does not affect batter outcomes
-      \item Lead distance does not affect outcomes on batted balls (note: domain knowledge tells us that secondary lead is more important than the lead distance we observe, which is first pitcher movement)
-      \item We are ignoring the possibility that the runner takes off for 2B on a pickoff attempt
-      \item Other than pickoff attempts, no other disengagements occur
-    \end{itemize}
+    First, we consider a single-agent model in which the runner has agency in deciding their lead distance, and the pitcher's decision of whether to attempt a pickoff is probabilistic, conditioned on the lead distance.
+This framework leads to the most actionable recommendation for runners under the current conditions of pitcher behavior.
+We estimate that the average team could expect to increase their expected offensive output by approximately 17 runs over a full season by implementing the optimal lead distance policy.
+Based on these results, we propose the Two-Foot Rule: We recommend that a runner on first base increase their lead distance by two feet after each successive disengagement.
+This easily implemented recommendation is a good approximation to the optimal lead distance policy across a wide array of contexts (outs and ball-strike count) and pitcher/runner skills.
 
-  \section{Acknowledgments}
+    We go a step further and consider the consequences of granting the pitcher agency to decide whether to attempt a pickoff given the runner's lead distance.
+We give a game theoretic treatment to this two-agent model in which the runner follows a maximin strategy (with respect to rest-of-inning run expectancy) so that the pitcher is indifferent to attempting a pickoff or throwing a pitch.
+Under the game theoretic equilibrium, runners are more aggressive with their lead distance and pitchers are more aggressive with their pickoff attempts than the current behavior we observe.
+While baseball analytics sometimes garners negative attention for diminishing the aesthetic appeal of the game \parencite{sheinin_analytics_2025}, this is an example for which the analytically recommended strategy could be more aesthetically pleasing.
+
+    Our results come with some simplifying assumptions that are necessary limitations given the data available.
+We assume that observed lead distance fully describes the runner's behavior and that they cannot disguise their aggressiveness by leaning in one direction.
+In reality, the runner may achieve a greater (or lesser) effective lead distance than we observed by leaning toward second base (or first base).
+We also assume that lead distance only impacts pickoff and stolen base outcomes, not the outcomes of batted balls.
+For example, we assume that a greater lead distance does not improve the runner's probability of advancing from first base to third base on a single.
+Baseball domain knowledge suggests this is a reasonable assumption because while a pitch is being thrown, the runner moves to a ``secondary'' lead distance which is more important for determining advancement outcomes than primary lead distance (at first pitcher movement).
+Finally, we assume that, other than pickoff attempts, no other disengagements occur.
+In reality, the pitcher may derive value from stepping off of the pitching rubber if they are short of breath or need to resolve a miscommunication with the catcher.
+
+    In addition to the lead distance data used in the current work, which is based on the runner's center of mass, MLB has begun collecting 18-point pose tracking data on runners and fielders at 50 frames per second \parencite{jedlovec_introducing_2020}.
+If these data were publicly available, they would enable more granular modeling of the cat-and-mouse game between pitcher and runner.
+For example, with information about the runner's leaning posture and reaction time, future work might show that a mixed strategy for the pitcher's pickoff attempt decision achieves the best result by forcing a neutral posture for the runner and delaying the runner's reaction.
+For now, we present the Two-Foot Rule as a simple rule of thumb that provides easily digestible and actionable guidance to runners.
+
+  \section*{Acknowledgments}
 
     Major League Baseball trademarks and copyrights are used with permission of MLB Advanced Media, L.P. All rights reserved.
+
+  \section*{Code and Data Availability}
+
+    All code related to this paper is available at \url{https://github.com/jfhahn2/pickoff-game-theory}.
+Lead distance data are not publicly available and were obtained from Major League Baseball.
+All other data are publicly available and may be downloaded using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
 

### 2025-08-03 scripts/estimate_models.R
Complete the first full draft of the paper (#4)
* switch from natbib to biblatex

* add code missing from generate_results.R

* switch paper to new light mode figures

* switch paper to new tables

* force locations of figures and tables

* biblatex can handle underscores in URLs

* rotate lead by count tables wide

* move lead-by-outs table to generate_results.R

* remove old bibliography style file

* transpose tab:lead-by-outs

* write transposed lead-by-count tables to file

* remove deprecated code for unsuccuessful pickoff transitions

* add code for reporting average lead distance by sc

@@ -160,21 +160,6 @@ succesful_pick <- function(State) {
   return(paste0("00", substr(State, 3, 3), " 0 ", substr(State, 7, 8), as.numeric(outs)+1))
 }
 
-# Function for what happens on unsuccesful pickoff
-# Assumes no pickoff errors or other runners moving
-unsuccesful_pick <- function(State) {
-  if (substr(State, 1, 2) != "10") {
-    return(State)
-  }
-  dis <- substr(State, 5, 5) 
-  if (dis == 2) {
-    return(paste0("010 0 ", substr(State, 7, 9)))
-  } else {
-    return(paste0(substr(State, 1, 4), as.numeric(dis) + 1, substr(State, 6, 9)))
-  }
-  
-}
-
 # Call function on all states
 all_states <- data.frame(State = all_possible_states)
 T_sp <- all_states %>% mutate(New_State = sapply(State, succesful_pick), Freq = 1)
@@ -185,19 +170,6 @@ all_transitions_tsp <- state_grid %>% left_join(T_sp, by = c("State", "New_State
 # Create final transition prob matrix
 P_SP <-  all_transitions_tsp %>% group_by(State) %>% mutate(Total = sum(Freq)) %>% ungroup() %>% mutate(Prob = Freq / Total)
 P_SP <- P_SP %>% mutate(Prob = ifelse(is.na(Prob), ifelse(New_State == State, 1, 0), Prob)) 
-
-
-# P^UP MATRIX (Failed Pick)
-
-# Call function on all states
-T_up <- all_states %>% mutate(New_State = sapply(State, unsuccesful_pick), Freq = 1)
-
-# join state grid with transition probs
-all_transitions_tup <- state_grid %>% left_join(T_up, by = c("State", "New_State")) %>% mutate(Freq = ifelse(is.na(Freq), 0 , Freq))
-
-# Create final transition prob matrix
-P_UP <-  all_transitions_tup %>% group_by(State) %>% mutate(Total = sum(Freq)) %>% ungroup() %>% mutate(Prob = Freq / Total)
-P_UP <- P_UP %>% mutate(Prob = ifelse(is.na(Prob), ifelse(New_State == State, 1, 0), Prob)) 
 
 
 # P^UP MATRIX (Failed Pick)
@@ -831,22 +803,6 @@ for (i in 1:nrow(skill_grid)) {
 }
 
 
-# SUMMARY TABLES ----
-
-
[…]

### 2025-08-03 scripts/generate_results.R

@@ -1,6 +1,6 @@
 
 fig_make <- TRUE
-fig_mode <- "dark"
+fig_mode <- "light"
 
 
 # Read data ----
@@ -44,6 +44,10 @@ lead_distance_1b <- rbind(second_open_22, second_open_23) |>
   ) |>
   dplyr::filter(!is.na(year_dis))
 
+lead_distance_1b |>
+  dplyr::group_by(year_dis) |>
+  dplyr::summarize(mean(`lead_distance_1st Base`))
+
 breaks <- seq(from = 4, to = 16, by = 2)
 
 if (fig_make) {
@@ -63,7 +67,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(x = "Lead Distance", y = "Density") +
     ggplot2::coord_cartesian(xlim = c(3, 16)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.2, 0.8))
   print(plot)
   dev.off()
@@ -126,7 +130,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(title = "Pickoff Attempt", x = "Lead Distance", y = "Probability") +
     ggplot2::coord_cartesian(xlim = c(3, 16), ylim = c(0, 0.7)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.4, 0.7))
   print(plot)
   dev.off()
@@ -177,7 +181,7 @@ if (fig_make) {
     ) +
     ggplot2::labs(title = "Pickoff Success", x = "Lead Distance", y = "Probability") +
     ggplot2::coord_cartesian(xlim = c(3, 16), ylim = c(0, 0.7)) +
-    sputil::theme_sleek() +
+    sputil::theme_sleek(mode = fig_mode) +
     ggplot2::theme(legend.position.inside = c(0.3, 0.7))
   print(plot)
   dev.off()
@@ -204,7 +208,7 @@ if (fig_make) {
       y = "Effect on SB Success (log-odds)"
     ) +
     ggplot2::coord_cartesian(ylim = c(-0.6, 0.6)) +
-    sputil::theme_sleek()
+    sputil::theme_sleek(mode = fig_mode)
   print(plot)
   dev.off()
 }
@@ -230,7 +234,7 @@ if (fig_make) {
       y = "Effect on SB Success (log-odds)"
     ) +
[…]

### 2025-08-12 articles/arxiv/paper.tex
Update project on Overleaf

@@
 \usepackage{amsfonts, amsmath, float, graphicx}
+
+\graphicspath {{../../output/figures/}}
 
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/leads_overall_light.pdf}
+        \includegraphics[width = 0.8\textwidth]{leads_overall_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.4\textwidth]{../../output/figures/prob_po_attempt_light.pdf}
-        \includegraphics[width = 0.4\textwidth]{../../output/figures/prob_po_success_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{prob_po_attempt_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{prob_po_success_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/prob_sb_success_light.pdf}
+        \includegraphics[width = 0.8\textwidth]{prob_sb_success_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.49\textwidth]{../../output/figures/effect_catcher_light.pdf}
-        \includegraphics[width = 0.49\textwidth]{../../output/figures/effect_runner_light.pdf}
+        \includegraphics[width = 0.49\textwidth]{effect_catcher_light.pdf}
+        \includegraphics[width = 0.49\textwidth]{effect_runner_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{../../output/figures/finding_optimal_lead_light.pdf}
+        \includegraphics[width = 0.8\textwidth]{finding_optimal_lead_light.pdf}
         \caption{

### 2025-09-12 articles/arxiv/paper.tex
Siva's edits

@@
-\documentclass{article}
+\documentclass[11pt]{article}
 
@@
 
-\usepackage{amsfonts, amsmath, float, graphicx}
+\usepackage{amsmath, amssymb, amsthm, amsfonts, color, enumerate, framed, graphicx, mathrsfs, mathtools, mdframed, natbib, subcaption, tabularx, theoremref, titlesec, verbatim, xfrac, soul, mathtools, wrapfig, tikz, float}
 
 \graphicspath {{../../output/figures/}}
+
+\newtheorem*{theorem*}{Theorem}
+\newtheorem{theorem}{Theorem}
+\newtheorem*{lemma*}{Lemma}
+\newtheorem{lemma}{Lemma}
+\newtheorem*{conjecture*}{Conjecture}
+\newtheorem{conjecture}{Conjecture}
+\newtheorem{proposition}{Proposition}
+\newtheorem*{assumption*}{Assumption}
+\newtheorem{assumption}{Assumption}
+\newtheorem*{definition*}{Definition}
+\newtheorem{definition}{Definition}
+
+\usepackage[colorinlistoftodos]{todonotes}
+\usepackage[normalem]{ulem}
+
+\newcommand{\argmax}{\operatornamewithlimits{argmax}}
+\newcommand{\argmin}{\operatornamewithlimits{argmin}}
+\newcommand{\defeq}{\stackrel{\text{def}}{=}}
+\newcommand{\abs}[1]{\left\lvert#1\right\rvert}
 
@@
 
+    \section{The baseball inning as a two-player stochastic game}
+    \label{sec:sequential_model}
+    We first describe a single-player setup where only the first-base runner has agency.
+This is modeled as an infinite-horizon Markov decision process (MDP).
+We then extend this model to a two-player case where the first-base runner and the pitcher have agencies.
+The two-player case is modeled as an infinite-horizon zero sum stochastic game.
+\todo{Explain how the single-player case, despite being simple and restrictive than the two-player case, still provides valuable and actionable insights.} Since the state space for the single- and two-player models are the same, we first describe the (common) state space before explaining the specifics of each model.
+
+    {\bf State space.} The game state is given by $s = (b, c, d, o)$, where:
+    \begin{align*}
+      b &\in \{0, 1\}^3 \mbox{ represents which bases are occupied by runners};\\
+      c &\in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ represents the ball-strike count};\\
+      d &\in \{0, 1, 2\} \mbox{ represents the number of disengagements already used}; and\\
+      o &\in \{0, 1, 2\} \mbox{ represents the number of outs}.
+    \end{align*}
+    The total number of non-terminal states is $2^3 \times (4 \times 3) \times 3 \times 3 = 864$.
+There is one terminal state (end of inning), denoted $\Delta$, bringing the total number of states to 865.
+We use $\mathcal{S}$ to denote the set of all states ( including the terminal state), and $S_t$ the state at play $t$ in a (fixed) innings.
+Traditionally, the state is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states) \parencite{bukiet_markov_1997}.
+Our construction of the state extends them by further including the number of disengagements as a component of the state, thereby allowing us to analyze the game under the new pickoff limit rule.
+
+    \subsection{Single-player Markov decision process}
+    \label{sec:mdp_model}
+    This section considers the setup where agency is allowed only for the first-base runner. We model this scenario as a MDP. The other components of this MDP are as follows.
+
+    {\bf State space.} The state space for the MDP is the one described in the beginning of Section \ref{sec:sequential_model}.
+
+    {\bf Action space.} We limit the agency to states in which first base is occupied and the other bases are unoccupied.
+For such states, the action is the lead distance in feet chosen by the first-base runner from the first base.
+We formally describe it as follows.
+Define ${\cal L} \defeq [0,20]$.
+The action space of the runner, $A_\text{R}(s) \ \forall s \in {\cal S}$, is then defined as
+    \begin{align}
+    \label{eqn:runner_action_sapce}
+        A_\text{R}(s) \defeq \begin{cases}
+            {\cal L} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta.        \end{cases}
+    \end{align}    
+     Restricting the agency to states in which first base is occupied and the other bases are unoccupied, i.e., $b = (1, 0, 0)$, limits our scope to situations in which the runner might steal second base (encompassing 71\% of stolen base attempts in 2023).
+This obviously excludes the first-and-third scenario ($b = (1, 0, 1)$) in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
+
+     {\bf Single-stage reward function.} The single-stage reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition.
+This reward function does not depend on the action taken by the runner, and is given as
+     \begin{align}
+         \label{eqn:runner_reward}
+         r(s'|s,a) \defeq r(s'|s) \ \forall s,s' \in {\cal S}, a \in A_\text{R}(s),
+     \end{align}
+     where
+        \begin{align}
+          \label{eqn:common_reward}
+          r(s'|s) \defeq  \begin{cases}
+              (g(b) + o) - (g(b') + o') + \mathbb{I}\{c' = (0, 0),\, d' = 0\} \ \forall s,s' \in {\cal S} - \{\Delta\}\\
+              0 \text{ if } s = \Delta,
+          \end{cases} 
+        \end{align}
+        where $g: \{0, 1\}^3 \rightarrow \{0, 1, 2, 3\}$ counts the number of runners on base.
+This expression for the single-stage reward function counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance.
+For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state.
+To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state.
+These intermediary states transition to the terminal end-of-inning state with probability 1.
+
+    {\bf Transition probabilities.} \todo[inline]{Since the derivation of transition probabilities is involved, write it in a separate section and point to that section.
+Describe here in at most a couple of sentences how those probabilities are computed.
+Highlight which states transition to the absorbing state with positive probability.}
+
+    The MDP is then defined by the tuple ${\cal M} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{r(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}, \\  \{p(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}\rangle$.
+Using this MDP, the progression of an inning can be captured as follows.
+At the beginning of a play, the game is in state $s \in {\cal S}$.
+Upon observing this state, an action $a \in A_\text{R}(s)$ is picked.
+The game then transitions to state $s'$ with probability $p(s'|s,a)$ and the batting team collects a reward $r(s'|s,a)$.
+A play then begins from $s'$ and the process repeats.
+The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
+
+   A (deterministic stationary Markovian) policy $\pi$ for the runner is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi(s) \in A_\text{R}(s)$.
+That is, $\pi(s)$ is the action chosen in state $s$.
+\sout{In other words, a policy is a contingency plan (in this specific context) employed by the batting team.} Denote the set of all policies by $\Pi_\text{R}$.
+Note that $\Pi_\text{R}$ is uncountable but compact.
+It is intuitive that an inning in a baseball eventually halts.
+The following assumption formalizes this intuition.
+   \begin{assumption}\label{assum:mdp_halts}
+       \normalfont
+      The MDP ${\cal M}$ satisfies
+      $
+      \lim_{t \to \infty}\sup_{\pi \in \Pi, s \in {\cal S}}\mathbb{P}^\pi[S_t = \Delta | S_0 = s] = 1.
+      $
+   \end{assumption}   
+  \noindent The above assumption states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
+
+   The value of a policy $\pi$ when starting from state $s \in {\cal S}$ equals  
+   \begin{equation}
+       \label{eqn:mdp_pol_val}
+       V^{\pi}(s) \defeq \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
+   \end{equation}
+  That is, the value of the policy is the the expected total runs scored in an inning.
+Under Assumption \ref{assum:mdp_halts}, $V^\pi(s)$ exists for all $\pi \in \Pi$ and $s \in {\cal S}$.
+To see that, since $r(s'|s) \geq 0$ for all $s,s' \in {\cal S}$, we have by the monotone convergence theorem
+  \[
+  V^\pi(s) = \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] = \lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall \pi \in \Pi, s \in {\cal S}.
+  \]
+   Using Assumption \ref{assum:mdp_halts} and the fact that $\sup_{s,s' \in {\cal S}}\abs{r(s'|s)} < \infty$ (due to finiteness of ${\cal S}$), it is easy to establish that $\lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right]$ exists for all $\pi \in \Pi, s \in {\cal S}$.
+Furthermore, the limit is uniformly bounded over $\Pi \times {\cal S}$.
+The value of the MDP is defined as
+   \begin{equation}
+        \label{eqn:mdp_val}
+       V^*(s) \defeq \max_{\pi \in \Pi_\text{R}}V^\pi(s) \ \forall s \in S.
+   \end{equation}
+   The function $V^*$, known as the optimal value function, is the maximum expected runs scored starting from state $s \in {\cal S}$.
+A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called the optimal policy.
+
+   We present the following definitions that will be needed to characterize the optimal value function and the optimal policy.
+
+   \begin{definition}\label{defn:mdp_operators}
+    \normalfont
+    \begin{enumerate}
+        \item For any $\pi \in \Pi_\text{R}$, the evaluation operator $\Phi_\pi:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
+        \[
+        (\Phi_\pi V)(s) \defeq \mathbb{E}_{s' \sim p(\cdot|s,\pi(s))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+        \]
+        \item The optimality operator $\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
+        \[
+        (\Phi V)(s) \defeq \max_{a \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a)}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+        \]
+    \end{enumerate}       
+   \end{definition}
+
+   \begin{theorem}[\parencite{bertsekas1991analysis}]\label{thm:mdp_optimality_condition}
+       \normalfont
+      Suppose Assumption \ref{assum:mdp_halts} holds. Then
+      \begin{enumerate}
+          \item For any policy $\pi$, its value $V^\pi$ is the unique fixed point of the operator $\Phi_\pi$.
+          \item The optimal value function $V^*$ is the unique fixed point of the operator $\Phi$. Furthermore, $\pi^*$ is an optimal policy if and only if $\Phi_{\pi^*}V^* = \Phi V^*$.
+      \end{enumerate}
+   \end{theorem}
+% \todo[inline]{The above theorem also requires $p(s'|s,a)$ to be continuous in $a$ for any $s,s' \in S$.
+This is true for fixed-effects model since $x \mapsto e^x/1+e^x$ is continuous.
+I do not know about random-effects model.
+[[SP: $a$ comes into the model as a fixed effect (the random effects are for player identities), so, yes, $p(s'|s,a)$ is continuous in $a$.]]}
+
+The above theorem enables the characterization of $V^*(s)$ as in the following lemma.
+
+\begin{lemma}\label{lem:mdp_value_function_characterization}
+    \normalfont
+    Suppose Assumption \ref{assum:mdp_halts} holds. Then $V^*$ is the unique solution to the following system of equations.
+    \begin{align*}
+        V(s) &= \max_{\ell \in {\cal L}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
+        V(s) &= \sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
+        V(s) &= 0 \quad \text{ if } s = \Delta.
+    \end{align*}
+    \begin{proof}
+        Immediate from Theorem \ref{thm:mdp_optimality_condition}.
+    \end{proof}
+\end{lemma}
+
+An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{bertsekas1991analysis}.
+We use policy iteration  in this work to compute the optimal policy which is as follows.
+Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi^k$ by applying the updates
+        \begin{align}
+          \label{eqn:update-value-single-agent}
+          V^{k+1}(s) &= \sum_{s' \in \mathcal{S}} p(s' | s,\, \pi^k(s)) [r(s, s') + V^k(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
+          \label{eqn:update-policy-single-agent}
+          \pi^{k+1}(s) &= \argmax_{a \in A_\text{R}(s)} \sum_{s' \in \mathcal{S}} p(s' | s,\, a) [r(s, s') + V^{k+1}(s')] && \forall s \in \mathcal{S}.
+        \end{align}
+    The iterates generated by the policy iteration satisfy $V^k \uparrow V^*$ and every limit point of $\pi^k$ is an optimal policy \parencite{bertsekas1991analysis}.
+
+    \subsection{Two-player Stochastic Game}
+    \label{sec:stochastic_game}
+    In this section, we extend the single-agent model by allowing the pitcher to choose whether to attempt a pickoff.
+This is modeled as a two-player zero sum stochastic game, where the runner first chooses a lead distance, and then the pitcher decides whether to attempt a pickoff after observing the runner's lead distance.
+The components of this stochastic game are as follows.
+
+    {\bf State space.} The state space  is the same as described in the beginning of Section \ref{sec:sequential_model}.
+
+    {\bf Action space.} As in the single-player model of Section \ref{sec:mdp_model}, we limit actions (of runner as well as pitcher) to states in which only the first base is occupied.
+The (first-base) runner chooses a lead distance, and then the pitcher chooses a pickoff action $p \in \mathcal{P} \defeq \{0,1\}$, where $p = 1$ corresponds to a pickoff attempt and $0$ otherwise.
+The action space of the runner, $A_\text{R}(s) \ \forall s \in {\cal S}$, is the same as \eqref{eqn:runner_action_sapce}.
+The pitcher's action space $A_{\text{P}}(s,a_\text{R}) \ \forall s \in {\cal S}, a_\text{R} \in A_\text{R}(s)$ is defined as
+    \begin{align}
+    \label{eqn:pitcher_action_sapce}
+        A_{\text{P}}(s,a_\text{R}) \defeq \begin{cases}
+            {\cal P} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0), \ a_\text{R} \in A_\text{R}(s) \\
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta, \ a_\text{R} \in A_\text{R}(s).        \end{cases}
+    \end{align}
+
+    {\bf Single-stage reward function.} The single-stage reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition.
+Like in the single-player model, this reward function does not depend on the action taken by the runner, and is given
+     \begin{align}
+         \label{eqn:runner_pitcher_reward}
+         r(s'|s,a_\text{R},a_\text{P}) \defeq r(s'|s) \ \forall s,s' \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})
+     \end{align}
+     where $r(s'|s) \forall s,s' \in {\cal S}$ is as defined in \eqref{eqn:common_reward}.
+For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state.
+To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state.
+These intermediary states transition to the terminal end-of-inning state with probability 1.
+
+     {\bf Transition probabilities.} \todo[inline]{Since the derivation of transition probabilities is involved, write it in a separate section and point to that section.
+Describe here in at most a couple of sentences how those probabilities are computed.
+Highlight which states transition to the absorbing state with positive probability.}
+
+     The stochastic game is then defined by the tuple ${\cal G} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{A_\text{P}(s,a_\text{R})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s)},\\ \{r(s'|s,a_\text{R},a_\text{P})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})}^{s' \in {\cal S}}, \{p(s'|s,a_\text{R},a_\text{P})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s, a_\text{R})}^{s' \in {\cal S}}\rangle$.
+The progression of an inning is then modeled as follows.
+At the beginning of a play, the game is in state $s \in {\cal S}$.
+Upon observing this state, the batting team picks action $a_\text{R} \in A_\text{R}(s)$.
+The pitcher then picks an action $a_\text{P} \in A_\text{P}(s,a_\text{R})$.
+The game then transitions to state $s'$ with probability $p(s'|s,a_\text{R},a_\text{P})$ and the batting team collects a reward $r(s'|s,a_\text{R},a_\text{P})$.
+A play then begins from $s'$ and the process repeats.
+The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning while the goal of the pitcher is to minimize the expected total rewards obtained by the batting team.
+
+   A (deterministic stationary Markovian) policy $\pi_\text{R}$ for the runner is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi_\text{R}(s) \in A_\text{R}(s)$.
+That is, $\pi_\text{R}(s)$ is the action chosen in state $s$.
+Similarly, a (deterministic stationary Markovian) policy $\pi_\text{P}$ for the pitcher is a mapping of the form $\{(s,a) \in {\cal S} \times \cup_{s \in {\cal S}}A_\text{R}(s) \vert a \in A_\text{R}(s)\} \ni (x,a) \mapsto A_\text{P}(x,a)$ with the interpretation that $\pi_\text{P}(x,a_\text{R})$ is the action chosen by the pitcher when the state is $s$ and the runner picks the action $a_\text{R}$.
+Denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
+   Note that $\Pi_\text{R}$ and $\Pi_\text{P}$ are uncountable but compact. It is intuitive that an innings in a baseball eventually halts. The following assumption formalizes this intuition.
+   \begin{assumption}\label{assum:game_halts}
+       \normalfont
+      The stochastic game ${\cal G}$ satisfies
+      $
+      \lim_{t \to \infty}\sup_{\pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}}\mathbb{P}^{\pi_\text{R}, \pi_\text{P}}[S_t = \Delta | S_0 = s] = 1.$
+   \end{assumption}
+
+   The value of the policy pair $(\pi_\text{R},\pi_\text{P}) \in \Pi_\text{R} \times \Pi_\text{P}$ is given by
+    \begin{equation}
+       \label{eqn:sg_pol_val}
+       \widetilde V^{\pi_\text{R},\pi_\text{P}}(s) \defeq \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
+   \end{equation}
+   Similar to the single-player case, it follows that Assumption \ref{assum:game_halts} along with the finiteness of $\cal S$ implies that $\widetilde V^{\pi_\text{R},\pi_\text{P}}(s)$ exists for all $(\pi_\text{R},\pi_\text{P}) \in \Pi_\text{R} \times \Pi_\text{P}, s \in {\cal S}$ and is uniformly bounded over $\Pi_\text{R} \times \Pi_\text{P} \times {\cal S}$.
+The value of the stochastic game is then defined as
+   \begin{equation}
+        \label{eqn:sg_val}
+       \widetilde V^*(s) \defeq \max_{\pi_\text{R} \in \Pi_\text{R}}\min_{\pi_\text{P} \in \Pi_\text{P}}\widetilde V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in S.
+   \end{equation}
+   The function $\widetilde V^*$ will be referred to as the value of the game.
+A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $\widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = \widetilde V^*(s) \ \forall s \in {\cal S}$ is called the optimal strategy or equilibrium strategy.
+Note that $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an optimal strategy if and only if $\widetilde V^{\pi_\text{R},\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}}(s) \ \forall \pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}$.
+
+   We present the following definitions that will be needed to characterize the optimal value function and the optimal policy.
+
+   \begin{definition}\label{defn:sg_operators}
+    \normalfont
+    \begin{enumerate}
+        \item For any $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, the evaluation operators $\widetilde\Phi_{\pi_\text{R}}:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ and $\widetilde\Phi_{\pi_\text{P}}:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ are defined as
+        \begin{align*}
+        (\widetilde\Phi_{\pi_\text{R}} V)(s) &\defeq \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}(s))}\mathbb{E}_{s' \sim p(\cdot|s,\pi_\text{R}(s),a_\text{P})}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}\\
+        (\widetilde\Phi_{\pi_\text{P}} V)(s) &\defeq \max_{a_\text{R} \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},\pi_\text{P}(s,a_\text{R}))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+        \end{align*}
+        \item The optimality operator $\widetilde\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
+        \[
+        (\widetilde\Phi V)(s) \defeq \max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+        \]
+    \end{enumerate}       
+   \end{definition}
+
+   \begin{theorem}[\parencite{patek1999stochastic}]\label{thm:sg_optimality_condition}
+       \normalfont
+      Suppose Assumption \ref{assum:game_halts} holds.
+Then the value of the game $\widetilde V^*$ is the unique fixed point of the operator $\widetilde \Phi$.
+Furthermore, $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an optimal strategy if and only if $\widetilde\Phi_{\pi_\text{R}^*}\widetilde V^* = \widetilde\Phi \widetilde V^* = \widetilde\Phi_{\pi_\text{P}^*}\widetilde V^*$.
+   \end{theorem}
+
+   The above theorem enables the characterization of $\widetilde V^*(s)$ as in the following lemma.
+
+\begin{lemma}\label{lem:sg_value_characterization}
+    \normalfont
+    Suppose Assumption \ref{assum:game_halts} holds. Then $\widetilde V^*$ is the unique solution to the following system of equations.
+    \begin{align*}
+        V(s) &= \max_{\ell \in {\cal L}}\min_{a \in {\cal P}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell,a) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
+        V(s) &= \sum_{s' \in {\cal      S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
+        V(s) &=  0 \quad \text{ if } s = \Delta.
+    \end{align*}
+    \begin{proof}
+        Immediate from Theorem \ref{thm:sg_optimality_condition}.
+    \end{proof}
+\end{lemma}
+
+   An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{patek1999stochastic}.
+We use policy iteration  in this work to compute the optimal policy which is as follows.
+Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi_\text{R}^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi_\text{R}^k$ by applying the updates
+        \begin{align}
+          \label{eqn:update-value-two-agent}
+          V^{k+1}(s) &= \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}^k(s))} \sum_{s' \in \mathcal{S}} [r(s, s') +  V^k(s')]p(s'|s,\pi_\text{R}^k(s),a_\text{P}) && \forall s \in \mathcal{S}, \mbox{ and}\\
+          \label{eqn:update-policy-two-agent}
+          \pi_\text{R}^{k+1}(s) &= \argmax_{a_\text{R} \in A_\text{R}(s)} \left\{\min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s, s') + V^{k+1}(s')] p(s'|s,a_\text{R},a_\text{P})\right\} && \forall s \in \mathcal{S}.
+        \end{align}
+    The iterates generated by the policy iteration satisfy $V^k \uparrow \widetilde V^*$\parencite{patek1999stochastic}.
+
   \section{Results}

### 2025-09-12 articles/arxiv/paper.tex
Powers re-organizes paper

@@
 
-  \section{Data}
-
-    \subsection{MLB Stats API}
-    \label{sec:statsapi}
-
-      Most of the data used for this article were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
-The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
-For each play, we observe the identities of the pitcher, catcher and runners involved.
-Before and after each play, we observe the state of the game, described by the number of outs in the inning; which bases are occupied by runners; the ball-strike count on the batter; and the number of disengagements already used by the pitcher.
-On each pickoff attempt, we observe whether it is successful (runner is out) or unsuccessful (runner is safe).
-On each pitch, we observe whether the runner attempts a stolen base; and if so, we observe whether they are successful (runner is safe) or unsuccessful (runner is out).
-On each pitch terminating a plate appearance (a batting attempt by a batter), we observe the plate appearance outcome, such as strikeout, groundout, single or home run.
-
-    \subsection{Baseball Savant}
-    \label{sec:baseballsavant}
-
-      In addition to the play-by-play data described in Section \ref{sec:statsapi}, we manually downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
-For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
-This is defined as the average of the top 10 percent of throw speeds when the catcher is attempting to catch a base stealer.
-For each runner-season, we observe {\it sprint speed} (feet per second), which measures the runner's top running speed in their fastest one-second window.
-This is defined as the average of the top 67 percent of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
-
-    \subsection{Proprietary Statcast Data}
-
-      In addition to the publicly available data described in Sections \ref{sec:statsapi} and \ref{sec:baseballsavant}, this article relies on one key piece of proprietary information provided by MLB Advanced Media, estimated using computer vision on video data from several camera angles.
-For each play, observe each runner's {\it lead distance} (feet), which is how far they are ranging from their base at the time of the pitcher's first movement.
-This is defined as the distance between the runner's center of mass (projected onto the line between the runner's current base and their next base) and the nearest edge of the runner's current base.
-When the pitcher throws a pitch, runners will typically extend their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a {\it secondary lead}), which is relevant for runner advancement on batted balls but not on stolen base attempts.
-Our data include only the lead distance at the time of the pitcher's first movement, not the secondary lead.
-
-  \section{Methods}
-
-    \subsection{Notation}
-
-    Throughout this section, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
-    \begin{align*}
-      b_i & \in \{0, 1\}^3 \mbox{ denotes which bases (1B, 2B, 3B, respectively) are occupied;}\\
-      (c_i^B, c_i^S) \equiv c_i & \in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ denotes the ball-strike count;}\\
-      d_i & \in \{0, 1, 2\} \mbox{ denotes the number of disengagements already used; and}\\
-      o_i & \in \{0, 1, 2\} \mbox{ denotes the number of outs.}
-    \end{align*}
-    We use $\mathcal{H}^R$, $\mathcal{H}^P$, and $\mathcal{H}^C$ to denote the sets of runners, pitchers and catchers, respectively. On play $i$:
-    \begin{align*}
-      h_i^R & \in \mathcal{H}^R \mbox{ is the runner;}\\
-      h_i^P & \in \mathcal{H}^P \mbox{ is the pitcher;}\\
-      h_i^C & \in \mathcal{H}^C \mbox{ is the catcher;}\\
-      z_i^R & \in \mathbb{R^+} \mbox{ is the sprint speed of runner $h_i^R$; and}\\
-      z_i^C & \in \mathbb{R^+} \mbox{ is the arm strength of catcher $h_i^C$.}
-    \end{align*}
-    When the play begins, we use the following notation to describe the actions taken by the players:
-    \begin{align}
-    \label{eqn:runner-outcome}
-      \begin{split}
-        \ell_i  \in &~ \mathbb{R}^+ \mbox{ denotes the lead distance (in feet) taken by the first-base runner;}\\
-        p_i \in &~ \{0, 1\} \mbox{ denotes whether the pitcher attempts (1) or does not attempt (0) a pickoff; and}\\
-        r_i \in &~ \{\mbox{P}^+,\, \mbox{P}^-,\, \mbox{S}^+,\, \mbox{S}^-,\, \mbox{N}\} \equiv \mathcal{R} \mbox{ denotes the runner outcome, where}\\
-          & \mbox{P}^+  \mbox{ represents a successful pickoff attempt (runner is out);}\\
-          & \mbox{P}^-  \mbox{ represents an unsuccessful pickoff attempt (runner is safe);}\\
-          & \mbox{S}^+  \mbox{ represents a successful stolen base attempt (runner is safe);}\\
-          & \mbox{S}^-  \mbox{ represents an usuccessful stolen base attempt (runner is out); and}\\
-          & \mbox{N}~   \mbox{ represents no runner action (a pitch with no stolen base attempt).}
-%        p_i & \in \{0, 1\} \mbox{ indicates whether the pitcher attempts a pickoff on the first-base runner; and}\\
-%        s_i & \in \{0, 1\} \mbox{ indicates whether the first-base runner attempts a stolen base.}
-      \end{split}
-    \end{align}
-    If there is no runner on first base, then $\ell_i$ and $r_i$ are undefined.% We use a superscript to denote the success/failure of pickoff and stolen base attempts: 
-%    \begin{align*}
-%      p_i^+ & \in \{0, 1\} \mbox{ denotes whether the pickoff attempt is successful (i.e. the runner is tagged out); and}\\
-%      s_i^+ & \in \{0, 1\} \mbox{ denotes whether the stolen base attempt is successful.}
-%    \end{align*}
-
-    \subsection{Probability model for runner outcomes}
-    \label{sec:prob-runner-outcome}
-
-      As itemized in (\ref{eqn:runner-outcome}), there are five possible outcomes for the runner with respect to the run game.
-To estimate each of these probabilities conditional on player identities, count and lead distance, we model four probabilities.
-Using $R_i$ to denote the random variable governing the probability distribution over $r_i$,
-      \begin{align*}
-        \pi_i    &= \mathbb{P}(R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff attempt probability;}\\
-        \pi_i^+  &= \mathbb{P}(R_i \in \{\mbox{P}^+\} \mid R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff success probability;}\\
-        \psi_i   &= \mathbb{P}(R_i \in \{\mbox{S}^+,\, \mbox{S}^-\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-, N\}) \mbox{ is the stolen base attempt probability; and}\\
-        \psi_i^+ &= \mathbb{P}(R_i \in \{\mbox{S}^+\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-\}) \mbox{ is the stolen base success probability.}
-      \end{align*}
-      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
-
-      \subsubsection{Pickoff attempt probability}
-      \label{sec:prob-po-attempt}
-
-        We model $\pi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), and lead distance; and with a random effect for the pitcher.
-        \begin{align}
-          \label{eqn:prob-po-attempt}
-          \begin{split}
-            \log\left(\frac{\pi_i}{1 - \pi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + \beta^L \ell_i + \gamma^P_{h_i^P}\\
-            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
-          \end{split}
-        \end{align}
-        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$ and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$.
-The number of random effects is $|\mathcal{H}^R| + |\mathcal{H}^P|$.
-
-      \subsubsection{Pickoff success probability}
-      \label{sec:prob-po-success}
-
-        We model $\pi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher.
-These were the only variables for which we found substantive effects.
-        \begin{align}
-          \label{eqn:prob-po-success}
-          \begin{split}
-            \log\left(\frac{\pi_i^+}{1 - \pi_i^+}\right) &= \alpha + \beta^L \ell_i + \gamma^P_{h_i^P},\\
-            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
-          \end{split}
-        \end{align}
-        This model has three fixed, unknown parameters: the intercept $\alpha$, the slope $\beta^L$, and the variance parameters $\sigma^2_P$. The number of random effects is $|\mathcal{H}^P|$.
-
-      \subsubsection{Stolen base attempt probability}
-      \label{sec:prob-sb-attempt}
-
-        Modeling the runner's decision to attempt a stolen base is complicated.
-One may think of this as a deterministic decision made by the runner based on the expected run value with and without attempting a steal.
-Such a model aligns very poorly with real decisions made by runners.
-We posit that the runner's ability to attempt a stolen base is influenced by external factors such as the runner's perception of the pitcher's body language timing between pitches.
-For this reason, we model stolen base attempts as a stochastic process rather than a deterministic decision.
-
-        Deliberately, we exclude lead distance from the stolen base attempt model.
-While lead distance does help explain the probability of a stolen base attempt, this can cause problems with the framing of the runner's decision.
-When including lead distance, we find situations in which the expected value of a stolen base attempt is negative, and the optimal lead distance is zero feet to minimize the probability of a stolen base attempt.
-This is a poor reflection of reality because the runner always has the power to choose not to attempt a stolen base.
-We obtain a more fit-for-purpose model by excluding lead distance.
-
-        We model $\psi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
-        \begin{align}
-          \label{eqn:prob-sb-attempt}
-          \begin{split}
-            \log\left(\frac{\psi_i}{1 - \psi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
-            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
-            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
-            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
-          \end{split}
-        \end{align}
-
-        We interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the effect of runner $h_i^R$ on the attempt probability, combining the fixed effect of the runner's sprint speed with the random effect of their identity.
-By including a fixed effect for sprint speed, we are effectively regularizing the runner effects toward priors based on their sprint speeds.
-The interpretation of $(\beta^C z_i^C + \gamma^C_{h_i^C})$ for catchers is similar: Each catcher has an estimated effect regularized toward a prior based on their arm strength.
-
-      \subsubsection{Stolen base success probability}
-      \label{sec:prob-sb-success}
-
-        We model $\psi_i^+$ using a mixed-effects logistic regression with fixed effects for lead distance, runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
-        \begin{align}
-          \label{eqn:prob-sb-success}
-          \begin{split}
-            \log\left(\frac{\psi_i^+}{1 - \psi_i^+}\right) &= \alpha + \beta^L \ell_i + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
-            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
-            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
-            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
-          \end{split}
-        \end{align}
-        As in (\ref{eqn:prob-sb-attempt}), we interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the runner effect regularized toward a prior based on sprint speed and $(\beta^C z_i^C + \gamma^C_{h_i^C})$ as the catcher effect regularized toward a prior based on arm strength.
-
-    \subsection{The baseball inning as a Markov decision process}
-
-      \subsubsection{Markov chain model}
-
-        Modeling the progression of a baseball inning is a well-trodden application of Markov chains in sports \parencite{bukiet_markov_1997}.
-Traditionally, the state space is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states).
-To understand the game theory under the new pickoff limit rule, we take this a step further and include the number of disengagements in the state space.
-
-        {\it State space.} The game state is given by $s = (b, c, d, o)$, where:
-        \begin{align*}
-          b &\in \{0, 1\}^3 \mbox{ represents which bases are occupied by runners};\\
-          c &\in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ represents the ball-strike count};\\
-          d &\in \{0, 1, 2\} \mbox{ represents the number of disengagements already used}; and\\
-          o &\in \{0, 1, 2\} \mbox{ represents the number of outs}.
-        \end{align*}
-        The total number of non-terminal states is $2^3 \times (4 \times 3) \times 3 \times 3 = 864$.
-There is one terminal state (end of inning), bringing the total number of states to 865.
-We use $\mathcal{S}$ to denote the set of all states, including the terminal state; we use $s_i = (b_i, c_i, d_i, o_i)$ to denote the starting state for play $i$; and we use $S_i$ to denote the random variable governing the probability distribution over $s_i$.
-
-        {\it Transition probabilities.} The probability of transition to state $s'$ from starting state $s$ is given by the transition probability function $\mathbb{P}(S_{i+1} = s' \mid S_i = s)$.
-Because the present work focuses on the consequences of decisions surrounding the runner on first base, it is helpful to break down these transition probabilities by conditioning on the runner outcome:
-        \begin{align}
-          \label{eqn:transition-prob}
-          \mathbb{P}(S_{i+1} = s' \mid S_i = s) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s, R_i = r).
-        \end{align}
-
-        To estimate $\mathbb{P}(R_i = r \mid S_i = s)$, we may use the empirical frequency of each runner outcome $r$ in each state $s$.
-In Section \ref{sec:single-agent-mdp} below, we will replace these empirical frequencies with the probabilities estimated in Section \ref{sec:prob-runner-outcome}.
-
-        To estimate $\mathbb{P}(S_{i+1} = s' \mid S_i = s, R_i = r)$, we use conditional empirical frequencies of the ending state $s'$ given the starting state $s$ and the runner outcome $r$.
-However, we pool across the starting disengagements $d$ when calculating these empirical frequencies.
-Our assumption is that, given the runner event, the starting disengagements $d$ does not impact the transition probabilities between the reduced states $\tilde s \equiv (b, c, \cdot, o)$ and $\tilde s' \equiv (b', c', \cdot, o')$.
-By pooling data across different values of $d$, we mitigate small sample size issues when estimating transition probabilities from rare states, such as having a runner on third base with an 0-0 count, two disengagements and zero outs.
-
-        The pooled empirical transition probability conditions on $e = d' - d$ instead of conditioning on $d$ and $d'$.
-Using $E_{i}$ to denote the random variable governing the probability distribution over $d_{i + 1} - d_i$,
-        \begin{align*}
-          Q(\tilde s, r, \tilde s', e)
-            &= \hat{\mathbb{P}}\left(\tilde S_{i+1} = \tilde s',\, E_i = e \mid \tilde S_i = \tilde s,\, R_i = r\right)\\
-            &= \frac
-              {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\} \cdot \mathbb{I}\left\{\tilde s_{i+1} = \tilde s',\, d_{i+1} - d_i = e\right\}}
-              {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\}}.
-        \end{align*}
-        Finally, we estimate the full-state transition probabilities as
-        \begin{align*}
-          \hat{\mathbb{P}}(S_{i+1} = (b', c', d', o') \mid S_i = (b, c, d, o), R_i = r) = Q\big((b, c, \cdot, o),\, r,\, (b', c', \cdot, o'),\, d' - d\big).
-        \end{align*}
-
-% NOTE (SP): I don't think we need the paragraph below, but I'm holding onto it for now in case we decide we do need it.
-%        To convert the reduced-state estimated transition probabilities to full-state transition probabilities, we leverage the fact that the number of disengagements $d'$ in the ending state is a deterministic function of the starting disengagements $d$, the runner outcome $r$ and  the ending count $c'$:
-%        \begin{align}
-%          \label{eqn:disengagement-transition}
-%          d' = D(d, c', r)
-%            \equiv \begin{cases}
-%            \hfil d & \mbox{if } r \in \{\mbox{N}\} \mbox{ and } c' \ne (0, 0)\\
-%            \hfil 0 & \mbox{if } r \in \{\mbox{N}\} \mbox{ and } c' = (0, 0)\\
-%            d + 1   & \mbox{if } r \in \{\mbox{P}^-\} \mbox{ and } d < 2\\
-%            \hfil 0 & \mbox{if } r \in \{\mbox{P}^-\} \mbox{ and } d = 2\\
-%            \hfil 0 & \mbox{if } r \in \{\mbox{P}^+,\, \mbox{S}^+,\, \mbox{S}^-\}\\
-%          \end{cases}.
-%        \end{align}
-
-      \subsubsection{Single-agent Markov decision process}
-      \label{sec:single-agent-mdp}
-
-        The MDP is an extension of the Markov chain model that allows for deterministic actions which change transition probabilities between states.
-In this section we allow for the runner to choose a lead distance $\ell$, and we model the pitcher's decision whether to attempt a pickoff as a stochastic process (see Section \ref{sec:prob-po-attempt}).
-We call this the {\it single-agent} model because only the runner has agency in this model of the game.
-We limit this action to states in which first base is occupied and the other bases are unoccupied, i.e.
-$b = (1, 0, 0)$.
-This limits our scope to situations in which the runner might steal second base (encompassing 71\% of stolen base attempts in 2023) and excludes the first-and-third scenario $b = (1, 0, 1)$, in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
-
-        {\it Action space.} In states for which $b = (1, 0, 0)$, the first-base runner must choose a lead distance $\ell \in [0, 20] \equiv \mathcal{L}$, representing distance in feet from first base.
-For computational purposes, we discretize this action space by rounding $\ell$ to the nearest tenth of a foot, meaning that there are 201 possible actions.
-
-        {\it Transition probabilities.} The probability of transition to ending state $s'$ from starting state $s$ given lead distance $\ell$ is $P(s' \mid s,\, \ell) = \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = \ell)$.
-To estimate state transition probabilities that depend on lead distance, we modify (\ref{eqn:transition-prob}) by conditioning on the lead distance:
-        \begin{align}
-          \label{eqn:transition-prob-lead}
-          P(s' \mid s,\, \ell) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r).
-        \end{align}
-        In this equation, the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance.
-The procedure for estimating $\mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r)$ is described in Section \ref{sec:single-agent-mdp}.
-To estimate $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell)$, we use the models from Section \ref{sec:prob-runner-outcome}.
-Specifically, $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$ is given by
-        \begin{align*}
-          \begin{cases}
-              \hfil \pi_i(\ell) \cdot \pi_i^+(\ell)                             & \mbox{if } r = \mbox{P}^+\\
-              \hfil \pi_i(\ell) \cdot (1 - \pi_i^+(\ell))                       & \mbox{if } r = \mbox{P}^-\\
-              \hfil (1 - \pi_i(\ell)) \cdot \psi_i \cdot \psi_i^+(\ell)         & \mbox{if } r = \mbox{S}^+\\
-              \hfil (1 - \pi_i(\ell)) \cdot \psi_i \cdot (1 - \psi_i^+(\ell))   & \mbox{if } r = \mbox{S}^-\\
-              \hfil (1 - \pi_i(\ell)) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
-          \end{cases}.
-        \end{align*}
-        Recall that the stolen base attempt probability $\psi_i$ does not depend on $\ell$, as described in Section \ref{eqn:prob-sb-attempt}.
-
-        {\it Reward function.} The reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition.
-Note that this reward function does not depend on the action taken by the runner, so we can write $R(s, \ell, s') = R(s, s')$.
-If $s = (b, c, d, o)$ and $s' = (b', c', d', o')$, then:
-        \begin{align}
-          \label{eqn:reward}
-          R(s, s') = (g(b') + o') - (g(b) + o) - \mathbb{I}\{c' = (0, 0),\, d' = 0\},
-        \end{align}
-        where $g: \{0, 1\}^3 \rightarrow \{0, 1, 2, 3\}$ counts the number of runners on base.
-This expression for the reward function counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance.
-For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state.
-To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state.
-These intermediary states transition to the terminal end-of-inning state with probability 1.
-
-        {\it Policy iteration.} A policy function $\pi : \mathcal{S} \rightarrow \mathcal{L}$ maps the state space to the action space, defining the lead distance to be taken by the runner in each state.
-The policy function $\pi$ specifies transition probabilities $P(s' \mid s,\, \ell = \pi(s))$ and induces a probability distribution over the sequence of states $\{S_t\}_{t=1}^\infty$.
-The value function $V_\pi : \mathcal{S} \rightarrow \mathbb{R}^+$ is the expected sum of rewards with respect to this probability distribution, which we interpret as the expected number of runs scored in the inning, starting from state $s$:
-        \begin{align}
-          \label{eqn:run-expectancy}
-          V_\pi(s) = \mathbb{E}_\pi\left[\left.\sum_{t = 0}^\infty R(S_{t}, S_{t+1})~\right|~S_0 = s\right].
-        \end{align}
-
-        The relevant question for the runner is, ``What lead distance should I choose in each situation to maximize run scoring overall?'' We find the policy function $\pi^*$ which maximizes (\ref{eqn:run-expectancy}) via policy iteration \parencite{bellman_dynamic_1957} by initializing $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, and then iteratively applying the updates
-        \begin{align}
-          \label{eqn:update-value-single-agent}
-          \hat V_{i+1}(s) &= \sum_{s' \in \mathcal{S}} P(s' | s,\, \hat\pi_i(s)) [R(s, s') + \hat V_i(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
-          \label{eqn:update-policy-single-agent}
-          \hat\pi_{i+1}(s) &= \arg\max_{\ell \in \mathcal{L}} \sum_{s' \in \mathcal{S}} P(s' | s,\, \ell) [R(s, s') + \hat V_{i+1}(s')] && \forall s \in \mathcal{S},
-        \end{align}
-        until $\max_{s\in\mathcal{S}}|\pi_{i+1}(s) - \pi_i(s)|$ converges to zero.\footnote{Do we need to prove that this will converge to the globally optimal policy?}
-
-      \subsubsection{Two-agent Markov decision process}
-      \label{sec:two-agent-mdp}
-
-        In this section we extend the single-agent model by allowing the pitcher to choose whether to attempt a pickoff.
-We call this the {\it two-agent} model because both the runner and the pitcher have agency in this model of the game.
-Specifically, we model this as a two-stage game, where the runner first chooses a lead distance, and then the pitcher decides whether to attempt a pickoff after observing the runner's lead distance.
-
-        {\it Action space.} As in the single-agent model, we limit actions to states in which only first base is occupied.
-The runner chooses a lead distance $\ell \in \mathcal{L}$, and then the pitcher chooses a pickoff action $p \in \{0, 1\} \equiv \mathcal{P}$, where $p = 1$ corresponds to a pickoff attempt.
-
-        {\it Transition probabilities.} The probability of transition to ending state $s'$ from starting state $s$ given lead distance $\ell$ and pickoff decision $p$ is $P(s' \mid s,\, \ell,\, p) = \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$.
-We modify (\ref{eqn:transition-prob-lead}) by conditioning additionally on the pickoff decision:
-        \begin{align}
-          \label{eqn:transition-prob-lead-pickoff}
-          P(s' \mid s, \ell, p) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r),
-        \end{align}
-        where $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = \ell,\, p_i = p)$ is given by
-        \begin{align*}
-            \begin{cases}
-                \hfil p \cdot \pi_i^+(\ell)                             & \mbox{if } r = \mbox{P}^+\\
-                \hfil p \cdot (1 - \pi_i^+(\ell))                       & \mbox{if } r = \mbox{P}^-\\
-                \hfil (1 - p) \cdot \psi_i \cdot \psi_i^+(\ell)         & \mbox{if } r = \mbox{S}^+\\
-                \hfil (1 - p) \cdot \psi_i \cdot (1 - \psi_i^+(\ell))   & \mbox{if } r = \mbox{S}^-\\
-                \hfil (1 - p) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
-            \end{cases}.
-        \end{align*}
-
-        {\it Reward function.} The reward function (\ref{eqn:reward}) is the same for the two-agent model as for the single-agent model.
-In this game, however, the runner wants to maximize the reward, and the pitcher wants to minimize the reward.
-
-        {\it Policy iteration.} In the policy update step (\ref{eqn:update-policy-single-agent}) of policy iteration for the single-agent model, we seek the policy which maximizes the expected one-step sum of reward and resulting value, given an estimated value function.
-In the two-agent model, there are game theoretic considerations.
-Given the estimated value function $\hat V$ and the runner's chosen lead distance $\ell$, the pitcher chooses $p$ to minimize the expected one-step sum of reward and resulting value:
-        \begin{align*}
-          \arg\min_{p\in\mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' \mid s,\, \ell,\, p) [R(s, s') + \hat V(s')].
-        \end{align*}
-        Knowing that this is the optimal behavior for the pitcher, the runner chooses their lead distance to maximize the resulting run value:
-        \begin{align*}
-          \arg\max_{\ell \in \mathcal{L}}\left\{\min_{p\in\mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' \mid s,\, \ell,\, p) [R(s, s') + \hat V(s')]\right\}.
-        \end{align*}
-
-        The optimal policy for the pitcher is clear: between the choices of attempting a pickoff or not, make the choice that results in the lower run expectation.
-To find the optimal policy $\pi^* : \mathcal{S} \rightarrow \mathcal{L}$ for the runner, we again apply policy iteration.
-Starting with $\hat V_0(s) = 0$ and $\hat \pi_0(s) = 0$ for all $s \in \mathcal{S}$, we iteratively apply the updates below until the policy converges.\footnote{TODO: Are we guaranteed convergence to the optimal policy in the two-agent case?}
-
-        \begin{align}
-          \label{eqn:update-value-two-agent}
-          \hat V_{i+1}(s) &= \min_{p \in \mathcal{P}} \sum_{s' \in \mathcal{S}} P(s' | s,\, \hat\pi_i(s)) [R(s, s') + \hat V_i(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
-          \label{eqn:update-policy-two-agent}
-          \hat\pi_{i+1}(s) &= \arg\max_{\ell \in \mathcal{L}} \left\{\min_{p \in \mathcal{P}}\sum_{s' \in \mathcal{S}} P(s' | s,\, \ell,\, p) [R(s, s') + \hat V_{i+1}(s')]\right\} && \forall s \in \mathcal{S}.
-        \end{align}
-
-        Given this runner policy, the pitcher is indifferent between throwing a pitch and attempting a pickoff.
-
     \section{The baseball inning as a two-player stochastic game}
-    \label{sec:sequential_model}
+    \label{sec:sequential-model}
     We first describe a single-player setup where only the first-base runner has agency.
@@
 The two-player case is modeled as an infinite-horizon zero sum stochastic game.
-\todo{Explain how the single-player case, despite being simple and restrictive than the two-player case, still provides valuable and actionable insights.} Since the state space for the single- and two-player models are the same, we first describe the (common) state space before explaining the specifics of each model.
+While the single-player case is more simple and restrictive than the two-player case, it still provides valuable and actionable insights for runners because pitchers are unlikely to adopt an optimal policy in the short term.
+Since the state space for the single- and two-player models are the same, we first describe the (common) state space before explaining the specifics of each model.
 
@@
     \subsection{Single-player Markov decision process}
-    \label{sec:mdp_model}
+    \label{sec:mdp-model}
     This section considers the setup where agency is allowed only for the first-base runner. We model this scenario as a MDP. The other components of this MDP are as follows.
 
-    {\bf State space.} The state space for the MDP is the one described in the beginning of Section \ref{sec:sequential_model}.
+    {\bf State space.} The state space for the MDP is the one described in the beginning of Section \ref{sec:sequential-model}.
 
@@
 
-    {\bf Transition probabilities.} \todo[inline]{Since the derivation of transition probabilities is involved, write it in a separate section and point to that section.
-Describe here in at most a couple of sentences how those probabilities are computed.
-Highlight which states transition to the absorbing state with positive probability.}
+    {\bf Transition probabilities.} The transition probability function $p(s'|s,a)$ depends on the action taken by the runner.
+We describe in detail our derivation of transition probabilities for the single-player Markov decision process in Section \ref{sec:transition-mdp-model}.
+In summary form, we use generalized linear mixed-effects models to estimate runner outcome (e.g., unsuccessful pickoff attempt or successful stolen base attempt) probabilities which depend on lead distance, and we use empirical frequencies to estimate state transition probabilities conditional on runner outcome probabilities.
+The state transition probabilities depend on lead distance only through the runner outcome probabilities.
+Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
 
@@
 
-    \subsection{Two-player Stochastic Game}
-    \label{sec:stochastic_game}
+    \subsection{Two-player stochastic game}
+    \label{sec:stochastic-game}
     In this section, we extend the single-agent model by allowing the pitcher to choose whether to attempt a pickoff.
@@
 
-    {\bf State space.} The state space  is the same as described in the beginning of Section \ref{sec:sequential_model}.
+    {\bf State space.} The state space  is the same as described in the beginning of Section \ref{sec:sequential-model}.
 
-    {\bf Action space.} As in the single-player model of Section \ref{sec:mdp_model}, we limit actions (of runner as well as pitcher) to states in which only the first base is occupied.
+    {\bf Action space.} As in the single-player model of Section \ref{sec:mdp-model}, we limit actions (of runner as well as pitcher) to states in which only the first base is occupied.
 The (first-base) runner chooses a lead distance, and then the pitcher chooses a pickoff action $p \in \mathcal{P} \defeq \{0,1\}$, where $p = 1$ corresponds to a pickoff attempt and $0$ otherwise.
@@
 
-     {\bf Transition probabilities.} \todo[inline]{Since the derivation of transition probabilities is involved, write it in a separate section and point to that section.
-Describe here in at most a couple of sentences how those probabilities are computed.
-Highlight which states transition to the absorbing state with positive probability.}
+    {\bf Transition probabilities.} The transition probability function $p(s'|s,a_\text{R},a_\text{P})$ depends on the actions taken by the pitcher and runner.
+We describe in detail our derivation of transition probabilities for the two-player stochastic game in Section \ref{sec:transition-stochastic-game}.
+These transition probabilities are the same as for the single-player Markov decision process, except that the runner outcome probabilities are conditioned on the pitcher's pickoff attempt decision, as well as the lead distance.
+For example, if the pitcher does not attempt a pickoff, then the successful pickoff outcome and the unsuccessful pickoff outcome have probability zero.
+Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
 
@@
 
-  \section{Results}
+  \section{Transition probability model}
+  \label{sec:transition-probability-model}
 
-    \subsection{Observed runner behavior}
+    In this section, we describe how we use empirical data to estimate state transition probabilities which we use for the single-player Markov decision process and the two-player stochastic game described in Section \ref{sec:sequential-model}.
+Throughou, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
+    \begin{align*}
+      b_i & \in \{0, 1\}^3 \mbox{ denotes which bases (1B, 2B, 3B, respectively) are occupied;}\\
+      (c_i^B, c_i^S) \equiv c_i & \in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ denotes the ball-strike count;}\\
+      d_i & \in \{0, 1, 2\} \mbox{ denotes the number of disengagements already used; and}\\
+      o_i & \in \{0, 1, 2\} \mbox{ denotes the number of outs.}
+    \end{align*}
+    We use $\mathcal{H}^R$, $\mathcal{H}^P$, and $\mathcal{H}^C$ to denote the sets of runners, pitchers and catchers, respectively. On play $i$:
+    \begin{align*}
+      h_i^R & \in \mathcal{H}^R \mbox{ is the runner;}\\
+      h_i^P & \in \mathcal{H}^P \mbox{ is the pitcher;}\\
+      h_i^C & \in \mathcal{H}^C \mbox{ is the catcher;}\\
+      z_i^R & \in \mathbb{R^+} \mbox{ is the sprint speed of runner $h_i^R$; and}\\
+      z_i^C & \in \mathbb{R^+} \mbox{ is the arm strength of catcher $h_i^C$.}
+    \end{align*}
+    When the play begins, we use the following notation to describe the actions taken by the players:
+    \begin{align}
+    \label{eqn:runner-outcome}
+      \begin{split}
+        \ell_i  \in &~ \mathbb{R}^+ \mbox{ denotes the lead distance (in feet) taken by the first-base runner;}\\
+        p_i \in &~ \{0, 1\} \mbox{ denotes whether the pitcher attempts (1) or does not attempt (0) a pickoff; and}\\
+        r_i \in &~ \{\mbox{P}^+,\, \mbox{P}^-,\, \mbox{S}^+,\, \mbox{S}^-,\, \mbox{N}\} \equiv \mathcal{R} \mbox{ denotes the runner outcome, where}\\
+          & \mbox{P}^+  \mbox{ represents a successful pickoff attempt (runner is out);}\\
+          & \mbox{P}^-  \mbox{ represents an unsuccessful pickoff attempt (runner is safe);}\\
+          & \mbox{S}^+  \mbox{ represents a successful stolen base attempt (runner is safe);}\\
+          & \mbox{S}^-  \mbox{ represents an usuccessful stolen base attempt (runner is out); and}\\
+          & \mbox{N}~   \mbox{ represents no runner action (a pitch with no stolen base attempt).}
+      \end{split}
+    \end{align}
+    If there is no runner on first base, then $\ell_i$ and $r_i$ are undefined.
+
+    \subsection{Probability model for runner outcomes}
+    \label{sec:prob-runner-outcome}
+
+      As itemized in (\ref{eqn:runner-outcome}), there are five possible outcomes for the runner with respect to the run game.
+To estimate each of these probabilities conditional on player identities, count and lead distance, we model four probabilities.
+Using $R_i$ to denote the random variable governing the probability distribution over $r_i$,
+      \begin{align*}
+        \pi_i    &= \mathbb{P}(R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff attempt probability;}\\
+        \pi_i^+  &= \mathbb{P}(R_i \in \{\mbox{P}^+\} \mid R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff success probability;}\\
+        \psi_i   &= \mathbb{P}(R_i \in \{\mbox{S}^+,\, \mbox{S}^-\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-, N\}) \mbox{ is the stolen base attempt probability; and}\\
+        \psi_i^+ &= \mathbb{P}(R_i \in \{\mbox{S}^+\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-\}) \mbox{ is the stolen base success probability.}
+      \end{align*}
+      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
+
+      \subsubsection{Pickoff attempt probability}
+      \label{sec:prob-po-attempt}
+
+        We model $\pi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), and lead distance; and with a random effect for the pitcher.
+        \begin{align}
+          \label{eqn:prob-po-attempt}
+          \begin{split}
+            \log\left(\frac{\pi_i}{1 - \pi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + \beta^L \ell_i + \gamma^P_{h_i^P}\\
+            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
+          \end{split}
+        \end{align}
+        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$ and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$.
+The number of random effects is $|\mathcal{H}^R| + |\mathcal{H}^P|$.
+
+      \subsubsection{Pickoff success probability}
+      \label{sec:prob-po-success}
+
+        We model $\pi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher.
+These were the only variables for which we found substantive effects.
+        \begin{align}
+          \label{eqn:prob-po-success}
+          \begin{split}
+            \log\left(\frac{\pi_i^+}{1 - \pi_i^+}\right) &= \alpha + \beta^L \ell_i + \gamma^P_{h_i^P},\\
+            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
+          \end{split}
+        \end{align}
+        This model has three fixed, unknown parameters: the intercept $\alpha$, the slope $\beta^L$, and the variance parameters $\sigma^2_P$. The number of random effects is $|\mathcal{H}^P|$.
+
+      \subsubsection{Stolen base attempt probability}
+      \label{sec:prob-sb-attempt}
+
+        Modeling the runner's decision to attempt a stolen base is complicated.
+One may think of this as a deterministic decision made by the runner based on the expected run value with and without attempting a steal.
+Such a model aligns very poorly with real decisions made by runners.
+We posit that the runner's ability to attempt a stolen base is influenced by external factors such as the runner's perception of the pitcher's body language timing between pitches.
+For this reason, we model stolen base attempts as a stochastic process rather than a deterministic decision.
+
+        Deliberately, we exclude lead distance from the stolen base attempt model.
+While lead distance does help explain the probability of a stolen base attempt, this can cause problems with the framing of the runner's decision.
+When including lead distance, we find situations in which the expected value of a stolen base attempt is negative, and the optimal lead distance is zero feet to minimize the probability of a stolen base attempt.
+This is a poor reflection of reality because the runner always has the power to choose not to attempt a stolen base.
+We obtain a more fit-for-purpose model by excluding lead distance.
+
+        We model $\psi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
+        \begin{align}
+          \label{eqn:prob-sb-attempt}
+          \begin{split}
+            \log\left(\frac{\psi_i}{1 - \psi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
+            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
+            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
+            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
+          \end{split}
+        \end{align}
+
+        We interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the effect of runner $h_i^R$ on the attempt probability, combining the fixed effect of the runner's sprint speed with the random effect of their identity.
+By including a fixed effect for sprint speed, we are effectively regularizing the runner effects toward priors based on their sprint speeds.
+The interpretation of $(\beta^C z_i^C + \gamma^C_{h_i^C})$ for catchers is similar: Each catcher has an estimated effect regularized toward a prior based on their arm strength.
+
+      \subsubsection{Stolen base success probability}
+      \label{sec:prob-sb-success}
+
+        We model $\psi_i^+$ using a mixed-effects logistic regression with fixed effects for lead distance, runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
+        \begin{align}
+          \label{eqn:prob-sb-success}
+          \begin{split}
+            \log\left(\frac{\psi_i^+}{1 - \psi_i^+}\right) &= \alpha + \beta^L \ell_i + (\beta^R z_i^R + \gamma^R_{h_i^R}) + \gamma^P_{h_i^P} + (\beta^C z_i^C + \gamma^C_{h_i^C}),\\
+            \gamma^R_{h} &\sim \mathcal{N}(0, \sigma^2_R) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^R,\\
+            \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P,\\
+            \gamma^C_{h} &\sim \mathcal{N}(0, \sigma^2_C) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^C.
+          \end{split}
+        \end{align}
+        As in (\ref{eqn:prob-sb-attempt}), we interpret $(\beta^R z_i^R + \gamma^R_{h_i^R})$ as the runner effect regularized toward a prior based on sprint speed and $(\beta^C z_i^C + \gamma^C_{h_i^C})$ as the catcher effect regularized toward a prior based on arm strength.
+
+    \subsection{Transition probabilities for the single-player Markov decision process}
+    \label{sec:transition-mdp-model}
+
+      To derive the transition probability function $p(s'|s,a)$, we use the law of total probability to combine the runner outcome probabilities estimated in this section with the conditional transition probability given the runner outcome:
+      \begin{align}
+        \label{eqn:transition-mdp-model}
+        p(s'|s,\, a) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r).
+      \end{align}
+      In this equation, the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance.
+
+      To estimate the first probability inside the summation of equation (\ref{eqn:transition-mdp-model}), we use the models from Section \ref{sec:prob-runner-outcome}.
+Specifically, $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a)$ is given by
+      \begin{align*}
+        \begin{cases}
+            \hfil \pi_i(a) \cdot \pi_i^+(a)                             & \mbox{if } r = \mbox{P}^+\\
+            \hfil \pi_i(a) \cdot (1 - \pi_i^+(a))                       & \mbox{if } r = \mbox{P}^-\\
+            \hfil (1 - \pi_i(a)) \cdot \psi_i \cdot \psi_i^+(a)         & \mbox{if } r = \mbox{S}^+\\
+            \hfil (1 - \pi_i(a)) \cdot \psi_i \cdot (1 - \psi_i^+(a))   & \mbox{if } r = \mbox{S}^-\\
+            \hfil (1 - \pi_i(a)) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
+        \end{cases}.
+      \end{align*}
+      Recall that $\psi_i$ does not depend on $a$, as described in Section \ref{sec:prob-sb-attempt}.
+
+      To estimate $\mathbb{P}(S_{i+1} = s' \mid S_i   = s, R_i = r)$, we use conditional empirical frequencies of the ending state $s'$ given the starting state $s$ and the runner outcome $r$.
+However, we pool across the starting disengagements $d$ when calculating these empirical frequencies.
+Our assumption is that, given the runner event, the starting disengagements $d$ does not impact the transition probabilities between the reduced states $\tilde s \equiv (b, c, \cdot, o)$ and $\tilde s' \equiv (b', c', \cdot, o')$.
+By pooling data across different values of $d$, we mitigate small sample size issues when estimating transition probabilities from rare states, such as having a runner on third base with an 0-0 count, two disengagements and zero outs.
+
+      The pooled empirical transition probability conditions on $e = d' - d$ instead of conditioning on $d$ and $d'$.
+Using $E_{i}$ to denote the random variable governing the probability distribution over $d_{i + 1} - d_i$,
+      \begin{align*}                                  
+        Q(\tilde s, r, \tilde s', e)                  
+          &= \hat{\mathbb{P}}\left(\tilde S_{i+1} =   \tilde s',\, E_i = e \mid \tilde S_i = \tilde s,\, R_i = r\right)\\
+          &= \frac
+            {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\} \cdot \mathbb{I}\left\{\tilde s_{i+1} = \tilde s',\, d_{i+1} - d_i = e\right\}}
+            {\sum_{i=1}^n\mathbb{I}\left\{\tilde s_i = \tilde s,\, r_i = r\right\}}.
+      \end{align*}
+      Finally, we estimate the full-state transition probabilities as
+      \begin{align*}
+        \hat{\mathbb{P}}(S_{i+1} = (b', c', d', o') \mid S_i = (b, c, d, o), R_i = r) = Q\big((b, c, \cdot, o),\, r,\, (b', c', \cdot, o'),\, d' - d\big).
+      \end{align*}
+
+    \subsection{Transition probabilities for the two-player stochastic game}
+    \label{sec:transition-stochastic-game}
+
+      To derive the transition probability function $p(s'|s,a_\text{R},a_\text{P})$, we modify equation (\ref{eqn:transition-mdp-model}) to condition the runner outcome probabilities on the pitcher's pickoff attempt decision as well as the runner's lead distance decision:
+      \begin{align}
+        \label{eqn:transition-stochastic-game}
+        p(s'|s,\, a) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R}, p_i = a_\text{P}) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r),
+      \end{align}
+      where $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P})$ is given by
+      \begin{align*}
+          \begin{cases}
+              \hfil a_\text{P} \cdot \pi_i^+(a_\text{R})                             & \mbox{if } r = \mbox{P}^+\\
+              \hfil a_\text{P} \cdot (1 - \pi_i^+(a_\text{R}))                       & \mbox{if } r = \mbox{P}^-\\
+              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot \psi_i^+(a_\text{R})         & \mbox{if } r = \mbox{S}^+\\
+              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot (1 - \psi_i^+(a_\text{R}))   & \mbox{if } r = \mbox{S}^-\\
+              \hfil (1 - a_\text{P}) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
+          \end{cases}.
+      \end{align*}
+      The procedure for estimating $\mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r)$ is the same as in Section \ref{sec:transition-mdp-model}.
+
+  \section{Numerical Results}
+
+    In this section, we present the results of applying the single-player Markov decision process model and the two-player stochastic game model to real MLB data.
+Most of the data were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
+The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
+For each play, we observe the identities of the pitcher, catcher and runners involved.
+Before and after each play, we observe the state of the game, described by the number of outs in the inning; which bases are occupied by runners; the ball-strike count on the batter; and the number of disengagements already used by the pitcher.
+On each pickoff attempt, we observe whether it is successful (runner is out) or unsuccessful (runner is safe).
+On each pitch, we observe whether the runner attempts a stolen base; and if so, we observe whether they are successful (runner is safe) or unsuccessful (runner is out).
+On each pitch terminating a plate appearance (a batting attempt by a batter), we observe the plate appearance outcome, such as strikeout, groundout, single or home run.
+
+    In addition to the play-by-play data, we manually downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
+This is defined as the average of the top 10 percent of throw speeds when the catcher is attempting to catch a base stealer.
+For each runner-season, we observe {\it sprint speed} (feet per second), which measures the runner's top running speed in their fastest one-second window.
+This is defined as the average of the top 67 percent of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
+
+    In addition to the publicly available data described above, we rely on one key piece of proprietary information provided by MLB Advanced Media, estimated using computer vision on video data from several camera angles.
+For each play, observe each runner's {\it lead distance} (feet), which is how far they are ranging from their base at the time of the pitcher's first movement.
+This is defined as the distance between the runner's center of mass (projected onto the line between the runner's current base and their next base) and the nearest edge of the runner's current base.
+When the pitcher throws a pitch, runners will typically extend their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a secondary lead), which is relevant for runner advancement on batted balls but not on stolen base attempts.
+Our data include only the lead distance at the time of the pitcher's first movement, not the secondary lead.
 
@@
 
-    \subsection{Probability model for runner outcomes}
+    \subsection{Runner outcome probability model}
 
@@
 
-    \subsection{Single-agent Markov decision process}
+    \subsection{Single-player Markov decision process}
 
@@
 
-    \subsection{Two-agent Markov decision process}
+    \subsection{Two-player stochastic game}
 
@@
 While the single-agent model is the most immediately useful framework for runner decision-making in the current environment, we also consider how pitchers might respond using the same model to optimize their decision-making.
-In the two-agent model, the runner first chooses a lead distance, and the pitcher then decides whether to attempt a pickoff, as detailed in Section \ref{sec:two-agent-mdp}.
+In the two-agent model, the runner first chooses a lead distance, and the pitcher then decides whether to attempt a pickoff, as detailed in Section \ref{sec:sequential-model}.
 The optimal strategy for the runner is maximin because they choose the lead distance which maximizes the minimum run expectancy across the pitcher's two options (pickoff or pitch).
@@
 For example, we assume that a greater lead distance does not improve the runner's probability of advancing from first base to third base on a single.
-Baseball domain knowledge suggests this is a reasonable assumption because while a pitch is being thrown, the runner moves to a ``secondary'' lead distance which is more important for determining advancement outcomes than primary lead distance (at first pitcher movement).
+Baseball domain knowledge suggests this is a reasonable assumption because while a pitch is being thrown, the runner moves to a secondary lead distance which is more important for determining advancement outcomes than primary lead distance (at first pitcher movement).
 Finally, we assume that, other than pickoff attempts, no other disengagements occur.

### 2025-10-13 articles/arxiv/paper.tex
overleaf edits

@@
 The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
+    \todo[inline]{Justify expected runs as the reward rather than win probability.}
 
@@
   \noindent The above assumption states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
+  \todo[inline]{Justify in simple terms that inning in a baseball eventually halts.}
 
@@
 These intermediary states transition to the terminal end-of-inning state with probability 1.
+     \todo[inline]{Justify the paths/states through which the terminal state can be reached in an inning.}
 
-    {\bf Transition probabilities.} The transition probability function $p(s'|s,a_\text{R},a_\text{P})$ depends on the actions taken by the pitcher and runner.
+    {\bf Transition probabilities.} The transition probability function $p(s'|s,a_\text{R},a_\text{P})$ depends on the actions taken by the runner and pitcher.
 We describe in detail our derivation of transition probabilities for the two-player stochastic game in Section \ref{sec:transition-stochastic-game}.
@@
 
-   A (deterministic stationary Markovian) policy $\pi_\text{R}$ for the runner is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi_\text{R}(s) \in A_\text{R}(s)$.
+   A (deterministic stationary Markovian) runner policy $\pi_\text{R}$ is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi_\text{R}(s) \in A_\text{R}(s)$.
 That is, $\pi_\text{R}(s)$ is the action chosen in state $s$.
-Similarly, a (deterministic stationary Markovian) policy $\pi_\text{P}$ for the pitcher is a mapping of the form $\{(s,a) \in {\cal S} \times \cup_{s \in {\cal S}}A_\text{R}(s) \vert a \in A_\text{R}(s)\} \ni (x,a) \mapsto A_\text{P}(x,a)$ with the interpretation that $\pi_\text{P}(x,a_\text{R})$ is the action chosen by the pitcher when the state is $s$ and the runner picks the action $a_\text{R}$.
+Similarly, a (deterministic stationary Markovian) pitcher policy $\pi_\text{P}$ is a mapping of the form $\{(s,a) \in {\cal S} \times \cup_{s \in {\cal S}}A_\text{R}(s) \vert a \in A_\text{R}(s)\} \ni (x,a) \mapsto A_\text{P}(x,a)$ with the interpretation that $\pi_\text{P}(x,a_\text{R})$ is the action chosen by the pitcher when the state is $s$ and the runner picks the action $a_\text{R}$.
 Denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
@@
     In this section, we describe how we use empirical data to estimate state transition probabilities which we use for the single-player Markov decision process and the two-player stochastic game described in Section \ref{sec:sequential-model}.
-Throughou, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
+Throughout, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
     \begin{align*}
@@
 
-  \section{Numerical Results}
+  \section{Numerical results}
 

### 2025-10-14 scripts/estimate_models.R
refine encoding of covariates in runner outcome models

@@ -1,3 +1,6 @@
+
+fit_glmer_models <- FALSE
+
 # DATA SET UP ----
 
 library(tidyverse)
@@ -79,7 +82,13 @@ sb_att_1b <- sb_attempts %>% filter(!is.na(run1b) & is.na(run2b) & is.na(run3b))
 sb_threats <- sb_att_1b %>% count(pre_runner_1b_id) %>% filter(n >= 3) %>% pull(pre_runner_1b_id)
 pickoff_var_1b_threats <- pickoff_var_1b %>% filter(pre_runner_1b_id %in% sb_threats) %>% filter(pre_balls < 3 | pre_strikes < 2 | pre_outs < 2)
 pickoff_att_1b_threats <- pickoff_att_1b %>% filter(pre_runner_1b_id %in% sb_threats) %>% filter(pre_balls < 3 | pre_strikes < 2 | pre_outs < 2)
-sb_var_1b_threats <- sb_var_1b %>% filter(pre_runner_1b_id %in% sb_threats) %>% filter(pre_balls < 3 | pre_strikes < 2 | pre_outs < 2)
+sb_var_1b_threats <- sb_var_1b |>
+  dplyr::filter(
+    pre_runner_1b_id %in% sb_threats,
+    pre_balls < 3 | pre_strikes < 2 | pre_outs < 2,
+    pre_disengagements < 3
+  ) |>
+  dplyr::mutate(pre_disengagements = as.factor(pre_disengagements))
 sb_att_1b_threats <- sb_att_1b %>% filter(pre_runner_1b_id %in% sb_threats) %>% filter(pre_balls < 3 | pre_strikes < 2 | pre_outs < 2)
 
 
@@ -298,49 +307,71 @@ sb_att_1b_threats22 <- sb_att_1b22 %>% filter(pre_runner_1b_id %in% sb_threats22
 
 
 
-all_sb_att1b <- rbind(sb_att_1b_threats, sb_att_1b_threats22)
+all_sb_att1b <- rbind(sb_att_1b_threats, sb_att_1b_threats22) |>
+  dplyr::mutate(
+    pitcher_id = ifelse(pitcher_id %in% pitchers_in_sample, pitcher_id, "p1"),
+    run1b = ifelse(run1b %in% runners_in_sample & run1b %in% sb_threats, run1b, "r1"),
+    is_stolen_base = ifelse(runner_going == TRUE & post_balls == 4, TRUE, is_stolen_base),
[…]

### 2025-10-14 scripts/generate_results.R
export model summaries to latex table

@@ -76,9 +76,74 @@ if (fig_make) {
 
 # Plot probability model summaries ----
 
-fit_po_attempt <- readRDS("models/fit_po_attempt.rds")
-fit_po_success <- readRDS("models/fit_po_success.rds")
-fit_sb_success <- readRDS("models/fit_sb_success.rds")
+fit_po_attempt <- readRDS("output/models/fit_po_attempt.rds")
+fit_po_success <- readRDS("output/models/fit_po_success.rds")
+fit_sb_attempt <- readRDS("output/models/fit_sb_attempt.rds")
+fit_sb_success <- readRDS("output/models/fit_sb_success.rds")
+
+# Table
+
+extract_model_fit <- function(model) {
+  ranef <- lme4::VarCorr(model) |>
+    tibble::as_tibble() |>
+    dplyr::mutate(type = "random", variable = grp, `Std. Error` = sdcor) |>
+    dplyr::select(type, variable, `Std. Error`)
+  summary(model)$coefficients |>
+    tibble::as_tibble(rownames = "variable") |>
+    dplyr::mutate(type = "fixed") |>
+    dplyr::bind_rows(ranef) |>
+    dplyr::mutate(
+      estimate = paste(
+        ifelse(type == "random", "", glue::glue("${sprintf('%.3f', Estimate)}$")),
+        glue::glue("$\\pm{sprintf('%.3f', `Std. Error`)}$")
+      )
+    ) |>
+    dplyr::select(type, variable, estimate)
+}
+
+table_po_attempt <- extract_model_fit(fit_po_attempt) |>
+  dplyr::rename(po_attempt = estimate)
+table_po_success <- extract_model_fit(fit_po_success) |>
+  dplyr::rename(po_success = estimate)
+table_sb_attempt <- extract_model_fit(fit_sb_attempt) |>
+  dplyr::rename(sb_attempt = estimate)
+table_sb_success <- extract_model_fit(fit_sb_success) |>
+  dplyr::rename(sb_success = estimate)
+
+table_po_attempt |>
+  dplyr::full_join(table_po_success, by = c("type", "variable")) |>
+  dplyr::full_join(table_sb_attempt, by = c("type", "variable")) |>
+  dplyr::full_join(table_sb_success, by = c("type", "variable")) |>
+  dplyr::arrange(type) |>
+  dplyr::mutate(
+    variable = dplyr::case_when(
+      variable == "(Intercept)" ~ "{\\it Intercept}",
[…]

### 2025-10-14 articles/arxiv/paper.tex
Report summaries for runner outcome models (#5)
* refine encoding of covariates in runner outcome models

* export model summaries to latex table

* add model summary table to paper

@@
 
-\usepackage{amsmath, amssymb, amsthm, amsfonts, color, enumerate, framed, graphicx, mathrsfs, mathtools, mdframed, natbib, subcaption, tabularx, theoremref, titlesec, verbatim, xfrac, soul, mathtools, wrapfig, tikz, float}
+\usepackage{amsmath, amssymb, amsthm, amsfonts, color, enumerate, framed, graphicx, mathrsfs, mathtools, mdframed, subcaption, tabularx, theoremref, titlesec, verbatim, xfrac, soul, mathtools, wrapfig, tikz, float}
 
@@
 
-      Figures \ref{fig:prob-pickoff} and \ref{fig:prob-sb-success} illustrate the relationship between lead distance and runner outcome probabilities, as estimated using the mixed-effects logistic regression models detailed in Section \ref{sec:prob-runner-outcome}.
-For all three models, the direction of the relationship between lead distance and outcome probability is as expected.
+      Table \ref{tab:model-summary} summarizes the results of estimating the runner outcome probability models detailed in Section \ref{sec:prob-runner-outcome}.
+Figures \ref{fig:prob-pickoff} and \ref{fig:prob-sb-success} illustrate the relationship between lead distance and runner outcome probabilities for the models which include lead distance.
+For each of these three models, the direction of the relationship between lead distance and outcome probability is as expected.
+
+      \begin{table}[H]
+        \hspace{-6mm}
+        \input{../../output/tables/model_summary.tex}
+        \caption{\it Summary of estimated runner outcome probability models.
+We estimated four different generalized (binomial) linear mixed-effects models for pickoff (PO) attempt and success and for stolen base (SB) attempt and success.
+For fixed effects we report the estimated effect with one standard error, and for random effects we report the estimated standard deviation term.
+For 2022, we encoded disengagements as always zero to reflect no pickoff limit.}
+        \label{tab:model-summary}
+      \end{table}
 

### 2025-10-14 articles/arxiv/paper.tex
SP revised based on AS feedback

@@
     In 2023, Major League Baseball (MLB) introduced new several rule changes to incentivize base stealing and more exciting plays.
-Among these changes, bases became wider (slighly reducing running distance for a stolen base), and the pitcher was limited to three pickoff attempts.
-The pickoff limit creates an interesting cat-and-mouse game between pitcher and runner because each failed pickoff attempt adds pressure on the pitcher to avoid using the next pickoff attempt.
+Among these changes, and the pitcher was limited to three pickoff attempts.
+This pickoff limit creates an interesting cat-and-mouse game between pitcher and runner because each failed pickoff attempt adds pressure on the pitcher to avoid using the next pickoff attempt.
 The runner can increase this pressure by extending the distance of their leadoff toward the next base, increasing their chances of successfully stealing the base.
+We model this dynamic as a sequential two-player stochastic game in which the runner first chooses a lead distance and then the pitcher chooses whether to attempt a pickoff.
+We prove that this game has a unique equilibrium strategy which we attain via policy iteration.
 Using pitch-by-pitch lead distance data obtained from MLB, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
-Using these conditional transition probabilities, we estimate optimal lead distance via two separate Markov decision process models: single-agent (runner only) and two-agent (game theoretic).
-From these results, we present an actionable rule of thumb: the Two-Foot Rule.
-We recommend that a runner on first base increase their lead distance by two feet after each successive pickoff attempt.
+With these conditional transition probabilities, we estimate an optimal lead distance policy for the runner.
+The results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner on first base increase their lead distance by two feet after each successive pickoff attempt.
   \end{abstract}
@@
 
-    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out.
-Major League Baseball (MLB) stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
-On the micro level, it improves a team's win probability to reduce stolen base attempts that have negative expected outcomes, but on the macro level, the reduction in stolen base attempts affects the entertainment value of the game.
+    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out \parencite{major_league_baseball_baseball_2025}.
+Whether attempting a stolen base or not, runners typically take a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, in the direction of the next base.
+The longer the lead, the shorter the distance needed to run to reach the next base.
+The pitcher's primary deterrent against long leads is a {\it pickoff attempt}, in which they throw the ball to a fielder at the runner's base, attempting a tag before the runner returns safely to the base.
 
-    Acknowledging this and other trends, MLB introduced several rule changes taking effect in the 2023 regular season, with the goal of making the game more exciting \parencite{castrovince_pitch_2023}.
+    MLB stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
+Before 2023, pickoff attempts could cause indefinite delays in the progress of a game because unsuccessful pickoff attempts do not advance the state of the game.
+Acknowledging this and other trends, MLB introduced several rule changes starting in 2023, with the goal of making the game more exciting \parencite{castrovince_pitch_2023}.
 With these changes, a pitch clock limits the time between pitches, and there are restrictions on infielder positioning, curtailing defensive shifts.
-Two changes are particularly relevant to stolen base attempts: Bases are now wider (reducing the distance between first and second base by 4.5 inches), and there are now count limitations on pickoff attempts, a previously unlimited deterrent of stolen bases.
+Two changes are particularly relevant to stolen base attempts: Bases are now wider (reducing the distance between first and second base by 4.5 inches), and there are now limitations on pickoff attempts, or more formally, {\it disengagements}.
 
-    Whether attempting a stolen base or not, runners typically take a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, in the direction of the next base.
-The longer the lead, the shorter the distance need to run on a stolen base attempt.
-On a pickoff attempt, instead of throwing a pitch, the pitcher will disengage from the pitching rubber and throw the ball behind the runner to a fielder covering the base, attempting to tag the runner out.
+    Under the new rules, pitchers are limited to three disengagements.
+When attempting a pickoff, instead of throwing a pitch, the pitcher disengages from the pitching rubber and throws the ball to the runner's base.
 Under the new rules, if the pitcher disengages from the pitching rubber three times without successfully picking off the runner, then the runner automatically advances to the next base.
+Disengagements also include any time the pitcher steps off the rubber without attempting a pickoff.
 
@@
 
-    \textcite{downey_pick_2015} previously analyzed pickoff throws and stolen base attempts as a mixed strategy game without using lead distance data.
+    \textcite{downey_pick_2015} analyzed pickoff throws and stolen base attempts as a mixed strategy game without using lead distance data.
 They modeled the two-player game as a simultaneous decision in which the pitcher chooses whether to attempt a pickoff and the runner chooses whether to attempt a stolen base, showing that this theoretical game has a mixed-strategy Nash equilibrium.
-Under this equilibrium, left-handed pitchers (possessing better pickoff moves than their right-handed counterparts) would attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data.
+Under this equilibrium, left-handed pitchers (possessing better pickoff moves than their right-handed counterparts) attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data.
 In a later analysis, \textcite{downey_pressure_2019} argued that pitchers can effectively implement a mixed strategy in low-pressure situations but struggle to randomize their decisions as pressure increases, showing a higher auto-correlation for sequential pickoff decisions in high-pressure situations.
@@
 
-    Several studies in sports economics literature have examined the extent to which professional athletes play minimax strategy.
+    Several studies in the sports economics literature have examined the extent to which professional athletes play minimax strategy.
 \textcite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save.
@@
 
+    The history of quantitative analysis in baseball has deep roots in operations research \parencite{lindsey_investigation_1963}.
 The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model \parencite{bukiet_markov_1997}.
@@
 This is modeled as an infinite-horizon Markov decision process (MDP).
-We then extend this model to a two-player case where the first-base runner and the pitcher have agencies.
-The two-player case is modeled as an infinite-horizon zero sum stochastic game.
-While the single-player case is more simple and restrictive than the two-player case, it still provides valuable and actionable insights for runners because pitchers are unlikely to adopt an optimal policy in the short term.
+We then extend this model to a two-player case where the first-base runner and the pitcher have agency.
+The two-player case is modeled as an infinite-horizon zero-sum stochastic game.
+While the single-player case is simple and more restrictive than the two-player case, it still provides valuable and actionable insights for runners.
 Since the state space for the single- and two-player models are the same, we first describe the (common) state space before explaining the specifics of each model.
@@
 There is one terminal state (end of inning), denoted $\Delta$, bringing the total number of states to 865.
-We use $\mathcal{S}$ to denote the set of all states ( including the terminal state), and $S_t$ the state at play $t$ in a (fixed) innings.
+We use $\mathcal{S}$ to denote the set of all states (including the terminal state), and $S_t$ the state at play $t$ in a (fixed) inning.
 Traditionally, the state is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states) \parencite{bukiet_markov_1997}.
@@
 The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
-    \todo[inline]{Justify expected runs as the reward rather than win probability.}
+In other words, we assume the objective is to maximize runs rather than maximizing the probability of winning the game.
+The former objective is an approximation of the latter objective which allowed us to simplify the state space.
+This is a reasonable approximation before the final innings of the game, when it is more important to consider the score.
 
@@
   \noindent The above assumption states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
-  \todo[inline]{Justify in simple terms that inning in a baseball eventually halts.}
+This is a reasonable assumption because innings do not continue forever.
 
@@
    The function $V^*$, known as the optimal value function, is the maximum expected runs scored starting from state $s \in {\cal S}$.
-A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called the optimal policy.
+A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called an optimal policy.
 
-   We present the following definitions that will be needed to characterize the optimal value function and the optimal policy.
+   We present the following definitions that will be needed to characterize the optimal value function and an optimal policy.
 
@@
 
-The above theorem enables the characterization of $V^*(s)$ as in the following lemma.
+Theorem \ref{thm:mdp_optimality_condition} enables the characterization of $V^*(s)$ as in the following lemma.
 
@@
 Denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
-   Note that $\Pi_\text{R}$ and $\Pi_\text{P}$ are uncountable but compact. It is intuitive that an innings in a baseball eventually halts. The following assumption formalizes this intuition.
+   Note that $\Pi_\text{R}$ and $\Pi_\text{P}$ are uncountable but compact. It is intuitive that an innings in a baseball eventually halts. Assumption \ref{assum:game_halts} formalizes this intuition.
    \begin{assumption}\label{assum:game_halts}
@@
     In this section, we describe how we use empirical data to estimate state transition probabilities which we use for the single-player Markov decision process and the two-player stochastic game described in Section \ref{sec:sequential-model}.
-Throughout, we use the following notation to describe the data observed at the beginning of each play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt):
-    \begin{align*}
-      b_i & \in \{0, 1\}^3 \mbox{ denotes which bases (1B, 2B, 3B, respectively) are occupied;}\\
-      (c_i^B, c_i^S) \equiv c_i & \in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ denotes the ball-strike count;}\\
-      d_i & \in \{0, 1, 2\} \mbox{ denotes the number of disengagements already used; and}\\
-      o_i & \in \{0, 1, 2\} \mbox{ denotes the number of outs.}
-    \end{align*}
-    We use $\mathcal{H}^R$, $\mathcal{H}^P$, and $\mathcal{H}^C$ to denote the sets of runners, pitchers and catchers, respectively. On play $i$:
+Throughout, we use $s_i = (b_i, c_i, d_i, o_i)$ to denote the state of the game at the beginning of each observed play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt).
+We use $\mathcal{H}^R$, $\mathcal{H}^P$, and $\mathcal{H}^C$ to denote the sets of runners, pitchers and catchers, respectively.
+On play $i$:
     \begin{align*}
@@
     \end{align*}
-    When the play begins, we use the following notation to describe the actions taken by the players:
+
+    When the play begins, we use $\ell_i \in \mathbb{R}^+$ to denote the lead distance (in feet) taken by the first-base runner; and we use $p_i \in \{0, 1\}$ to denote whether the pitcher attempts (1) or does not attempt (0) a pickoff.
+We use $r_i \in \{\mbox{P}^+,\, \mbox{P}^-,\, \mbox{S}^+,\, \mbox{S}^-,\, \mbox{N}\} \equiv \mathcal{R}$ to denote the runner outcome, where
     \begin{align}
@@
       \begin{split}
-        \ell_i  \in &~ \mathbb{R}^+ \mbox{ denotes the lead distance (in feet) taken by the first-base runner;}\\
-        p_i \in &~ \{0, 1\} \mbox{ denotes whether the pitcher attempts (1) or does not attempt (0) a pickoff; and}\\
-        r_i \in &~ \{\mbox{P}^+,\, \mbox{P}^-,\, \mbox{S}^+,\, \mbox{S}^-,\, \mbox{N}\} \equiv \mathcal{R} \mbox{ denotes the runner outcome, where}\\
         & \mbox{P}^+  \mbox{ represents a successful pickoff attempt (runner is out);}\\
@@
       We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
+Specifically, we use logistic regression because it is appropriate for binary outcomes.
+The mixed-effects models allow the incorporation of player random effects along with fixed effects for game state variables and lead distance.
 
@@
         We model $\pi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher.
-These were the only variables for which we found substantive effects.
         \begin{align}
@@
 For this reason, we model stolen base attempts as a stochastic process rather than a deterministic decision.
+One might consider modeling the stolen base attempt decision as part of the action space for the runner.
+However, this decision relies on information not available in our data---namely, how well the runner reads the pitcher's body language and reacts to the pitcher's movement.
+More sophisticated modeling of this decision is an opportunity for future work.
 
@@
       \end{align}
-      In this equation, the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance.
+      In Equation (\ref{eqn:transition-mdp-model}), the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance.
 
-      To estimate the first probability inside the summation of equation (\ref{eqn:transition-mdp-model}), we use the models from Section \ref{sec:prob-runner-outcome}.
+      To estimate the first probability inside the summation of Equation (\ref{eqn:transition-mdp-model}), we use the models from Section \ref{sec:prob-runner-outcome}.
 Specifically, $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a)$ is given by
@@
 
-      To derive the transition probability function $p(s'|s,a_\text{R},a_\text{P})$, we modify equation (\ref{eqn:transition-mdp-model}) to condition the runner outcome probabilities on the pitcher's pickoff attempt decision as well as the runner's lead distance decision:
+      To derive the transition probability function $p(s'|s,a_\text{R},a_\text{P})$, we modify Equation (\ref{eqn:transition-mdp-model}) to condition the runner outcome probabilities on the pitcher's pickoff attempt decision as well as the runner's lead distance decision:
       \begin{align}
@@
 
-    In this section, we present the results of applying the single-player Markov decision process model and the two-player stochastic game model to real MLB data.
+    In this section, we present the results of applying the single-player Markov decision process model and the two-player stochastic game model to MLB data.
 Most of the data were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
@@
 
-    In addition to the play-by-play data, we manually downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+    In addition to the play-by-play data, we downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
 For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
-This is defined as the average of the top 10 percent of throw speeds when the catcher is attempting to catch a base stealer.
+This is defined by Baseball Savant as the average of the top 10 percent (to capture peak performance) of throw speeds when the catcher is attempting to catch a base stealer.
 For each runner-season, we observe {\it sprint speed} (feet per second), which measures the runner's top running speed in their fastest one-second window.
-This is defined as the average of the top 67 percent of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
+This is defined by Baseball Savant as the average of the top 67 percent (to exclude low-effort plays) of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
 
@@
 For each of these three models, the direction of the relationship between lead distance and outcome probability is as expected.
+For the pickoff success probability model, we only included a fixed effect for lead distance and a random effect for pitcher because these were the only variables for which we found substantive effects.
 
@@
 In practice, the purpose of this MDP is to optimize runner decision-making given current conditions of pitcher behavior, anticipating that pitcher behavior will take some time to react to changes in runner behavior.
+This model is more immediately actionable for runners than the two-player stochastic game because it is based on the present observed dynamic of pitcher behavior.
 
@@
 Three separate curves show the run expectancy for differing numbers of prior disengagements, assuming 0 balls; 0 strikes; 0 outs; third base unoccupied; and median player effects for pitcher, catcher and runner.
-Run expectancy is defined as the expectation of the value function $V_{\pi^*}$ (evaluated at the resulting state) corresponding to the optimal policy $\pi^*$ defined by (\ref{eqn:run-expectancy}).
+Run expectancy is defined as the expectation of the optimal value function $V^*$ (evaluated at the resulting state), defined by (\ref{eqn:mdp_val}).
 The vertical line corresponding to each curve marks the value at which the curve is maximized.
@@
 For example, with information about the runner's leaning posture and reaction time, future work might show that a mixed strategy for the pitcher's pickoff attempt decision achieves the best result by forcing a neutral posture for the runner and delaying the runner's reaction.
+As highlighted in Section \ref{sec:prob-sb-attempt}, another opportunity for future work is improved modeling of the runner's decision to attempt a stolen base, which we assumed to be stochastic and independent of lead distance.
 For now, we present the Two-Foot Rule as a simple rule of thumb that provides easily digestible and actionable guidance to runners.

### 2025-12-18 scripts/generate_results.R
refine names in results tables

@@ -118,18 +118,18 @@ table_po_attempt |>
   dplyr::mutate(
     variable = dplyr::case_when(
       variable == "(Intercept)" ~ "{\\it Intercept}",
-      variable == "lead1b" ~ "Lead Distance (feet)",
+      variable == "lead1b" ~ "Lead Distance (ft)",
       variable == "pre_balls" ~ "Balls",
       variable == "pre_strikes" ~ "Strikes",
       variable == "pre_outs" ~ "Outs",
       variable == "year2023" ~ "Year (2023 vs. 2022)",
       variable == "pre_disengagements1" ~ "Disengagements (1 vs. 0)",
       variable == "pre_disengagements2" ~ "Disengagements (2 vs. 0)",
-      variable == "sprint_speed" ~ "Sprint Speed (feet per second)",
-      variable == "arm_strength" ~ "Arm Strength (miles per hour)",
+      variable == "sprint_speed" ~ "Runner Sprint Speed (ft/s)",
+      variable == "arm_strength" ~ "Catcher Arm Strength (mi/h)",
       variable == "pitcher_id" ~ "Pitcher",
-      variable == "run1b" ~ "Runner",
       variable == "fielder_2_id" ~ "Catcher",
+      variable == "run1b" ~ "Runner",
     )
   ) |>
   tibble::add_row(variable = "{\\it Fixed Effects}", .after = 1) |>
@@ -418,7 +418,7 @@ sorted |>
   dplyr::select(count, dis, lead1b) |>
   tidyr::pivot_wider(names_from = count, values_from = lead1b, names_prefix = "count") |>
   sputil::write_latex_table(
-    file = "output/tables/lead_by_count.tex",
+    file = "output/tables/lead_by_count_one_player.tex",
     prefix_rows = "Prior & \\multicolumn{12}{c}{Count}",
     colnames = c(
       "Disengagements",
@@ -435,7 +435,7 @@ sorted |>
   dplyr::select(outs, dis, lead1b) |>
   tidyr::pivot_wider(names_from = dis, values_from = lead1b, names_prefix = "dis_") |>
   sputil::write_latex_table(
-    file = "output/tables/lead_by_outs.tex",
+    file = "output/tables/lead_by_outs_one_player.tex",
     prefix_rows = " & \\multicolumn{3}{c}{Disengagements}",
     colnames = c("Outs", "0", "1", "2"),
     align = "r|ccc",
@@ -484,7 +484,7 @@ sorted_two |>
[…]

### 2025-12-18 articles/arxiv/paper.tex

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem*{corollary*}{Corollary}
+\newtheorem{corollary}{Corollary}
 \newtheorem*{lemma*}{Lemma}
@@
 
-\title{The Two-Foot Rule: A game theoretic analysis of the pickoff limit in Major League Baseball}
+\title{
+  Lead distance under a pickoff limit in Major League Baseball:\\A sequential game model
+}
 
@@
   \begin{abstract}
-    In 2023, Major League Baseball (MLB) introduced new several rule changes to incentivize base stealing and more exciting plays.
-Among these changes, and the pitcher was limited to three pickoff attempts.
-This pickoff limit creates an interesting cat-and-mouse game between pitcher and runner because each failed pickoff attempt adds pressure on the pitcher to avoid using the next pickoff attempt.
-The runner can increase this pressure by extending the distance of their leadoff toward the next base, increasing their chances of successfully stealing the base.
-We model this dynamic as a sequential two-player stochastic game in which the runner first chooses a lead distance and then the pitcher chooses whether to attempt a pickoff.
-We prove that this game has a unique equilibrium strategy which we attain via policy iteration.
-Using pitch-by-pitch lead distance data obtained from MLB, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
-With these conditional transition probabilities, we estimate an optimal lead distance policy for the runner.
-The results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner on first base increase their lead distance by two feet after each successive pickoff attempt.
+    Recently, Major League Baseball (MLB)  limited  pitcher to three pickoff attempts, creating a cat-and-mouse game between pitcher and runner.
+Each failed  attempt adds pressure on the pitcher to avoid using another, and the runner can intensify this pressure by extending their leadoff toward the next base.
+We model this dynamic as a  two-player zero-sum sequential game in which the runner first chooses a lead distance, and then the pitcher chooses whether to attempt a pickoff.
+We establish optimality characterizations for the game and present variants of value iteration and policy iteration to solve the game.
+Using  lead distance data, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
+We compute the game-theoretic equilibria under the two-player model, as well as the optimal runner policy under a simplified one-player Markov decision process model.
+In the one-player setting, our results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
   \end{abstract}
+
+% In 2023, Major League Baseball (MLB) introduced several new rule changes to make the game more exciting for fans.
+Among these changes, the pitcher was limited to three pickoff attempts, creating a cat-and-mouse game between pitcher and runner.
+Each failed pickoff attempt adds pressure on the pitcher to avoid using another pickoff attempt, and the runner can intensify this pressure by extending their leadoff toward the next base.
+We model this dynamic as a sequential two-player zero-sum sequential game in which the runner first chooses a lead distance, and then the pitcher chooses whether to attempt a pickoff.
+We establish optimality characterizations for our sequential game and present variants of value iteration and policy iteration to solve the game.
+Using MLB lead distance data, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
+We compute the game-theoretic equilibrium policies under the two-player model using policy iteration, as well as the optimal runner policy under a reduced one-player Markov decision process model.
+In the one-player setup, our results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
 
@@
     On May 17, 2025, the Chicago Cubs hosted the Chicago White Sox in the second game of their Crosstown Series.
-In the bottom of the first inning, White Sox pitcher Sean Burke faced Cubs batter Michael Bush with two outs and Cubs runner Kyle Tucker on first base.
+In the bottom of the first inning, White Sox pitcher Sean Burke faced Cubs batter Michael Busch with two outs and Cubs runner Kyle Tucker on first base.
 Before the 1-1 pitch, Burke pivoted and threw to first base behind Tucker, who retreated safely before a tag.
-At the time, Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball among base stealers who had not been caught \parencite{fangraphs_major_2025}.
+Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball among base stealers who had not been caught \parencite{fangraphs_major_2025}.
 On the White Sox broadcast, color commentator Steve Stone explained to the audience, ``The only problem is now there's two throws over there, and Tucker's gonna get a much bigger lead.'' On the next pitch, Tucker stole his 11$^{th}$ base of the season, the first Burke had allowed.
@@
 
-    MLB stolen base totals peaked in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
+    MLB stolen base totals surged in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
 Before 2023, pickoff attempts could cause indefinite delays in the progress of a game because unsuccessful pickoff attempts do not advance the state of the game.
@@
 When attempting a pickoff, instead of throwing a pitch, the pitcher disengages from the pitching rubber and throws the ball to the runner's base.
-Under the new rules, if the pitcher disengages from the pitching rubber three times without successfully picking off the runner, then the runner automatically advances to the next base.
-Disengagements also include any time the pitcher steps off the rubber without attempting a pickoff.
+If the pitcher disengages from the pitching rubber three times without successfully picking off the runner, then all runners automatically advance to the next base.
+Disengagements also include any time the pitcher steps off the rubber without attempting a pickoff, which we disregard in the present work.
+The disengagement count resets if a runner advances or if a plate appearance ends.
 
@@
 Each unsuccessful pickoff attempt disincentivizes the pitcher from attempting another pickoff, which incentivizes the runner to increase their lead distance.
-As explained by Stone before the Tucker stolen base against the White Sox, players and coaches understand this intuitively.
-But how much should the runner extend their lead?
-And how does the answer change depending on runner speed and other context?
-
-    In the present work, we conduct the first public analysis of this game.
-We model the baseball inning as a Markov decision process (MDP) and solve for (a) the optimal policy for runner lead distance given current pitcher pickoff behavior and (b) the game theoretic equilibrium behavior in the two-player game between runner and pitcher.
-We present the practically actionable rule-of-thumb that an average runner should increase their lead distance by two feet after each disengagement.
+As indicated by Stone before the Tucker stolen base against the White Sox, players and coaches understand this intuitively.
+But how much should the runner extend their lead, and when should the pitcher attempt a pickoff?
+Our work introduces a formal model to address this timely question and the consequences arising from the implementation of our recommendations.
+
+    {\color{blue}{Section \ref{sec:sg-model} uses a two-payer zero-sum sequential game to model the evolving interaction between the runner and the pitcher.
+Our model captures all the vital aspects of the play while retaining analytical and computational tractability.
+Our model is closely related to the class of stochastic/dynamic Stackelberg games \parencite{goktas2022zero,li2017review,lopez2022stationary,vasal2020stochastic,vorobeychik2012computing}, which combines the elements of Stackelberg and stochastic games.
+The main difference between our infinite-horizon model and the literature on stochastic Stackelberg games is that we do not discount the rewards.
+The rationale for this modeling choice is that the rewards in our model correspond to the runs scored on a play, and runs scored on later plays count the same as those scored on earlier ones.
+We establish a maximin Bellman equation for our sequential game and propose a variant of value iteration and policy iteration to determine the optimal strategies for runner and the pitcher.}}
+
+    In addition to developing a model which captures the important features of the pitcher-runner game, we analyze the model using recently available real-world MLB data.
+The most innovative component of our implementation is the construction of transition probabilities.
+The outcome of each play is influenced by factors outside of our state variable, such as the sprint speed of the runner and the arm strength of the catcher.
+As a result, constructing transition probabilities using empirical conditional estimates (which use only the information from the current state and the actions of the player) is likely to be less accurate compared with a method that incorporates the player-specific factors.
+
+    Section \ref{sec:transition-probability-model} uses a two-step procedure that includes the player-specific factors to estimate the transition probabilities.
+In the first step, we use a mixed-effects logistic regression model that calculates the probability of various runner outcomes taking into account the actions of the players and the player-specific information.
+Next, we construct empirical estimates of the next state conditioned upon the current state, runner outcomes, and the actions of the player.
+Finally, we determine the transition probability via these two probabilities.
+
+    Section \ref{sec:results} presents a numerical study on recent real-world data which illustrates that, relative to the optimal policy, observed pitcher and runner decisions are conservative.
+Pitchers attempt fewer pickoffs than optimal, and runners take shorter leads than optimal.
+Relative to observed behavior, optimal play would favor the offense slightly.
+Moreover, when restricting out model to a single-player setup which allows agency only for the runner, the results reveal an actionable recommendation: The Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
+
+%    We use a two-step process to first estimate runner outcome probabilities and then use empirical transition probabilities conditioned on the runner outcome.
+For the runner outcome, we improve upon empirical probabilities by using multilevel models\footnote{citation needed} to share information across players using observable as well as unobservable player skills.
+Our numerical study on recent real-world data illustrates that, under the optimal policy, actions played by pitcher and runner would favor the runner, relative to observed decisions.
+The results obtained by restricting our model to a single-player setup, that allows agency only for the runner, reveals an actionable recommendation: The Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
 
@@
 They modeled the two-player game as a simultaneous decision in which the pitcher chooses whether to attempt a pickoff and the runner chooses whether to attempt a stolen base, showing that this theoretical game has a mixed-strategy Nash equilibrium.
-Under this equilibrium, left-handed pitchers (possessing better pickoff moves than their right-handed counterparts) attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data.
+Under this equilibrium, left-handed pitchers (typically possessing better pickoff moves than their right-handed counterparts) attempt fewer pickoffs and see fewer stolen bases attempted against them, a result validated by empirical data.
 In a later analysis, \textcite{downey_pressure_2019} argued that pitchers can effectively implement a mixed strategy in low-pressure situations but struggle to randomize their decisions as pressure increases, showing a higher auto-correlation for sequential pickoff decisions in high-pressure situations.
@@
 
-    The history of quantitative analysis in baseball has deep roots in operations research \parencite{lindsey_investigation_1963}.
-The Markov process is a very popular model in sports analytics, particularly in baseball, where the game's highly structured progression lends it well to a state space model \parencite{bukiet_markov_1997}.
+    The history of quantitative analysis in baseball has deep roots in operations research (e.g., \cite{lindsey_investigation_1963}; \cite{freeze_analysis_1974}; \cite{pankin_evaluating_1978}; \cite{russell_devising_1994}; \cite{bukiet_markov_1997}).
+The Markov process is particularly well suited for baseball, where the game's highly structured progression lends it well to an intuitive state space model \parencite{sokol_intuitive_2004}.
+\textcite{hirotsu_markov_2003} used a Markov model to inform substitution (pinch hitting) decision-making in baseball.
+The Markov process is also popular across sport analytics.
 \textcite{hirotsu_using_2002} used a four-state Markov model for soccer to inform substitution decision-making based on how substitutions impact transition probabilities between states.
-\textcite{hirotsu_markov_2003} used similar methodology to inform substitution (pinch hitting) decision-making in baseball.
 Recently, \textcite{chan_points_2021} used a Markov model to evaluate team performance in football.
@@
 They showed that the spread of pitcher effects is slightly greater than the spread of catcher effects.
-That analysis did not have data on runner sprint speed and catcher arm strength.
+That analysis did not use data on runner sprint speed and catcher arm strength.
 More recently, \textcite{stanley_modeling_2023} modeled stolen base success probability using both logistic regression and a random forest, including runner sprint speed and catcher arm strength as features but excluding player identities.
@@
 
-    \section{The baseball inning as a two-player stochastic game}
-    \label{sec:sequential-model}
-    We first describe a single-player setup where only the first-base runner has agency.
-This is modeled as an infinite-horizon Markov decision process (MDP).
-We then extend this model to a two-player case where the first-base runner and the pitcher have agency.
-The two-player case is modeled as an infinite-horizon zero-sum stochastic game.
-While the single-player case is simple and more restrictive than the two-player case, it still provides valuable and actionable insights for runners.
-Since the state space for the single- and two-player models are the same, we first describe the (common) state space before explaining the specifics of each model.
-
-    {\bf State space.} The game state is given by $s = (b, c, d, o)$, where:
+    \section{The two-player sequential game model}
+    \label{sec:sg-model}
+    In this section, we present our two-player zero-sum sequential game model under the pickoff limit rule.
+Though our sequential game model describes only a portion of an inning, it captures all the vital aspects that enable the generation of valuable and actionable insights for the runner and pitcher.
+Before describing the specifics of our model, we note that our model assumes a representative pitcher and a representative runner for the whole inning, rather than capturing full lineups and substitutions of players with varying skills.
+Without this assumption, we will end up with a nonstationary (infinite-horizon) two-player sequential game model, which appears to be significantly more challenging to analyze when compared with the stationary version.
+The four components of our sequential game are as follows.
+
+    {\bf 1. State space.} We use $s=(b,c,d,o) \in {\cal X}$ to capture a play, where     
+    % The mid-inning game state $s \in \mathcal{M}$ is given by $s = (b, c, d, o)$, where:
     \begin{align*}
@@
     \end{align*}
-    The total number of non-terminal states is $2^3 \times (4 \times 3) \times 3 \times 3 = 864$.
-There is one terminal state (end of inning), denoted $\Delta$, bringing the total number of states to 865.
-We use $\mathcal{S}$ to denote the set of all states (including the terminal state), and $S_t$ the state at play $t$ in a (fixed) inning.
+    The number of elements in ${\cal X}$ equals $2^3 \times (4 \times 3) \times 3 \times 3 = 864$.
+We use $\mathcal P$ to denote the set of four penultimate states and $\Delta$ to denote the terminal end-of-inning state.
+The purpose of the penultimate states is to keep track of the number of runs that are scored on the transition to the terminal state.
+A penultimate state $s \in \mathcal{P} = \{0, 1, 2, 3\}$ represents the number of runs scored on the final play of the inning.
+We use $\mathcal{S} \equiv \mathcal{X} \cup \mathcal{P} \cup \{\Delta\}$ to denote the set of all states, and $S_t$ to denote the state at play $t$ in a (fixed) inning.
 Traditionally, the state is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states) \parencite{bukiet_markov_1997}.
@@
 
-    \subsection{Single-player Markov decision process}
-    \label{sec:mdp-model}
-    This section considers the setup where agency is allowed only for the first-base runner. We model this scenario as a MDP. The other components of this MDP are as follows.
-
-    {\bf State space.} The state space for the MDP is the one described in the beginning of Section \ref{sec:sequential-model}.
-
-    {\bf Action space.} We limit the agency to states in which first base is occupied and the other bases are unoccupied.
-For such states, the action is the lead distance in feet chosen by the first-base runner from the first base.
-We formally describe it as follows.
-Define ${\cal L} \defeq [0,20]$.
-The action space of the runner, $A_\text{R}(s) \ \forall s \in {\cal S}$, is then defined as
+    {\bf 2.
+Action space.} We limit the agency of runners as well as pitchers to states in which only the first base is occupied (and the other bases are unoccupied).
+Though restricting the agency to states in which only the first base is occupied (i.e., the component $b$ in the state variable equals $(1,0,0)$) limits the scope, the situations in which the runner might steal the second base comprises roughly $71\%$ of stolen base attempts in 2023.
+We emphasize that this  excludes the scenario with $b = (1, 0, 1)$ in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
+We leave this for future research.
+
+    For states in which only the first base is occupied, the first-base runner's action is the lead distance chosen in feet from the first base.
+The pitcher has only two actions: attempt a pickoff or throw a pitch.
+Formally, the action space of the first-base runner, $A_{\text{R}}$, and the action space of the pitcher, $A_{\text{P}}$, are as follows.
+
     \begin{align}
     \label{eqn:runner_action_sapce}
-        A_\text{R}(s) \defeq \begin{cases}
+        A_\text{R}(s) &\defeq \begin{cases}
             {\cal L} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta.        \end{cases}
-    \end{align}    
-     Restricting the agency to states in which first base is occupied and the other bases are unoccupied, i.e., $b = (1, 0, 0)$, limits our scope to situations in which the runner might steal second base (encompassing 71\% of stolen base attempts in 2023).
-This obviously excludes the first-and-third scenario ($b = (1, 0, 1)$) in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
-
-     {\bf Single-stage reward function.} The single-stage reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition.
-This reward function does not depend on the action taken by the runner, and is given as
-     \begin{align}
-         \label{eqn:runner_reward}
-         r(s'|s,a) \defeq r(s'|s) \ \forall s,s' \in {\cal S}, a \in A_\text{R}(s),
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta, 
+        \end{cases} \ \forall s \in {\cal S}\\
+        \nonumber
+        \text{ where } {\cal L} &\defeq [0,20] \text{ is the set of allowable lead distances for the first-base runner.}
+    \end{align}    
+    \begin{align}
+    \label{eqn:pitcher_action_sapce}
+        A_{\text{P}}(s,a_\text{R}) &\defeq \begin{cases}
+            {\cal P} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0) \\
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta,         \end{cases} \ \forall s \in {\cal S}, \ a_\text{R} \in A_\text{R}(s)\\
+            \nonumber
+        \text{ where } {\cal P} &\defeq \{0,1\}, \text{ with } p = 1 \text{ corresponding to a pickoff attempt and } 0 \text{  otherwise.}
+    \end{align} 
+    Here, $\delta$ is a dummy element that captures the lack of agency.
+We note that the state of our stochastic game progresses even when no actions are played by both the players.
+This will be modeled through appropriate transition probabilities.
+
+    {\bf 3.
+Transition probabilities.} The uncertainty in our sequential game is modeled through appropriate transition probabilities.
+Basically, quantities describe the conditional probability of transitioning to a future state given the sequence of previously visited states and the actions played in those states.
+Our sequential game model assumes that the transition probabilities satisfy the Markov property, i.e., conditional on the entire history sequence, the probability of transitioning to a next state depends only on the current state and the actions played in that state.
+Following the convention in the literature, we denote the transition probability as $p(\cdot|s,a_\text{R},a_\text{P}) \ \forall s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})$.
+Section \ref{sec:transition-probability-model} details our specific construction of transition probabilities.
+At a high level, we use generalized linear mixed-effects models to estimate runner outcome (e.g., unsuccessful pickoff attempt or successful stolen base attempt) probabilities which depend on lead distance and the pitcher's pickoff attempt decision.
+This along with appropriate conditional empirical frequencies will be used to construct the transition probabilities.
+    % SP: We are deleting this sentence because we can't remember why we need it
+    %Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
+\todo{Siva: This last sentence needs to go to the section on transition probabilities.}
+
+    {\bf 4.
+Reward function.} The final component of our sequential game is the reward function, which models the payoff received by the players.
+Since we use a  zero-sum sequential game, the payoff obtained by one player comes at the expense of the other player.
+We choose the number of runs scored as the reward function.
+This design choice is more suitable for the early and middle innings of a game and less suitable for the later innings, particularly the bottom of the ninth or extra innings, when the offense knows how many runs they need to score to win the game.
+For any state $s \in {\cal S}$, the reward function captures the number of runs scored by the batting team in transitioning from state $s$ to $s'$ given the actions of the first-base runner and the pitcher.
+Formally,
+    \begin{align}
+      \label{eqn:runner_pitcher_reward}
+        r(s'|s,a_\text{R},a_\text{P}) \defeq r(s'|s) \ \forall s,s' \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R}),
     \end{align}
@@
       r(s'|s) \defeq  \begin{cases}
-              (g(b) + o) - (g(b') + o') + \mathbb{I}\{c' = (0, 0),\, d' = 0\} \ \forall s,s' \in {\cal S} - \{\Delta\}\\
-              0 \text{ if } s = \Delta,
+          (g(b) + o) - (g(b') + o') + \mathbb{I}\{c' = (0, 0),\, d' = 0\} & \mbox{ if } s,\, s' \in \mathcal{X}\\
+          \hfil s' & \text{ if } s \in \mathcal{X},\, s' \in \mathcal{P}\\
+          \hfil 0 & \text{ if } s' = \Delta,
       \end{cases} 
@@
     where $g: \{0, 1\}^3 \rightarrow \{0, 1, 2, 3\}$ counts the number of runners on base.
-This expression for the single-stage reward function counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance.
-For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state.
-To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state.
-These intermediary states transition to the terminal end-of-inning state with probability 1.
-
-    {\bf Transition probabilities.} The transition probability function $p(s'|s,a)$ depends on the action taken by the runner.
-We describe in detail our derivation of transition probabilities for the single-player Markov decision process in Section \ref{sec:transition-mdp-model}.
-In summary form, we use generalized linear mixed-effects models to estimate runner outcome (e.g., unsuccessful pickoff attempt or successful stolen base attempt) probabilities which depend on lead distance, and we use empirical frequencies to estimate state transition probabilities conditional on runner outcome probabilities.
-The state transition probabilities depend on lead distance only through the runner outcome probabilities.
-Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
-
-    The MDP is then defined by the tuple ${\cal M} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{r(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}, \\  \{p(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}\rangle$.
-Using this MDP, the progression of an inning can be captured as follows.
-At the beginning of a play, the game is in state $s \in {\cal S}$.
-Upon observing this state, an action $a \in A_\text{R}(s)$ is picked.
-The game then transitions to state $s'$ with probability $p(s'|s,a)$ and the batting team collects a reward $r(s'|s,a)$.
+Note that our reward function in \eqref{eqn:runner_pitcher_reward} depends only on the current state and the next state, and not on the actions of the first-base runner or the pitcher.
+The formula in \eqref{eqn:common_reward} counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance.
+Since the number of runs scored on the last play can not be uniquely determined from the current state, we use the penultimate states to track the number of runs scored on the last play.
+We note that our transition probabilities will be constructed such that these penultimate states transition to the terminal end-of-inning state with probability $1$.
+Since the reward function \eqref{eqn:runner_pitcher_reward} depends only on the current and the next state, it looks like it is independent of the actions of the players.
+However, the next state to which our sequential game transitions is (probabilistically) determined by the actions of both the players through the transition probabilities, thereby making the reward function indirectly depend on the actions of both the players.
+
+    Our sequential game models the progression of an inning as follows.
+Let $s \in {\cal S}$ denote the state of the sequential game from which a play begins.
+Upon observing this state, the first-base runner picks an action $a_\text{R} \in A_\text{R}(s)$.
+The pitcher observes the state $s$ and the first-base runner's action $a_\text{R}$, and then picks an action $a_\text{P} \in A_\text{P}(s,a_\text{R})$.
+After playing $a_\text{R}$ and $a_\text{P}$, the sequential game then transitions into state $s'$ with probability $p(s'|s,a_\text{R},a_\text{P})$ and the batting team collects a reward $r(s'|s,a_\text{R},a_\text{P})$ in the zero-sum game.
 A play then begins from $s'$ and the process repeats.
-The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
-In other words, we assume the objective is to maximize runs rather than maximizing the probability of winning the game.
-The former objective is an approximation of the latter objective which allowed us to simplify the state space.
-This is a reasonable approximation before the final innings of the game, when it is more important to consider the score.
-
-   A (deterministic stationary Markovian) policy $\pi$ for the runner is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi(s) \in A_\text{R}(s)$.
-That is, $\pi(s)$ is the action chosen in state $s$.
-\sout{In other words, a policy is a contingency plan (in this specific context) employed by the batting team.} Denote the set of all policies by $\Pi_\text{R}$.
-Note that $\Pi_\text{R}$ is uncountable but compact.
-It is intuitive that an inning in a baseball eventually halts.
-The following assumption formalizes this intuition.
-   \begin{assumption}\label{assum:mdp_halts}
-       \normalfont
-      The MDP ${\cal M}$ satisfies
-      $
-      \lim_{t \to \infty}\sup_{\pi \in \Pi, s \in {\cal S}}\mathbb{P}^\pi[S_t = \Delta | S_0 = s] = 1.
-      $
-   \end{assumption}   
-  \noindent The above assumption states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
-This is a reasonable assumption because innings do not continue forever.
-
-   The value of a policy $\pi$ when starting from state $s \in {\cal S}$ equals  
-   \begin{equation}
-       \label{eqn:mdp_pol_val}
-       V^{\pi}(s) \defeq \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
-   \end{equation}
-  That is, the value of the policy is the the expected total runs scored in an inning.
-Under Assumption \ref{assum:mdp_halts}, $V^\pi(s)$ exists for all $\pi \in \Pi$ and $s \in {\cal S}$.
-To see that, since $r(s'|s) \geq 0$ for all $s,s' \in {\cal S}$, we have by the monotone convergence theorem
-  \[
-  V^\pi(s) = \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] = \lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall \pi \in \Pi, s \in {\cal S}.
-  \]
-   Using Assumption \ref{assum:mdp_halts} and the fact that $\sup_{s,s' \in {\cal S}}\abs{r(s'|s)} < \infty$ (due to finiteness of ${\cal S}$), it is easy to establish that $\lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right]$ exists for all $\pi \in \Pi, s \in {\cal S}$.
-Furthermore, the limit is uniformly bounded over $\Pi \times {\cal S}$.
-The value of the MDP is defined as
-   \begin{equation}
-        \label{eqn:mdp_val}
-       V^*(s) \defeq \max_{\pi \in \Pi_\text{R}}V^\pi(s) \ \forall s \in S.
-   \end{equation}
-   The function $V^*$, known as the optimal value function, is the maximum expected runs scored starting from state $s \in {\cal S}$.
-A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called an optimal policy.
-
-   We present the following definitions that will be needed to characterize the optimal value function and an optimal policy.
-
-   \begin{definition}\label{defn:mdp_operators}
-    \normalfont
-    \begin{enumerate}
-        \item For any $\pi \in \Pi_\text{R}$, the evaluation operator $\Phi_\pi:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
-        \[
-        (\Phi_\pi V)(s) \defeq \mathbb{E}_{s' \sim p(\cdot|s,\pi(s))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-        \]
-        \item The optimality operator $\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
-        \[
-        (\Phi V)(s) \defeq \max_{a \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a)}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-        \]
-    \end{enumerate}       
-   \end{definition}
-
-   \begin{theorem}[\parencite{bertsekas1991analysis}]\label{thm:mdp_optimality_condition}
-       \normalfont
-      Suppose Assumption \ref{assum:mdp_halts} holds. Then
-      \begin{enumerate}
-          \item For any policy $\pi$, its value $V^\pi$ is the unique fixed point of the operator $\Phi_\pi$.
-          \item The optimal value function $V^*$ is the unique fixed point of the operator $\Phi$. Furthermore, $\pi^*$ is an optimal policy if and only if $\Phi_{\pi^*}V^* = \Phi V^*$.
-      \end{enumerate}
-   \end{theorem}
-% \todo[inline]{The above theorem also requires $p(s'|s,a)$ to be continuous in $a$ for any $s,s' \in S$.
-This is true for fixed-effects model since $x \mapsto e^x/1+e^x$ is continuous.
-I do not know about random-effects model.
-[[SP: $a$ comes into the model as a fixed effect (the random effects are for player identities), so, yes, $p(s'|s,a)$ is continuous in $a$.]]}
-
-Theorem \ref{thm:mdp_optimality_condition} enables the characterization of $V^*(s)$ as in the following lemma.
-
-\begin{lemma}\label{lem:mdp_value_function_characterization}
-    \normalfont
-    Suppose Assumption \ref{assum:mdp_halts} holds. Then $V^*$ is the unique solution to the following system of equations.
-    \begin{align*}
-        V(s) &= \max_{\ell \in {\cal L}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-        V(s) &= \sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
-        V(s) &= 0 \quad \text{ if } s = \Delta.
-    \end{align*}
-    \begin{proof}
-        Immediate from Theorem \ref{thm:mdp_optimality_condition}.
-    \end{proof}
-\end{lemma}
-
-An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{bertsekas1991analysis}.
-We use policy iteration  in this work to compute the optimal policy which is as follows.
-Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi^k$ by applying the updates
-        \begin{align}
-          \label{eqn:update-value-single-agent}
-          V^{k+1}(s) &= \sum_{s' \in \mathcal{S}} p(s' | s,\, \pi^k(s)) [r(s, s') + V^k(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
-          \label{eqn:update-policy-single-agent}
-          \pi^{k+1}(s) &= \argmax_{a \in A_\text{R}(s)} \sum_{s' \in \mathcal{S}} p(s' | s,\, a) [r(s, s') + V^{k+1}(s')] && \forall s \in \mathcal{S}.
-        \end{align}
-    The iterates generated by the policy iteration satisfy $V^k \uparrow V^*$ and every limit point of $\pi^k$ is an optimal policy \parencite{bertsekas1991analysis}.
-
-    \subsection{Two-player stochastic game}
-    \label{sec:stochastic-game}
-    In this section, we extend the single-agent model by allowing the pitcher to choose whether to attempt a pickoff.
-This is modeled as a two-player zero sum stochastic game, where the runner first chooses a lead distance, and then the pitcher decides whether to attempt a pickoff after observing the runner's lead distance.
-The components of this stochastic game are as follows.
-
-    {\bf State space.} The state space  is the same as described in the beginning of Section \ref{sec:sequential-model}.
-
-    {\bf Action space.} As in the single-player model of Section \ref{sec:mdp-model}, we limit actions (of runner as well as pitcher) to states in which only the first base is occupied.
-The (first-base) runner chooses a lead distance, and then the pitcher chooses a pickoff action $p \in \mathcal{P} \defeq \{0,1\}$, where $p = 1$ corresponds to a pickoff attempt and $0$ otherwise.
-The action space of the runner, $A_\text{R}(s) \ \forall s \in {\cal S}$, is the same as \eqref{eqn:runner_action_sapce}.
-The pitcher's action space $A_{\text{P}}(s,a_\text{R}) \ \forall s \in {\cal S}, a_\text{R} \in A_\text{R}(s)$ is defined as
-    \begin{align}
-    \label{eqn:pitcher_action_sapce}
-        A_{\text{P}}(s,a_\text{R}) \defeq \begin{cases}
-            {\cal P} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0), \ a_\text{R} \in A_\text{R}(s) \\
-            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta, \ a_\text{R} \in A_\text{R}(s).        \end{cases}
-    \end{align}
-
-    {\bf Single-stage reward function.} The single-stage reward function for transitioning from state $s$ to state $s'$ is given by the number of runs scored in the transition.
-Like in the single-player model, this reward function does not depend on the action taken by the runner, and is given
-     \begin{align}
-         \label{eqn:runner_pitcher_reward}
-         r(s'|s,a_\text{R},a_\text{P}) \defeq r(s'|s) \ \forall s,s' \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})
-     \end{align}
-     where $r(s'|s) \forall s,s' \in {\cal S}$ is as defined in \eqref{eqn:common_reward}.
-For transitions to the terminal end-of-inning state, the number of runs scored may not be uniquely determined by the starting state.
-To handle this case, we introduce intermediary states which track the number of runs scored on the last play, only for transitions to the terminal end-of-inning state.
-These intermediary states transition to the terminal end-of-inning state with probability 1.
-     \todo[inline]{Justify the paths/states through which the terminal state can be reached in an inning.}
-
-    {\bf Transition probabilities.} The transition probability function $p(s'|s,a_\text{R},a_\text{P})$ depends on the actions taken by the runner and pitcher.
-We describe in detail our derivation of transition probabilities for the two-player stochastic game in Section \ref{sec:transition-stochastic-game}.
-These transition probabilities are the same as for the single-player Markov decision process, except that the runner outcome probabilities are conditioned on the pitcher's pickoff attempt decision, as well as the lead distance.
-For example, if the pitcher does not attempt a pickoff, then the successful pickoff outcome and the unsuccessful pickoff outcome have probability zero.
-Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
-
-    The stochastic game is then defined by the tuple ${\cal G} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{A_\text{P}(s,a_\text{R})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s)},\\ \{r(s'|s,a_\text{R},a_\text{P})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})}^{s' \in {\cal S}}, \{p(s'|s,a_\text{R},a_\text{P})\}_{s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s, a_\text{R})}^{s' \in {\cal S}}\rangle$.
-The progression of an inning is then modeled as follows.
-At the beginning of a play, the game is in state $s \in {\cal S}$.
-Upon observing this state, the batting team picks action $a_\text{R} \in A_\text{R}(s)$.
-The pitcher then picks an action $a_\text{P} \in A_\text{P}(s,a_\text{R})$.
-The game then transitions to state $s'$ with probability $p(s'|s,a_\text{R},a_\text{P})$ and the batting team collects a reward $r(s'|s,a_\text{R},a_\text{P})$.
-A play then begins from $s'$ and the process repeats.
-The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning while the goal of the pitcher is to minimize the expected total rewards obtained by the batting team.
-
-   A (deterministic stationary Markovian) runner policy $\pi_\text{R}$ is a mapping ${\cal S} \mapsto \cup_{s \in {\cal S}}A_\text{R}(s)$ such that $\pi_\text{R}(s) \in A_\text{R}(s)$.
-That is, $\pi_\text{R}(s)$ is the action chosen in state $s$.
-Similarly, a (deterministic stationary Markovian) pitcher policy $\pi_\text{P}$ is a mapping of the form $\{(s,a) \in {\cal S} \times \cup_{s \in {\cal S}}A_\text{R}(s) \vert a \in A_\text{R}(s)\} \ni (x,a) \mapsto A_\text{P}(x,a)$ with the interpretation that $\pi_\text{P}(x,a_\text{R})$ is the action chosen by the pitcher when the state is $s$ and the runner picks the action $a_\text{R}$.
-Denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
-   Note that $\Pi_\text{R}$ and $\Pi_\text{P}$ are uncountable but compact. It is intuitive that an innings in a baseball eventually halts. Assumption \ref{assum:game_halts} formalizes this intuition.
-   \begin{assumption}\label{assum:game_halts}
-       \normalfont
-      The stochastic game ${\cal G}$ satisfies
-      $
-      \lim_{t \to \infty}\sup_{\pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}}\mathbb{P}^{\pi_\text{R}, \pi_\text{P}}[S_t = \Delta | S_0 = s] = 1.$
-   \end{assumption}
-
-   The value of the policy pair $(\pi_\text{R},\pi_\text{P}) \in \Pi_\text{R} \times \Pi_\text{P}$ is given by
+The goal of the batting team is to maximize (through the actions of the first-base runner) the expected total rewards over the plays in the inning while the goal of the fielding team is to minimize (through the actions of the pitcher) the expected total rewards obtained by the batting team.
+
+    It is intuitive that any sound strategy for the first-base runner and the pitcher depends on the state of the sequential game (for the pitcher, it also depends on the actions played by the first-base runner).
+This is formally captured using the concept of policy.
+A (deterministic, stationary) policy for the first-base runner is a decision-rule that prescribes the action to be played in every possible state.
+In this context, stationarity refers to the fact that the same action is played in a state, regardless of when this state occurs.
+In other words, the action depends only on the state and not the timing of when the state occurs.
+Similarly, a (deterministic, stationary) policy for the pitcher is a decision-rule that prescribes the action to be played for every possible state and the corresponding actions of the first-base runner in those states.
+A mathematical treatment of policies is presented in Section \ref{sec:appendix}.
+We denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
+
+    The concept of policies allows us to mathematically characterize the runs scored in an inning.
+Suppose the first-base runner and the pitcher employ policies $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, respectively.
+Assuming the play starts from a state $s \in {\cal S}$, the expected total runs scored (by the batting team) in an inning then equals
     \begin{equation}
        \label{eqn:sg_pol_val}
-       \widetilde V^{\pi_\text{R},\pi_\text{P}}(s) \defeq \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
+       V^{\pi_\text{R},\pi_\text{P}}(s) \defeq \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S},
    \end{equation}
-   Similar to the single-player case, it follows that Assumption \ref{assum:game_halts} along with the finiteness of $\cal S$ implies that $\widetilde V^{\pi_\text{R},\pi_\text{P}}(s)$ exists for all $(\pi_\text{R},\pi_\text{P}) \in \Pi_\text{R} \times \Pi_\text{P}, s \in {\cal S}$ and is uniformly bounded over $\Pi_\text{R} \times \Pi_\text{P} \times {\cal S}$.
-The value of the stochastic game is then defined as
+   where $\mathbb{E}^{\pi_\text{R},\pi_\text{P}}$ denotes the expectation with respect to the distribution induced by $\pi_\text{R}$ and $\pi_{\text{P}}$.
+Observe that, unlike the typical sequential games studied in the literature, we do not discount the rewards.
+This is because runs scored on later plays count the same as runs scored on earlier plays.
+Though the undiscounted case is more realistic in our setup, it might lead to technical issues since $V^{\pi_\text{R},\pi_\text{P}}(s)$ in \eqref{eqn:sg_pol_val} can potentially be infinite.
+However, it is obvious that an inning in a baseball eventually halts.
+This means that the infinite series appearing inside the expectation in the right-hand side of \eqref{eqn:sg_pol_val} can be truncated at some finite $T \in \mathbb{N}$, and hence we can expect $V^{\pi_\text{R},\pi_\text{P}}(s)$ to take finite values.
+This reasoning is rigorously justified in Section \ref{sec:appendix}.
+Therefore, in the rest of the paper, we take $V^{\pi_\text{R},\pi_\text{P}}(s)$ to be a finite number for all $\pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}$, and $s \in {\cal S}$.
+
+    The goal of the batting team is to select a policy that will maximize the expected total runs scored by them.
+At the same time, the fielding team tries to minimize the expected total runs scored by the batting team by employing a suitable policy.
+Taking the rational behavior of both the players into account, the maximum runs scored by the batting team when a play begins from state $s \in {\cal S}$ is given as
    \begin{equation}
         \label{eqn:sg_val}
-       \widetilde V^*(s) \defeq \max_{\pi_\text{R} \in \Pi_\text{R}}\min_{\pi_\text{P} \in \Pi_\text{P}}\widetilde V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in S.
+        V^*(s) \defeq \max_{\pi_\text{R} \in \Pi_\text{R}}\min_{\pi_\text{P} \in \Pi_\text{P}}V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in S.
    \end{equation}
-   The function $\widetilde V^*$ will be referred to as the value of the game.
-A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $\widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = \widetilde V^*(s) \ \forall s \in {\cal S}$ is called the optimal strategy or equilibrium strategy.
-Note that $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an optimal strategy if and only if $\widetilde V^{\pi_\text{R},\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}}(s) \ \forall \pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}$.
-
-   We present the following definitions that will be needed to characterize the optimal value function and the optimal policy.
-
-   \begin{definition}\label{defn:sg_operators}
-    \normalfont
-    \begin{enumerate}
-        \item For any $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, the evaluation operators $\widetilde\Phi_{\pi_\text{R}}:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ and $\widetilde\Phi_{\pi_\text{P}}:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ are defined as
-        \begin{align*}
-        (\widetilde\Phi_{\pi_\text{R}} V)(s) &\defeq \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}(s))}\mathbb{E}_{s' \sim p(\cdot|s,\pi_\text{R}(s),a_\text{P})}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}\\
-        (\widetilde\Phi_{\pi_\text{P}} V)(s) &\defeq \max_{a_\text{R} \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},\pi_\text{P}(s,a_\text{R}))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-        \end{align*}
-        \item The optimality operator $\widetilde\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
-        \[
-        (\widetilde\Phi V)(s) \defeq \max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-        \]
-    \end{enumerate}       
-   \end{definition}
-
-   \begin{theorem}[\parencite{patek1999stochastic}]\label{thm:sg_optimality_condition}
-       \normalfont
-      Suppose Assumption \ref{assum:game_halts} holds.
-Then the value of the game $\widetilde V^*$ is the unique fixed point of the operator $\widetilde \Phi$.
-Furthermore, $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an optimal strategy if and only if $\widetilde\Phi_{\pi_\text{R}^*}\widetilde V^* = \widetilde\Phi \widetilde V^* = \widetilde\Phi_{\pi_\text{P}^*}\widetilde V^*$.
-   \end{theorem}
-
-   The above theorem enables the characterization of $\widetilde V^*(s)$ as in the following lemma.
-
-\begin{lemma}\label{lem:sg_value_characterization}
-    \normalfont
-    Suppose Assumption \ref{assum:game_halts} holds. Then $\widetilde V^*$ is the unique solution to the following system of equations.
-    \begin{align*}
-        V(s) &= \max_{\ell \in {\cal L}}\min_{a \in {\cal P}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell,a) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-        V(s) &= \sum_{s' \in {\cal      S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
-        V(s) &=  0 \quad \text{ if } s = \Delta.
-    \end{align*}
-    \begin{proof}
-        Immediate from Theorem \ref{thm:sg_optimality_condition}.
-    \end{proof}
-\end{lemma}
-
-   An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{patek1999stochastic}.
-We use policy iteration  in this work to compute the optimal policy which is as follows.
-Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi_\text{R}^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi_\text{R}^k$ by applying the updates
-        \begin{align}
-          \label{eqn:update-value-two-agent}
-          V^{k+1}(s) &= \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}^k(s))} \sum_{s' \in \mathcal{S}} [r(s, s') +  V^k(s')]p(s'|s,\pi_\text{R}^k(s),a_\text{P}) && \forall s \in \mathcal{S}, \mbox{ and}\\
-          \label{eqn:update-policy-two-agent}
-          \pi_\text{R}^{k+1}(s) &= \argmax_{a_\text{R} \in A_\text{R}(s)} \left\{\min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s, s') + V^{k+1}(s')] p(s'|s,a_\text{R},a_\text{P})\right\} && \forall s \in \mathcal{S}.
-        \end{align}
-    The iterates generated by the policy iteration satisfy $V^k \uparrow \widetilde V^*$\parencite{patek1999stochastic}.
+   We call the function $V^*$ the \emph{optimal value function}.
+In our context, $V^*$ is the maximum expected total runs scored in an inning by the batting team, which, by the zero-sum property of the game, equals the minimum expected total runs incurred by the fielding team.
+A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called equilibrium policy.
+That is, $\pi_\text{R}^*$ is the policy for the batting team that maximizes the  expected total runs scored by them, and $\pi_\text{P}^*$ is the policy for the fielding team that minimizes the expected total runs incurred by them.
+In other words, $\pi_\text{R}^*$ is the optimal policy for the batting team and $\pi_\text{P}^*$ is the optimal policy for the fielding team.
+   % Note that any unilateral deviation from $(\pi_\text{R}^*,\pi_\text{P}^*)$ is suboptimal
+   % \todo{AS: Tone this down in case optimal not unique? Or do we have an expansive notion of suboptimal?}
+   % for the respective players.
+
+   A characterization of optimality conditions for \eqref{eqn:sg_val} including a dynamic programming equation is presented in Section \ref{sec:appendix}.
+In our computational experiments in Section \ref{sec:results}, we used value iteration to solve \eqref{eqn:sg_val}.
+The value iteration algorithm is as follows.
+Starting from an arbitrary $V^0 \in \mathbb{R}^{\abs{\cal S}}$ with $V^0(\Delta) = 0$, iteratively obtain $V^k$ via the following updates
+        \begin{align}
+          \label{eqn:vi_algorithm}
+          V^k(s) &\defeq \max_{a_\text{R} \in A_\text{R}(s)} \min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s'|s) + V^{k-1}(s')] p(s'|s,a_\text{R},a_\text{P}) \  \forall s \in \mathcal{S}.
+        \end{align}
+    Lemma \ref{lem:vi_convergence} proves that the $V^k \to  V^*$.
+The proof of that lemma establishes a convergence rate of the form $\abs{V^k(s) - V^*(s)} \leq O(\rho^k)$ for all $s \in {\cal S}$, where $\rho \in (0,1)$ is a constant.
+
+    We note that fixing a pitcher policy reduces our two-player sequential game to a single-player MDP whose objective is to optimize the lead distance for the first-base runner.
+Optimality criterion similar to the two-player sequential game also exists for MDP with undiscounted rewards \parencite[Section 7.2]{bertsekas2012dynamic}.
+Numerical experiments using this decision  model for a specific pitcher policy are presented in Section \ref{sec:two-foot-rule}.
+These results may provide more immediately actionable insights for lead distance selection, as we discuss in Section \ref{sec:discussion}.
 
@@
 
-    In this section, we describe how we use empirical data to estimate state transition probabilities which we use for the single-player Markov decision process and the two-player stochastic game described in Section \ref{sec:sequential-model}.
+    In this section, we describe how we estimate state transition probabilities which we use for the two-player sequential game described in Section \ref{sec:sg-model}.
 Throughout, we use $s_i = (b_i, c_i, d_i, o_i)$ to denote the state of the game at the beginning of each observed play $i \in \{1, ..., n\}$ (which can be a pitch or a pickoff attempt).
@@
       \begin{align*}
-        \pi_i    &= \mathbb{P}(R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff attempt probability;}\\
-        \pi_i^+  &= \mathbb{P}(R_i \in \{\mbox{P}^+\} \mid R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff success probability;}\\
+        \phi_i    &= \mathbb{P}(R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff attempt probability;}\\
+        \phi_i^+  &= \mathbb{P}(R_i \in \{\mbox{P}^+\} \mid R_i \in \{\mbox{P}^+,\, \mbox{P}^-\}) \mbox{ is the pickoff success probability;}\\
         \psi_i   &= \mathbb{P}(R_i \in \{\mbox{S}^+,\, \mbox{S}^-\} \mid R_i \in \{\mbox{S}^+,\, \mbox{S}^-, N\}) \mbox{ is the stolen base attempt probability; and}\\
@@
 
-        We model $\pi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), and lead distance; and with a random effect for the pitcher.
+        We model $\phi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), and lead distance; and with a random effect for the pitcher.
         \begin{align}
@@
           \begin{split}
-            \log\left(\frac{\pi_i}{1 - \pi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + \beta^L \ell_i + \gamma^P_{h_i^P}\\
+            \log\left(\frac{\phi_i}{1 - \phi_i}\right) &= \alpha + \beta^B c_i^B + \beta^S c_i^S + \beta^O o_i + \beta^D_{d_i} + \beta^L \ell_i + \gamma^P_{h_i^P}\\
             \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
@@
 
-        We model $\pi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher.
+        We model $\phi_i^+$ using a mixed-effects logistic regression with a fixed effect for lead distance and a random effect for pitcher.
         \begin{align}
@@
           \begin{split}
-            \log\left(\frac{\pi_i^+}{1 - \pi_i^+}\right) &= \alpha + \beta^L \ell_i + \gamma^P_{h_i^P},\\
+            \log\left(\frac{\phi_i^+}{1 - \phi_i^+}\right) &= \alpha + \beta^L \ell_i + \gamma^P_{h_i^P},\\
             \gamma^P_{h} &\sim \mathcal{N}(0, \sigma^2_P) \hspace{4mm} \mbox{\it i.i.d.} \hspace{4mm} \forall h \in \mathcal{H}^P.
@@
 
-    \subsection{Transition probabilities for the single-player Markov decision process}
-    \label{sec:transition-mdp-model}
-
-      To derive the transition probability function $p(s'|s,a)$, we use the law of total probability to combine the runner outcome probabilities estimated in this section with the conditional transition probability given the runner outcome:
-      \begin{align}
-        \label{eqn:transition-mdp-model}
-        p(s'|s,\, a) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r).
-      \end{align}
-      In Equation (\ref{eqn:transition-mdp-model}), the runner outcome depends on the lead distance, but we assume that given the runner outcome, the state transition is independent of the lead distance.
-
-      To estimate the first probability inside the summation of Equation (\ref{eqn:transition-mdp-model}), we use the models from Section \ref{sec:prob-runner-outcome}.
-Specifically, $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a)$ is given by
-      \begin{align*}
+    \subsection{Probability model for state transitions}
+    \label{sec:prob-state-transition}
+
+      We first describe the estimation of transition probabilities for the two-player sequential game, and then we present transition probabilities for the single-player MDP as a reduction.
+In the case where $a_\text{R} = a_\text{P} = \delta$, we simply use the empirical transition probabilities between states.
+Otherwise, we use the law of total probability to decompose the transition probability function $p(s'|s,a_\text{R},a_\text{P})$ conditionally on the runner outcome.
+The runner outcome depends on runner and pitcher actions, but we assume that given the runner outcome, the state transition does not depend on the actions.
+      \begin{align}
+        \label{eqn:prob-state-transition-two-player}
+        p(s'|s,\, a_\text{R},\, a_\text{P}) &= \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P}) \notag\\
+          &= \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P}) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P},\, R_i = r) \notag\\
+          &= \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P}) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r).
+      \end{align}
+      The first factor within the summation of (\ref{eqn:prob-state-transition-two-player}) is the runner outcome probability, and the second factor is the conditional transition probability given the runner outcome.
+To estimate runner outcome probability, we use the models from Section \ref{sec:prob-runner-outcome}.
+Specifically, assuming $a_\text{R} \ne \delta$ and $a_\text{P} \ne \delta$, then $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P})$ is given by
+      \begin{align}
+        \label{eqn:prob-runner-outcome-two-player}
+        \begin{split}
           \begin{cases}
-            \hfil \pi_i(a) \cdot \pi_i^+(a)                             & \mbox{if } r = \mbox{P}^+\\
-            \hfil \pi_i(a) \cdot (1 - \pi_i^+(a))                       & \mbox{if } r = \mbox{P}^-\\
-            \hfil (1 - \pi_i(a)) \cdot \psi_i \cdot \psi_i^+(a)         & \mbox{if } r = \mbox{S}^+\\
-            \hfil (1 - \pi_i(a)) \cdot \psi_i \cdot (1 - \psi_i^+(a))   & \mbox{if } r = \mbox{S}^-\\
-            \hfil (1 - \pi_i(a)) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
+              \hfil a_\text{P} \cdot \phi_i^+(a_\text{R})                             & \mbox{if } r = \mbox{P}^+\\
+              \hfil a_\text{P} \cdot (1 - \phi_i^+(a_\text{R}))                       & \mbox{if } r = \mbox{P}^-\\
+              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot \psi_i^+(a_\text{R})         & \mbox{if } r = \mbox{S}^+\\
+              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot (1 - \psi_i^+(a_\text{R}))   & \mbox{if } r = \mbox{S}^-\\
+              \hfil (1 - a_\text{P}) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
           \end{cases}.
-      \end{align*}
-      Recall that $\psi_i$ does not depend on $a$, as described in Section \ref{sec:prob-sb-attempt}.
+        \end{split}
+      \end{align}
+      Recall that $\psi_i$ does not depend on $a_\text{R}$, as described in Section \ref{sec:prob-sb-attempt}.
 
@@
 
-    \subsection{Transition probabilities for the two-player stochastic game}
-    \label{sec:transition-stochastic-game}
-
-      To derive the transition probability function $p(s'|s,a_\text{R},a_\text{P})$, we modify Equation (\ref{eqn:transition-mdp-model}) to condition the runner outcome probabilities on the pitcher's pickoff attempt decision as well as the runner's lead distance decision:
-      \begin{align}
-        \label{eqn:transition-stochastic-game}
-        p(s'|s,\, a) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R}, p_i = a_\text{P}) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r),
-      \end{align}
-      where $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R},\, p_i = a_\text{P})$ is given by
-      \begin{align*}
+      In Section \ref{sec:two-foot-rule} we present additional numerical results for the optimal runner policy assuming that the pitcher follows a fixed, exploitable policy (matching observed behavior).
+In this case, we follow the same estimation method for the transition probabilities except that the function $p(s'|s,a_\text{R})$ depends only on the runner's action.
+For this model, we replace (\ref{eqn:prob-state-transition-two-player}) with
+      \begin{align}
+        \label{eqn:prob-state-transition-one-player}
+        p(s'|s,\, a_\text{R}) = \sum_{r \in \mathcal{R}} \mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R}) \cdot \mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r),
+      \end{align}
+      where $\mathbb{P}(R_i = r \mid S_i = s,\, \ell_i = a_\text{R})$ is given by
+      \begin{align}
+        \label{eqn:prob-runner-outcome-one-player}
+        \begin{split}
           \begin{cases}
-              \hfil a_\text{P} \cdot \pi_i^+(a_\text{R})                             & \mbox{if } r = \mbox{P}^+\\
-              \hfil a_\text{P} \cdot (1 - \pi_i^+(a_\text{R}))                       & \mbox{if } r = \mbox{P}^-\\
-              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot \psi_i^+(a_\text{R})         & \mbox{if } r = \mbox{S}^+\\
-              \hfil (1 - a_\text{P}) \cdot \psi_i \cdot (1 - \psi_i^+(a_\text{R}))   & \mbox{if } r = \mbox{S}^-\\
-              \hfil (1 - a_\text{P}) \cdot (1 - \psi_i)                        & \mbox{if } r = \mbox{N}\\
+              \hfil \phi_i(a_\text{R}) \cdot \phi_i^+(a_\text{R})                             & \mbox{if } r = \mbox{P}^+\\
+              \hfil \phi_i(a_\text{R}) \cdot (1 - \phi_i^+(a_\text{R}))                       & \mbox{if } r = \mbox{P}^-\\
+              \hfil (1 - \phi_i(a_\text{R})) \cdot \psi_i \cdot \psi_i^+(a_\text{R})         & \mbox{if } r = \mbox{S}^+\\
+              \hfil (1 - \phi_i(a_\text{R})) \cdot \psi_i \cdot (1 - \psi_i^+(a_\text{R}))   & \mbox{if } r = \mbox{S}^-\\
+              \hfil (1 - \phi_i(a_\text{R})) \cdot (1 - \psi_i)                              & \mbox{if } r = \mbox{N}\\
           \end{cases}.
-      \end{align*}
-      The procedure for estimating $\mathbb{P}(S_{i+1} = s' \mid S_i = s,\, R_i = r)$ is the same as in Section \ref{sec:transition-mdp-model}.
-
-  \section{Numerical results}
-
-    In this section, we present the results of applying the single-player Markov decision process model and the two-player stochastic game model to MLB data.
+        \end{split}
+      \end{align}
+      Note that (\ref{eqn:prob-runner-outcome-one-player}) matches (\ref{eqn:prob-runner-outcome-two-player}) except that we have replaced the pitcher's action $a_\text{P}$ with a pickoff attempt probability $\phi_i(a_\text{R})$ which depends on the runner's action.
+
+  \section{Numerical study}
+  \label{sec:results}
+
+    In this section, we present the results of applying the two-player stochastic game model and a single-player MDP model and to MLB data.
 Most of the data were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
@@
 
-    In addition to the play-by-play data, we downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
+    We downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
 For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
@@
     In addition to the publicly available data described above, we rely on one key piece of proprietary information provided by MLB Advanced Media, estimated using computer vision on video data from several camera angles.
-For each play, observe each runner's {\it lead distance} (feet), which is how far they are ranging from their base at the time of the pitcher's first movement.
-This is defined as the distance between the runner's center of mass (projected onto the line between the runner's current base and their next base) and the nearest edge of the runner's current base.
+For each play with a runner on first base, observe the runner's {\it lead distance} (feet), which is how far they are ranging from first base at the time of the pitcher's first movement.
+This is defined as the distance between the runner's center of mass (projected onto the line between the first base and second base) and the nearest edge of first base.
 When the pitcher throws a pitch, runners will typically extend their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a secondary lead), which is relevant for runner advancement on batted balls but not on stolen base attempts.
@@
     \subsection{Runner outcome probability model}
+    \label{sec:results-multilevel-model}
 
@@
 The modeled probability remains low even for extremely long leads.
-We have very limited data on lead distances beyond 14 feet, especially with fewer than two prior disengagements, so extrapolation much beyond 14 feet is likely not reliable.
+MLB has very limited data on lead distances beyond 14 feet, especially with fewer than two prior disengagements, so extrapolation much beyond 14 feet is likely not reliable.
 The ball-strike count and the identity of the pitcher also influence the pickoff attempt probability, but the primary drivers are lead distance and the number prior disengagements.
@@
         \centering
-        \includegraphics[width = 0.8\textwidth]{prob_sb_success_light.pdf}
+        \includegraphics[width = 0.7\textwidth]{prob_sb_success_light.pdf}
         \caption{
@@
         \centering
-        \includegraphics[width = 0.49\textwidth]{effect_catcher_light.pdf}
-        \includegraphics[width = 0.49\textwidth]{effect_runner_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{effect_catcher_light.pdf}
+        \includegraphics[width = 0.4\textwidth]{effect_runner_light.pdf}
         \caption{
@@
 
-    \subsection{Single-player Markov decision process}
-
-      In the single-agent MDP, the only decision-maker is the runner, and they choose a policy for lead distance as a function of the game state, with the objective of maximizing the expected number of runs in the inning.
-In practice, the purpose of this MDP is to optimize runner decision-making given current conditions of pitcher behavior, anticipating that pitcher behavior will take some time to react to changes in runner behavior.
-This model is more immediately actionable for runners than the two-player stochastic game because it is based on the present observed dynamic of pitcher behavior.
-
-      Figure \ref{fig:finding-optimal-lead} illustrates how the runner would choose an optimal lead distance for different numbers of prior disengagements in an example scenario.
+    \subsection{Two-player sequential game}
+    \label{sec:results-two-player}
+
+      Using the transition probability function estimated on the empirical data, we implemented the policy iteration algorithm presented in Section \ref{sec:sg-model}.
+Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and prior disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and prior disengagements.
+This optimal lead distance policy maximizes runs scored while taking into account the rational behavior of both players, as defined by (\ref{eqn:sg_val}).
+For this runner policy, the pitcher is indifferent to attempting a pickoff or not.
+
+      \begin{table}[H]
+        \centering
+        \input{../../output/tables/lead_by_count_two_player.tex}
+        \caption{
+          \it The optimal lead distance (in feet) by count and prior disengagements under the two-player stochastic game, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
+        }
+        \label{tab:lead-by-count-two-player}
+      \end{table}
+
+      The most striking observation is that the optimal lead distance is significantly longer with two prior disengagements (14.8--16.1 feet) than in other states.
+These recommended lead distances are near the edge of observed lead distances in the dataset but not outside of the observed range.
+Overall, the optimal runner policy is more aggressive than observed behavior.
+Recall from Section \ref{sec:results-multilevel-model} that the average lead distance after zero, one, and two prior disengagements was 9.6 feet, 10.3 feet, and 11.0 feet, respectively.
+
+      \begin{table}[H]
+        \centering
+        \input{../../output/tables/lead_by_outs_two_player.tex}
+        \caption{
+          \it The optimal lead distance (in feet) by outs and prior disengagements under the two-player stochastic game, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
+        }
+        \label{tab:lead-by-outs-two-player}
+      \end{table}
+
+      We observe that the lead distance generally (although not strictly) shortens as the count deepens (i.e., as the numbers of balls and strikes increase).
+This makes intuitive sense because the plate appearance is closer to ending, which relieves the pressure on the pitcher to avoid attempting a pickoff.
+The lead distance generally (although not strictly) increases as the number of outs increases.
+This makes intuitive sense because the runner is less likely to score without more aggressive action.
+In other words, the riskier lead distance is more palatable because the runner has a lower baseline probability of scoring from first base with two outs.
+The non-monotonicity of the optimal lead distance policy may be a consequence of estimating transition probabilities using empirical data.
+
+      Under the game-theoretic equilibrium, pitchers would be more aggressive with attempting pickoffs when runners take long leads.
+Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 8.3\% of opportunities with only first base occupied.
+When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.8\% of the time---this would be 100\% in the two-player model.
+When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 5.0\% of the time---this would be 0\% in the two-player model.
+Overall, pitchers attempted pickoffs on 5.8\% of opportunities, less than the recommended 8.3\%.
+
+      Despite the fact that pitchers would attempt more pickoffs under the game-theoretic equilibrium, Tables \ref{tab:lead-by-count-two-player} and \ref{tab:lead-by-outs-two-player} recommend longer leads for runners, particularly with two prior disengagements.
+The optimal strategy for the runner is maximin because they choose the lead distance which maximizes the minimum run expectancy across the pitcher's two options (pickoff or pitch).
+The pitcher will never attempt a pickoff when it is suboptimal to do so.
+The optimal behavior for the runner is to force their hand (or rather, to make the pitcher indifferent to their two choices).
+
+    \subsection{One-player MDP and the Two-Foot Rule}
+    \label{sec:two-foot-rule}
+
+      Pitchers may be unlikely to adapt their strategy to the optimal policy in the short term.
+Runner actions are already observable by the defense, and current pitcher behavior seems more conservative than optimal, as noted in Section \ref{sec:results-two-player}.
+After all, the pitcher has the additional cognitive burden of contending with the batter.
+In light of this, we consider the results of policy iteration applied to the one-player MDP.
+In this model, we assume the pitcher's pickoff attempt decision is a stochastic policy modeled using empirical data.
+Because the pitcher's action is not optimal, the runner can exploit it.
+Table \ref{tab:lead-by-count-one-player} reports how the optimal runner policy varies by count and prior disengagements, and Table \ref{tab:lead-by-outs-one-player} reports how the optimal runner policy varies by outs and prior disengagements.
+
+      \begin{table}[H]
+        \centering
+        \input{../../output/tables/lead_by_count_one_player.tex}
+        \caption{
+          \it The optimal lead distance (in feet) by count and prior disengagements under the single-player Markov decision process, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
+        }
+        \label{tab:lead-by-count-one-player}
+      \end{table}
+
+      For the most part, the trends in the optimal lead distance policy in the one-player MDP are similar to that of the two-player game.
+One exception is that optimal lead distance increases with the number of balls in the count because pitchers are less likely to attempt pickoffs (Table \ref{tab:model-summary}).
+The relationship between prior disengagements and optimal lead distance is also different.
+We observe that increasing prior disengagements from zero to one generally increases the optimal lead distance by approximately 1.5 feet.
+Increasing prior disengagements from one to two generally increases the optimal lead distance by approximately 2.5 feet.
+Based on these results, we recommend the {\it Two-Foot Rule}: that runners increase their lead by two feet after each disengagement, a very simple and actionable guideline for athletes to follow.
+
+      \begin{table}[H]
+        \centering
+        \input{../../output/tables/lead_by_outs_one_player.tex}
+        \caption{
+          \it The optimal lead distance (in feet) by outs and prior disengagements under the single-player Markov decision process, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
+        }
+        \label{tab:lead-by-outs-one-player}
+      \end{table}
+
+% SP: I don't think we need this given the intended OR audience
+%      Figure \ref{fig:finding-optimal-lead} illustrates how the runner would choose an optimal lead distance for different numbers of prior disengagements in an example scenario.
 Short lead distances have very low pickoff probabilities and low stolen base probabilities.
@@
 The optimal lead distances for 0, 1 and 2 disengagements are 10.0 feet, 11.5 feet and 14.1 feet, respectively.
-
-      \begin{figure}[H]
-        \centering
-        \includegraphics[width = 0.8\textwidth]{finding_optimal_lead_light.pdf}
-        \caption{
-          \it Rest-of-inning run expectancy for different choices of lead distance.
+%      
+%      \begin{figure}[H]
+%        \centering
+%        \includegraphics[width = 0.8\textwidth]{finding_optimal_lead_light.pdf}
+%        \caption{
+%          \it Rest-of-inning run expectancy for different choices of lead distance.
 Three separate curves show the run expectancy for differing numbers of prior disengagements, assuming 0 balls; 0 strikes; 0 outs; third base unoccupied; and median player effects for pitcher, catcher and runner.
@@
 The vertical line corresponding to each curve marks the value at which the curve is maximized.
-        }
-        \label{fig:finding-optimal-lead}
-      \end{figure}
-
-      Policy iteration, as detailed by equations (\ref{eqn:update-value-single-agent}) and (\ref{eqn:update-policy-single-agent}), yields a policy $\pi^*$ optimizing rest-of-inning run expectancy.
-Tables \ref{tab:lead-by-count}, \ref{tab:lead-by-runners-outs} and \ref{tab:lead-by-players} report the optimal lead distance policy $\pi^*$ for different game states and player skills.
-
-      Table \ref{tab:lead-by-count} shows how optimal lead distance varies by ball-strike count and number of prior disengagements, for an average runner facing an average battery (pitcher and catcher) with no outs.
-We observe that count has a mild influence on optimal lead distance, and the number of prior disengagements has a strong influence.
-As the pitcher nears the disengagement limit, the runner has more incentive to extend their lead distance because pickoff attempts are less likely and a free base advancement (from a third unsuccessful pickoff attempt) is more likely.
-
-      \begin{table}[H]
-        \centering
-        \input{../../output/tables/lead_by_count.tex}
-        \caption{
-          \it The optimal lead distance (in feet) by count and prior disengagements for an average runner on first base, facing an average battery (pitcher and catcher) with no outs and third base empty.
-This lead distance policy maximizes the expected runs to end of inning as modeled by a the single-agent MDP.
-        }
-        \label{tab:lead-by-count}
-      \end{table}
-
-      Across counts, the recommended increase in lead distance is generally close to 1.5 feet after the first disengagement and 2.5 feet after the second disengagement.
-We observe this to be a good approximation across different runner/battery skills and different numbers of outs.
-Based on these results, we recommend the {\it Two-Foot Rule}: that runners increase their lead by two feet after each disengagement, a very simple and actionable guideline for athletes to follow.
-
-      Table \ref{tab:lead-by-runners-outs} shows how optimal lead distance varies by number of outs, for an average runner facing an average battery with a 0-0 count.
-Again, we observe that each successive disengagement leads to an increase in recommended lead distance of approximately two feet.
-The recommended leads for zero outs and one out are virtually identical, and for two outs they are slightly longer.
-With two outs, stolen bases are more likely and pickoff attempts are less likely, so runners can be more aggressive.
-
-      \begin{table}[H]
-        \centering
-        \input{../../output/tables/lead_by_outs.tex}
-        \caption{
-          \it The optimal lead distance (in feet) by outs and prior disengagements for an average runner on first base, facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
-This lead distance policy maximizes the expected runs to end of inning as modeled by the single-agent MDP.
-        }
-        \label{tab:lead-by-runners-outs}
-      \end{table}
-
-      Table \ref{tab:lead-by-players} shows how optimal lead distance varies by battery and runner skill, with no outs and a 0-0 count.
-The battery skill accounts for catcher arm strength and for catcher and pitcher random effects.
-The composite runner skill accounts for sprint speed and for runner random effect.
-Against batteries more skilled at controlling the run game, the optimal lead distance is shorted.
+%        }
+%        \label{fig:finding-optimal-lead}
+%      \end{figure}
+
+      Table \ref{tab:lead-by-players} shows the robustness of the Two-Foot Rule under different assumptions regarding player talent.
+Against batteries more skilled at controlling the run game, the optimal lead distance is shorter.
 For more skilled runners, the optimal lead distance is longer.
-Regardless of the skill levels, the optimal policy for the runner is to increase their lead by 1.5--2.5 feet with each successive disengagement.
-We observe that the Two-Foot Rule is a good approximation for different combinations of battery and runner skill, not just for average players.
+Regardless of the skill levels, the rule of thumb holds up well.
+The most notable exception to the Two-Foot Rule is when facing a 10th-percentile battery after the first disengagement.
+In this situation, we recommend a long lead before the first disengagement and an increase of roughly one foot after the first disengagement.
 
@@
           \it Optimal lead distance (in feet) by battery/runner skill and prior disengagements, for an 0-0 count with no outs and third base empty.
-Player skill is held constant across runner outcome models; for example, the hypothetical 90$^{th}$-percentile battery represents the 90$^{th}$ percentile at attempting pickoffs, successfully executing pickoffs, suppressing stolen base attempts, and catching would-be base stealers.
+For example, the hypothetical 90$^{th}$-percentile battery represents the 90$^{th}$ percentile at attempting pickoffs, successfully executing pickoffs, suppressing stolen base attempts, and catching would-be base stealers.
         }
@@
 
-      Table \ref{tab:actual-vs-rec} compares observed runner behavior against the optimal policy learned by the MDP.
+% SP: Remove this result to reduce the imbalance between two-player results and one-player results because we have chosen to emphasizse the two-player model more.
+%      Table \ref{tab:actual-vs-rec} compares observed runner behavior against the optimal policy learned by the MDP.
 For each pitch with a runner on first base and second base empty, we compare the runner's lead distance with the recommended lead distance for an average runner against an average battery.
@@
 On average, batters increase their leads by 0.6--0.7 feet afer each disengagement, significantly less than the Two-Foot Rule would recommend.
-
-      \begin{table}[H]
-        \centering
-        \input{../../output/tables/actual_vs_rec_lead.tex}
-        \caption{
-          \it Comparison of actual versus recommended lead distance by number of prior disengagements.
-The recommended lead distance is based on the single-agent MDP model and accounts for count, outs and prior disengagements, assuming average battery and runner skill.
-        }
-        \label{tab:actual-vs-rec}
-      \end{table}
-
-      Our recommendations differ from runner behavior, and we can approximate the difference in run value between runner behavior and our recommendations.
-In 2023, the scoring expectation was 0.516 runs per inning.
-Under the optimal lead distance policy $\pi^*$, the corresponding value of the start-of-inning state is 0.528 runs, an increase of 0.012 runs per inning.
-While that may seem insigificant at first, it adds up to approximately 17 runs over 162 games of nine innings each.
-To add that many expected runs in free agency would cost roughly \$17 million \parencite{clemens_what_2021}, which speaks to a significant (but not outlandish) impact.
-
-    \subsection{Two-player stochastic game}
-
-      All of the results shown so far correspond to our single-agent model, in which the runner chooses a lead distance and all other outcomes are probabilistic based on their choice.
-While the single-agent model is the most immediately useful framework for runner decision-making in the current environment, we also consider how pitchers might respond using the same model to optimize their decision-making.
-In the two-agent model, the runner first chooses a lead distance, and the pitcher then decides whether to attempt a pickoff, as detailed in Section \ref{sec:sequential-model}.
-The optimal strategy for the runner is maximin because they choose the lead distance which maximizes the minimum run expectancy across the pitcher's two options (pickoff or pitch).
-
-      In the two-agent model, pitchers would be more aggressive with attempting pickoffs when runners take long leads.
-Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 8.3\% of opportunities with only first base occupied.
-When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.8\% of the time---this would be 100\% in the two-agent model.
-When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 5.0\% of the time---this would be 0\% in the two-agend model.
-Overall, pitchers attempted pickoffs on 5.8\% of opportunities, less than the recommmended 8.3\%.
-
-      Table \ref{tab:count-twoagent} reports the optimal lead distance policy by count when both runner and pitcher are behaving optimally under the two-agent model.
-In this setting, the optimal lead distance is longer than in the single-agent model (Table \ref{tab:lead-by-count}), especially early in the count and with two prior disengagements.
-The longer lead distances are somewhat counterintuitive because we expect more pickoff attempts from pitchers when behaving optimally.
-The explanation is that in the two-agent model, the pitcher will never attempt a pickoff when it is suboptimal to do so.
-The optimal behavior for the runner is to force their hand (or rather, to make the pitcher indifferent to their two choices).
-Interestingly, the Two-Foot Rule breaks down in this setting.
-Averaged across counts, we recommend extending the lead by roughly 0.7 feet after the first disengagement and by much more after the second disengagement---roughly 4.2 feet!
-However, this advice will only be actionable if pitchers change their behavior to be more in line with the modeled optimal policy.
-
-      \begin{table}[H]
-        \centering
-        \input{../../output/tables/count_two_agent.tex}
-        \caption{
-          \it Based on the count in each row, each entry shows the optimal lead distance (in feet) for a runner on 1B depending on the number of disengagements for our two-agent model.
-        }
-        \label{tab:count-twoagent}
-      \end{table}
-
-      If both runners and pitchers adapted their behavior to match the game theoretic optimal policy under the two-agent model, the corresponding run value of the start-of-inning state is 0.520 runs per inning.
-Relative to the current status quo, this reflects an increase of 0.004 runs per inning, meaning a very slight increase to overall offense.
-Over 162 games of nine innings each, we would expect each team to score approximately 5 more runs over a full season.
-
-  \section{Discussion}
+%
+%      \begin{table}[H]
+%        \centering
+%        \input{../../output/tables/actual_vs_rec_lead.tex}
+%        \caption{
+%          \it Comparison of actual versus recommended lead distance by number of prior disengagements.
+The recommended lead distance is based on the one-player MDP model and accounts for count, outs and prior disengagements, assuming average battery and runner skill.
+%        }
+%        \label{tab:actual-vs-rec}
+%      \end{table}
+
+  \section{Discussion and strategic insights}
+  \label{sec:discussion}
+
+    In comparing the optimal lead distance policy under the two-player stochastic game model (Table \ref{tab:lead-by-count-two-player}) and the single-player MDP (Table \ref{tab:lead-by-count-one-player}), we observe substantial differences between the optimal lead distance policy under a game-theoretic equilibrium and the optimal lead distance policy given observed pitcher behavior.
+For most states, the former lead distance is longer than the latter.
+The two-player optimal policy for the runner is to extend their lead until the pitcher is indifferent between attempting a pickoff or not.
+By contrast, the one-player optimal policy corresponds to a pickoff attempt probability below 10\%.
+Because the pitcher sometimes attempts low-success pickoffs, the runner does not need to extend their lead further.
+
+    In the two-player model we recommend long leads (beyond 15 feet) after two prior disengagements.
+These distances are near the edge of the range of leads we observe but not outside the range.
+To develop intuition for this policy, consider the optimal lead distance of 15.1 feet in a 3-2 count after two prior disengagements (Table \ref{tab:lead-by-count-two-player}).
+If an average pitcher attempts a pickoff in this scenario, we estimate a 20.5\% probability that the runner is out.
+Otherwise, the runner advances to second base.
+By contrast, if the pitcher throws a pitch and if the runner attempts a stolen base (14.0\% probability), we estimate an 82.8\% probability that the stolen base is successful.
+In other words, the pitcher is indifferent to throwing a pitch or attempting a pickoff because the probability of an unsuccessful pickoff attempt is similar to the probability of a successful stolen base (conditioned on a steal attempt).
+
+    Because a third failed pickoff attempt results directly in free baserunner advancement, the pitcher is strongly disincentivized from attempting a pickoff after two prior disengagements.
+In the two-player model, this is reflected in a drastic increase in the optimal lead distance after the second disengagement (3--5 feet, Table \ref{tab:lead-by-count-two-player}).
+In the one-player model, we observe a more steady increase in optimal lead distance as a function of prior disengagements because observed pitcher pickoff attempt beheavior steadily decreases as a function of prior disengagements (see Figure \ref{fig:prob-pickoff}).
+
+    Given observed lead distance behavior, offenses scored 0.516 runs per inning in 2023.
+In other words, if we model the inning as a Markov reward process using empirical transition probabilities between states, the value of the start-of-inning state is 0.516 runs.
+Under the two-player game, the value of the start-of-inning state is 0.520 runs, which corresponds to approximately 5 more runs over a full season of 162 games of nine innings each.
+Therefore, offenses stand to benefit more in the transition from observed behavior to the game-theoretic equilibrium.
+Under the one-player MDP, the value of the start-of-inning state is 0.528 runs, which corresponds to approximately 17 more runs over a full season, relative to observed behavior.
+Naturally, scoring expectation is higher in the one-player MDP because the runner exploits the sub-optimal strategy of the pitcher.
+To add 17 expected runs in free agency would cost roughly \$17 million \parencite{clemens_what_2021}, which speaks to a significant (but not outlandish) impact.
+This value may have increased since the last study we found to have estimated it.
+
+    When putting these results into practice for an individual team, we recommended implementing the optimal lead distance of the one-player MDP rather than the two-player game.
+In the short term, pitchers are unlikely to immediately adapt their behavior to the optimal policy, and runners can exploit this.
+Because the pickoff attempt action is observable, runners can adjust their lead distance policy if they observe pitchers making more optimal pickoff attempt decisions.
+Therefore, the actionable strategic insight from our results is the Two-Foot Rule: Extend your lead by two feet after each pickoff attempt.
+This is a simple and easily communicated rule of thumb that approximates the optimal policy for runners given observed pitcher behavior.
+
+  \section{Conclusions}
 
     Under the new pickoff rules, baseball players and coaches understand intuitively that runners want to increase their lead distance after each successive disengagement by the pitcher---but by how much?
-In this paper, we formalize the cat-and-mouse game between pitcher and runner as a Markov decision process and estimate the optimal lead distance policy for a runner on first base with other bases empty, given the number of outs, the ball-strike count, and the number of prior disengagements by the pitcher.
+In this paper, we formalize the cat-and-mouse game between pitcher and runner as a two-player stochastic game and estimate the optimal lead distance policy for a runner on first base with other bases empty, given the number of outs, the ball-strike count, and the number of prior disengagements by the pitcher.
 We model the probabilities of runner outcomes (pickoff attempt/success, stolen base attempt/success) as functions of lead distance, context, and player skill.
@@
 
-    First, we consider a single-agent model in which the runner has agency in deciding their lead distance, and the pitcher's decision of whether to attempt a pickoff is probabilistic, conditioned on the lead distance.
+    We consider a two-player sequential game model in which the runner first decides their lead distance and then the pitcher (upon observing the runner's lead distance) then decides whether to attempt a pickoff or throw a pitch.
+We show that the game-theoretic optimal policy for the runner is a maximin strategy (with respect to rest-of-inning run expectancy) so that the pitcher is indifferent to attempting a pickoff or throwing a pitch.
+We estimate model parameters using empirical data from the 2023 MLB season and find that under the game-theoretic equilibrium, runners would be more aggressive with their lead distances than the current behavior we observe.
+While baseball analytics sometimes garners negative attention for diminishing the aesthetic appeal of the game \parencite{sheinin_analytics_2025}, this is an example for which the analytically recommended strategy could be more aesthetically pleasing.
+In other sports, stochastic and dynamic models have shown that decision-makers have been traditionally too conservative \parencite{romer_firms_2006, beaudoin_strategies_2010, yam_what_2019}, but this is a new result for leadoff and pickoff attempt decisions baseball.
+
+    We go a step further and consider the opportunity for runners to exploit the observed suboptimal behavior by pitchers.
+We consider a one-player MDP model in which the pitcher's agency in the pickoff attempt decision is replaced by a stochastic process, modeled on empirical data as a function of the runner's lead distance.
 This framework leads to the most actionable recommendation for runners under the current conditions of pitcher behavior.
 We estimate that the average team could expect to increase their expected offensive output by approximately 17 runs over a full season by implementing the optimal lead distance policy.
-Based on these results, we propose the Two-Foot Rule: We recommend that a runner on first base increase their lead distance by two feet after each successive disengagement.
+Based on these results, we propose the Two-Foot Rule: Extend your lead by two feet after each pickoff attempt.
 This easily implemented recommendation is a good approximation to the optimal lead distance policy across a wide array of contexts (outs and ball-strike count) and pitcher/runner skills.
-
-    We go a step further and consider the consequences of granting the pitcher agency to decide whether to attempt a pickoff given the runner's lead distance.
-We give a game theoretic treatment to this two-agent model in which the runner follows a maximin strategy (with respect to rest-of-inning run expectancy) so that the pitcher is indifferent to attempting a pickoff or throwing a pitch.
-Under the game theoretic equilibrium, runners are more aggressive with their lead distance and pitchers are more aggressive with their pickoff attempts than the current behavior we observe.
-While baseball analytics sometimes garners negative attention for diminishing the aesthetic appeal of the game \parencite{sheinin_analytics_2025}, this is an example for which the analytically recommended strategy could be more aesthetically pleasing.
 
@@
 In reality, the pitcher may derive value from stepping off of the pitching rubber if they are short of breath or need to resolve a miscommunication with the catcher.
+Including all these extensions is an interesting future work.
+Another future direction is to model an entire roster, which is a nonstationary two-player stochastic game and introduces enormous technical challenges.
 
@@
 As highlighted in Section \ref{sec:prob-sb-attempt}, another opportunity for future work is improved modeling of the runner's decision to attempt a stolen base, which we assumed to be stochastic and independent of lead distance.
-For now, we present the Two-Foot Rule as a simple rule of thumb that provides easily digestible and actionable guidance to runners.
 
@@
 
-    Major League Baseball trademarks and copyrights are used with permission of MLB Advanced Media, L.P. All rights reserved.
-
-  \section*{Code and Data Availability}
-
-    All code related to this paper is available at \url{https://github.com/jfhahn2/pickoff-game-theory}.
+    Major League Baseball trademarks and copyrights are used with permission of MLB Advanced Media, L.P.
+All rights reserved.
+The authors thank Mallesh Pai of Rice University for constructive discussions during conceptualization of the problem and early implementation of the solution.
+
+  \section*{Code and data availability}
+
+    Code related to this paper is available at \url{https://github.com/jfhahn2/pickoff-game-theory}.
 Lead distance data are not publicly available and were obtained from Major League Baseball.
@@
 
+\section{Appendix: Technical results}
+\label{sec:appendix}
+We present a formal description of the sequential game described in Section \ref{sec:sg-model} and the related optimality characterizations.
+
+Recall from Section \ref{sec:sg-model}, the action space of the runner and pitcher are denoted by $A_{\text{R}}$ and $A_\text{P}$, respectively.
+The sets $\mathbb{F}_\text{R}, \mathbb{K}_\text{P}$, and $\mathbb{F}_\text{P}$ defined below will be used to rigorously define a policy.
+    \begingroup
+    \allowdisplaybreaks
+    \begin{align*}
+        \mathbb{F}_\text{R} &\defeq \left\{f: {\cal S} \to \cup_{s \in {\cal S}}A_{\text{R}}(s) \big| f(s) \in A_{\text{R}}(s)\right\}\\
+        \mathbb{K}_\text{P} &\defeq \left\{(s,a_\text{R})\big| s \in {\cal S}, a_\text{R} \in A_\text{R}(s)\right\}\\
+        \mathbb{F}_\text{P} &\defeq \left\{f: \mathbb{K}_\text{P} \to \cup_{(s,a_\text{R}) \in \mathbb{K}_\text{P}} A_\text{P}(s,a_\text{R}) \big| f(s,a_\text{R}) \in A_\text{P}(s,a_\text{R})\right\}.
+    \end{align*}
+    \endgroup
+
+   \noindent The set $\mathbb{F}_\text{R}$ is the set of all deterministic decision rules for the runner.
+Recall that a deterministic decision rule for the pitcher takes as input a state and a runner action, and maps it to an allowable action for the pitcher.
+The set $\mathbb{K}_\text{P}$ includes the set of all possible state and runner action pairs.
+The set $\mathbb{F}_\text{P}$ then captures the set of all deterministic decision rules for the pitcher.
+A deterministic, Markovian policy for the runner is a sequence $\pi = (\pi_0,\pi_1,\ldots,)$ such that $\pi_t \in \mathbb{F}_\text{R}$ for all integers $t \geq 0$.
+A deterministic, stationary policy for the runner is a sequence $\pi = (\pi_0,\pi_1,\ldots,)$ such that $\pi_t = \bar\pi \in \mathbb{F}_\text{R}$ for all integers $t \geq 0$.
+That is a stationary policy employs the same decision rule in all time-stages.
+Similarly, a deterministic, Markovian policy for the pitcher is a sequence $\tilde\pi = (\tilde\pi_0,\tilde\pi_1,\ldots,)$ such that $\tilde\pi_t \in \mathbb{F}_\text{P}$ for all integers $t \geq 0$.
+Finally, a deterministic, stationary policy for the pitcher is a sequence $\tilde\pi = (\tilde\pi_0,\tilde\pi_1,\ldots,)$ such that $\tilde\pi_t = \hat\pi \in \mathbb{F}_\text{P}$ for all integers $t \geq 0$.
+Consistent with the notation used in Section \ref{sec:sg-model}, we denote the set of deterministic, stationary policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
+Following this style of notation, we denote the set of deterministic, Markovian policies for the pitcher and runner by $\overline\Pi_\text{R}$ and $\overline\Pi_\text{P}$, respectively.
+Clearly, $\Pi_\text{R} \subseteq \overline\Pi_\text{R}$ and $\Pi_\text{P} \subseteq \overline\Pi_\text{P}$.
+
+   The optimal value function is defined as
+   \begingroup
+   \allowdisplaybreaks
+   \begin{equation}
+        \label{eqn:sg_val_extended}
+        V^*(s) \defeq \max_{\pi_\text{R} \in \overline\Pi_\text{R}}\min_{\pi_\text{P} \in \overline\Pi_\text{P}}V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in S.
+   \end{equation}
+   \endgroup
+   The policy space in the right-hand side of \eqref{eqn:sg_val_extended} is over the larger class of deterministic, Markovian policies rather than deterministic, stationary policies as in \eqref{eqn:sg_val}.
+With an abuse of notation, we overload $V^*$ to denote the optimal value function in \eqref{eqn:sg_val} and \eqref{eqn:sg_val_extended}.
+The reason for this choice is that under Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous}, Lemma \ref{lem:stationary_pol_optimal} proves that there exists  deterministic, stationary policies for the runner and pitcher that attain the respective maximum and minimum in \eqref{eqn:sg_val_extended}.
+That is the policy spaces in the right-hand side of \eqref{eqn:sg_val_extended} can be restricted to the smaller class of deterministic, stationary policies as in \eqref{eqn:sg_val} without affecting the optimal value in \eqref{eqn:sg_val_extended}.
+
+   As alluded in Section \ref{sec:sg-model}, we assume that an inning eventually halts. This can be mathematically captured via the following condition.
+   \begin{assumption}\label{assum:game_halts}
+       There exists $m \in \mathbb{N}$ such that 
+       \begingroup
+       \allowdisplaybreaks
+       \begin{align*}
+            \rho &\defeq \sup_{\pi_\text{R} \in \overline\Pi_\text{R}, \pi_\text{P} \in \overline\Pi_\text{P}, s \in {\cal S}} \rho(\pi_\text{R},\pi_\text{P},s) < 1,\\
+            \text{ where }       \rho(\pi_\text{R},\pi_\text{P},s) &\defeq \mathbb{P}^{\pi_\text{R}, \pi_\text{P}}[S_m \neq \Delta|S_0=s] \ \forall \pi_\text{R} \in \overline\Pi_\text{R}, \, \overline\pi_\text{P} \in \Pi_\text{P}, \, s \in {\cal S}.
+       \end{align*}
+       \endgroup
+   \end{assumption}
+   \noindent Assumption \ref{assum:game_halts} states that regardless of the policies and the starting state, there is a positive probability (at least $1-\rho$) that an inning ends after no more than $m \in \mathbb{N}$ plays.
+
+   Under Assumption \ref{assum:game_halts}, we prove that the expected number of runs scored by the batter, $V^{\pi_\text{R},\pi_\text{P}}$, in  \eqref{eqn:sg_pol_val} is well-defined (i.e., the limit exists) and finite.
+Our proof is based on standard arguments used in the analyses of undiscounted problems \parencite[Section 7.2]{bertsekas2012dynamic}.
+We present it here for completeness.
+
+   \begin{lemma}
+       \label{lem:sg_pol_val_well_defined}
+       Suppose Assumption \ref{assum:game_halts} holds.
+Then $V^{\pi_\text{R},\pi_\text{P}}(s)$ in \eqref{eqn:sg_pol_val} is well-defined for all $\pi_\text{R} \in \overline\Pi_\text{R}, \pi_\text{P} \in \overline\Pi_\text{P}$, and $s \in {\cal S}$.
+Also, $V^{\pi_\text{R},\pi_\text{P}}$ is uniformly bounded over $\overline\Pi_\text{R} \times \overline\Pi_\text{P} \times {\cal S}$.
+   \end{lemma}
+   \begin{proof}
+       Fix a $\pi_\text{R} \in \overline\Pi_\text{R}, \pi_\text{P} \in \overline\Pi_\text{P}$, and $s \in {\cal S}$. Then 
+       \begingroup
+       \allowdisplaybreaks
+       \begin{align*}
+           V^{\pi_\text{R},\pi_\text{P}}(s) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right]\\
+           &= \lim_{t \to \infty} \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{j=0}^{t-1} r(S_{j+1}|S_j) \Big| S_0 = s\right].
+       \end{align*}
+       \endgroup
+       The first equality is the  definition of $V^{\pi_\text{R},\pi_\text{P}}(s)$ in \eqref{eqn:sg_pol_val}.
+The second equality follows by the monotone convergence theorem since the rewards are nonnegative.
+Under Assumption \ref{assum:game_halts}, it follows by standard Markov chain arguments that
+       \[
+       \mathbb{P}^{\pi_\text{R}, \pi_\text{P}}[S_{km} \neq \Delta|S_0=s] \leq \rho^k \ \forall k \in \mathbb{N}.
+       \]
+       Hence, the expected total rewards earned between periods $km$ and $(k+1)m - 1$ is upper bounded by $m\rho^k \max_{s,s' \in {\cal S}}r(s'|s)$. Therefore
+       \[
+       V^{\pi_\text{R},\pi_\text{P}}(s) \leq \sum_{k=0}^\infty m\rho^k \max_{s,s' \in {\cal S}}r(s'|s) = \frac{m \max_{s,s' \in {\cal S}}r(s'|s)}{1-\rho},
+       \]
+       thereby establishing the existence of $V^{\pi_\text{R},\pi_\text{P}}(s)$.
+The uniform bound follows since the above upper bound on $V^{\pi_\text{R},\pi_\text{P}}(s)$ is independent of $\pi_\text{R},\pi_\text{P}$, and $s$.
+   \end{proof}
+
+   % The function $\widetilde V^*$ will be referred to as the value of the game.
+A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $\widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = \widetilde V^*(s) \ \forall s \in {\cal S}$ is called the optimal strategy or equilibrium strategy.
+Note that $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an optimal strategy if and only if $\widetilde V^{\pi_\text{R},\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}}(s) \ \forall \pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}$.
+
+    We present optimality characterizations in Lemmas \ref{lem:vi_convergence} to \ref{lem:policy_iteration}.
+These lemmas and their proofs are similar to the undiscounted problem in the single-player case \parencite[Section 7.2]{bertsekas2012dynamic} and the zero-sum stochastic game with simultaneous moves \parencite{patek_stochastic_1999}.
+We state and prove our lemmas for completeness and to outline the minor differences.
+  \begin{definition}\label{defn:sg_operators}
+    \normalfont
+    Consider the sequential game defined in Section \ref{sec:sg-model}. Fix a $u \in \mathbb{R}^{\abs{\cal S}}$ with $u(\Delta) = 0$.
+    \begin{enumerate}
+        \item For any $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, the evaluation operator $T_{\pi_\text{R},\pi_{\text{P}}}$ on $\mathbb{R}^{\abs{\cal S}}$ is defined as
+        \begingroup
+        \allowdisplaybreaks
+        \begin{align}
+        \label{eqn:eval_operator}
+        (T_{\pi_\text{R},\pi_{\text{P}}} u)(s) &\defeq \mathbb{E}_{s' \sim p(\cdot|s,\pi_\text{R}(s),\pi_\text{P}(s,\pi_\text{R}(s)))}\left[r(s'|s) + u(s')\right] \ \forall s \in {\cal S}.
+        \end{align}
+        \endgroup
+        \item The optimality operator $T$ on $\mathbb{R}^{\abs{\cal S}}$ is defined as
+        \begingroup
+        \allowdisplaybreaks
+        \begin{align}
+        \label{eqn:bellman_operator}
+        (T u)(s) &\defeq \max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + u(s')\right] \ \forall s \in {\cal S}.
+        \end{align}
+        \endgroup
+    \end{enumerate}       
+   \end{definition}
+
+      \begin{lemma}
+       \label{lem:vi_convergence}
+       Consider the sequential game defined in Section \ref{sec:sg-model}. Fix a $u_0 \in \mathbb{R}^{\abs{\cal S}}$ with $u_0(\Delta) = 0$. Suppose Assumption \ref{assum:game_halts} holds. Then
+       \begin{enumerate}[(a)]
+           \item $\lim_{k \to \infty}u_k(s) = V^*(s)$ for all $s \in {\cal S}$, where $u_{k} \defeq Tu_{k-1} \ \forall k \in \mathbb{N}$.
+           \item For any $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, we have $\lim_{k \to \infty}\tilde u_k(s) = V^{\pi_\text{R},\pi_\text{P}}(s)$ for all $s \in {\cal S}$, where $\tilde u_k \defeq T_{\pi_\text{R},\pi_\text{P}}\tilde u_{k-1} \ \forall k \in \mathbb{N}$ and $\tilde u_0 \defeq u_0$.
+       \end{enumerate}
+   \end{lemma}
+   \begin{proof}
+       Consider part $(a)$.
+Let $\pi_\text{R} = (\mu_0,\mu_1,\ldots,)$ and $\pi_\text{P} = (\nu_0,\nu_1,\ldots)$ be any  deterministic, Markovian (may not be stationary) policies for the runner and pitcher, respectively.
+Let $m \in \mathbb{N}$ be as defined in Assumption \ref{assum:game_halts}.
+Fix $s_0 \in {\cal S}$.
+For any $K \in \mathbb{N}$, we have
+       \begingroup
+       \allowdisplaybreaks
+       \begin{align}
+       \nonumber
+           V^{\pi_\text{R},\pi_\text{P}}(s_0) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
+           \nonumber
+           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + \sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
+           \nonumber
+           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t)\Big| S_0 = s_0\right] + \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
+           \nonumber
+           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) - u_0(S_{mK-1})\Big| S_0 = s_0\right] +\\
+           \nonumber
+           &\qquad \qquad \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
+           \nonumber
+           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \Big| S_0 = s_0\right] -\\
+           \label{eqn:vi_convergence_temp1}
+           &\qquad \qquad \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[ u_0(S_{mK-1})\Big| S_0 = s_0\right] + \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right].
+       \end{align}
+       \endgroup
+       Using similar arguments as in the proof of Lemma \ref{lem:sg_pol_val_well_defined}, it follows that $\mathbb{E}^{\pi_\text{R},\pi_\text{P}}\big[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\big] \leq \sum_{k=K}^\infty m \rho^k \max_{s,s' \in {\cal S}}r(s'|s) = m\rho^K\max_{s,s' \in {\cal S}}r(s'|s)/(1-\rho)$.
+Next, $\abs{\mathbb{E}^{\pi_\text{R},\pi_\text{P}}\big[ u_0(S_{mK-1})\Big| S_0 = s_0\big]} = \abs{\sum_{s' \in {\cal S}}\mathbb{P}^{\pi_\text{R},\pi_\text{P}}[S_{mK-1}=s']u_0(s')} \leq \sum_{s' \in {\cal S}-\Delta}\mathbb{P}^{\pi_\text{R},\pi_\text{P}}[S_{mK-1}=s'] \left(\max_{s' \in {\cal S}-\Delta}\abs{u_0(s')}\right) + \mathbb{P}^{\pi_\text{R},\pi_\text{P}}[S_{mK-1}=\Delta]u_0(\Delta) \leq \rho^K \max_{s' \in {\cal S}-\Delta}\abs{u_0(s')}$, where the last inequality is due to Assumption \ref{assum:game_halts}.
+Utilizing these two bounds in \eqref{eqn:vi_convergence_temp1} gives us
+       \begin{align*}
+       &\mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \Big| S_0 = s_0\right] -\rho^K \max_{s' \in {\cal S}-\Delta}\abs{u_0(s')} \leq  V^{\pi_\text{R},\pi_\text{P}}(s_0) \leq\\
+       &\qquad \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \Big| S_0 = s_0\right] +\rho^K \max_{s' \in {\cal S}-\Delta}\abs{u_0(s')} + \frac{m\rho^K\max_{s,s' \in {\cal S}}r(s'|s)}{(1-\rho)}.
+       \end{align*}
+       From the definition of the operator $T$ in \eqref{eqn:bellman_operator}, it follows that $\max_{\pi_{\text{R}}}\min_{\pi_\text{P}}\mathbb{E}^{\pi_\text{R},\pi_\text{P}}\big[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \big| S_0 = s_0\big] = u_k$.
+Hence, taking the minimum over all pitcher policies followed by taking the maximum over all runner policies in the above chain and then passing to limit $K \to \infty$ gives us $\lim_{K \to \infty}u_{Km}(s_0) = V^*(s_0)$.
+From Assumption \ref{assum:game_halts}, it follows that $\abs{u_{Km+t} - u_{Km}} \leq m \rho^K \max_{s,s' \in {\cal S}}r(s'|s)$ for all $t=1,\ldots,m$.
+This along with $\lim_{K \to \infty}u_{Km}(s_0) = V^*(s_0)$ establishes our claim that $\lim_{k \to \infty}u_k(s_0) = V^*(s_0)$.
+
+       The proof of part $(b)$ follows via similar arguments as that of part $(a)$.
+   \end{proof}
+
+   \begin{assumption}
+       \label{assum:transition_pby_continuous}
+       The transition probability $p(\cdot|s,a_\text{R},a_\text{P})$ satisfies the following conditions.
+Fix $s \in {\cal S}$.
+Consider sequences $\{a_\text{R}^k\} \in A_\text{R}(s)$ and $\{a_\text{P}^k\} \in A_\text{P}(s,a_\text{R}^k)$ such that $\lim_{k \to \infty}a_\text{R}^k = a_\text{R} \in A_\text{R}(s)$ and  $\lim_{k \to \infty}a_\text{P}^k = a_\text{P} \in A_\text{P}(s,a_\text{R})$.
+Then $\lim_{k \to \infty}p(s'|s,a_\text{R}^k,a_\text{P}^k) = p(s'|s,a_\text{R},a_\text{P})$ for all $s' \in {\cal S}$.
+   \end{assumption}
+   \noindent {\color{blue}{Assumption \ref{assum:transition_pby_continuous} is a continuity condition imposed on the transition probabilities when viewed as a function of the runner and pitcher actions.
+Lemmas \ref{lem:bellman_equations} to \ref{lem:policy_iteration} rely upon this assumption to prove their stated claims.
+Our construction of transition probabilities in Section \ref{sec:transition-probability-model} satisfy this assumption.
+Note that this assumption trivially holds if the action space of the runner $A_\text{R}(s)$ is finite for all $s \in {\cal S}$.}}
+
+   \begin{lemma}
+       \label{lem:bellman_equations}
+       Consider the sequential game defined in Section \ref{sec:sg-model}. Suppose Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous} hold. Then
+       \begin{enumerate}[(a)]
+           \item The optimal value function $V^*$ is the unique solution to $u = Tu$, i.e., 
+       \begingroup
+       \allowdisplaybreaks
+       \begin{equation}
+           \label{eqn:bellman_equation}
+           V^*(s) = \max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + V^*(s')\right] \ \forall s \in {\cal S}.
+       \end{equation}
+       \endgroup
+       \item For any $\pi_\text{R} \in \Pi_\text{R}$ and $\pi_\text{P} \in \Pi_\text{P}$, the function $V^{\pi_\text{R},\pi_\text{P}}$ is the unique solution to $u = T_{\pi_\text{R},\pi_\text{P}}u$, i.e.,
+       \begingroup
+       \allowdisplaybreaks
+       \begin{equation}
+       \label{eqn:bellman_eval}
+           V^{\pi_\text{R},\pi_\text{P}}(s) =   \mathbb{E}_{s' \sim p(\cdot|s,\pi_\text{R}(s),\pi_\text{P}(s,\pi_\text{R}(s)))}\left[r(s'|s) + V^{\pi_\text{R},\pi_\text{P}}(s')\right] \ \forall s \in {\cal S}.
+       \end{equation}
+       \endgroup
+       \end{enumerate}
+   \end{lemma}
+   \begin{proof}
+       We first establish part $(a)$.
+Consider arbitrary $u_0 \in \mathbb{R}^{\abs{\cal S}}$ with $u_0(\Delta) = 0$ and define $u_{k} = Tu_{k-1} \ \forall k \in \mathbb{N}$.
+By Lemma \ref{lem:vi_convergence}, $\lim_{k \to \infty}u_k(s) = V^*(s)$ for all $s \in {\cal S}$.
+Passing to limit $k \to \infty$ in $u_k = Tu_{k-1}$,  we have for all $s \in {\cal S}$
+       \begingroup
+       \allowdisplaybreaks
+       \begin{align}
+       \nonumber
+       V^*(s) &= \lim_{k\to\infty}\max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + u_{k-1}(s')\right]\\
+       \label{eqn:bellman_equation_temp1}
+       &= \lim_{k \to \infty}\max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\left[r(s'|s) + u_{k-1}(s')\right]p(s'|s,a_\text{R},a_\text{P}).
+       \end{align}
+       \endgroup
+       We prove that $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ converges to $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[\\r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},a_\text{P})$ uniformly over $A_\text{R}(s)$ for each $s \in {\cal S}$.
+
+       Fix a $s \in {\cal S}$.
+If $s = (b,c,o,d)$ is such that $b \neq (1,0,0)$, then from \eqref{eqn:runner_action_sapce} and \eqref{eqn:pitcher_action_sapce}  $A_\text{R}(s) = \{\delta\}$ and $ A_\text{P}(a_\text{R},s) = \{\delta\}$.
+As a result, $p(\cdot|s,a_\text{R},a_\text{P}) = p(\cdot|s)$, and hence the convergence is independent of $a_\text{R}$ and $a_\text{P}$.
+Suppose the component $b$ equals $(1,0,0)$.
+Let $\{a_\text{R}^k\} \in A_\text{R}(s)$ be such that $\lim_{k \to \infty}a_\text{R}^k = a_\text{R} \in A_\text{R}(s)$.
+Then from  \eqref{eqn:pitcher_action_sapce}, $A_\text{P}(a_\text{R}^k,s) = \{0,1\}$ for all $k \in \mathbb{N}$.
+Therefore
+       \begingroup
+       \allowdisplaybreaks
+       \begin{align*}
+       &\lim_{k \to \infty}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R}^k)}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R}^k,a_\text{P}) \\
+       &= \lim_{k \to \infty}\min\Bigg(\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R}^k,0),\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R}^k,1)\Bigg)\\
+       &= \min\Bigg(\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},0),\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},1)\Bigg)\\
+       &= \min_{a_\text{P} \in A_\text{P}(s,a_\text{R}^k)}\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},a_\text{P}),
+       \end{align*}
+       \endgroup
+       where the second equality follows from Assumption \ref{assum:transition_pby_continuous} and the convergence of $u_{k-1}$ to $V^*$ established in Lemma \ref{lem:vi_convergence}.
+We have established the continuous convergence of the sequence of functions $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ over $A_\text{R}(s)$.
+The set $A_\text{R}(s)$ in \eqref{eqn:runner_action_sapce}   is compact.
+Hence, by Theorem 5 in Subsection X, Section 21 in  \parencite{kuratowski2014topology}, $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ converges to $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},a_\text{P})$ uniformly over $A_\text{R}(s)$.
+
+       Utilizing this uniform convergence in \eqref{eqn:bellman_equation_temp1} gives us 
+       \begin{align*}
+       &\lim_{k \to \infty}\max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\left[r(s'|s) + u_{k-1}(s')\right]p(s'|s,a_\text{R},a_\text{P})\\
+       &= \max_{a_\text{R} \in A_\text{R}(s)}\lim_{k \to \infty}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\left[r(s'|s) + u_{k-1}(s')\right]p(s'|s,a_\text{R},a_\text{P})\\
+       &= \max_{a_\text{R} \in A_\text{R}(s)}\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\mathbb{E}_{s' \sim p(\cdot|s,a_\text{R},a_\text{P})}\left[r(s'|s) + V^*(s')\right],
+       \end{align*}
+       thereby establishing \eqref{eqn:bellman_equation}. The uniqueness of $V^*$ follows from the value iteration convergence result established in Lemma \ref{lem:vi_convergence}.
+
+       Regarding part $(b)$, for each $s \in {\cal S}$, set $A_\text{R}(s) = \{\pi_\text{R}(s)\}$ and $A_\text{P}(s,\pi_\text{R}(s)) = \{\pi_\text{P}(s,\pi_\text{R}(s))\}$.
+The claim then  follows from part $(a)$.
+   \end{proof}
+
+   \begin{lemma}
+       \label{lem:stationary_pol_optimal}
+       Consider the sequential game defined in Section \ref{sec:sg-model}.
+Suppose Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous} hold.
+Then a stationary pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an equilibrium policy if and only if $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
+   \end{lemma}
+   \begin{proof}
+       Suppose $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
+Then \eqref{eqn:bellman_equation} becomes
+       \[
+       V^*(s) = \mathbb{E}_{s' \sim p(\cdot|s,\pi_\text{R}^*(s),\pi_\text{P}^*(s,\pi_\text{R}^*(s)))}\left[r(s'|s) + V^*(s')\right] = (T_{\pi_\text{R}^*,\pi_\text{P}^*}V^*)(s)  \ \forall s \in {\cal S},
+       \]
+       where the second equality follows by the definition of $T_{\pi_\text{R}^*,\pi_\text{P}^*}$ in \eqref{eqn:eval_operator}.
+By Lemma \ref{lem:bellman_equations}(b), we then get $V^*(s)=V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \ \forall s \in {\cal S}$.
+Hence, $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an equilibrium policy.
+Conversely, suppose $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an equilibrium policy.
+Then $V^*(s) = V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \ \forall s \in {\cal S}$.
+From parts $(a)$ and $(b)$ of Lemma  \ref{lem:bellman_equations}, it follows that $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
+   \end{proof}
+
+   \begin{lemma}
+       \label{lem:policy_iteration}
+       Consider the sequential game defined in Section \ref{sec:sg-model} and suppose Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous} hold.
+Starting from any $\pi_\text{R}^0 \in \Pi_\text{R}$, define the following updates for all $k \geq 0$.
+       \begin{align*}
+          % \label{eqn:update-value-two-player}
+          V^{\pi_{\text{R}}^k}(s) &\defeq \min_{\pi_\text{P} \in \Pi_\text{P}} V^{\pi_\text{R}^k,\pi_\text{P}}(s) && \forall s \in \mathcal{S}, \mbox{ and}\\
+          \label{eqn:update-policy-two-player}
+          \pi_\text{R}^{k+1}(s) &\in \argmax_{a_\text{R} \in A_\text{R}(s)} \left\{\min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s'|s) + V^{\pi_{\text{R}}^k}(s')] p(s'|s,a_\text{R},a_\text{P})\right\} && \forall s \in \mathcal{S}.
+        \end{align*}
+        Then $V^{\pi_\text{R}^k} \uparrow V^*$. 
+   \end{lemma}
+   \begin{proof}
+       For any $\pi_\text{R} \in \Pi_\text{R}$, define the operator $T_{\pi_\text{R}}$ on $\mathbb{R}^{\abs{\cal S}}$ as 
+       \begingroup
+       \allowdisplaybreaks
+       \begin{equation}
+           \label{eqn:policy_iteration_temp1}
+           (T_{\pi_\text{R}} u )(s) \defeq \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}(s))}\sum_{s' \in S}\left[r(s'|s) + u(s')\right]p(s'|s,\pi_\text{R}(s),a_\text{P}) \ \forall s \in {\cal S}.
+       \end{equation}
+       Since $\pi_\text{R}$ is stationary, the best response of the pitcher corresponding to $\pi_\text{R}$ reduces to solving a MDP.
+Denote the value of this MDP as $V^{\pi_\text{R}}$, i.e., $V^{\pi_\text{R}}(s) \defeq \min_{\pi_\text{P} \in \Pi_\text{P}} V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in {\cal S}$.
+Using similar arguments as in the proof of Lemma \ref{lem:vi_convergence}, it follows that $\lim_{k \to \infty}T_{\pi_\text{R}}^k u = V^{\pi_\text{R}}$, where $T_{\pi_\text{R}}^k$ is the composition of $T_{\pi_\text{R}}$ taken with itself $k$ times.
+Also, using arguments similar to the proof of Lemma \ref{lem:bellman_equations}, it follows that $V^{\pi_\text{R}}$ is the unique fixed point of $T_{\pi_\text{R}}$.
+
+       Note that the policy sequence $\pi_\text{R}^k$ generated by our algorithm satisfies  
+       \begingroup
+       \allowdisplaybreaks
+       \begin{equation}
+       \label{eqn:pi_iteration_temp2}
+       T_{\pi_\text{R}^{k+1}}V^{\pi_\text{R}^k} = TV^{\pi_\text{R}^k} \ \forall k \in \mathbb{N} \cup \{0\}.
+       \end{equation}
+       \endgroup
+       Next, we observe that the sequence of functions $V^{\pi_\text{R}^k}$ correspond to the optimal value function of the pitcher's best response MDP.
+Hence, $V^{\pi_\text{R}^k} = T_{\pi_\text{R}^k} V^{\pi_\text{R}^k}$ for all integers $k \geq 0$.
+Since $Tu \geq T_{\pi_\text{R}}u$ for all $\pi_\text{R} \in \Pi_\text{R}$ and $u \in \mathbb{R}^{\abs{\cal S}}$, it follows that $V^{\pi_\text{R}^k} = T_{\pi_\text{R}^k} V^{\pi_\text{R}^k} \leq T V^{\pi_\text{R}^k}$ for all integers $k \geq 0$.
+Utilizing this in \eqref{eqn:pi_iteration_temp2} gives us
+       \begingroup
+       \allowdisplaybreaks
+       \[
+           T_{\pi_\text{R}^{k+1}}V^{\pi_\text{R}^k} = TV^{\pi_\text{R}^k} \geq V^{\pi_\text{R}^k} \ \forall k \in \mathbb{N} \cup \{0\}.
+       \]
+       \endgroup
+       Repeatedly applying $T_{\pi_\text{R}^{k+1}}$ in the above chain of expression, we get for any $j \in \mathbb{N}$ that $T_{\pi_\text{R}^{k+1}}^j V^{\pi_\text{R}^k} \geq TV^{\pi_\text{R}^k} \geq V^{\pi_\text{R}^k} \ \forall k \in \mathbb{N} \cup \{0\}$.
+Passing to limit $j \to \infty$ and noting that $\lim_{j \to \infty}T_{\pi_\text{R}^{k+1}}^j V^{\pi_\text{R}^k} = V^{\pi_\text{R}^{k+1}}$ gives us
+       \begingroup
+       \allowdisplaybreaks
+       \begin{equation}
+       \label{eqn:policy_iteration_temp2}
+           V^{\pi_\text{R}^{k+1}} \geq TV^{\pi_\text{R}^k} \geq V^{\pi_\text{R}^k} \ \forall k \in \mathbb{N} \cup \{0\}.
+       \end{equation}
+       \endgroup
+       By Lemma \ref{lem:sg_pol_val_well_defined}, it follows that $V^{\pi_\text{R}}$ is uniformly bounded over $\Pi_\text{R}$ and ${\cal S}$.
+Hence, $V^{\pi_\text{R}^k} \uparrow \bar V \in \mathbb{R}^{\abs{\cal S}}$.
+We prove that $\bar V = V^*$.
+Using arguments similar to the ones used in the proof of Lemma \ref{lem:bellman_equations}, it follows that $T$ is a  continuous operator on $\mathbb{R}^{\abs{\cal S}}$, i.e., $T u^k \to T u$ for all $u^k \to u$.
+Hence, passing to limit $k \to \infty$ in \eqref{eqn:policy_iteration_temp2} gives us $T \bar V = \bar V$.
+From Lemma \ref{lem:bellman_equations}(a), it follows that $\bar V = V^*$, thereby completing the proof.
+   \end{proof}
+
+   If the state- and action-spaces are all finite, the the convergence in Lemma \ref{lem:policy_iteration} happens in a finite number of steps.
+In that case, the policy obtained in the last iterate, say, $\pi_\text{R}^*$, is optimal for the runner.
+The pitcher's optimal policy (best response), $\pi_\text{P}^*$, can then be obtained as
+   \begingroup
+   \allowdisplaybreaks
+   \[
+   \pi_\text{P}^*(s,\pi_\text{R}^*(s)) \in \argmin_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}^*(s))}[r(s'|s) + V^*(s')] p(s'|s,\pi_\text{R}^*(s),a_\text{P}) \ \forall s \in {\cal S}, \pi_\text{R}^*(s) \in A_\text{R}(s).
+   \]
+   \endgroup
+   We use this approach in our computational experiments in Section \ref{sec:results}, after an appropriate discretization of the runner's action space.
+
+%    \begin{theorem}[\parencite{patek_stochastic_1999}]\label{thm:sg_optimality_condition}
+%        \normalfont
+%       Suppose Assumption \ref{assum:game_halts} holds.
+Then the value of the game $\widetilde V^*$ is the unique fixed point of the operator $\widetilde \Phi$.
+Furthermore, $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an optimal strategy if and only if $\widetilde\Phi_{\pi_\text{R}^*}\widetilde V^* = \widetilde\Phi \widetilde V^* = \widetilde\Phi_{\pi_\text{P}^*}\widetilde V^*$.
+%    \end{theorem}
+
+%    The above theorem enables the characterization of $\widetilde V^*(s)$ as in the following lemma.
+
+% \begin{corollary}\label{corr:sg_value_characterization}
+%     \normalfont
+%     Suppose Assumption \ref{assum:game_halts} holds. Then $\widetilde V^*$ is the unique solution to the following system of equations.
+%     \begin{align*}
+%         V(s) &= \max_{\ell \in {\cal L}}\min_{a \in {\cal P}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell,a) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
+%         V(s) &= \sum_{s' \in {\cal      S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
+%         V(s) &=  0 \quad \text{ if } s = \Delta.
+%     \end{align*}
+%     \begin{proof}
+%         Immediate from Theorem \ref{thm:sg_optimality_condition}.
+%     \end{proof}
+% \end{corollary}
+
+   % An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{patek_stochastic_1999}.
+We use policy iteration  in this work to compute the optimal policy which is as follows.
+Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi_\text{R}^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi_\text{R}^k$ by applying the updates
+   %      \begin{align}
+   %        \label{eqn:update-value-two-player}
+   %        V^{k+1}(s) &= \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}^k(s))} \sum_{s' \in \mathcal{S}} [r(s, s') +  V^k(s')]p(s'|s,\pi_\text{R}^k(s),a_\text{P}) && \forall s \in \mathcal{S}, \mbox{ and}\\
+   %        \label{eqn:update-policy-two-player}
+   %        \pi_\text{R}^{k+1}(s) &= \argmax_{a_\text{R} \in A_\text{R}(s)} \left\{\min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s, s') + V^{k+1}(s')] p(s'|s,a_\text{R},a_\text{P})\right\} && \forall s \in \mathcal{S}.
+   %      \end{align}
+   %  The \footnote{Andrew - don't like (13) - argmax is a set, but pi is an element of the set.
+Can we use $\in$ instead of =?
+Likely true for other arg maxes} iterates generated by the policy iteration satisfy $V^k \uparrow \widetilde V^*$\parencite{patek_stochastic_1999}.
+
+      % \subsection{Single-player setup}
+   % For a fixed stationary policy of the pitcher, the stochastic game reduces to a (single-player) MDP.
+Lemmas \ref{lem:vi_convergence} to \ref{lem:policy_iteration} then reduce to the corresponding MDP versions.
+We omit stating them for brevity.
+See \cite[Section 7.2]{bertsekas2012dynamic} for details.
+
+%    The MDP is then defined by the tuple ${\cal M} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{r(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}, \\  \{p(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}\rangle$.
+Using this MDP, the progression of an inning can be captured as follows.
+At the beginning of a play, the game is in state $s \in {\cal S}$.
+Upon observing this state, an action $a \in A_\text{R}(s)$ is picked.
+The game then transitions to state $s'$ with probability $p(s'|s,a)$ and the batting team collects a reward $r(s'|s,a)$.
+A play then begins from $s'$ and the process repeats.
+The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
+In other words, we assume the objective is to maximize runs rather than maximizing the probability of winning the game.
+The former objective is an approximation of the latter objective which allowed us to simplify the state space.
+This is a reasonable approximation before the final innings of the game, when it is more important to consider the score.
+
+%       \begin{assumption}\label{assum:mdp_halts}
+%        \normalfont
+%       The MDP ${\cal M}$ satisfies
+%       $
+%       \lim_{t \to \infty}\sup_{\pi \in \Pi, s \in {\cal S}}\mathbb{P}^\pi[S_t = \Delta | S_0 = s] = 1.
+%       $
+%    \end{assumption}   
+%   \noindent  Assumption\footnote{Andrew - can we put the assumption in words, and the formal math in the text below?
+Similar comment for Assumption 2 later}   \ref{assum:mdp_halts} states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
+This is a reasonable assumption because an inning does not continue forever.
+
+%    The value of a policy $\pi$ when starting from state $s \in {\cal S}$ equals  
+%    \begin{equation}
+%        \label{eqn:mdp_pol_val}
+%        V^{\pi}(s) \defeq \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
+%    \end{equation}
+%   That is, the value of the policy is the the expected total runs scored in an inning.
+Under Assumption \ref{assum:mdp_halts}, $V^\pi(s)$ exists for all $\pi \in \Pi$ and $s \in {\cal S}$.
+To see that, since $r(s'|s) \geq 0$ for all $s,s' \in {\cal S}$, we have by the monotone convergence theorem
+%   \[
+%   V^\pi(s) = \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] = \lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall \pi \in \Pi, s \in {\cal S}.
+%   \]
+%    Using Assumption \ref{assum:mdp_halts} and the fact that $\sup_{s,s' \in {\cal S}}\abs{r(s'|s)} < \infty$ (due to finiteness of ${\cal S}$), it is easy to establish that $\lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right]$ exists for all $\pi \in \Pi, s \in {\cal S}$.
+Furthermore, the limit is uniformly bounded over $\Pi \times {\cal S}$.
+The value of the MDP is defined as
+%    \begin{equation}
+%         \label{eqn:mdp_val}
+%        V^*(s) \defeq \max_{\pi \in \Pi_\text{R}}V^\pi(s) \ \forall s \in S.
+%    \end{equation}
+%    The function $V^*$, known as the optimal value function, is the maximum expected runs scored starting from state $s \in {\cal S}$.
+A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called an optimal policy.
+
+%    We present the following definitions that will be needed to characterize the optimal value function and an optimal policy.
+
+%    \begin{definition}\label{defn:mdp_operators}
+%     \normalfont
+%     \begin{enumerate}
+%         \item For any $\pi \in \Pi_\text{R}$, the evaluation operator $\Phi_\pi:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
+%         \[
+%         (\Phi_\pi V)(s) \defeq \mathbb{E}_{s' \sim p(\cdot|s,\pi(s))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+%         \]
+%         \item The optimality operator $\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
+%         \[
+%         (\Phi V)(s) \defeq \max_{a \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a)}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
+%         \]
+%     \end{enumerate}       
+%    \end{definition}
+
+%    \begin{theorem}[\parencite{bertsekas_analysis_1991}]\label{thm:mdp_optimality_condition}
+%        \normalfont
+%       Suppose Assumption \ref{assum:mdp_halts} holds. Then
+%       \begin{enumerate}
+%           \item For any policy $\pi$, its value $V^\pi$ is the unique fixed point of the operator $\Phi_\pi$.
+%           \item The optimal value function $V^*$ is the unique fixed point of the operator $\Phi$. Furthermore, $\pi^*$ is an optimal policy if and only if $\Phi_{\pi^*}V^* = \Phi V^*$.
+%       \end{enumerate}
+%    \end{theorem}
+%  \todo[inline]{The above theorem also requires $p(s'|s,a)$ to be continuous in $a$ for any $s,s' \in S$.
+This is true for fixed-effects model since $x \mapsto e^x/1+e^x$ is continuous.
+I do not know about random-effects model.
+[[SP: $a$ comes into the model as a fixed effect (the random effects are for player identities), so, yes, $p(s'|s,a)$ is continuous in $a$.]]}
+
+% Theorem \ref{thm:mdp_optimality_condition} enables the characterization of $V^*(s)$ as in the following lemma.
+
+% \begin{lemma}\label{lem:mdp_value_function_characterization}
+%     \normalfont
+%     Suppose \footnote{Andrew - Don't love this as a lemma - corollary seems more appropriate} Assumption \ref{assum:mdp_halts} holds.
+Then $V^*$ is the unique solution to the following system of equations.
+%     \begin{align*}
+%         V(s) &= \max_{\ell \in {\cal L}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
+%         V(s) &= \sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
+%         V(s) &= 0 \quad \text{ if } s = \Delta.
+%     \end{align*}
+%     \begin{proof}
+%         Immediate from Theorem \ref{thm:mdp_optimality_condition}.
+%     \end{proof}
+% \end{lemma}
+
+% An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{bertsekas_analysis_1991}.
+We use policy iteration  in this work to compute the optimal policy which is as follows.
+Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi^k$ by applying the updates
+%         \begin{align}
+%           \label{eqn:update-value-single-agent}
+%           V^{k+1}(s) &= \sum_{s' \in \mathcal{S}} p(s' | s,\, \pi^k(s)) [r(s, s') + V^k(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
+%           \label{eqn:update-policy-single-agent}
+%           \pi^{k+1}(s) &= \argmax_{a \in A_\text{R}(s)} \sum_{s' \in \mathcal{S}} p(s' | s,\, a) [r(s, s') + V^{k+1}(s')] && \forall s \in \mathcal{S}.
+%         \end{align}
+%     The iterates generated by the policy iteration satisfy $V^k \uparrow V^*$ and every limit point of $\pi^k$ is an optimal policy \parencite{bertsekas_analysis_1991}.
+
   \printbibliography

### 2025-12-18 scripts/generate_results.R
fix convergence threshold for value iteration

@@ -360,7 +360,7 @@ if (fig_make) {
 # Plot MDP example ----
 
 lead_value <- transition_values |>
-  dplyr::left_join(old_re_table, by = c("New_State" = "State")) |>
+  dplyr::left_join(value_old, by = c("New_State" = "State")) |>
   dplyr::group_by(State, lead1b) |>
   dplyr::summarize(RE = sum(TotalProb * (RunsScored + RE)), n = mean(n), .groups = "drop") |>
   dplyr::filter(State %in% c("100 0 000", "100 1 000", "100 2 000")) |>
@@ -402,7 +402,7 @@ if (fig_make) {
 
 # Write results tables ----
 
-sorted <- old_run_1b |>
+sorted <- policy_old |>
   dplyr::mutate(
     bases = substr(State, 1, 3),
     dis = substr(State, 5,5),
@@ -467,7 +467,7 @@ skill_grid |>
     hline.after = c(0, 3, 6)
   )
 
-sorted_two <- old_run_1b_two |>
+sorted_two <- policy_old_two |>
   dplyr::mutate(
     bases = substr(State, 1, 3),
     dis = substr(State, 5,5),

### 2025-12-18 scripts/estimate_models.R
Fix convergence threshold for value iteration (#6)
* refine names in results tables

* fix convergence threshold for value iteration

@@ -1,37 +1,38 @@
 
 fit_glmer_models <- FALSE
+value_iteration_threshold <- 1e-4   # well-accepted default
 
 # DATA SET UP ----
 
 library(tidyverse)
 library(lme4)
 
-event_2022 <- read_csv("input/data/event/2022.csv")
-event_2023 <- read_csv("input/data/event/2023.csv")
+event_2022 <- data.table::fread("input/data/event/2022.csv")
+event_2023 <- data.table::fread("input/data/event/2023.csv")
 
-lead_2022 <- read_csv("input/data/lead_distance/2022.csv")
-lead_2023 <- read_csv("input/data/lead_distance/2023.csv")
+lead_2022 <- data.table::fread("input/data/lead_distance/2022.csv")
+lead_2023 <- data.table::fread("input/data/lead_distance/2023.csv")
 
-pitch_2022 <- read_csv("input/data/pitch/2022.csv") |>
-  select(play_id, description)
-pitch_2023 <- read_csv("input/data/pitch/2023.csv") |>
-  select(play_id, description)
+pitch_2022 <- data.table::fread("input/data/pitch/2022.csv") |>
+  dplyr::select(play_id, description)
+pitch_2023 <- data.table::fread("input/data/pitch/2023.csv") |>
+  dplyr::select(play_id, description)
 
-play_2022 <- read_csv("input/data/play/2022.csv")
-play_2023 <- read_csv("input/data/play/2023.csv")
+play_2022 <- data.table::fread("input/data/play/2022.csv")
+play_2023 <- data.table::fread("input/data/play/2023.csv")
 
-event_map <- read_csv("input/data/batter_event.csv")
-pitch_map <- read_csv("input/data/batter_pitch.csv")
+event_map <- data.table::fread("input/data/batter_event.csv")
+pitch_map <- data.table::fread("input/data/batter_pitch.csv")
 
-arm_strength_2023 <- read_csv("input/data/arm_strength/2023.csv") |>
-  select(player_id, player_name, arm_strength, sb_attempts)
-arm_strength_2022 <- read_csv("input/data/arm_strength/2022.csv") |>
-  select(player_id, player_name, arm_strength, sb_attempts)
+arm_strength_2023 <- data.table::fread("input/data/arm_strength/2023.csv") |>
+  dplyr::select(player_id, player_name, arm_strength, sb_attempts)
[…]

### 2025-12-18 scripts/generate_results.R

@@ -118,18 +118,18 @@ table_po_attempt |>
   dplyr::mutate(
     variable = dplyr::case_when(
       variable == "(Intercept)" ~ "{\\it Intercept}",
-      variable == "lead1b" ~ "Lead Distance (feet)",
+      variable == "lead1b" ~ "Lead Distance (ft)",
       variable == "pre_balls" ~ "Balls",
       variable == "pre_strikes" ~ "Strikes",
       variable == "pre_outs" ~ "Outs",
       variable == "year2023" ~ "Year (2023 vs. 2022)",
       variable == "pre_disengagements1" ~ "Disengagements (1 vs. 0)",
       variable == "pre_disengagements2" ~ "Disengagements (2 vs. 0)",
-      variable == "sprint_speed" ~ "Sprint Speed (feet per second)",
-      variable == "arm_strength" ~ "Arm Strength (miles per hour)",
+      variable == "sprint_speed" ~ "Runner Sprint Speed (ft/s)",
+      variable == "arm_strength" ~ "Catcher Arm Strength (mi/h)",
       variable == "pitcher_id" ~ "Pitcher",
-      variable == "run1b" ~ "Runner",
       variable == "fielder_2_id" ~ "Catcher",
+      variable == "run1b" ~ "Runner",
     )
   ) |>
   tibble::add_row(variable = "{\\it Fixed Effects}", .after = 1) |>
@@ -360,7 +360,7 @@ if (fig_make) {
 # Plot MDP example ----
 
 lead_value <- transition_values |>
-  dplyr::left_join(old_re_table, by = c("New_State" = "State")) |>
+  dplyr::left_join(value_old, by = c("New_State" = "State")) |>
   dplyr::group_by(State, lead1b) |>
   dplyr::summarize(RE = sum(TotalProb * (RunsScored + RE)), n = mean(n), .groups = "drop") |>
   dplyr::filter(State %in% c("100 0 000", "100 1 000", "100 2 000")) |>
@@ -402,7 +402,7 @@ if (fig_make) {
 
 # Write results tables ----
 
-sorted <- old_run_1b |>
+sorted <- policy_old |>
   dplyr::mutate(
     bases = substr(State, 1, 3),
     dis = substr(State, 5,5),
@@ -418,7 +418,7 @@ sorted |>
   dplyr::select(count, dis, lead1b) |>
   tidyr::pivot_wider(names_from = count, values_from = lead1b, names_prefix = "count") |>
[…]

### 2025-12-18 articles/arxiv/paper.tex
Updates from Overleaf

@@
 Before the 1-1 pitch, Burke pivoted and threw to first base behind Tucker, who retreated safely before a tag.
-Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball among base stealers who had not been caught \parencite{fangraphs_major_2025}.
+Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball (MLB) among base stealers who had not been caught \parencite{fangraphs_major_2025}.
 On the White Sox broadcast, color commentator Steve Stone explained to the audience, ``The only problem is now there's two throws over there, and Tucker's gonna get a much bigger lead.'' On the next pitch, Tucker stole his 11$^{th}$ base of the season, the first Burke had allowed.
@@
 The main difference between our infinite-horizon model and the literature on stochastic Stackelberg games is that we do not discount the rewards.
-The rationale for this modeling choice is that the rewards in our model correspond to the runs scored on a play, and runs scored on later plays count the same as those scored on earlier ones.
+The rationale for this modeling choice is that we choose our optimization objective to be expected runs in the inning, which by definition includes no discount.
 We establish a maximin Bellman equation for our sequential game and propose a variant of value iteration and policy iteration to determine the optimal strategies for runner and the pitcher.}}
@@
 
-      The most striking observation is that the optimal lead distance is significantly longer with two prior disengagements (14.8--16.1 feet) than in other states.
+      The most striking observation is that the optimal lead distance is significantly longer with two prior disengagements (14.8--15.8 feet) than in other states.
 These recommended lead distances are near the edge of observed lead distances in the dataset but not outside of the observed range.
@@
       Under the game-theoretic equilibrium, pitchers would be more aggressive with attempting pickoffs when runners take long leads.
-Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 8.3\% of opportunities with only first base occupied.
-When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.8\% of the time---this would be 100\% in the two-player model.
-When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 5.0\% of the time---this would be 0\% in the two-player model.
-Overall, pitchers attempted pickoffs on 5.8\% of opportunities, less than the recommended 8.3\%.
+Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 11.4\% of opportunities with only first base occupied.
+When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.4\% of the time---this would be 100\% in the two-player model.
+When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 4.9\% of the time---this would be 0\% in the two-player model.
+Overall, pitchers attempted pickoffs on 5.6\% of opportunities, less than the recommended 11.4\%.
 
@@
 In other words, if we model the inning as a Markov reward process using empirical transition probabilities between states, the value of the start-of-inning state is 0.516 runs.
-Under the two-player game, the value of the start-of-inning state is 0.520 runs, which corresponds to approximately 5 more runs over a full season of 162 games of nine innings each.
+Under the two-player game, the value of the start-of-inning state is 0.520 runs, which corresponds to approximately 6 more runs over a full season of 162 games of nine innings each.
 Therefore, offenses stand to benefit more in the transition from observed behavior to the game-theoretic equilibrium.
-Under the one-player MDP, the value of the start-of-inning state is 0.528 runs, which corresponds to approximately 17 more runs over a full season, relative to observed behavior.
+Under the one-player MDP, the value of the start-of-inning state is 0.524 runs, which corresponds to approximately 12 more runs over a full season, relative to observed behavior.
 Naturally, scoring expectation is higher in the one-player MDP because the runner exploits the sub-optimal strategy of the pitcher.
-To add 17 expected runs in free agency would cost roughly \$17 million \parencite{clemens_what_2021}, which speaks to a significant (but not outlandish) impact.
+To add 12 expected runs in free agency would cost roughly \$12 million \parencite{clemens_what_2021}, which speaks to a significant (but not outlandish) impact.
 This value may have increased since the last study we found to have estimated it.
@@
 This framework leads to the most actionable recommendation for runners under the current conditions of pitcher behavior.
-We estimate that the average team could expect to increase their expected offensive output by approximately 17 runs over a full season by implementing the optimal lead distance policy.
+We estimate that the average team could expect to increase their expected offensive output by approximately 12 runs over a full season by implementing the optimal lead distance policy.
 Based on these results, we propose the Two-Foot Rule: Extend your lead by two feet after each pickoff attempt.

### 2025-12-18 output/tables/lead_by_players.tex
Updates from Overleaf

@@
 % latex table generated in R 4.5.0 by xtable 1.8-4 package
-% Sat Aug  2 17:11:21 2025
+% Thu Dec 18 00:04:13 2025
   \hline
-10th & 10th & 10.5 & 11.8 & 14.1 \\ 
-  10th & 50th & 11.2 & 12.3 & 14.3 \\ 
-  10th & 90th & 11.8 & 12.7 & 14.5 \\ 
+10th & 10th & 10.4 & 11.7 & 14.1 \\ 
+  10th & 50th & 11.0 & 12.1 & 14.2 \\ 
+  10th & 90th & 11.9 & 12.8 & 14.6 \\ 
    \hline
-50th & 10th & 9.4 & 10.9 & 13.7 \\ 
-  50th & 50th & 9.9 & 11.4 & 13.8 \\ 
-  50th & 90th & 10.6 & 11.9 & 14.0 \\ 
+50th & 10th & 9.4 & 11.1 & 13.8 \\ 
+  50th & 50th & 10.0 & 11.4 & 13.9 \\ 
+  50th & 90th & 11.0 & 12.2 & 14.2 \\ 
    \hline
-90th & 10th & 8.1 & 9.8 & 12.5 \\ 
-  90th & 50th & 8.6 & 10.1 & 12.6 \\ 
-  90th & 90th & 9.2 & 10.5 & 12.7 \\ 
+90th & 10th & 8.2 & 9.9 & 12.6 \\ 
+  90th & 50th & 8.7 & 10.2 & 12.6 \\ 
+  90th & 90th & 9.7 & 10.9 & 12.9 \\ 
 \end{tabular}

### 2025-12-23 articles/arxiv/paper.tex
final(?) read throughs

@@
 
-\usepackage{amsmath, amssymb, amsthm, amsfonts, color, enumerate, framed, graphicx, mathrsfs, mathtools, mdframed, subcaption, tabularx, theoremref, titlesec, verbatim, xfrac, soul, mathtools, wrapfig, tikz, float}
+\usepackage{amsmath, amssymb, amsthm, amsfonts, authblk, color, enumerate, framed, graphicx, mathrsfs, mathtools, mdframed, subcaption, tabularx, theoremref, titlesec, verbatim, xfrac, soul, mathtools, wrapfig, tikz, float}
 
@@
 \newcommand{\abs}[1]{\left\lvert#1\right\rvert}
+\newcommand{\revision}[1]{{\color{orange} #1}}
 
@@
 }
+\author[1]{Scott Powers}
+\author[2]{Sivaramakrishnan Ramani}
+\author[3]{Jacob Hahn}
+\author[2]{Andrew J. Schaefer}
+\affil[1]{Department of Sport Management, Rice University}
+\affil[2]{Dept of Computational Applied Mathematics and Operations Research, Rice University}
+\affil[3]{Department of Statistics, Rice University}
 
@@
   \begin{abstract}
-    Recently, Major League Baseball (MLB)  limited  pitcher to three pickoff attempts, creating a cat-and-mouse game between pitcher and runner.
+    Recently, Major League Baseball (MLB) limited pitchers to three pickoff attempts, creating a cat-and-mouse game between pitcher and runner.
 Each failed attempt adds pressure on the pitcher to avoid using another, and the runner can intensify this pressure by extending their leadoff toward the next base.
@@
 Using lead distance data, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
-We compute the game-theoretic equilibria under the two-player model, as well as the optimal runner policy under a simplified one-player Markov decision process model.
+We compute the game-theoretic equilibria under the two-player model, as well as the optimal runner policy under a simplified one-player Markov decision process (MDP) model.
 In the one-player setting, our results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
@@
 Before the 1-1 pitch, Burke pivoted and threw to first base behind Tucker, who retreated safely before a tag.
-Tucker had 10 stolen bases in 10 attempts on the young season, leading Major League Baseball (MLB) among base stealers who had not been caught \parencite{fangraphs_major_2025}.
+Tucker had 10 stolen bases in 10 attempts on the young season, the most of any base stealer who had not been caught \parencite{fangraphs_major_2025}.
 On the White Sox broadcast, color commentator Steve Stone explained to the audience, ``The only problem is now there's two throws over there, and Tucker's gonna get a much bigger lead.'' On the next pitch, Tucker stole his 11$^{th}$ base of the season, the first Burke had allowed.
@@
 
-    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the ball to the base for a fielder to tag the runner out \parencite{major_league_baseball_baseball_2025}.
-Whether attempting a stolen base or not, runners typically take a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, in the direction of the next base.
+    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the to a fielder at the base to tag the runner out \parencite{mlb_baseball_2025}.
+Whether attempting a stolen base or not, the runner takes a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, toward the next base.
 The longer the lead, the shorter the distance needed to run to reach the next base.
@@
     MLB stolen base totals surged in the 1980s and 1990s but then waned in the 2000s and 2010s as teams cut back on the play, which generally requires a high success rate in order to have a positive impact on run scoring \parencite{tango_book_2007}.
-Before 2023, pickoff attempts could cause indefinite delays in the progress of a game because unsuccessful pickoff attempts do not advance the state of the game.
-Acknowledging this and other trends, MLB introduced several rule changes starting in 2023, with the goal of making the game more exciting \parencite{castrovince_pitch_2023}.
+Before 2023, pickoff attempts could delay games because unsuccessful pickoff attempts do not advance the state of the game.
+Acknowledging this and other trends, MLB introduced several rule changes in 2023, with the goal of making the game more exciting \parencite{castrovince_pitch_2023}.
 With these changes, a pitch clock limits the time between pitches, and there are restrictions on infielder positioning, curtailing defensive shifts.
@@
 
-    {\color{blue}{Section \ref{sec:sg-model} uses a two-payer zero-sum sequential game to model the evolving interaction between the runner and the pitcher.
-Our model captures all the vital aspects of the play while retaining analytical and computational tractability.
+    Section \ref{sec:sg-model} uses a two-payer zero-sum sequential game to model the evolving interaction between the runner and the pitcher.
+Our model captures the vital aspects of the play while retaining analytical and computational tractability.
 Our model is closely related to the class of stochastic/dynamic Stackelberg games \parencite{goktas2022zero,li2017review,lopez2022stationary,vasal2020stochastic,vorobeychik2012computing}, which combines the elements of Stackelberg and stochastic games.
-The main difference between our infinite-horizon model and the literature on stochastic Stackelberg games is that we do not discount the rewards.
-The rationale for this modeling choice is that we choose our optimization objective to be expected runs in the inning, which by definition includes no discount.
-We establish a maximin Bellman equation for our sequential game and propose a variant of value iteration and policy iteration to determine the optimal strategies for runner and the pitcher.}}
+The main difference between our infinite-horizon model and the stochastic Stackelberg games literature is that we do not discount the rewards.
+The rationale for this modeling choice is that our optimization objective---expected runs in the inning---by definition includes no discount.
+We establish a maximin Bellman equation for our sequential game and propose a variant of value iteration and policy iteration to determine the optimal strategies for runner and pitcher.
 
     In addition to developing a model which captures the important features of the pitcher-runner game, we analyze the model using recently available real-world MLB data.
-The most innovative component of our implementation is the construction of transition probabilities.
+The most innovative component of our parameterization is the construction of transition probabilities.
 The outcome of each play is influenced by factors outside of our state variable, such as the sprint speed of the runner and the arm strength of the catcher.
@@
 In the first step, we use a mixed-effects logistic regression model that calculates the probability of various runner outcomes taking into account the actions of the players and the player-specific information.
-Next, we construct empirical estimates of the next state conditioned upon the current state, runner outcomes, and the actions of the player.
-Finally, we determine the transition probability via these two probabilities.
+Next, we construct an empirical probability distribution over the next state conditioned upon the current state, the runner outcomes, and the actions of the players.
+Finally, we determine the transition probability as the product of these two probabilities.
 
     Section \ref{sec:results} presents a numerical study on recent real-world data which illustrates that, relative to the optimal policy, observed pitcher and runner decisions are conservative.
-Pitchers attempt fewer pickoffs than optimal, and runners take shorter leads than optimal.
+In current practice, pitchers attempt fewer pickoffs than optimal, and runners take shorter leads than optimal.
 Relative to observed behavior, optimal play would favor the offense slightly.
-Moreover, when restricting out model to a single-player setup which allows agency only for the runner, the results reveal an actionable recommendation: The Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
+Moreover, when restricting our model to a single-player setup which allows agency only for the runner, the results reveal an actionable recommendation: the Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
 
@@
     Several studies in the sports economics literature have examined the extent to which professional athletes play minimax strategy.
-\textcite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save.
+\textcite{palacios-huerta_professionals_2003} found evidence consistent with minimax equilibrium behavior in the mixed strategy game of the soccer penalty kick, in which the kicker chooses to shoot the ball left or right and the goalkeeper simultaneously chooses to dive left or right for the save.
 \textcite{palacios-huerta_experientia_2008} followed this with a laboratory experiment featuring professional soccer players and college students simulating the penalty kick decision using cards, showing that the professionals played minimax but the students did not.
@@
 The Markov process is particularly well suited for baseball, where the game's highly structured progression lends it well to an intuitive state space model \parencite{sokol_intuitive_2004}.
-\textcite{hirotsu_markov_2003} used a Markov model to inform substitution (pinch hitting) decision-making in baseball.
+\textcite{hirotsu_markov_2003} used a Markov model to inform substitution decision-making (specifically, pinch hitting) in baseball.
 The Markov process is also popular across sport analytics.
@@
 Recently, \textcite{chan_points_2021} used a Markov model to evaluate team performance in football.
-More specifically, the MDP has become particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \parencite{nadimpalli_when_2013}, baseball \parencite{hirotsu_using_2019}, basketball \parencite{sandholtz_markov_2020}, darts \parencite{baird_optimising_2020} and football \parencite{biro_reinforcement_2022}.
-
-    An intermediate result of the present work is a probability model for stolen base attempts and successes (as well as an analysis how those probabilities changed with the 2023 MLB rule changes), a topic which has not received much attention.
+More specifically, the MDP has become a particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \parencite{nadimpalli_when_2013}, baseball \parencite{hirotsu_using_2019}\todo{pull this out with its own sentence at the end}, basketball \parencite{sandholtz_markov_2020}, darts \parencite{baird_optimising_2020} and football \parencite{biro_reinforcement_2022}.
+
+    An intermediate result of the present work is a probability model for stolen base attempts and successes (as well as an analysis of how those probabilities changed with the 2023 MLB rule changes), a topic which has not received much attention.
 \textcite{loughin_assessing_2008} estimated linear mixed-effects models for stolen base attempts and successes, using random effects for pitcher and catcher (not the runner, notably) and fixed effects for balls, strikes, outs, score, venue and pitcher hand.
@@
     \label{sec:sg-model}
-    In this section, we present our two-player zero-sum sequential game model under the pickoff limit rule.
-Though our sequential game model describes only a portion of an inning, it captures all the vital aspects that enable the generation of valuable and actionable insights for the runner and pitcher.
+    In this section, we present our two-player zero-sum sequential game that models the dynamic between runner and pitcher under the pickoff limit rule.
+Though our sequential game describes only a portion of an inning, it captures the vital aspects that enable the generation of valuable and actionable insights for the runner and the pitcher.
 Before describing the specifics of our model, we note that our model assumes a representative pitcher and a representative runner for the whole inning, rather than capturing full lineups and substitutions of players with varying skills.
-Without this assumption, we will end up with a nonstationary (infinite-horizon) two-player sequential game model, which appears to be significantly more challenging to analyze when compared with the stationary version.
+Without this assumption, we would end up with a nonstationary (infinite-horizon) two-player sequential game, which appears to be significantly more challenging to analyze when compared with the stationary version.
 The four components of our sequential game are as follows.
@@
     \begin{align*}
-      b &\in \{0, 1\}^3 \mbox{ represents which bases are occupied by runners};\\
-      c &\in \{0, 1, 2, 3\} \times \{0, 1, 2\} \mbox{ represents the ball-strike count};\\
-      d &\in \{0, 1, 2\} \mbox{ represents the number of disengagements already used}; and\\
-      o &\in \{0, 1, 2\} \mbox{ represents the number of outs}.
+      b &\in \{0, 1\}^3 \text{ represents which bases are occupied by runners};\\
+      c &\in \{0, 1, 2, 3\} \times \{0, 1, 2\} \text{ represents the ball-strike count};\\
+      d &\in \{0, 1, 2\} \text{ represents the number of disengagements already used; and}\\
+      o &\in \{0, 1, 2\} \text{ represents the number of outs}.
     \end{align*}
@@
 The purpose of the penultimate states is to keep track of the number of runs that are scored on the transition to the terminal state.
-A penultimate state $s \in \mathcal{P} = \{0, 1, 2, 3\}$ represents the number of runs scored on the final play of the inning.
-We use $\mathcal{S} \equiv \mathcal{X} \cup \mathcal{P} \cup \{\Delta\}$ to denote the set of all states, and $S_t$ to denote the state at play $t$ in a (fixed) inning.
+A penultimate state $s \in \mathcal{P} \equiv \{0, 1, 2, 3\}$ represents the number of runs scored on the final play of the inning.
+We use $\mathcal{S} \equiv \mathcal{X} \cup \mathcal{P} \cup \{\Delta\}$ to denote the set of all states and $S_t$ to denote the state at play $t$ in a (fixed) inning.
 Traditionally, the state is defined by the bases occupied and the number of outs (24 non-terminal states) or by those and the ball-strike count (288 non-terminal states) \parencite{bukiet_markov_1997}.
@@
 
-    {\bf 2.
-Action space.} We limit the agency of runners as well as pitchers to states in which only the first base is occupied (and the other bases are unoccupied).
-Though restricting the agency to states in which only the first base is occupied (i.e., the component $b$ in the state variable equals $(1,0,0)$) limits the scope, the situations in which the runner might steal the second base comprises roughly $71\%$ of stolen base attempts in 2023.
-We emphasize that this  excludes the scenario with $b = (1, 0, 1)$ in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home.
+    {\bf 2. Action space.} We limit the agency of runners and pitchers to states in which only first base is occupied 
+    (i.e., the component $b$ in the state variable equals $(1,0,0)$).
+Although this restriction limits the scope, these states covered roughly $71\%$ of stolen base attempts in 2023.
+We emphasize that this  excludes the scenario with $b = (1, 0, 1)$ in which the first-base runner will often attempt a stolen base to lure a throw from the catcher, giving the third-base runner a chance to steal home and establishing a different dynamic \parencite{miller_first-and-third_2024}.
 We leave this for future research.
 
-    For states in which only the first base is occupied, the first-base runner's action is the lead distance chosen in feet from the first base.
+    For states in which only first base is occupied, the first-base runner's action is the lead distance from first base, chosen in feet.
 The pitcher has only two actions: attempt a pickoff or throw a pitch.
@@
             {\cal L} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta, 
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s \in {\cal P} \cup \{\Delta\}
         \end{cases} \ \forall s \in {\cal S}\\
@@
             {\cal P} \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0) \\
-            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s = \Delta,         \end{cases} \ \forall s \in {\cal S}, \ a_\text{R} \in A_\text{R}(s)\\
+            \{\delta\} \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0), \text{ or } s \in {\cal P} {\cup} \{\Delta\},         \end{cases} \ \forall s \in {\cal S}, \ a_\text{R} \in A_\text{R}(s)\\
             \nonumber
@@
     Here, $\delta$ is a dummy element that captures the lack of agency.
-We note that the state of our stochastic game progresses even when no actions are played by both the players.
-This will be modeled through appropriate transition probabilities.
+We note that the state of our sequential game progresses even when no actions are played by both players.
+We will model this through appropriate transition probabilities.
 
     {\bf 3.
-Transition probabilities.} The uncertainty in our sequential game is modeled through appropriate transition probabilities.
-Basically, quantities describe the conditional probability of transitioning to a future state given the sequence of previously visited states and the actions played in those states.
-Our sequential game model assumes that the transition probabilities satisfy the Markov property, i.e., conditional on the entire history sequence, the probability of transitioning to a next state depends only on the current state and the actions played in that state.
+Transition probabilities.} The uncertainty in our sequential game is modeled through transition probabilities, i.e., the conditional probability of transitioning to a future state given the sequence of previously visited states and the actions played in those states.
+Our sequential game assumes that the transition probabilities satisfy the Markov property.
 Following the convention in the literature, we denote the transition probability as $p(\cdot|s,a_\text{R},a_\text{P}) \ \forall s \in {\cal S}, a_\text{R} \in A_\text{R}(s), a_\text{P} \in A_\text{P}(s,a_\text{R})$.
-Section \ref{sec:transition-probability-model} details our specific construction of transition probabilities.
-At a high level, we use generalized linear mixed-effects models to estimate runner outcome (e.g., unsuccessful pickoff attempt or successful stolen base attempt) probabilities which depend on lead distance and the pitcher's pickoff attempt decision.
-This along with appropriate conditional empirical frequencies will be used to construct the transition probabilities.
+Section \ref{sec:transition-probability-model} details our specific construction of transition probabilities, which uses generalized linear mixed-effects models to estimate runner outcome (e.g., unsuccessful pickoff attempt or successful stolen base attempt) probabilities which depend on lead distance and the pitcher's pickoff decision, as well as player-specific factors.
+These, along with appropriate conditional empirical frequencies, will be used to construct the transition probabilities.
     % SP: We are deleting this sentence because we can't remember why we need it
@@
 Note that our reward function in \eqref{eqn:runner_pitcher_reward} depends only on the current state and the next state, and not on the actions of the first-base runner or the pitcher.
-The formula in \eqref{eqn:common_reward} counts the total number of baserunners and outs before and after the transition and subtracts one if the transition corresponds to the end of a plate appearance.
-Since the number of runs scored on the last play can not be uniquely determined from the current state, we use the penultimate states to track the number of runs scored on the last play.
+The formula in \eqref{eqn:common_reward} infers the number of runs scored by counting the offensive players captured by the states before and after the transition.
+If the transition corresponds to the end of a plate appearance (i.e.
+$c' = (0, 0),\, d' = 0$), then we account for one more offensive player (the batter) in the next state.
+Since the number of runs scored on the last play cannot be uniquely determined from the current state, we use the penultimate states to track the number of runs scored on the last play.
 We note that our transition probabilities will be constructed such that these penultimate states transition to the terminal end-of-inning state with probability $1$.
-Since the reward function \eqref{eqn:runner_pitcher_reward} depends only on the current and the next state, it looks like it is independent of the actions of the players.
-However, the next state to which our sequential game transitions is (probabilistically) determined by the actions of both the players through the transition probabilities, thereby making the reward function indirectly depend on the actions of both the players.
+Since the reward function \eqref{eqn:runner_pitcher_reward} depends only on the current and the next state, it appears to be independent of the actions of the players.
+However, the next state to which our sequential game transitions is (probabilistically) determined by the actions of both the players through the transition probabilities, thereby making the rewards depend indirectly on the actions of both players.
 
@@
 After playing $a_\text{R}$ and $a_\text{P}$, the sequential game then transitions into state $s'$ with probability $p(s'|s,a_\text{R},a_\text{P})$ and the batting team collects a reward $r(s'|s,a_\text{R},a_\text{P})$ in the zero-sum game.
-A play then begins from $s'$ and the process repeats.
+A play then begins from $s'$, and the process repeats.
 The goal of the batting team is to maximize (through the actions of the first-base runner) the expected total rewards over the plays in the inning while the goal of the fielding team is to minimize (through the actions of the pitcher) the expected total rewards obtained by the batting team.
 
-    It is intuitive that any sound strategy for the first-base runner and the pitcher depends on the state of the sequential game (for the pitcher, it also depends on the actions played by the first-base runner).
+    Intuitively, any sound strategy for the first-base runner and the pitcher depends on the state of the sequential game (for the pitcher, it also depends on the actions played by the first-base runner).
 This is formally captured using the concept of policy.
 A (deterministic, stationary) policy for the first-base runner is a decision-rule that prescribes the action to be played in every possible state.
-In this context, stationarity refers to the fact that the same action is played in a state, regardless of when this state occurs.
-In other words, the action depends only on the state and not the timing of when the state occurs.
 Similarly, a (deterministic, stationary) policy for the pitcher is a decision-rule that prescribes the action to be played for every possible state and the corresponding actions of the first-base runner in those states.
-A mathematical treatment of policies is presented in Section \ref{sec:appendix}.
-We denote the set of all policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
+A mathematical treatment of policies is presented in the Appendix.
+We denote the set of policies for the runner and pitcher by $\Pi_\text{R}$ and $\Pi_\text{P}$, respectively.
 
@@
 Observe that, unlike the typical sequential games studied in the literature, we do not discount the rewards.
-This is because runs scored on later plays count the same as runs scored on earlier plays.
+This is because our optimization objective is rest-of-inning expected runs, which by definition does not discount runs scored on later plays.
 Though the undiscounted case is more realistic in our setup, it might lead to technical issues since $V^{\pi_\text{R},\pi_\text{P}}(s)$ in \eqref{eqn:sg_pol_val} can potentially be infinite.
-However, it is obvious that an inning in a baseball eventually halts.
-This means that the infinite series appearing inside the expectation in the right-hand side of \eqref{eqn:sg_pol_val} can be truncated at some finite $T \in \mathbb{N}$, and hence we can expect $V^{\pi_\text{R},\pi_\text{P}}(s)$ to take finite values.
-This reasoning is rigorously justified in Section \ref{sec:appendix}.
+However, all innings eventually end.
+As a result, the infinite series appearing inside the expectation in the right-hand side of \eqref{eqn:sg_pol_val} can be approximated with arbitrary precision by truncating the sum at some finite $T \in \mathbb{N}$, and hence we can expect $V^{\pi_\text{R},\pi_\text{P}}(s)$ to take finite values.
+This reasoning is rigorously justified in Lemma \ref{lem:sg_pol_val_well_defined} in the Appendix.
 Therefore, in the rest of the paper, we take $V^{\pi_\text{R},\pi_\text{P}}(s)$ to be a finite number for all $\pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}$, and $s \in {\cal S}$.
@@
     The goal of the batting team is to select a policy that will maximize the expected total runs scored by them.
-At the same time, the fielding team tries to minimize the expected total runs scored by the batting team by employing a suitable policy.
+At the same time, the fielding team minimizes the expected total runs scored by the batting team by employing a suitable policy.
 Taking the rational behavior of both the players into account, the maximum runs scored by the batting team when a play begins from state $s \in {\cal S}$ is given as
@@
    \end{equation}
-   We call the function $V^*$ the \emph{optimal value function}.
+   We call $V^*$ the \emph{optimal value function}.
 In our context, $V^*$ is the maximum expected total runs scored in an inning by the batting team, which, by the zero-sum property of the game, equals the minimum expected total runs incurred by the fielding team.
-A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called equilibrium policy.
-That is, $\pi_\text{R}^*$ is the policy for the batting team that maximizes the  expected total runs scored by them, and $\pi_\text{P}^*$ is the policy for the fielding team that minimizes the expected total runs incurred by them.
-In other words, $\pi_\text{R}^*$ is the optimal policy for the batting team and $\pi_\text{P}^*$ is the optimal policy for the fielding team.
+A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called an equilibrium policy.
+That is, $\pi_\text{R}^*$ is a policy for the batting team that maximizes the expected total runs scored by them, and $\pi_\text{P}^*$ is a policy for the fielding team that minimizes the expected total runs incurred by them.
+In other words, $\pi_\text{R}^*$ is a optimal policy for the batting team and $\pi_\text{P}^*$ is an optimal policy for the fielding team.
    % Note that any unilateral deviation from $(\pi_\text{R}^*,\pi_\text{P}^*)$ is suboptimal
@@
 
-   A characterization of optimality conditions for \eqref{eqn:sg_val} including a dynamic programming equation is presented in Section \ref{sec:appendix}.
-In our computational experiments in Section \ref{sec:results}, we used value iteration to solve \eqref{eqn:sg_val}.
+   A characterization of optimality conditions for \eqref{eqn:sg_val} including a dynamic programming equation is presented in the Appendix.
+In our computational experiments in Section \ref{sec:results}, we use value iteration to solve \eqref{eqn:sg_val}.
 The value iteration algorithm is as follows.
@@
         \end{align}
-    Lemma \ref{lem:vi_convergence} proves that the $V^k \to  V^*$.
-The proof of that lemma establishes a convergence rate of the form $\abs{V^k(s) - V^*(s)} \leq O(\rho^k)$ for all $s \in {\cal S}$, where $\rho \in (0,1)$ is a constant.
+     {\bf Theorem 1} (Informal) {\it 
+     The iterates $V^k$ generated in \eqref{eqn:vi_algorithm} converge to the optimal value function $V^*$.}
+
+     \noindent This result is proved in  in the Appendix.
+Furthermore, the proof of Theorem 1 also establishes a convergence rate of the form $\abs{V^k(s) - V^*(s)} \leq O(\rho^k)$ for all $s \in {\cal S}$ and a $\rho \in (0,1)$.
 
@@
       As itemized in (\ref{eqn:runner-outcome}), there are five possible outcomes for the runner with respect to the run game.
-To estimate each of these probabilities conditional on player identities, count and lead distance, we model four probabilities.
+To estimate each of these probabilities conditional on player identities, count, and lead distance, we model four probabilities.
 Using $R_i$ to denote the random variable governing the probability distribution over $r_i$,
@@
       \end{align*}
-      We estimate each of these probabilities with generalized linear mixed-effects models, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
-Specifically, we use logistic regression because it is appropriate for binary outcomes.
-The mixed-effects models allow the incorporation of player random effects along with fixed effects for game state variables and lead distance.
+      We estimate each of these probabilities with generalized linear mixed-effects models \parencite{stroup_generalized_2012}, implemented in R using the package lme4 \parencite{bates_fitting_2015}.
+Specifically, we use logistic regression because the outcomes are binary.
+Mixed-effects models allow us to incorporate player random effects along with fixed effects for game state variables and lead distance.
+Player random effects are particularly appropriate for this setting because they yield player skill estimates which go beyond measurable characteristics like sprint speed and are regularized towards a common mean.
 
@@
         \end{align}
-        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$ and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$.
+        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$, and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$.
 The number of random effects is $|\mathcal{H}^R| + |\mathcal{H}^P|$.
@@
 One may think of this as a deterministic decision made by the runner based on the expected run value with and without attempting a steal.
-Such a model aligns very poorly with real decisions made by runners.
-We posit that the runner's ability to attempt a stolen base is influenced by external factors such as the runner's perception of the pitcher's body language timing between pitches.
+Such a model aligns poorly with real decisions made by runners.
+We posit that the runner's ability to attempt a stolen base is influenced by external factors such as the runner's perception of the pitcher's body language and timing between pitches.
 For this reason, we model stolen base attempts as a stochastic process rather than a deterministic decision.
@@
 
-        We model $\psi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
+        We model $\psi_i$ using a mixed-effects logistic regression with fixed effects for balls, strikes, outs, disengagements (as a categorical), runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher, and catcher.
         \begin{align}
@@
 
-        We model $\psi_i^+$ using a mixed-effects logistic regression with fixed effects for lead distance, runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher and catcher.
+        We model $\psi_i^+$ using a mixed-effects logistic regression with fixed effects for lead distance, runner sprint speed, and catcher arm strength; and with random effects for runner, pitcher, and catcher.
         \begin{align}
@@
 
-    In this section, we present the results of applying the two-player stochastic game model and a single-player MDP model and to MLB data.
+    In this section, we present the results of applying the two-player sequential game model and a single-player MDP model to MLB data.
 Most of the data were downloaded from the publicly available MLB Stats API using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
 The dataset comprises 722,899 plays from the 2022 MLB regular season and 726,868 plays from the 2023 MLB regular season, where each play is either a pitch or a pickoff attempt (2.0\% pickoff attempts in 2022, 1.2\% in 2023).
-For each play, we observe the identities of the pitcher, catcher and runners involved.
-Before and after each play, we observe the state of the game, described by the number of outs in the inning; which bases are occupied by runners; the ball-strike count on the batter; and the number of disengagements already used by the pitcher.
-On each pickoff attempt, we observe whether it is successful (runner is out) or unsuccessful (runner is safe).
-On each pitch, we observe whether the runner attempts a stolen base; and if so, we observe whether they are successful (runner is safe) or unsuccessful (runner is out).
-On each pitch terminating a plate appearance (a batting attempt by a batter), we observe the plate appearance outcome, such as strikeout, groundout, single or home run.
+For each play, we observe the identities of the pitcher, catcher, and runners involved.
+Before and after each play, we observe the state of the game (bases occupied, count, disengagements, and outs).
+On each pickoff attempt, we observe whether it is successful.
+On each pitch, we observe whether the runner attempts a stolen base; and if so, we observe whether they are successful.
+On each pitch terminating a plate appearance, we observe the plate appearance outcome, such as strikeout, groundout, single, or home run.
 
     We downloaded publicly available player-season level data from Baseball Savant \parencite{baseball_savant_statcast_2024} describing measurable performance characteristics of catchers and runners in each season, both estimated using computer vision on video data from several camera angles.
-For each catcher-season, we observe {\it arm strength} (miles per hour), which measures how hard the catcher throws the ball.
-This is defined by Baseball Savant as the average of the top 10 percent (to capture peak performance) of throw speeds when the catcher is attempting to catch a base stealer.
-For each runner-season, we observe {\it sprint speed} (feet per second), which measures the runner's top running speed in their fastest one-second window.
-This is defined by Baseball Savant as the average of the top 67 percent (to exclude low-effort plays) of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
+For each catcher-season, we observe arm strength (miles per hour), which Baseball Savant defines as the average of the top 10 percent (to capture peak performance) of throw speeds when the catcher is attempting to catch a base stealer.
+For each runner-season, we observe sprint speed (feet per second), which Baseball Savant defines as the average of the top 67 percent (to exclude low-effort plays) of peak running speeds on qualified runs: (1) home-to-first runs on weakly hit batted balls and (2) runs of two or more bases on balls hit into play (excluding runs from second base on extra-base hits).
 
     In addition to the publicly available data described above, we rely on one key piece of proprietary information provided by MLB Advanced Media, estimated using computer vision on video data from several camera angles.
-For each play with a runner on first base, observe the runner's {\it lead distance} (feet), which is how far they are ranging from first base at the time of the pitcher's first movement.
-This is defined as the distance between the runner's center of mass (projected onto the line between the first base and second base) and the nearest edge of first base.
-When the pitcher throws a pitch, runners will typically extend their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a secondary lead), which is relevant for runner advancement on batted balls but not on stolen base attempts.
+For each play with a runner on first base, we observe the runner's lead distance (feet), which is defined as the distance between the runner's center of mass (projected onto the line between the first base and second base) and the nearest edge of first base, at the time of the pitcher's first movement.
+When the pitcher throws a pitch, the runner extends their lead by several feet from the time of the pitcher's first movement to the time of pitch release (known as taking a secondary lead), which is relevant for runner advancement on batted balls but not on stolen base attempts.
 Our data include only the lead distance at the time of the pitcher's first movement, not the secondary lead.
@@
 In 2022, the average lead distance was 9.6 feet.
-In 2023, the average lead distance under zero, one, and two disengagements was 9.6 feet, 10.3 feet and 11.0 feet, respectively.
+In 2023, the average lead distance under zero, one, and two disengagements was 9.6 feet, 10.3 feet, and 11.0 feet, respectively.
 
@@
 In 2022, there was no limit on the number of disengagements.
-The vertical lines mark the means of the distributions.
       }
@@
 For each of these three models, the direction of the relationship between lead distance and outcome probability is as expected.
-For the pickoff success probability model, we only included a fixed effect for lead distance and a random effect for pitcher because these were the only variables for which we found substantive effects.
 
@@
       Figure \ref{fig:prob-pickoff} illustrates how pickoff attempt probability and success probability vary with distance.
-We observe that pickoff attempt rates were lower in 2023 than in 2022, especially as the number of prior disengagement increased.
+We observe that pickoff attempt rates were lower in 2023 than in 2022, especially as the number of prior disengagements increased.
 The modeled probability remains low even for extremely long leads.
@@
 Interestingly, there is a large gap between the 90$^{th}$-percentile pitcher and the median pitcher but a small gap between the median pitcher and the 10$^{th}$-percentile pitcher, demonstrating a right skew in the distribution of pitcher pickoff skill.
-Earlier exploration indicated that game state and runner identity did not provide predictive value for pickoff success given lead distance.
 
@@
           \it Pickoff attempt probability (left) and success probability (right) modeled as functions of lead distance.
-The left figure corresponds to the model detailed from Section \ref{sec:prob-po-attempt}, assuming 0 balls, 0 strikes, 0 outs and a median pitcher random effect.
+The left figure corresponds to the model detailed from Section \ref{sec:prob-po-attempt}, assuming zero balls, zero strikes, zero outs, and a median pitcher random effect.
 The right figure corresponds to the model from Section \ref{sec:prob-po-success}, which includes only an intercept, a fixed effect for lead distance and a random effect for pitcher identity.
@@
 
-      The stolen base attempt probability model does not include lead distance, as explained in Section \ref{sec:prob-sb-attempt}.
-As a result, this model has only a secondary effect on the lead distance policies learned in later sections.
-
-      Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as function of lead distance, split by season and by runner skill (based on sprint speed and runner random effect).
+      Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as a function of lead distance, split by season and by runner skill (accounting for sprint speed and runner random effect).
 We observe that, conditioned on lead distance, even 10$^{th}$-percentile runners in 2023 had higher success probabilities than the median runner in 2022, which is likely attributable to rule changes making it easier to steal bases in 2023.
 Within the range covering most lead distances, the difference between the 90$^{th}$-percentile runner and the median runner is just under 5 percentage points of success probability, similar to the difference between the median runner and the 10$^{th}$-percentile runner.
-Base-out state and {\it battery} (i.e.
+Base-out state and battery (i.e.
 pitcher and catcher) skill also influence stolen base success probability although they are not shown in this visualization.
@@
           \it Stolen base success probability modeled as function of lead distance, split by season and runner effect (combining the fixed effect of runner sprint speed and the random effect of runner identity).
-This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-sb-success}, assuming 0 balls, 0 strikes, 0 outs and a median value for the combined effect of pitcher identity (random effect), catcher arm strength (fixed effect) and catcher identity (random effect).
+This figure shows the probability estimated by the model detailed in Section \ref{sec:prob-sb-success}, assuming zero balls, zero strikes, zero outs, and a median value for the combined effect of pitcher identity (random effect), catcher arm strength (fixed effect), and catcher identity (random effect).
         }
@@
 And to what extent do players over- or under-perform what is expected from their measurable characteristics?
-The combination of the fixed effect and the random effect gives us an estimate (which we call the {\it combined effect}) of each player's skill, regularized toward what is expected from their measurable characteristics.
+The combination of the fixed effect and the random effect gives us an estimate (which we call the combined effect) of each player's skill, regularized toward what is expected from their measurable characteristics.
 
@@
 
-      Using the transition probability function estimated on the empirical data, we implemented the policy iteration algorithm presented in Section \ref{sec:sg-model}.
+      Using the transition probabilities estimated on the empirical data, we implemented the value iteration algorithm presented in Section \ref{sec:sg-model}, discretizing the runner's action space to the nearest tenth of a foot in lead distance.
 Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and prior disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and prior disengagements.
@@
         \caption{
-          \it The optimal lead distance (in feet) by count and prior disengagements under the two-player stochastic game, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
+          \it The optimal lead distance (in feet) by count and prior disengagements under the two-player sequential game, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
         }
@@
 
-      The most striking observation is that the optimal lead distance is significantly longer with two prior disengagements (14.8--15.8 feet) than in other states.
+      The most striking observation is that the optimal lead distance is significantly longer with two prior disengagements (14.9--15.8 feet) than in other states.
 These recommended lead distances are near the edge of observed lead distances in the dataset but not outside of the observed range.
@@
         \caption{
-          \it The optimal lead distance (in feet) by outs and prior disengagements under the two-player stochastic game, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
+          \it The optimal lead distance (in feet) by outs and prior disengagements under the two-player sequential game, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
         }
@@
       Under the game-theoretic equilibrium, pitchers would be more aggressive with attempting pickoffs when runners take long leads.
-Based on observed lead distances in 2023, we estimate that a pickoff attempt would reduce run expectancy relative to a pitch in 11.4\% of opportunities with only first base occupied.
-When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.4\% of the time---this would be 100\% in the two-player model.
-When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 4.9\% of the time---this would be 0\% in the two-player model.
+Based on observed lead distances in 2023, we estimate that a pickoff attempt would be the optimal action in 11.4\% of opportunities with only first base occupied.
+When a pickoff attempt would have been recommended, pitchers attempted pickoffs 10.4\% of the time---100\% would be optimal.
+When a pickoff attempt would not have been recommended, pitchers attempted pickoffs 4.9\% of the time---0\% would be optimal.
 Overall, pitchers attempted pickoffs on 5.6\% of opportunities, less than the recommended 11.4\%.
@@
 After all, the pitcher has the additional cognitive burden of contending with the batter.
-In light of this, we consider the results of policy iteration applied to the one-player MDP.
-In this model, we assume the pitcher's pickoff attempt decision is a stochastic policy modeled using empirical data.
+In light of this, we consider the results of value iteration applied to the one-player MDP.
+In this model, we assume the pitcher's pickoff decision is a stochastic policy modeled using empirical data.
 Because the pitcher's action is not optimal, the runner can exploit it.
@@
         \caption{
-          \it The optimal lead distance (in feet) by count and prior disengagements under the single-player Markov decision process, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
+          \it The optimal lead distance (in feet) by count and prior disengagements under the single-player MDP, assuming an average runner on first base facing an average battery (pitcher and catcher) with no outs and third base empty.
         }
@@
         \caption{
-          \it The optimal lead distance (in feet) by outs and prior disengagements under the single-player Markov decision process, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
+          \it The optimal lead distance (in feet) by outs and prior disengagements under the single-player MDP, assuming an average runner on first base facing an average battery (pitcher and catcher) with an 0-0 count and third base empty.
         }
@@
 Regardless of the skill levels, the rule of thumb holds up well.
-The most notable exception to the Two-Foot Rule is when facing a 10th-percentile battery after the first disengagement.
-In this situation, we recommend a long lead before the first disengagement and an increase of roughly one foot after the first disengagement.
+The greatest deviation from the Two-Foot Rule is when a 90th-percentile runner faces a 10th-percentile battery after the first disengagement.
+In this situation, we recommend a long lead before the first disengagement and an increase of only 0.9 feet afterward.
 
@@
 
-    In comparing the optimal lead distance policy under the two-player stochastic game model (Table \ref{tab:lead-by-count-two-player}) and the single-player MDP (Table \ref{tab:lead-by-count-one-player}), we observe substantial differences between the optimal lead distance policy under a game-theoretic equilibrium and the optimal lead distance policy given observed pitcher behavior.
+    In comparing the optimal lead distance policy under the two-player game (Table \ref{tab:lead-by-count-two-player}) and the single-player MDP (Table \ref{tab:lead-by-count-one-player}), we observe substantial differences between them.
 For most states, the former lead distance is longer than the latter.
-The two-player optimal policy for the runner is to extend their lead until the pitcher is indifferent between attempting a pickoff or not.
+The optimal policy for the runner under the two-player game is to extend their lead until the pitcher is indifferent between attempting a pickoff or not.
 By contrast, the one-player optimal policy corresponds to a pickoff attempt probability below 10\%.
@@
 
-    In the two-player model we recommend long leads (beyond 15 feet) after two prior disengagements.
+    In the two-player game we recommend long leads (beyond 15 feet) after two prior disengagements.
 These distances are near the edge of the range of leads we observe but not outside the range.
@@
 By contrast, if the pitcher throws a pitch and if the runner attempts a stolen base (14.0\% probability), we estimate an 82.8\% probability that the stolen base is successful.
-In other words, the pitcher is indifferent to throwing a pitch or attempting a pickoff because the probability of an unsuccessful pickoff attempt is similar to the probability of a successful stolen base (conditioned on a steal attempt).
+In other words, the pitcher is indifferent to throwing a pitch or attempting a pickoff because the probability of an unsuccessful pickoff attempt is similar to the probability of a successful stolen base (conditioned on a steal attempt), and the outcome is the same either way.
 
     Because a third failed pickoff attempt results directly in free baserunner advancement, the pitcher is strongly disincentivized from attempting a pickoff after two prior disengagements.
-In the two-player model, this is reflected in a drastic increase in the optimal lead distance after the second disengagement (3--5 feet, Table \ref{tab:lead-by-count-two-player}).
-In the one-player model, we observe a more steady increase in optimal lead distance as a function of prior disengagements because observed pitcher pickoff attempt beheavior steadily decreases as a function of prior disengagements (see Figure \ref{fig:prob-pickoff}).
+In the two-player game, this is reflected in a drastic increase in the optimal lead distance after the second disengagement (3--5 feet, Table \ref{tab:lead-by-count-two-player}).
+In the one-player MDP, we observe a more steady increase in optimal lead distance as a function of prior disengagements because observed pitcher pickoff attempt beheavior steadily decreases as a function of prior disengagements (see Figure \ref{fig:prob-pickoff}).
 
@@
 Under the two-player game, the value of the start-of-inning state is 0.520 runs, which corresponds to approximately 6 more runs over a full season of 162 games of nine innings each.
-Therefore, offenses stand to benefit more in the transition from observed behavior to the game-theoretic equilibrium.
+Therefore, offenses stand to benefit more in the transition from observed behavior to a game-theoretic equilibrium.
 Under the one-player MDP, the value of the start-of-inning state is 0.524 runs, which corresponds to approximately 12 more runs over a full season, relative to observed behavior.
@@
 In the short term, pitchers are unlikely to immediately adapt their behavior to the optimal policy, and runners can exploit this.
-Because the pickoff attempt action is observable, runners can adjust their lead distance policy if they observe pitchers making more optimal pickoff attempt decisions.
+Because the pickoff attempt action is observable, runners can adjust their lead distance policy if they observe pitchers making better pickoff decisions.
 Therefore, the actionable strategic insight from our results is the Two-Foot Rule: Extend your lead by two feet after each pickoff attempt.
@@
     Under the new pickoff rules, baseball players and coaches understand intuitively that runners want to increase their lead distance after each successive disengagement by the pitcher---but by how much?
-In this paper, we formalize the cat-and-mouse game between pitcher and runner as a two-player stochastic game and estimate the optimal lead distance policy for a runner on first base with other bases empty, given the number of outs, the ball-strike count, and the number of prior disengagements by the pitcher.
+In this paper, we formalize the cat-and-mouse game between pitcher and runner as a two-player sequential game and estimate the optimal lead distance policy for a runner on first base with other bases empty, given the number of outs, the ball-strike count, and the number of prior disengagements by the pitcher.
 We model the probabilities of runner outcomes (pickoff attempt/success, stolen base attempt/success) as functions of lead distance, context, and player skill.
@@
 
-    We consider a two-player sequential game model in which the runner first decides their lead distance and then the pitcher (upon observing the runner's lead distance) then decides whether to attempt a pickoff or throw a pitch.
+    We consider a two-player sequential game in which the runner first decides their lead distance and then the pitcher (upon observing the runner's lead distance) then decides whether to attempt a pickoff or throw a pitch.
 We show that the game-theoretic optimal policy for the runner is a maximin strategy (with respect to rest-of-inning run expectancy) so that the pitcher is indifferent to attempting a pickoff or throwing a pitch.
@@
 While baseball analytics sometimes garners negative attention for diminishing the aesthetic appeal of the game \parencite{sheinin_analytics_2025}, this is an example for which the analytically recommended strategy could be more aesthetically pleasing.
-In other sports, stochastic and dynamic models have shown that decision-makers have been traditionally too conservative \parencite{romer_firms_2006, beaudoin_strategies_2010, yam_what_2019}, but this is a new result for leadoff and pickoff attempt decisions baseball.
+In other sports, stochastic and dynamic models have shown that decision-makers have been traditionally too conservative \parencite{romer_firms_2006, beaudoin_strategies_2010, yam_what_2019}, but this is a new result for leadoff and pickoff decisions baseball.
 
     We go a step further and consider the opportunity for runners to exploit the observed suboptimal behavior by pitchers.
-We consider a one-player MDP model in which the pitcher's agency in the pickoff attempt decision is replaced by a stochastic process, modeled on empirical data as a function of the runner's lead distance.
+We consider a one-player MDP in which the pitcher's agency in the pickoff decision is replaced by a stochastic process, modeled on empirical data as a function of the runner's lead distance.
 This framework leads to the most actionable recommendation for runners under the current conditions of pitcher behavior.
@@
 We assume that observed lead distance fully describes the runner's behavior and that they cannot disguise their aggressiveness by leaning in one direction.
-In reality, the runner may achieve a greater (or lesser) effective lead distance than we observed by leaning toward second base (or first base).
+In reality, the runner may achieve a greater (or lesser) effective lead distance than we observe by leaning toward second base (or first base).
 We also assume that lead distance only impacts pickoff and stolen base outcomes, not the outcomes of batted balls.
 For example, we assume that a greater lead distance does not improve the runner's probability of advancing from first base to third base on a single.
-Baseball domain knowledge suggests this is a reasonable assumption because while a pitch is being thrown, the runner moves to a secondary lead distance which is more important for determining advancement outcomes than primary lead distance (at first pitcher movement).
+Baseball domain knowledge suggests this is a reasonable assumption because while a pitch is being thrown, the runner moves to a secondary lead distance which is more important than primary lead distance for determining advancement outcomes.
 Finally, we assume that, other than pickoff attempts, no other disengagements occur.
-In reality, the pitcher may derive value from stepping off of the pitching rubber if they are short of breath or need to resolve a miscommunication with the catcher.
-Including all these extensions is an interesting future work.
-Another future direction is to model an entire roster, which is a nonstationary two-player stochastic game and introduces enormous technical challenges.
+In reality, the pitcher may derive value from stepping off the pitching rubber if they are short of breath or need to resolve a miscommunication with the catcher.
+These extensions are opportunities for future work.
+Another future direction is to model an entire roster, which is a nonstationary two-player sequential game and introduces enormous technical challenges.
 
@@
 If these data were publicly available, they would enable more granular modeling of the cat-and-mouse game between pitcher and runner.
-For example, with information about the runner's leaning posture and reaction time, future work might show that a mixed strategy for the pitcher's pickoff attempt decision achieves the best result by forcing a neutral posture for the runner and delaying the runner's reaction.
+For example, with information about the runner's leaning posture and reaction time, future work might show that a mixed strategy for the pitcher's pickoff decision achieves the best result by forcing a neutral posture for the runner and delaying the runner's reaction.
 As highlighted in Section \ref{sec:prob-sb-attempt}, another opportunity for future work is improved modeling of the runner's decision to attempt a stolen base, which we assumed to be stochastic and independent of lead distance.
@@
     Code related to this paper is available at \url{https://github.com/jfhahn2/pickoff-game-theory}.
-Lead distance data are not publicly available and were obtained from Major League Baseball.
+Lead distance data are not publicly available and were obtained from MLB.
 All other data are publicly available and may be downloaded using the R package sabRmetrics \parencite{powers_sabrmetrics_2025}.
 
-\section{Appendix: Technical results}
-\label{sec:appendix}
+  \printbibliography
+
+  \newpage
+
+  \section*{Appendix: Technical results}
+    % \label{sec:appendix}
     We present a formal description of the sequential game described in Section \ref{sec:sg-model} and the related optimality characterizations.
@@
 With an abuse of notation, we overload $V^*$ to denote the optimal value function in \eqref{eqn:sg_val} and \eqref{eqn:sg_val_extended}.
-The reason for this choice is that under Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous}, Lemma \ref{lem:stationary_pol_optimal} proves that there exists  deterministic, stationary policies for the runner and pitcher that attain the respective maximum and minimum in \eqref{eqn:sg_val_extended}.
-That is the policy spaces in the right-hand side of \eqref{eqn:sg_val_extended} can be restricted to the smaller class of deterministic, stationary policies as in \eqref{eqn:sg_val} without affecting the optimal value in \eqref{eqn:sg_val_extended}.
+The reason for this choice is that under Assumptions \ref{assum:game_halts} and \ref{assum:transition_pby_continuous}, Proposition \ref{lem:stationary_pol_optimal} proves that there exists  deterministic, stationary policies for the runner and pitcher that attain the respective maximum and minimum in \eqref{eqn:sg_val_extended}.
+That is, the policy spaces in the right-hand side of \eqref{eqn:sg_val_extended} can be restricted to the smaller class of deterministic, stationary policies as in \eqref{eqn:sg_val} without affecting the optimal value in \eqref{eqn:sg_val_extended}.
 
@@
 
-   Under Assumption \ref{assum:game_halts}, we prove that the expected number of runs scored by the batter, $V^{\pi_\text{R},\pi_\text{P}}$, in  \eqref{eqn:sg_pol_val} is well-defined (i.e., the limit exists) and finite.
+       Under Assumption \ref{assum:game_halts}, we prove that the expected number of runs scored by the batter, $V^{\pi_\text{R},\pi_\text{P}}$, in  \eqref{eqn:sg_pol_val} is finite.
 Our proof is based on standard arguments used in the analyses of undiscounted problems \parencite[Section 7.2]{bertsekas2012dynamic}.
@@
            \begin{align*}
-           V^{\pi_\text{R},\pi_\text{P}}(s) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right]\\
-           &= \lim_{t \to \infty} \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{j=0}^{t-1} r(S_{j+1}|S_j) \Big| S_0 = s\right].
+               V^{\pi_\text{R},\pi_\text{P}}(s) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] = \lim_{t \to \infty} \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{j=0}^{t-1} r(S_{j+1}|S_j) \Big| S_0 = s\right].
            \end{align*}
@@
 
-    We present optimality characterizations in Lemmas \ref{lem:vi_convergence} to \ref{lem:policy_iteration}.
-These lemmas and their proofs are similar to the undiscounted problem in the single-player case \parencite[Section 7.2]{bertsekas2012dynamic} and the zero-sum stochastic game with simultaneous moves \parencite{patek_stochastic_1999}.
-We state and prove our lemmas for completeness and to outline the minor differences.
+        We present optimality characterizations in Theorem \ref{lem:vi_convergence} and Propositions \ref{lem:bellman_equations} to  \ref{lem:policy_iteration}.
+These results and their proofs are similar to the undiscounted problem in the single-player case \parencite[Section 7.2]{bertsekas2012dynamic} and the zero-sum stochastic game with simultaneous moves \parencite{patek_stochastic_1999}.
+We state and prove our results for completeness and to outline the minor differences.
       \begin{definition}\label{defn:sg_operators}
@@
 
-      \begin{lemma}
+          \begin{theorem}
            \label{lem:vi_convergence}
@@
            \end{enumerate}
-   \end{lemma}
+       \end{theorem}
        \begin{proof}
@@
            \nonumber
-           V^{\pi_\text{R},\pi_\text{P}}(s_0) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
+               V^{\pi_\text{R},\pi_\text{P}}(s_0) &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + \sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
                \nonumber
-           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + \sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
-           \nonumber
-           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t)\Big| S_0 = s_0\right] + \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
-           \nonumber
-           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) - u_0(S_{mK-1})\Big| S_0 = s_0\right] +\\
-           \nonumber
-           &\qquad \qquad \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right]\\
-           \nonumber
-           &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \Big| S_0 = s_0\right] -\\
-           \label{eqn:vi_convergence_temp1}
+               &= \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=0}^{mK-1} r(S_{t+1}|S_t) + u_0(S_{mK-1}) \Big| S_0 = s_0\right] -\\           \label{eqn:vi_convergence_temp1}
                &\qquad \qquad \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[ u_0(S_{mK-1})\Big| S_0 = s_0\right] + \mathbb{E}^{\pi_\text{R},\pi_\text{P}}\left[\sum_{t=mK}^\infty r(S_{t+1}|S_t) \Big| S_0 = s_0\right].
@@
 
+    Theorem \ref{lem:vi_convergence}(a) is value iteration for our sequential game.
+The computational experiments in Section \ref{sec:results} use this value iteration to solve our sequential game after a suitable discretization of the runner's action space.
+Apart from serving as an algorithmic device, Theorem \ref{lem:vi_convergence} also enables proving the optimality of deterministic, stationary policies.
+This result established in Proposition \ref{lem:stationary_pol_optimal} uses the claim in Proposition \ref{lem:bellman_equations}, which in turn relies on the below assumption.
+
        \begin{assumption}
@@
        \end{assumption}
-   \noindent {\color{blue}{Assumption \ref{assum:transition_pby_continuous} is a continuity condition imposed on the transition probabilities when viewed as a function of the runner and pitcher actions.
-Lemmas \ref{lem:bellman_equations} to \ref{lem:policy_iteration} rely upon this assumption to prove their stated claims.
+       \noindent Assumption \ref{assum:transition_pby_continuous} is a continuity condition imposed on the transition probabilities when viewed as a function of the runner and pitcher actions.
+Propositions  \ref{lem:bellman_equations} to \ref{lem:policy_iteration} rely upon this assumption to prove their stated claims.
 Our construction of transition probabilities in Section \ref{sec:transition-probability-model} satisfy this assumption.
-Note that this assumption trivially holds if the action space of the runner $A_\text{R}(s)$ is finite for all $s \in {\cal S}$.}}
-
-   \begin{lemma}
+Note that this assumption trivially holds if the action space of the runner $A_\text{R}(s)$ is finite for all $s \in {\cal S}$.
+
+       \begin{proposition}
            \label{lem:bellman_equations}
@@
            \end{enumerate}
-   \end{lemma}
+       \end{proposition}
        \begin{proof}
@@
 Consider arbitrary $u_0 \in \mathbb{R}^{\abs{\cal S}}$ with $u_0(\Delta) = 0$ and define $u_{k} = Tu_{k-1} \ \forall k \in \mathbb{N}$.
-By Lemma \ref{lem:vi_convergence}, $\lim_{k \to \infty}u_k(s) = V^*(s)$ for all $s \in {\cal S}$.
+By Theorem \ref{lem:vi_convergence}, $\lim_{k \to \infty}u_k(s) = V^*(s)$ for all $s \in {\cal S}$.
 Passing to limit $k \to \infty$ in $u_k = Tu_{k-1}$,  we have for all $s \in {\cal S}$
@@
            \endgroup
-       where the second equality follows from Assumption \ref{assum:transition_pby_continuous} and the convergence of $u_{k-1}$ to $V^*$ established in Lemma \ref{lem:vi_convergence}.
+           where the second equality follows from Assumption \ref{assum:transition_pby_continuous} and the convergence of $u_{k-1}$ to $V^*$ established in Theorem \ref{lem:vi_convergence}.
 We have established the continuous convergence of the sequence of functions $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ over $A_\text{R}(s)$.
@@
            \end{align*}
-       thereby establishing \eqref{eqn:bellman_equation}. The uniqueness of $V^*$ follows from the value iteration convergence result established in Lemma \ref{lem:vi_convergence}.
+           thereby establishing \eqref{eqn:bellman_equation}. The uniqueness of $V^*$ follows from the value iteration convergence result established in Theorem \ref{lem:vi_convergence}.
 
@@
 
-   \begin{lemma}
+       \begin{proposition}
            \label{lem:stationary_pol_optimal}
@@
 Then a stationary pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an equilibrium policy if and only if $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
-   \end{lemma}
+       \end{proposition}
        \begin{proof}
@@
            where the second equality follows by the definition of $T_{\pi_\text{R}^*,\pi_\text{P}^*}$ in \eqref{eqn:eval_operator}.
-By Lemma \ref{lem:bellman_equations}(b), we then get $V^*(s)=V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \ \forall s \in {\cal S}$.
+By Proposition  \ref{lem:bellman_equations}(b), we then get $V^*(s)=V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \ \forall s \in {\cal S}$.
 Hence, $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an equilibrium policy.
@@
 Then $V^*(s) = V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \ \forall s \in {\cal S}$.
-From parts $(a)$ and $(b)$ of Lemma  \ref{lem:bellman_equations}, it follows that $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
+From parts $(a)$ and $(b)$ of Proposition   \ref{lem:bellman_equations}, it follows that $\pi_\text{R}^*(s)$ and $\pi_\text{P}^*(s,\pi_\text{R}^*(s))$ attain the maximum and minimum in \eqref{eqn:bellman_equation} for all $s \in {\cal S}$.
        \end{proof}
 
-   \begin{lemma}
+    The next result states and proves the policy iteration algorithm for our two-player zero-sum sequential game.
+       \begin{proposition}
            \label{lem:policy_iteration}
@@
             Then $V^{\pi_\text{R}^k} \uparrow V^*$. 
-   \end{lemma}
+       \end{proposition}
        \begin{proof}
@@
 Denote the value of this MDP as $V^{\pi_\text{R}}$, i.e., $V^{\pi_\text{R}}(s) \defeq \min_{\pi_\text{P} \in \Pi_\text{P}} V^{\pi_\text{R},\pi_\text{P}}(s) \ \forall s \in {\cal S}$.
-Using similar arguments as in the proof of Lemma \ref{lem:vi_convergence}, it follows that $\lim_{k \to \infty}T_{\pi_\text{R}}^k u = V^{\pi_\text{R}}$, where $T_{\pi_\text{R}}^k$ is the composition of $T_{\pi_\text{R}}$ taken with itself $k$ times.
-Also, using arguments similar to the proof of Lemma \ref{lem:bellman_equations}, it follows that $V^{\pi_\text{R}}$ is the unique fixed point of $T_{\pi_\text{R}}$.
+Using similar arguments as in the proof of Theorem  \ref{lem:vi_convergence}, it follows that $\lim_{k \to \infty}T_{\pi_\text{R}}^k u = V^{\pi_\text{R}}$, where $T_{\pi_\text{R}}^k$ is the composition of $T_{\pi_\text{R}}$ taken with itself $k$ times.
+Also, using arguments similar to the proof of Proposition  \ref{lem:bellman_equations}, it follows that $V^{\pi_\text{R}}$ is the unique fixed point of $T_{\pi_\text{R}}$.
 
@@
 We prove that $\bar V = V^*$.
-Using arguments similar to the ones used in the proof of Lemma \ref{lem:bellman_equations}, it follows that $T$ is a  continuous operator on $\mathbb{R}^{\abs{\cal S}}$, i.e., $T u^k \to T u$ for all $u^k \to u$.
+Using arguments similar to the ones used in the proof of Proposition  \ref{lem:bellman_equations}, it follows that $T$ is a  continuous operator on $\mathbb{R}^{\abs{\cal S}}$, i.e., $T u^k \to T u$ for all $u^k \to u$.
 Hence, passing to limit $k \to \infty$ in \eqref{eqn:policy_iteration_temp2} gives us $T \bar V = \bar V$.
-From Lemma \ref{lem:bellman_equations}(a), it follows that $\bar V = V^*$, thereby completing the proof.
+From Proposition  \ref{lem:bellman_equations}(a), it follows that $\bar V = V^*$, thereby completing the proof.
        \end{proof}
 
-   If the state- and action-spaces are all finite, the the convergence in Lemma \ref{lem:policy_iteration} happens in a finite number of steps.
+       If the state- and action-spaces are all finite, the convergence in th policy iteration algorithm happens in a finite number of steps.
 In that case, the policy obtained in the last iterate, say, $\pi_\text{R}^*$, is optimal for the runner.
@@
        \endgroup
-   We use this approach in our computational experiments in Section \ref{sec:results}, after an appropriate discretization of the runner's action space.
 
@@
 
-  \printbibliography
-
 \end{document}

### 2025-12-23 articles/arxiv/paper.tex
delete vestigial latex comments
expand on the hirotsu and bickel (2019) reference

@@
 
-% In 2023, Major League Baseball (MLB) introduced several new rule changes to make the game more exciting for fans.
-Among these changes, the pitcher was limited to three pickoff attempts, creating a cat-and-mouse game between pitcher and runner.
-Each failed pickoff attempt adds pressure on the pitcher to avoid using another pickoff attempt, and the runner can intensify this pressure by extending their leadoff toward the next base.
-We model this dynamic as a sequential two-player zero-sum sequential game in which the runner first chooses a lead distance, and then the pitcher chooses whether to attempt a pickoff.
-We establish optimality characterizations for our sequential game and present variants of value iteration and policy iteration to solve the game.
-Using MLB lead distance data, we estimate generalized linear mixed-effects models for pickoff and stolen base outcome probabilities given lead distance, context, and player skill.
-We compute the game-theoretic equilibrium policies under the two-player model using policy iteration, as well as the optimal runner policy under a reduced one-player Markov decision process model.
-In the one-player setup, our results establish an actionable rule of thumb: the Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
-
   \section{Introduction}
@@
 
-%    We use a two-step process to first estimate runner outcome probabilities and then use empirical transition probabilities conditioned on the runner outcome.
-For the runner outcome, we improve upon empirical probabilities by using multilevel models\footnote{citation needed} to share information across players using observable as well as unobservable player skills.
-Our numerical study on recent real-world data illustrates that, under the optimal policy, actions played by pitcher and runner would favor the runner, relative to observed decisions.
-The results obtained by restricting our model to a single-player setup, that allows agency only for the runner, reveals an actionable recommendation: The Two-Foot Rule, which recommends that a runner increase their lead by two feet after each pickoff attempt.
-
     \subsection{Related work}
@@
 Recently, \textcite{chan_points_2021} used a Markov model to evaluate team performance in football.
-More specifically, the MDP has become a particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \parencite{nadimpalli_when_2013}, baseball \parencite{hirotsu_using_2019}\todo{pull this out with its own sentence at the end}, basketball \parencite{sandholtz_markov_2020}, darts \parencite{baird_optimising_2020} and football \parencite{biro_reinforcement_2022}.
+More specifically, the MDP has become a particularly popular model in recent years for analyzing decision-making in sports, with applications in tennis \parencite{nadimpalli_when_2013}, basketball \parencite{sandholtz_markov_2020}, darts \parencite{baird_optimising_2020} and football \parencite{biro_reinforcement_2022}.
+In one particularly relevant baseball application, \textcite{hirotsu_using_2019} used an MDP to model sacrifice bunt decisions on the basis of win probability, finding that the sacrifice bunt was more valuable than previously thought.
 
@@
     {\bf 1. State space.} We use $s=(b,c,d,o) \in {\cal X}$ to capture a play, where     
-    % The mid-inning game state $s \in \mathcal{M}$ is given by $s = (b, c, d, o)$, where:
     \begin{align*}
@@
 These, along with appropriate conditional empirical frequencies, will be used to construct the transition probabilities.
-    % SP: We are deleting this sentence because we can't remember why we need it
-    %Any state for which the number of baserunners plus the number of outs is at least three (i.e., $g(b) + o \ge 3$) may transition to the absorbing state with positive probability because there is always the possibility of recording enough outs on a single play to end the inning.
-\todo{Siva: This last sentence needs to go to the section on transition probabilities.}
 
@@
 In other words, $\pi_\text{R}^*$ is a optimal policy for the batting team and $\pi_\text{P}^*$ is an optimal policy for the fielding team.
-   % Note that any unilateral deviation from $(\pi_\text{R}^*,\pi_\text{P}^*)$ is suboptimal
-   % \todo{AS: Tone this down in case optimal not unique? Or do we have an expansive notion of suboptimal?}
-   % for the respective players.
 
@@
 
-% SP: I don't think we need this given the intended OR audience
-%      Figure \ref{fig:finding-optimal-lead} illustrates how the runner would choose an optimal lead distance for different numbers of prior disengagements in an example scenario.
-Short lead distances have very low pickoff probabilities and low stolen base probabilities.
-As lead distance increases, stolen base probability increases, but so does pickoff probability.
-For very long lead distances, the run expectancy drops off sharply as the pickoff probability increases sharply.
-Between those two extremes, the runner strikes a balance to maximize run expectancy.
-The optimal lead distances for 0, 1 and 2 disengagements are 10.0 feet, 11.5 feet and 14.1 feet, respectively.
-%      
-%      \begin{figure}[H]
-%        \centering
-%        \includegraphics[width = 0.8\textwidth]{finding_optimal_lead_light.pdf}
-%        \caption{
-%          \it Rest-of-inning run expectancy for different choices of lead distance.
-Three separate curves show the run expectancy for differing numbers of prior disengagements, assuming 0 balls; 0 strikes; 0 outs; third base unoccupied; and median player effects for pitcher, catcher and runner.
-Run expectancy is defined as the expectation of the optimal value function $V^*$ (evaluated at the resulting state), defined by (\ref{eqn:mdp_val}).
-The vertical line corresponding to each curve marks the value at which the curve is maximized.
-%        }
-%        \label{fig:finding-optimal-lead}
-%      \end{figure}
-
       Table \ref{tab:lead-by-players} shows the robustness of the Two-Foot Rule under different assumptions regarding player talent.
@@
       \end{table}
-
-% SP: Remove this result to reduce the imbalance between two-player results and one-player results because we have chosen to emphasizse the two-player model more.
-%      Table \ref{tab:actual-vs-rec} compares observed runner behavior against the optimal policy learned by the MDP.
-For each pitch with a runner on first base and second base empty, we compare the runner's lead distance with the recommended lead distance for an average runner against an average battery.
-With zero disengagements, the average lead is slightly less than the average recommended lead, and most leads are shorter than recommended.
-As the disengagements increase, this gap widens.
-On average, batters increase their leads by 0.6--0.7 feet afer each disengagement, significantly less than the Two-Foot Rule would recommend.
-%
-%      \begin{table}[H]
-%        \centering
-%        \input{../../output/tables/actual_vs_rec_lead.tex}
-%        \caption{
-%          \it Comparison of actual versus recommended lead distance by number of prior disengagements.
-The recommended lead distance is based on the one-player MDP model and accounts for count, outs and prior disengagements, assuming average battery and runner skill.
-%        }
-%        \label{tab:actual-vs-rec}
-%      \end{table}
 
@@
        \end{proof}
-
-       % The function $\widetilde V^*$ will be referred to as the value of the game.
-A policy pair $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ such that $\widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) = \widetilde V^*(s) \ \forall s \in {\cal S}$ is called the optimal strategy or equilibrium strategy.
-Note that $(\pi_\text{R}^*,\pi_\text{P}^*)$ is an optimal strategy if and only if $\widetilde V^{\pi_\text{R},\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}^*}(s) \leq \widetilde V^{\pi_\text{R}^*,\pi_\text{P}}(s) \ \forall \pi_\text{R} \in \Pi_\text{R}, \pi_\text{P} \in \Pi_\text{P}, s \in {\cal S}$.
 
@@
 
-    %    \begin{theorem}[\parencite{patek_stochastic_1999}]\label{thm:sg_optimality_condition}
-    %        \normalfont
-    %       Suppose Assumption \ref{assum:game_halts} holds.
-Then the value of the game $\widetilde V^*$ is the unique fixed point of the operator $\widetilde \Phi$.
-Furthermore, $(\pi_\text{R}^*,\pi_\text{P}^*) \in \Pi_\text{R} \times \Pi_\text{P}$ is an optimal strategy if and only if $\widetilde\Phi_{\pi_\text{R}^*}\widetilde V^* = \widetilde\Phi \widetilde V^* = \widetilde\Phi_{\pi_\text{P}^*}\widetilde V^*$.
-    %    \end{theorem}
-
-    %    The above theorem enables the characterization of $\widetilde V^*(s)$ as in the following lemma.
-
-    % \begin{corollary}\label{corr:sg_value_characterization}
-    %     \normalfont
-    %     Suppose Assumption \ref{assum:game_halts} holds. Then $\widetilde V^*$ is the unique solution to the following system of equations.
-    %     \begin{align*}
-    %         V(s) &= \max_{\ell \in {\cal L}}\min_{a \in {\cal P}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell,a) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-    %         V(s) &= \sum_{s' \in {\cal      S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
-    %         V(s) &=  0 \quad \text{ if } s = \Delta.
-    %     \end{align*}
-    %     \begin{proof}
-    %         Immediate from Theorem \ref{thm:sg_optimality_condition}.
-    %     \end{proof}
-    % \end{corollary}
-
-       % An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{patek_stochastic_1999}.
-We use policy iteration  in this work to compute the optimal policy which is as follows.
-Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi_\text{R}^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi_\text{R}^k$ by applying the updates
-       %      \begin{align}
-       %        \label{eqn:update-value-two-player}
-       %        V^{k+1}(s) &= \min_{a_\text{P} \in A_\text{P}(s,\pi_\text{R}^k(s))} \sum_{s' \in \mathcal{S}} [r(s, s') +  V^k(s')]p(s'|s,\pi_\text{R}^k(s),a_\text{P}) && \forall s \in \mathcal{S}, \mbox{ and}\\
-       %        \label{eqn:update-policy-two-player}
-       %        \pi_\text{R}^{k+1}(s) &= \argmax_{a_\text{R} \in A_\text{R}(s)} \left\{\min_{a_\text{p} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in \mathcal{S}}[r(s, s') + V^{k+1}(s')] p(s'|s,a_\text{R},a_\text{P})\right\} && \forall s \in \mathcal{S}.
-       %      \end{align}
-       %  The \footnote{Andrew - don't like (13) - argmax is a set, but pi is an element of the set.
-Can we use $\in$ instead of =?
-Likely true for other arg maxes} iterates generated by the policy iteration satisfy $V^k \uparrow \widetilde V^*$\parencite{patek_stochastic_1999}.
-
-          % \subsection{Single-player setup}
-       % For a fixed stationary policy of the pitcher, the stochastic game reduces to a (single-player) MDP.
-Lemmas \ref{lem:vi_convergence} to \ref{lem:policy_iteration} then reduce to the corresponding MDP versions.
-We omit stating them for brevity.
-See \cite[Section 7.2]{bertsekas2012dynamic} for details.
-
-    %    The MDP is then defined by the tuple ${\cal M} \defeq \langle {\cal S}, \{A_\text{R}(s)\}_{s \in {\cal S}}, \{r(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}, \\  \{p(s'|s,a)\}_{s \in {\cal S}, a \in A_\text{R}(s)}^{s' \in {\cal S}}\rangle$.
-Using this MDP, the progression of an inning can be captured as follows.
-At the beginning of a play, the game is in state $s \in {\cal S}$.
-Upon observing this state, an action $a \in A_\text{R}(s)$ is picked.
-The game then transitions to state $s'$ with probability $p(s'|s,a)$ and the batting team collects a reward $r(s'|s,a)$.
-A play then begins from $s'$ and the process repeats.
-The goal of the batting team is to maximize the expected total rewards collected over the plays in the inning.
-In other words, we assume the objective is to maximize runs rather than maximizing the probability of winning the game.
-The former objective is an approximation of the latter objective which allowed us to simplify the state space.
-This is a reasonable approximation before the final innings of the game, when it is more important to consider the score.
-
-    %       \begin{assumption}\label{assum:mdp_halts}
-    %        \normalfont
-    %       The MDP ${\cal M}$ satisfies
-    %       $
-    %       \lim_{t \to \infty}\sup_{\pi \in \Pi, s \in {\cal S}}\mathbb{P}^\pi[S_t = \Delta | S_0 = s] = 1.
-    %       $
-    %    \end{assumption}   
-    %   \noindent  Assumption\footnote{Andrew - can we put the assumption in words, and the formal math in the text below?
-Similar comment for Assumption 2 later}   \ref{assum:mdp_halts} states that there exists a natural number $T < \infty$ such that every inning halts within $T$ plays with arbitrarily high probability regardless of the policy employed or the starting state of the inning.
-This is a reasonable assumption because an inning does not continue forever.
-
-    %    The value of a policy $\pi$ when starting from state $s \in {\cal S}$ equals  
-    %    \begin{equation}
-    %        \label{eqn:mdp_pol_val}
-    %        V^{\pi}(s) \defeq \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall s \in {\cal S}.
-    %    \end{equation}
-    %   That is, the value of the policy is the the expected total runs scored in an inning.
-Under Assumption \ref{assum:mdp_halts}, $V^\pi(s)$ exists for all $\pi \in \Pi$ and $s \in {\cal S}$.
-To see that, since $r(s'|s) \geq 0$ for all $s,s' \in {\cal S}$, we have by the monotone convergence theorem
-    %   \[
-    %   V^\pi(s) = \mathbb{E}^\pi\left[\sum_{t=0}^\infty r(S_{t+1}|S_t) \Big| S_0 = s\right] = \lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right] \ \forall \pi \in \Pi, s \in {\cal S}.
-    %   \]
-    %    Using Assumption \ref{assum:mdp_halts} and the fact that $\sup_{s,s' \in {\cal S}}\abs{r(s'|s)} < \infty$ (due to finiteness of ${\cal S}$), it is easy to establish that $\lim_{T \to \infty} \mathbb{E}^\pi\left[\sum_{t=0}^T r(S_{t+1}|S_t) \Big| S_0 = s\right]$ exists for all $\pi \in \Pi, s \in {\cal S}$.
-Furthermore, the limit is uniformly bounded over $\Pi \times {\cal S}$.
-The value of the MDP is defined as
-    %    \begin{equation}
-    %         \label{eqn:mdp_val}
-    %        V^*(s) \defeq \max_{\pi \in \Pi_\text{R}}V^\pi(s) \ \forall s \in S.
-    %    \end{equation}
-    %    The function $V^*$, known as the optimal value function, is the maximum expected runs scored starting from state $s \in {\cal S}$.
-A policy $\pi^*$ such that $V^{\pi^*}(s) = V^*(s) \ \forall s \in {\cal S}$ is called an optimal policy.
-
-    %    We present the following definitions that will be needed to characterize the optimal value function and an optimal policy.
-
-    %    \begin{definition}\label{defn:mdp_operators}
-    %     \normalfont
-    %     \begin{enumerate}
-    %         \item For any $\pi \in \Pi_\text{R}$, the evaluation operator $\Phi_\pi:\mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
-    %         \[
-    %         (\Phi_\pi V)(s) \defeq \mathbb{E}_{s' \sim p(\cdot|s,\pi(s))}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-    %         \]
-    %         \item The optimality operator $\Phi: \mathbb{R}^{\abs{{\cal S}}} \to \mathbb{R}^{\abs{{\cal S}}}$ is defined as
-    %         \[
-    %         (\Phi V)(s) \defeq \max_{a \in A_\text{R}(s)}\mathbb{E}_{s' \sim p(\cdot|s,a)}\left[r(s'|s) + V(s')\right] \ \forall s \in {\cal S}.
-    %         \]
-    %     \end{enumerate}       
-    %    \end{definition}
-
-    %    \begin{theorem}[\parencite{bertsekas_analysis_1991}]\label{thm:mdp_optimality_condition}
-    %        \normalfont
-    %       Suppose Assumption \ref{assum:mdp_halts} holds. Then
-    %       \begin{enumerate}
-    %           \item For any policy $\pi$, its value $V^\pi$ is the unique fixed point of the operator $\Phi_\pi$.
-    %           \item The optimal value function $V^*$ is the unique fixed point of the operator $\Phi$. Furthermore, $\pi^*$ is an optimal policy if and only if $\Phi_{\pi^*}V^* = \Phi V^*$.
-    %       \end{enumerate}
-    %    \end{theorem}
-    %  \todo[inline]{The above theorem also requires $p(s'|s,a)$ to be continuous in $a$ for any $s,s' \in S$.
-This is true for fixed-effects model since $x \mapsto e^x/1+e^x$ is continuous.
-I do not know about random-effects model.
-[[SP: $a$ comes into the model as a fixed effect (the random effects are for player identities), so, yes, $p(s'|s,a)$ is continuous in $a$.]]}
-
-    % Theorem \ref{thm:mdp_optimality_condition} enables the characterization of $V^*(s)$ as in the following lemma.
-
-    % \begin{lemma}\label{lem:mdp_value_function_characterization}
-    %     \normalfont
-    %     Suppose \footnote{Andrew - Don't love this as a lemma - corollary seems more appropriate} Assumption \ref{assum:mdp_halts} holds.
-Then $V^*$ is the unique solution to the following system of equations.
-    %     \begin{align*}
-    %         V(s) &= \max_{\ell \in {\cal L}}\sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\ell) \quad  \text{ if } s = (b,c,d,o) \text{ and } b = (1,0,0)\\
-    %         V(s) &= \sum_{s' \in {\cal S}}\left[r(s'|s) + V(s')\right]p(s'|s,\delta) \quad \text{ if } s = (b,c,d,o) \text{ and } b \neq (1,0,0)\\
-    %         V(s) &= 0 \quad \text{ if } s = \Delta.
-    %     \end{align*}
-    %     \begin{proof}
-    %         Immediate from Theorem \ref{thm:mdp_optimality_condition}.
-    %     \end{proof}
-    % \end{lemma}
-
-    % An algorithmic device to obtain the optimal policy is value iteration or policy iteration \parencite{bertsekas_analysis_1991}.
-We use policy iteration  in this work to compute the optimal policy which is as follows.
-Initialize $V^0 \in \mathbb{R}^{\abs{{\cal S}}}$ and $\pi^0 \in \Pi_\text{R}$, and then iteratively obtain $V^k$ and $\pi^k$ by applying the updates
-    %         \begin{align}
-    %           \label{eqn:update-value-single-agent}
-    %           V^{k+1}(s) &= \sum_{s' \in \mathcal{S}} p(s' | s,\, \pi^k(s)) [r(s, s') + V^k(s')] && \forall s \in \mathcal{S}, \mbox{ and}\\
-    %           \label{eqn:update-policy-single-agent}
-    %           \pi^{k+1}(s) &= \argmax_{a \in A_\text{R}(s)} \sum_{s' \in \mathcal{S}} p(s' | s,\, a) [r(s, s') + V^{k+1}(s')] && \forall s \in \mathcal{S}.
-    %         \end{align}
-    %     The iterates generated by the policy iteration satisfy $V^k \uparrow V^*$ and every limit point of $\pi^k$ is an optimal policy \parencite{bertsekas_analysis_1991}.
-
 \end{document}

### 2025-12-23 articles/arxiv/paper.tex
fix figure/table typesetting, widows and orphans, etc.

@@
 
-    \begin{figure}[H]
+    \begin{figure}
       \centering
@@
 
-      \begin{table}[H]
-        \hspace{-6mm}
+      \begin{table}
+        \small
+        \center
         \input{../../output/tables/model_summary.tex}
@@
 
-      \begin{figure}[H]
+      \begin{figure}
         \centering
@@
 
-      Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as a function of lead distance, split by season and by runner skill (accounting for sprint speed and runner random effect).
+      Figure \ref{fig:prob-sb-success} shows stolen base success probability modeled as a function of lead distance, split by season and by runner skill.
 We observe that, conditioned on lead distance, even 10$^{th}$-percentile runners in 2023 had higher success probabilities than the median runner in 2022, which is likely attributable to rule changes making it easier to steal bases in 2023.

### 2025-12-24 articles/arxiv/paper.tex
clean up citations

@@
 \affil[1]{Department of Sport Management, Rice University}
-\affil[2]{Dept of Computational Applied Mathematics and Operations Research, Rice University}
+\affil[2]{Department of Computational Applied Mathematics and Operations Research, Rice University}
 \affil[3]{Department of Statistics, Rice University}
@@
 Our model captures the vital aspects of the play while retaining analytical and computational tractability.
-Our model is closely related to the class of stochastic/dynamic Stackelberg games \parencite{goktas2022zero,li2017review,lopez2022stationary,vasal2020stochastic,vorobeychik2012computing}, which combines the elements of Stackelberg and stochastic games.
+Our model is closely related to the class of stochastic/dynamic Stackelberg games \parencite{goktas_zero-sum_2022,li_review_2017,lopez_stationary_2023,vasal_sequential_2022,vorobeychik_computing_2012}, which combines the elements of Stackelberg and stochastic games.
 The main difference between our infinite-horizon model and the stochastic Stackelberg games literature is that we do not discount the rewards.
@@
     We note that fixing a pitcher policy reduces our two-player sequential game to a single-player MDP whose objective is to optimize the lead distance for the first-base runner.
-Optimality criterion similar to the two-player sequential game also exists for MDP with undiscounted rewards \parencite[Section 7.2]{bertsekas2012dynamic}.
+Optimality criterion similar to the two-player sequential game also exists for MDP with undiscounted rewards \parencite[Section 7.2]{bertsekas_dynamic_2012}.
 Numerical experiments using this decision  model for a specific pitcher policy are presented in Section \ref{sec:two-foot-rule}.
@@
        Under Assumption \ref{assum:game_halts}, we prove that the expected number of runs scored by the batter, $V^{\pi_\text{R},\pi_\text{P}}$, in  \eqref{eqn:sg_pol_val} is finite.
-Our proof is based on standard arguments used in the analyses of undiscounted problems \parencite[Section 7.2]{bertsekas2012dynamic}.
+Our proof is based on standard arguments used in the analyses of undiscounted problems \parencite[Section 7.2]{bertsekas_dynamic_2012}.
 We present it here for completeness.
@@
         We present optimality characterizations in Theorem \ref{lem:vi_convergence} and Propositions \ref{lem:bellman_equations} to  \ref{lem:policy_iteration}.
-These results and their proofs are similar to the undiscounted problem in the single-player case \parencite[Section 7.2]{bertsekas2012dynamic} and the zero-sum stochastic game with simultaneous moves \parencite{patek_stochastic_1999}.
+These results and their proofs are similar to the undiscounted problem in the single-player case \parencite[Section 7.2]{bertsekas_dynamic_2012} and the zero-sum stochastic game with simultaneous moves \parencite{patek_stochastic_1999}.
 We state and prove our results for completeness and to outline the minor differences.
@@
 The set $A_\text{R}(s)$ in \eqref{eqn:runner_action_sapce}   is compact.
-Hence, by Theorem 5 in Subsection X, Section 21 in  \parencite{kuratowski2014topology}, $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ converges to $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},a_\text{P})$ uniformly over $A_\text{R}(s)$.
+Hence, by Theorem 5 in Subsection X, Section 21 in  \parencite{kuratowski_topology_2014}, $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + u_{k-1}(s')\big]p(s'|s,a_\text{R},a_\text{P})$ converges to $\min_{a_\text{P} \in A_\text{P}(s,a_\text{R})}\sum_{s' \in {\cal S}}\big[r(s'|s) + V^*(s')\big]p(s'|s,a_\text{R},a_\text{P})$ uniformly over $A_\text{R}(s)$.
 

### 2025-12-24 articles/arxiv/paper.tex
fix typos

@@
 
-    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the to a fielder at the base to tag the runner out \parencite{mlb_baseball_2025}.
+    For background, the stolen base is an exciting play in which a runner takes off on the pitch, attempting to reach the next base before the catcher can throw the ball to a fielder at the base to tag the runner out \parencite{mlb_baseball_2025}.
 Whether attempting a stolen base or not, the runner takes a {\it lead} from their base, commonly positioning themselves 8--12 feet away from their base, toward the next base.
@@
 
-    Section \ref{sec:sg-model} uses a two-payer zero-sum sequential game to model the evolving interaction between the runner and the pitcher.
+    Section \ref{sec:sg-model} uses a two-player zero-sum sequential game to model the evolving interaction between the runner and the pitcher.
 Our model captures the vital aspects of the play while retaining analytical and computational tractability.
@@
 That is, $\pi_\text{R}^*$ is a policy for the batting team that maximizes the expected total runs scored by them, and $\pi_\text{P}^*$ is a policy for the fielding team that minimizes the expected total runs incurred by them.
-In other words, $\pi_\text{R}^*$ is a optimal policy for the batting team and $\pi_\text{P}^*$ is an optimal policy for the fielding team.
+In other words, $\pi_\text{R}^*$ is an optimal policy for the batting team and $\pi_\text{P}^*$ is an optimal policy for the fielding team.
 
@@
         & \mbox{S}^+  \mbox{ represents a successful stolen base attempt (runner is safe);}\\
-        & \mbox{S}^-  \mbox{ represents an usuccessful stolen base attempt (runner is out); and}\\
+        & \mbox{S}^-  \mbox{ represents an unsuccessful stolen base attempt (runner is out); and}\\
         & \mbox{N}~   \mbox{ represents no runner action (a pitch with no stolen base attempt).}
@@
         \end{align}
-        This model has eight fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$, and $\beta^L$; and the two variance parameters $\sigma^2_R$ and $\sigma^2_P$.
-The number of random effects is $|\mathcal{H}^R| + |\mathcal{H}^P|$.
+        This model has seven fixed, unknown parameters: the intercept $\alpha$; the five slopes $\beta^B$, $\beta^S$, $\beta^O$, $\beta^D$, and $\beta^L$; and the variance parameter $\sigma^2_P$.
+The number of random effects is $|\mathcal{H}^P|$.
 
@@
       \begin{table}
-        \small
+        \footnotesize
         \center
@@
 
-      \begin{table}[H]
+      \begin{table}
         \centering
@@
 
-      \begin{table}[H]
+      \begin{table}[b]
         \centering

### 2025-12-24 articles/arxiv/paper.tex
minor tweaks to allow arxiv submission

@@
 
-\graphicspath {{../../output/figures/}}
+\graphicspath {{figures/}}
+\newcommand{\tablepath}{tables/}
+\newcommand{\inputtable}[1]{\input{\tablepath #1}}
 
@@
         \center
-        \input{../../output/tables/model_summary.tex}
+        \inputtable{model_summary.tex}
         \caption{\it Summary of estimated runner outcome probability models.
@@
         \centering
-        \input{../../output/tables/lead_by_count_two_player.tex}
+        \inputtable{lead_by_count_two_player.tex}
         \caption{
@@
         \centering
-        \input{../../output/tables/lead_by_outs_two_player.tex}
+        \inputtable{lead_by_outs_two_player.tex}
         \caption{
@@
 
-      We observe that the lead distance generally (although not strictly) shortens as the count deepens (i.e., as the numbers of balls and strikes increase).
+      We observe that the lead distance generally, although not strictly, shortens as the count deepens (i.e., as the numbers of balls and strikes increase).
 This makes intuitive sense because the plate appearance is closer to ending, which relieves the pressure on the pitcher to avoid attempting a pickoff.
-The lead distance generally (although not strictly) increases as the number of outs increases.
+The lead distance generally, although not strictly, increases as the number of outs increases.
 This makes intuitive sense because the runner is less likely to score without more aggressive action.
@@
         \centering
-        \input{../../output/tables/lead_by_count_one_player.tex}
+        \inputtable{lead_by_count_one_player.tex}
         \caption{
@@
         \centering
-        \input{../../output/tables/lead_by_outs_one_player.tex}
+        \inputtable{lead_by_outs_one_player.tex}
         \caption{
@@
         \centering
-        \input{../../output/tables/lead_by_players.tex}
+        \inputtable{lead_by_players.tex}
         \caption{
@@
            \end{equation}
+           \endgroup
            Since $\pi_\text{R}$ is stationary, the best response of the pitcher corresponding to $\pi_\text{R}$ reduces to solving a MDP.

### 2025-12-24 articles/arxiv/paper.tex
fiddle with typesetting

@@
 \author[2]{Andrew J. Schaefer}
-\affil[1]{Department of Sport Management, Rice University}
-\affil[2]{Department of Computational Applied Mathematics and Operations Research, Rice University}
-\affil[3]{Department of Statistics, Rice University}
+\affil[1]{\small Department of Sport Management, Rice University}
+\affil[2]{\small Department of Computational Applied Mathematics and Operations Research, Rice University}
+\affil[3]{\small Department of Statistics, Rice University}
 
@@
       \begin{table}
-        \footnotesize
+        \small
         \center
@@
 
-      Using the transition probabilities estimated on the empirical data, we implemented the value iteration algorithm presented in Section \ref{sec:sg-model}, discretizing the runner's action space to the nearest tenth of a foot in lead distance.
-Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and prior disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and prior disengagements.
-This optimal lead distance policy maximizes runs scored while taking into account the rational behavior of both players, as defined by (\ref{eqn:sg_val}).
+      Using the transition probabilities estimated on the empirical data, we implemented the value iteration algorithm from Section \ref{sec:sg-model}, discretizing the runner's lead distance action space to the nearest tenth of a foot.
+Table \ref{tab:lead-by-count-two-player} reports how the optimal runner policy varies by count and disengagements, and Table \ref{tab:lead-by-outs-two-player} reports how the optimal runner policy varies by outs and disengagements.
+This runner policy maximizes runs scored while taking into account the rational behavior of both players, as defined by (\ref{eqn:sg_val}).
 For this runner policy, the pitcher is indifferent to attempting a pickoff or not.
@@
 
-      \begin{table}[b]
+      \begin{table}
         \centering
